---
title: "Data cleaning - Screening file"
author: "Paloma"
date: "2026-03-06"
output: workflowr::wflow_html
editor_options: 
  chunk_output_type: console
---

## Introduction

Data cleaning script for `01.SCREENING.csv`. The goal is to produce a clean,
analysis-ready dataset for the descriptive statistics table and regression
models. This script covers: duplicate merging, missing code assignment,
variable recoding, SES composite scoring, and a missingness summary.

Missing data coding scheme:

- 999 = Missing / lost / data entry error
- 333 = Pending (data exists but not yet entered)  
- 888 = Don't know  
- 777 = Refused / prefer not to answer  
- 666 = Illegible  
- 555 = Other (answer outside allowed categories)  
- 444 = Ambiguous  

No = 0, Yes = 1, Occasionally = 2 (unless variable-specific rules apply).
Sentinel codes are preserved in the master file and converted to NA only
within analysis scripts.

*Important modifications to include in manuscript*

- Age range was modified to 18-46 to include three participants
- Rainy season is 0, dry season is 1.

**Permanently excluded participants (removed from all databases and analyses):**

The following four participants are excluded at the start of this script and must
not appear in any analysis, table, or model:

| ID  | Reason |
|-----|--------|
| 10  | Does not meet the inclusion criteria: participant is 49 years old (eligibility range: 18–46). |
| 141 | Participant did not continuously reside in the study neighborhood; she moved between different neighborhoods during the study period, violating the residential stability criterion. |
| 164 | This ID number was skipped during data collection and was never assigned to a participant; it does not exist in the dataset. |
| 427 | Participant appeared not to understand the survey questions, suggesting a linguistic or cognitive limitation that compromises data validity. |

---

# Setup

```{r setup, message=FALSE, warning=FALSE, echo = FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(ggplot2)

data_path <- "./data"      # cleaned output
raw_path  <- "./data"      # raw input

MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

# IDs permanently excluded from all analyses — see Introduction for rationale
EXCLUDED_IDS <- c(10L, 141L, 164L, 427L)
```

---

# 1. Load raw data

```{r load, echo = FALSE}
df_raw <- read.csv(
  file.path(raw_path, "01.SCREENING.csv"),
  stringsAsFactors = FALSE,
  na.strings = c("")   # only truly blank cells become NA at read time
)

cat("Raw dimensions:", dim(df_raw), "\n")
cat("Unique IDs in raw file:", length(unique(df_raw$ID)), "\n")
cat("Duplicate IDs:", sum(duplicated(df_raw$ID)), "\n")
```

---

# 2. Standardize text before merging

Convert known missing-text patterns to NA so they don't interfere with
the merge logic. All other values are preserved.

```{r standardize-text}
na_strings <- c("N/A", "n/a", "NA", "na",
                "No hay niguna repuesta", "No hay ninguna repuesta")

# "pending" means data exists but was not entered — code as 333, not NA
pending_strings <- c("pending", "Pending", "PENDING")

df_raw <- df_raw %>%
  mutate(across(where(is.character), ~ {
    x <- str_trim(.)
    case_when(
      x %in% na_strings      ~ NA_character_,
      x %in% pending_strings ~ "333",
      TRUE                   ~ x
    )
  }))
```

---

## 2.1 Preserve raw values before any recoding

Raw copies are created here, before merging or recoding, so that
`_RAW` columns always reflect the original data entry.

```{r preserve-raw}
vars_to_preserve <- c("HLTH_TTM", "HLTH_MED", "HLTH_CPAIN",
                      "HLTH_CDIS", "HLTH_SMK", "D_EMP_TYPE", "D_CHLD", 
                      "SES_EDU_CAT")

df_raw <- df_raw %>%
  mutate(across(all_of(vars_to_preserve),
                ~ ., .names = "{.col}_RAW"))
```

---

# 3. Merge duplicate rows

For IDs with two rows, take the non-NA value when rows differ.
If both rows have different non-NA values for the same field, the
conflict is flagged with 444 (Ambiguous) so it can be reviewed manually.

```{r merge-duplicates}
merge_rows <- function(df_group) {
  if (nrow(df_group) == 1) return(df_group)
  
  merged <- df_group[1, , drop = FALSE]  # start with first row as template
  
  # Skip ID — it is the same by definition
  skip_cols <- c("ID")
  
  for (col in setdiff(names(df_group), skip_cols)) {
    v1 <- df_group[[col]][1]
    v2 <- df_group[[col]][2]
    
    both_na   <- is.na(v1) & is.na(v2)
    only_v2   <- is.na(v1) & !is.na(v2)
    both_same <- !is.na(v1) & !is.na(v2) &
                 (str_trim(tolower(as.character(v1))) ==
                  str_trim(tolower(as.character(v2))))
    conflict  <- !is.na(v1) & !is.na(v2) &
                 (str_trim(tolower(as.character(v1))) !=
                  str_trim(tolower(as.character(v2))))
    
    if (both_na)        { merged[[col]] <- NA }
    else if (only_v2)   { merged[[col]] <- v2 }
    else if (both_same) { merged[[col]] <- v1 }
    else if (conflict)  {
      # Free-text/notes columns: concatenate instead of flagging
      if (col %in% c("D_LOC_INFO", "D_HH_SIZE_INFO", "D_INFO",
                     "SES_INFO", "HLTH_CPAIN_CAT", "HLTH_CDIS_CAT")) {
        merged[[col]] <- paste(v1, v2, sep = " | ")
      } else {
        merged[[col]] <- "444"   # Ambiguous — review manually
      }
    }
  }
  return(merged)
}

df <- df_raw %>%
  group_by(ID) %>%
  group_modify(~ merge_rows(.x)) %>%
  ungroup()

cat("Rows after merging duplicates:", nrow(df), "\n")

# Report which fields were flagged as ambiguous
ambig_flags <- df %>%
  select(-ID) %>%
  summarise(across(everything(), ~ sum(. == "444", na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_ambiguous") %>%
  filter(n_ambiguous > 0)

cat("\nFields with ambiguous (444) values after merge:\n")
print(ambig_flags, n=nrow(ambig_flags))

# Print ID and conflicting values for manual review
for (v in ambig_flags$variable) {
  cat("\n---", v, "---\n")
  conflict_ids <- df %>%
    filter(.data[[v]] == "444") %>%
    pull(ID)
  print(df_raw %>%
    filter(ID %in% conflict_ids) %>%
    select(ID, all_of(v)))
}
```

## 3.1. Export conflicts for manual resolution

```{r export-conflicts}
conflict_rows <- list()

for (v in ambig_flags$variable) {
  conflict_ids <- df %>%
    filter(.data[[v]] == "444") %>%
    pull(ID)
  
  tmp <- df_raw %>%
    filter(ID %in% conflict_ids) %>%
    select(ID, all_of(v)) %>%
    mutate(across(c(ID, all_of(v)), as.character)) %>%
    group_by(ID) %>%
    summarise(value_1 = first(.data[[v]]),
              value_2 = last(.data[[v]]),
              .groups = "drop") %>%
    mutate(variable = v, resolution = NA_character_)
  
  conflict_rows[[v]] <- tmp
}

conflicts <- bind_rows(conflict_rows) %>%
  select(ID, variable, value_1, value_2, resolution)

# Uncomment once ready to export:
 write.csv(conflicts, file.path(data_path, "01.conflicts_to_review.csv"),
           row.names = FALSE)
 
 head(conflicts)
```

## 3.2. Incorporate resolutions back into dataset and check again

```{r apply-resolutions}
resolved <- read.csv(file.path(data_path, "01.conflicts_reviewed.csv"),
                     stringsAsFactors = FALSE)

for (i in seq_len(nrow(resolved))) {
  if (is.na(resolved$resolution[i])) next  # skip unresolved conflicts
  id  <- resolved$ID[i]
  v   <- resolved$variable[i]
  val <- resolved$resolution[i]
  df[df$ID == id, v] <- val
}

# Report which fields were flagged as ambiguous
ambig_flags <- df %>%
  select(-ID) %>%
  summarise(across(everything(), ~ sum(. == "444", na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_ambiguous") %>%
  filter(n_ambiguous > 0)

cat("\nFields with ambiguous (444) values after merge:\n")
print(ambig_flags, n=nrow(ambig_flags))

# Print ID and conflicting values for manual review
for (v in ambig_flags$variable) {
  cat("\n---", v, "---\n")
  conflict_ids <- df %>%
    filter(.data[[v]] == "444") %>%
    pull(ID)
  print(df_raw %>%
    filter(ID %in% conflict_ids) %>%
    select(ID, all_of(v)))
}
conflict_rows <- list()

for (v in ambig_flags$variable) {
  conflict_ids <- df %>%
    filter(.data[[v]] == "444") %>%
    pull(ID)
  
  tmp <- df_raw %>%
    filter(ID %in% conflict_ids) %>%
    select(ID, all_of(v)) %>%
    mutate(across(c(ID, all_of(v)), as.character)) %>%
    group_by(ID) %>%
    summarise(value_1 = first(.data[[v]]),
              value_2 = last(.data[[v]]),
              .groups = "drop") %>%
    mutate(variable = v, resolution = NA_character_)
  
  conflict_rows[[v]] <- tmp
}

conflicts <- bind_rows(conflict_rows) %>%
  select(ID, variable, value_1, value_2, resolution)

# Uncomment once ready to export:
 write.csv(conflicts, file.path(data_path, "01.conflicts_to_review_V2.csv"),
           row.names = FALSE)
 
 head(conflicts)



```

## 3.3 Incorporate more resolutions back into dataset 


```{r}
resolved <- read.csv(file.path(data_path, "01.conflicts_reviewed_V2.csv"),
                     stringsAsFactors = FALSE)

for (i in seq_len(nrow(resolved))) {
  if (is.na(resolved$resolution[i])) next  # skip unresolved conflicts
  id  <- resolved$ID[i]
  v   <- resolved$variable[i]
  val <- resolved$resolution[i]
  df[df$ID == id, v] <- val
}

table(df$HLTH_MED_RAW)
```

---

# 3.4. Exclude ineligible and invalid participants

Remove the four permanently excluded IDs before any further processing.
See the Introduction for the full rationale for each exclusion.

```{r exclude-participants}
df <- df %>% filter(!ID %in% EXCLUDED_IDS)

cat("Excluded IDs:", paste(EXCLUDED_IDS, collapse = ", "), "\n")
cat("Rows remaining after exclusion:", nrow(df), "\n")
```

---

# 4. Participant ID checks

```{r id-checks}
cat("Total unique participants:", nrow(df), "\n")

# Season 1: IDs 1-204 (rainy, Oct-Dec 2022)
# Season 2: IDs 301-497 (dry, Apr-Jun 2023)
missing_s1 <- setdiff(1:204, df$ID)
missing_s2 <- setdiff(301:497, df$ID)

cat("Missing IDs in Season 1 (1-204):",
    ifelse(length(missing_s1) == 0, "none",
           paste(missing_s1, collapse = ", ")), "\n")
cat("Missing IDs in Season 2 (301-497):",
    ifelse(length(missing_s2) == 0, "none",
           paste(missing_s2, collapse = ", ")), "\n")

# Age check: eligibility was 18-46
out_of_range_age <- df %>%
  filter(!is.na(D_AGE) & (as.numeric(D_AGE) < 18 | as.numeric(D_AGE) > 46)) %>%
  select(ID, D_YRBR, D_AGE)

cat("\nParticipants outside age eligibility range (18-46):\n")
print(out_of_range_age)

# Flag all non-NA data for out-of-range participants as 999
# NA cells are left as NA (not converted to 999)
out_of_range_ids <- out_of_range_age$ID

df <- df %>%
  mutate(across(-ID, ~ {
    x <- as.character(.)
    ifelse(ID %in% out_of_range_ids & !is.na(x), "999", x)
  }))
```

---

# 5. Assign season variable

```{r season}
df <- df %>%
  mutate(SEASON = as.factor(case_when(
    ID >= 1   & ID <= 204 ~ 0,   # 0 = rainy (reference in models)
    ID >= 301 & ID <= 497 ~ 1,   # 1 = dry
    TRUE ~ NA_real_
  )))

cat("Season distribution:\n")
print(table(df$SEASON, useNA = "always"))
```

---

# 6. Recode binary Yes/No variables

No = 0, Yes = 1. Responses with extra text (e.g., "Yes, Loratadina") are
recoded as 1 since the full text is captured in the corresponding _RAW column.
Sentinel codes are passed through unchanged.

```{r recode-binary}
recode_yesno <- function(x) {
  x <- str_trim(tolower(as.character(x)))
  case_when(
    is.na(x)                                          ~ NA_real_,
    x %in% as.character(MISSING_CODES)               ~ suppressWarnings(as.numeric(x)),
    x %in% c("no", "No", "no.", "no -div")                 ~ 0,
    str_starts(x, "yes") | str_starts(x, "si") | str_starts(x, "Dolor")  ~ 1,
    x %in% c("no sabe", "ns")                        ~ 888,
    x %in% c("refused")               ~ 777,
    TRUE                                              ~ 555
  )
}

binary_vars <- c("HLTH_TTM", "HLTH_MED", "HLTH_CPAIN", "HLTH_CDIS")

df <- df %>%
  mutate(across(all_of(binary_vars),
                recode_yesno, .names = "{.col}_CAT"))

# Visualization: RAW vs CLEAN for all binary variables
binary_long <- df %>%
  select(ID,
         all_of(paste0(binary_vars, "_RAW")),
         all_of(paste0(binary_vars))) %>%
  mutate(across(everything(), as.character)) 

```


```{r inspection}
df[,c("HLTH_MED", "HLTH_MED_RAW", "HLTH_MED_CAT")]
df[,c("HLTH_TTM", "HLTH_TTM_RAW", "HLTH_TTM_CAT")]
df[,c("HLTH_CPAIN", "HLTH_CPAIN_RAW", "HLTH_CPAIN_CAT")]
df[,c("HLTH_CDIS", "HLTH_CDIS_RAW", "HLTH_CDIS_CAT")]

table(df$HLTH_CPAIN_CAT)
table(df$HLTH_TTM_CAT)
temp <- df %>%
  select("HLTH_CPAIN_CAT", "HLTH_MED_CAT", "HLTH_TTM_CAT", "HLTH_CDIS_CAT")

temp2 <- temp %>%
  pivot_longer(everything(),
               names_to = "cat", 
               values_to = "vals") #%>% 
 # group_by(cat, vals) %>%
#  summarize(n = n())
ggplot(temp2, aes(x = cat, fill = as.factor(vals))) +
  geom_bar(position = "dodge")
```


---

# 7. Recode HLTH_SMK (smoking) → 3 categories

0 = No, 1 = Yes/Daily, 2 = Occasionally.
Sentinel codes are passed through unchanged.

```{r recode-smoking}
df <- df %>%
  mutate(
    HLTH_SMK_CLEAN = case_when(
      is.na(HLTH_SMK)                                                              ~ NA_real_,
      str_trim(tolower(as.character(HLTH_SMK))) %in% as.character(MISSING_CODES)  ~ suppressWarnings(as.numeric(HLTH_SMK)),
      str_trim(tolower(HLTH_SMK)) %in% c("no", "ya no- 10 atras", "fumaba", "no, hace 2 anios.")                                     ~ 0,
      str_trim(tolower(HLTH_SMK)) %in% c(
        "de vez en cuando", "a veces", "aveces",
        "casi nada", "3 al mes",
        "si de vez en cuando", "1 cada 3 dia",
        "a veces/ una vez por semana", "1 al mes","ocasionalmente",
        "ocasionalmente (1 vez x semana)", "c/6 meses", "a vez al anio", 
        "5 x mes", "a veces, 2-3 x semana", "casi nada (1 c/2 meses)", "1 c/ 2 o 3 meses")                                             ~ 2,
        str_starts(tolower(str_trim(HLTH_SMK)), "yes") |
        str_starts(tolower(str_trim(HLTH_SMK)), "si") |
        str_starts(tolower(str_trim(HLTH_SMK)), "1 al dia") |      str_starts(tolower(str_trim(HLTH_SMK)), "poquito- 3 x semana") |  str_starts(tolower(str_trim(HLTH_SMK)), "3 x semana") |
        str_starts(tolower(str_trim(HLTH_SMK)), "7 x dia")|
        str_starts(tolower(str_trim(HLTH_SMK)), "1 diario") ~ 1,
      TRUE                                                                         ~ 555
    ),
    HLTH_SMK_CAT = factor(HLTH_SMK_CLEAN,
                          levels = c(0, 1, 2, 555),
                          labels = c("No", "Yes", "Occasionally", "Other"))
  )

cat("Smoking distribution:\n")
print(table(df$HLTH_SMK_CAT, useNA = "always"))

cat("\nUnclassified smoking responses (555):\n")
df %>% filter(HLTH_SMK_CLEAN == 555) %>%
  select(ID, HLTH_SMK_RAW, HLTH_SMK_CLEAN) %>% print()

# Visualization: RAW vs CLEAN
smk_long <- df %>%
  select(ID, HLTH_SMK_RAW, HLTH_SMK_CLEAN) %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(-ID, names_to = "version", values_to = "value") %>%
  mutate(version = str_remove(version, "HLTH_SMK_"))

barplot(table(df$HLTH_SMK_CLEAN))
```

---

# 8. Recode D_EMP_TYPE → two variables

D_EMP_CLEAN: 0 = Not employed, 1 = Employed
D_EMP_FORM_CLEAN: 0 = Informal, 1 = Formal (NA if not employed)
Sentinel codes are passed through unchanged.

```{r recode-employment}

df <- df %>%
  mutate(
    D_EMP_TYPE_CLEAN = str_trim(tolower(D_EMP_TYPE)),
    D_EMP_CLEAN = case_when(
      is.na(D_EMP_TYPE_CLEAN)                                              ~ NA_real_,
      D_EMP_TYPE_CLEAN %in% as.character(MISSING_CODES)                   ~ suppressWarnings(as.numeric(D_EMP_TYPE_CLEAN)),
      D_EMP_TYPE_CLEAN %in% c("no", "no.", "niguna", "nada", "ninguna","no. ama de casa", "ama de casa", "ama de cas")   ~ 0,
      str_detect(D_EMP_TYPE_CLEAN, "no\\b") &
        !str_detect(D_EMP_TYPE_CLEAN,
                    "informal|formal|trabaj|vend|comerci")                 ~ 0,
      str_detect(D_EMP_TYPE_CLEAN,
                 "yes|si|formal|ayudante|tiene un|venta|temporal|informal|pestanias|comerci|vend|levado de selas|trabaj|administrat|limpieza|estilista|Taller de cortura|comerante") ~ 1,
      TRUE                                                                 ~ 555
    ),
    D_EMP_FORM_CLEAN = case_when(
      is.na(D_EMP_CLEAN) | D_EMP_CLEAN != 1                               ~ NA_real_,
      D_EMP_TYPE_CLEAN %in% as.character(MISSING_CODES)                   ~ suppressWarnings(as.numeric(D_EMP_TYPE_CLEAN)),
      str_detect(D_EMP_TYPE_CLEAN, "informal")                            ~ 0,
      str_detect(D_EMP_TYPE_CLEAN, "\\bformal\\b")                        ~ 1,
      str_detect(D_EMP_TYPE_CLEAN,
                 "vend|comerci|limpieza|chocolater|catalogo|ropa|dulces|postres|pintura|facebook") ~ 0,
      TRUE                                                                 ~ 555
    )
  )

print(df[,c("ID", "D_EMP_CLEAN","D_EMP_TYPE", "D_EMP_TYPE_CLEAN", "D_EMP_FORM_CLEAN")], n= nrow(df))
cat("\nEmployment (D_EMP_CLEAN):\n")
print(table(df$D_EMP_CLEAN, useNA = "always"))

cat("\nFormality (D_EMP_FORM_CLEAN, among employed):\n")
print(table(df$D_EMP_FORM_CLEAN, useNA = "always"))

cat("\nUnclassified employment responses (555):\n")
df %>%
  filter(D_EMP_CLEAN == 555 | D_EMP_FORM_CLEAN == 555) %>%
  select(ID, D_EMP_TYPE_RAW, D_EMP_CLEAN, D_EMP_FORM_CLEAN) %>%
  print()

# Visualization: RAW vs CLEAN for both employment variables
emp_long <- df %>%
  select(ID, D_EMP_CLEAN, D_EMP_FORM_CLEAN) %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(-ID, names_to = "variable", values_to = "value") %>%
  mutate(
    version  = if_else(variable == "D_EMP_TYPE_RAW", "RAW", "CLEAN"),
    variable = case_when(
     # variable == "D_EMP_TYPE_RAW"   ~ "D_EMP",
      variable == "D_EMP_CLEAN"      ~ "D_EMP",
      variable == "D_EMP_FORM_CLEAN" ~ "D_EMP_FORM"
    )
  )


ggplot(emp_long, aes(x = value, fill = value)) +
  geom_bar(position = "dodge") +
  facet_wrap(~ variable, scales = "free_x") +
  labs(title = "Employment variables distribution",
       x = "Value", y = "Count", fill = "Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---

# 9. Recode D_CHLD (parity)

Expected: integer 0-4. Non-numeric entries are flagged.
Sentinel codes are passed through unchanged.

```{r recode-parity}
df <- df %>%
  mutate(
    D_CHLD_CLEAN = case_when(
      is.na(D_CHLD)                                                          ~ NA_real_,
      str_trim(as.character(D_CHLD)) %in% as.character(MISSING_CODES)       ~ suppressWarnings(as.numeric(D_CHLD)),
      str_detect(tolower(as.character(D_CHLD)), "no hay nota")               ~ 444,
      str_detect(tolower(as.character(D_CHLD)), "^si")                       ~ 444,
      !is.na(suppressWarnings(as.numeric(D_CHLD)))                           ~ suppressWarnings(as.numeric(D_CHLD)),
      TRUE                                                                   ~ 555
    )
  )

cat("Parity distribution (D_CHLD_CLEAN):\n")
print(table(df$D_CHLD_CLEAN, useNA = "always"))

cat("\nFlagged parity values:\n")
df %>%
  filter(D_CHLD_CLEAN %in% c(333, 444, 555)) %>%
  select(ID, D_CHLD_RAW, D_CHLD_CLEAN) %>%
  print()

# Visualization: RAW vs CLEAN
chld_long <- df %>%
  select(ID, D_CHLD_RAW, D_CHLD_CLEAN) %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(-ID, names_to = "version", values_to = "value") %>%
  mutate(version = str_remove(version, "D_CHLD_"))

ggplot(chld_long, aes(x = value, fill = version)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("RAW" = "#d9534f", "CLEAN" = "#5bc0de")) +
  labs(title = "D_CHLD (parity): RAW vs CLEAN value distribution",
       x = "Value", y = "Count", fill = "Version") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---

# 10. Clean D_LOC_TIME (years in neighborhood)

```{r clean-loc-time}
df <- df %>%
  mutate(
    D_LOC_TIME_RAW   = D_LOC_TIME,
    D_LOC_TIME_CLEAN = case_when(
      is.na(D_LOC_TIME)                                                         ~ NA_real_,
      str_trim(as.character(D_LOC_TIME)) %in% as.character(MISSING_CODES)      ~ suppressWarnings(as.numeric(D_LOC_TIME)),
      !is.na(suppressWarnings(as.numeric(D_LOC_TIME)))                          ~ suppressWarnings(as.numeric(D_LOC_TIME)),
      TRUE                                                                      ~ 555
    )
  )

cat("D_LOC_TIME values flagged as 555 (non-numeric text):\n")
df %>%
  filter(D_LOC_TIME_CLEAN == 555) %>%
  select(ID, D_LOC_TIME_RAW, D_LOC_TIME_CLEAN) %>%
  print()
```

---

# 11. Compute SES composite score (NSE-AMAI)

Items: SES_EDU_SC, SES_BTHR_SC, SES_CAR_SC, SES_INT_SC, SES_WRK_SC, SES_BEDR_SC
Non-numeric entries (e.g. column-shift errors) are flagged as 666.

## 11.1. Recode SES_EDU_CAT
```{r recode-edu}

df <- df %>%
  mutate(
    SES_EDU_CAT_CLEAN = case_when(
      
      # ── Sentinel pass-through ──────────────────────────────────────────
      SES_EDU_CAT == "999" ~ "999",
      
      # ── Ambiguous / needs review ───────────────────────────────────────
      SES_EDU_CAT == "No hay"                                          ~ as.character(999),
      SES_EDU_CAT == "No hay una clara marca, necesita ver la grabacion" ~ as.character(444),
      SES_EDU_CAT == "no sabe si es completa o no para primaria"        ~ as.character(888),
      
      # ── Standard categories (case-insensitive via str_to_lower trick) ──
      SES_EDU_CAT == "no estudio/preescolar"        ~ "No formal education",
      SES_EDU_CAT == "Primaria incompleta"           ~ "Incomplete primary",
      SES_EDU_CAT == "Primaria completa"             ~ "Complete primary",
      SES_EDU_CAT == "Secundaria incompleta"         ~ "Incomplete secondary",
      SES_EDU_CAT %in% c(
        "Secundaria completa",
        "Secundaria completa. Tomo un per de curso (secretaicido y computear) sin concluir"
      )                                              ~ "Complete secondary",
      
      # ── Carrera comercial → Secundaria Completa ────────────────────────
      SES_EDU_CAT == "Carrera comercial"             ~ "Complete secondary",
      
      # ── Carrera tecnica (without bachillerato) → Preparatoria Incompleta
      SES_EDU_CAT %in% c(
        "Carrera tecnica",
        "Carrera tecnica completa",
        "carrera tecnica completa --> 35 puntos (for now: 3 march 2023)",
        "Carrera Tecnica: 35 pints ( 3 March 23)"
      )                                              ~ "Incomplete high school",
      
      # ── Carrera tecnica WITH bachillerato → Preparatoria Completa ──────
      SES_EDU_CAT %in% c(
        "carrera tecnica bachi",
        "Bachillerato, carrera tecnica completa. 35 (3 march 2023)"
      )                                              ~ "Complete high school",
      
      # ── Bachillerato ───────────────────────────────────────────────────
      SES_EDU_CAT == "Bachillerato trunco"           ~ "Incomplete high school",
      SES_EDU_CAT %in% c(
        "Bachillerato completo",
        "Bauhi? Completo"                            # likely "Bachillerato Completo"
      )                                              ~ "Complete high school",
      
      SES_EDU_CAT == "Preparatoria incompleta"       ~ "Incomplete high school",
      SES_EDU_CAT == "Preparatoria completa"         ~ "Complete high school",
      SES_EDU_CAT == "Licenciatura incompleta"       ~ "Incomplete university",
      SES_EDU_CAT == "Licenciatura completa"         ~ "Complete university",
      SES_EDU_CAT == "Postgrado"                     ~ "Postgraduate",
      
      # ── Fallback ───────────────────────────────────────────────────────
      TRUE ~ as.character(444)  # anything unmatched → flag for review
    ),
    
    # ── SES_EDU_SC update ─────────────────────────────────────────────────
    SES_EDU_SC_CLEAN = case_when(
      SES_EDU_CAT_CLEAN == "No formal education"       ~ 0,
      SES_EDU_CAT_CLEAN == "Incomplete primary"   ~ 10,
      SES_EDU_CAT_CLEAN == "Complete primary"     ~ 22,
      SES_EDU_CAT_CLEAN == "Incomplete secondary" ~ 23,
      SES_EDU_CAT_CLEAN == "Complete secondary"   ~ 31,
      SES_EDU_CAT_CLEAN == "Incomplete high school" ~ 35,
      SES_EDU_CAT_CLEAN == "Complete high school"   ~ 43,
      SES_EDU_CAT_CLEAN == "Incomplete university" ~ 59,
      SES_EDU_CAT_CLEAN == "Complete university"   ~ 73,
      SES_EDU_CAT_CLEAN == "Postgraduate"                ~ 101,
      TRUE ~ NA_real_   # sentinels and 444 get NA in the score column
    )
  )

# Check results
table(df$SES_EDU_CAT_CLEAN, useNA = "ifany")
df$SES_EDU_SC
edu_levels <- c(
  "No formal education",
  "Incomplete primary",
  "Complete primary",
  "Incomplete secondary",
  "Complete secondary",
  "Incomplete high school",
  "Complete high school",
  "Incomplete university",
  "Complete university",
  "Postgraduate"
)

df <- df %>%
  mutate(SES_EDU_CAT_CLEAN = factor(SES_EDU_CAT_CLEAN, levels = edu_levels, ordered = TRUE))

# Review flagged cases
df %>% filter(SES_EDU_CAT_CLEAN == "444") %>% 
  select(ID, SES_EDU_CAT, SES_EDU_CAT_CLEAN)

barplot(table(df$SES_EDU_CAT_CLEAN))
```

```{r ses-score}
ses_items <- c("SES_EDU_SC_CLEAN", "SES_BTHR_SC", "SES_CAR_SC",
               "SES_INT_SC", "SES_WRK_SC", "SES_BEDR_SC")

# Preserve originals
ses_raw <- df[ses_items]
names(ses_raw) <- paste0(ses_items, "_RAW")
df <- bind_cols(df, ses_raw)

# Convert to numeric; flag non-numeric entries as 666
df[ses_items] <- lapply(ses_items, function(col) {
  x     <- df[[col]]
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(!is.na(x) & is.na(x_num), 666, x_num)
})

# Compute total; NA if any item is missing or flagged
df <- df %>%
  rowwise() %>%
  mutate(
    SES_SC_Total = if_else(
      any(is.na(c_across(all_of(ses_items)))) |
        any(c_across(all_of(ses_items)) %in% MISSING_CODES),
      NA_real_,
      sum(c_across(all_of(ses_items)), na.rm = FALSE)
    )
  ) %>%
  ungroup()

# NSE-AMAI classification
df <- df %>%
  mutate(SES_CAT_AMAI = case_when(
    is.na(SES_SC_Total) ~ NA_character_,
    SES_SC_Total >= 205 ~ "AB",
    SES_SC_Total >= 166 ~ "C+",
    SES_SC_Total >= 135 ~ "C",
    SES_SC_Total >= 112 ~ "C-",
    SES_SC_Total >= 90  ~ "D+",
    SES_SC_Total >= 48  ~ "D-",
    TRUE                ~ "E"
  ))

cat("SES_SC_Total NAs:", sum(is.na(df$SES_SC_Total)), "\n")
SES_cat <- grep("^SES.*CAT$", names(df), value = TRUE)
cat("Participants with missing or flagged SES items:\n")
print(df[is.na(df$SES_SC_Total),
         c("ID", ses_items, paste0(ses_items, "_RAW"), SES_cat)])

# --- Long-format review file for student consensus -------------------------
# One row per (participant × SES item) that has an issue.
# The student fills in the `consensus` column with the correct numeric score,
# saves the file as 01.SES_consensus.csv, and section 11.2 applies it.

ses_scoring_hints <- c(
  SES_EDU_SC_CLEAN = "0=No formal, 10=Inc.primary, 22=Comp.primary, 23=Inc.secondary, 31=Comp.secondary, 35=Inc.high school, 43=Comp.high school, 59=Inc.university, 73=Comp.university, 101=Postgraduate",
  SES_BTHR_SC      = "Bathrooms score — see NSE-AMAI table",
  SES_CAR_SC       = "Cars score — see NSE-AMAI table",
  SES_INT_SC       = "Internet access score — see NSE-AMAI table",
  SES_WRK_SC       = "Primary earner education score — see NSE-AMAI table",
  SES_BEDR_SC      = "Bedrooms score — see NSE-AMAI table"
)

ses_review_ids <- df %>% filter(is.na(SES_SC_Total)) %>% pull(ID)

ses_long_clean <- df %>%
  filter(ID %in% ses_review_ids) %>%
  select(ID, all_of(ses_items)) %>%
  mutate(across(-ID, as.character)) %>%          # unify types before pivoting
  pivot_longer(cols = -ID, names_to = "variable", values_to = "current_value")

ses_long_raw <- df %>%
  filter(ID %in% ses_review_ids) %>%
  select(ID, all_of(paste0(ses_items, "_RAW"))) %>%
  mutate(across(-ID, as.character)) %>%          # unify types before pivoting
  pivot_longer(cols = -ID, names_to = "raw_variable", values_to = "raw_value") %>%
  mutate(variable = str_remove(raw_variable, "_RAW$")) %>%
  select(ID, variable, raw_value)

ses_to_review <- ses_long_clean %>%
  left_join(ses_long_raw, by = c("ID", "variable")) %>%
  mutate(
    issue = case_when(
      is.na(current_value)          ~ "Missing / not entered",
      current_value == "444"        ~ "Ambiguous — needs review",
      current_value == "999"        ~ "Lost / data entry error",
      current_value == "333"        ~ "Pending — data exists but not entered",
      current_value == "888"        ~ "Don't know",
      current_value == "777"        ~ "Refused",
      current_value == "666"        ~ "Illegible / column-shift error",
      current_value == "555"        ~ "Other — outside allowed categories",
      TRUE                          ~ NA_character_
    ),
    scoring_hint = ses_scoring_hints[variable],
    consensus    = NA_character_
  ) %>%
  filter(!is.na(issue)) %>%
  select(ID, variable, raw_value, current_value, issue, scoring_hint, consensus)

write.csv(ses_to_review,
          file.path(data_path, "01.SES_to_review.csv"),
          row.names = FALSE)

cat("Rows exported to 01.SES_to_review.csv:", nrow(ses_to_review), "\n")
print(ses_to_review)

cat("\nSES category distribution:\n")
print(table(df$SES_CAT_AMAI, useNA = "always"))
```

## 11.2. Apply SES consensus and recompute scores

Fill in the `consensus` column in `01.SES_to_review.csv` with the correct
numeric score for each flagged item, then save the file as
`01.SES_consensus.csv` and run this chunk to update the dataset.

```{r ses-apply-consensus}
ses_consensus_path <- file.path(data_path, "01.SES_consensus.csv")

if (!file.exists(ses_consensus_path)) {

  cat("Consensus file not found:", ses_consensus_path, "\n",
      "Fill in the 'consensus' column in 01.SES_to_review.csv,",
      "save as 01.SES_consensus.csv, then re-run this chunk.\n")

} else {

  consensus <- read.csv(ses_consensus_path, stringsAsFactors = FALSE)

  n_applied <- 0
  for (i in seq_len(nrow(consensus))) {
    if (is.na(consensus$consensus[i]) || consensus$consensus[i] == "") next
    id  <- consensus$ID[i]
    v   <- consensus$variable[i]
    val <- suppressWarnings(as.numeric(consensus$consensus[i]))
    if (!is.na(val)) {
      df[df$ID == id, v] <- val
      n_applied <- n_applied + 1
    } else {
      cat("Non-numeric consensus value skipped — ID:", id,
          "variable:", v, "value:", consensus$consensus[i], "\n")
    }
  }
  cat("Consensus values applied:", n_applied, "\n")

  # Recompute SES total and AMAI category
  df <- df %>%
    rowwise() %>%
    mutate(
      SES_SC_Total = if_else(
        any(is.na(c_across(all_of(ses_items)))) |
          any(c_across(all_of(ses_items)) %in% MISSING_CODES),
        NA_real_,
        sum(c_across(all_of(ses_items)), na.rm = FALSE)
      )
    ) %>%
    ungroup() %>%
    mutate(SES_CAT_AMAI = case_when(
      is.na(SES_SC_Total) ~ NA_character_,
      SES_SC_Total >= 205 ~ "AB",
      SES_SC_Total >= 166 ~ "C+",
      SES_SC_Total >= 135 ~ "C",
      SES_SC_Total >= 112 ~ "C-",
      SES_SC_Total >= 90  ~ "D+",
      SES_SC_Total >= 48  ~ "D-",
      TRUE                ~ "E"
    ))

  cat("SES_SC_Total NAs after consensus update:", sum(is.na(df$SES_SC_Total)), "\n")
  cat("\nSES category distribution after update:\n")
  print(table(df$SES_CAT_AMAI, useNA = "always"))

  # Report any items still unresolved
  ses_still_missing <- df %>%
    filter(is.na(SES_SC_Total)) %>%
    select(ID, all_of(ses_items))

  if (nrow(ses_still_missing) > 0) {
    cat("\nParticipants still missing SES total after consensus update:\n")
    print(ses_still_missing)
  } else {
    cat("\nAll SES totals resolved.\n")
  }
}
```

---

# 12. Missing data summary

```{r missing-summary}
# Exclude free-text notes columns and intermediate working columns
notes_cols <- c("D_LOC_INFO", "D_HH_SIZE_INFO", "D_INFO", "SES_INFO",
                "HLTH_CPAIN_CAT", "HLTH_CDIS_CAT", "D_EMP_DAYS",
                "W_HH_SIZE", "W_STORAGE", "D_LOC_CODE",
                "SES_EDU_SC", "SES_EDU_RAW")

analytic_vars <- setdiff(names(df), notes_cols)

missing_report <- df %>%
  select(all_of(analytic_vars)) %>%
  summarise(across(everything(), list(
    n_na       = ~ sum(is.na(.)),
    n_sentinel = ~ sum(. %in% as.character(MISSING_CODES), na.rm = TRUE)
  ))) %>%
  pivot_longer(everything(),
               names_to  = c("variable", ".value"),
               names_sep = "_(?=n_na|n_sentinel)") %>%
  mutate(
    n_total  = nrow(df),
    pct_na   = round(n_na / n_total * 100, 1),
    pct_flag = round(n_sentinel / n_total * 100, 1)
  ) %>%
  filter(n_na > 0 | n_sentinel > 0) %>%
  arrange(desc(pct_na))

cat("Missing data summary (analytic variables):\n")
print(missing_report, n = 50)

cat("\nVariables exceeding 5% true NA (flag for methods section):\n")
print(filter(missing_report, pct_na > 3), n =nrow(missing_report))
```

---

# 13. Save cleaned file

```{r save}

write.csv(df, file.path(data_path, "01.Screening_clean.csv"), row.names = FALSE)
cat("Saved. Final dimensions:", dim(df), "\n")

df$ID
```




# Standardize only continuous predictors
df_std <- df %>%
  mutate(across(c(D_AGE, SES_SC_Total, D_HH_SIZE, HW_TOTAL, PSS_TOTAL,
                  HRS_PER_WEEK, starts_with("HW_")),
                ~ scale(.)[,1], .names = "{.col}_z"))

# Re-run selected model with _z versions
m_final_std <- lm(PSS_TOTAL ~ D_AGE_z + SES_SC_Total_z + D_HH_SIZE_z + HW_TOTAL_z,
                  data = df_std)





# Pending

- Section 8 is not doing what it should - it is coding informal employment as formal. 
- Could rescue more SES (SES_to_review.csv)
- Could include data entry for D_LOC (second season has no colonia information)


