---
title: "Data merge & cleaning — Screening + HWISE + PSS"
author: "Paloma"
date: "2026-03-07"
output: workflowr::wflow_html
editor_options: 
  chunk_output_type: console
---

## Introduction

This script merges the cleaned screening file (`01.Screening_clean.csv`, produced by
`Data_cleaning.Rmd`) with the HWISE and PSS data (`02.HWISE_PSS.csv`). It covers:
item-level validation, score computation, the left-join merge, and a final
missingness check.

Missing data coding scheme (inherited from `Data_cleaning.Rmd`):

- 999 = Missing / lost / data entry error
- 333 = Pending (data exists but not yet entered)
- 888 = Don't know
- 777 = Refused / prefer not to answer
- 666 = Illegible
- 555 = Other / unclassifiable
- 444 = Ambiguous

Sentinel codes are preserved in the master file and converted to `NA` only within
analysis scripts.

**Permanently excluded participants:**

IDs **10, 141, 164, and 427** are removed at the start of this script and must not
appear in any output. Full rationale is documented in `01.data_cleaning.Rmd`:

- **ID 10**: Age 49 — outside the 18–46 eligibility range.
- **ID 141**: Did not continuously reside in the study neighborhood (moved between neighborhoods).
- **ID 164**: ID number skipped during data collection; participant never existed.
- **ID 427**: Appeared not to understand the survey questions (linguistic or cognitive limitation).

**Scale scoring rules:**

- **HWISE** (Household Water InSecurity Experiences): 12 items (`HW_WORRY` through
  `HW_SHAME`), each scored 0–3. Total range 0–36.
  
- **PSS-14** (Perceived Stress Scale, 14-item): items scored 0–4.
  Negatively-worded items (PSS4, PSS5, PSS6, PSS7, PSS9, PSS10, PSS13) are
  reverse-coded (0→4, 1→3, 2→2, 3→1, 4→0) before summing. Total range 0–56.

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(ggplot2)
library(psych)

data_path <- "./data"

MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

# IDs permanently excluded from all analyses — see Introduction for rationale
EXCLUDED_IDS <- c(10L, 141L, 164L, 427L)

# HWISE items (12 substantive items, scored 0-3)
HWISE_ITEMS <- c("HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
                 "HW_FOOD",  "HW_HANDS",  "HW_BODY",   "HW_DRINK",
                 "HW_ANGRY", "HW_SLEEP",  "HW_NONE",   "HW_SHAME")

# PSS-14 items: positively-worded (direct) vs negatively-worded (reversed)
PSS_DIRECT  <- paste0("PSS", c(1, 2, 3, 8, 11, 12, 14))
PSS_REVERSE <- paste0("PSS", c(4, 5, 6, 7, 9, 10, 13))
PSS_ITEMS   <- c(PSS_DIRECT, PSS_REVERSE)
```

---

# 1. Load data

```{r load}
screening <- read.csv(
  file.path(data_path, "01.Screening_clean.csv"),
  stringsAsFactors = FALSE,
  na.strings = c("")
)

hwise_pss <- read.csv(
  file.path(data_path, "02.HWISE_PSS.csv"),
  stringsAsFactors = FALSE,
  na.strings = c("")
)

cat("Screening rows/cols:", dim(screening), "\n")
cat("HWISE_PSS rows/cols:", dim(hwise_pss), "\n")

cat("\nScreening IDs:", length(unique(screening$ID)), "\n")
cat("HWISE_PSS IDs:", length(unique(hwise_pss$ID)), "\n")

# Remove permanently excluded participants from both files
screening <- screening %>% filter(!ID %in% EXCLUDED_IDS)
hwise_pss <- hwise_pss %>% filter(!ID %in% EXCLUDED_IDS)

cat("\nExcluded IDs:", paste(EXCLUDED_IDS, collapse = ", "), "\n")
cat("Screening rows after exclusion:", nrow(screening), "\n")
cat("HWISE_PSS rows after exclusion:", nrow(hwise_pss), "\n")
```

---

# 2. ID reconciliation

Before merging, identify IDs that exist in only one file.

```{r id-reconciliation}
ids_screen <- unique(screening$ID)
ids_hw     <- unique(hwise_pss$ID)

only_screen <- setdiff(ids_screen, ids_hw)
only_hw     <- setdiff(ids_hw, ids_screen)
both        <- intersect(ids_screen, ids_hw)

cat("# of IDs in both files:            ", length(both),        "\n")
cat("# of IDs only in screening:        ", length(only_screen), "\n")
cat("# of IDs only in HWISE_PSS:        ", length(only_hw),     "\n")

if (length(only_screen) > 0) {
  cat("  Screening-only IDs:", paste(only_screen, collapse = ", "), "\n")
}
if (length(only_hw) > 0) {
  cat("  HWISE_PSS-only IDs:", paste(only_hw, collapse = ", "), "\n")
}
```

---

# 3. Clean HWISE data

## 3.1 Preserve raw HWISE values

```{r hwise-preserve-raw}
hwise_pss <- hwise_pss %>%
  mutate(across(all_of(HWISE_ITEMS),
                ~ ., .names = "{.col}_RAW"))
```

## 3.2 Convert HWISE items to numeric

Expected values: 0, 1, 2, 3. Empty strings were already read as `NA`.
Any remaining non-numeric text is flagged as 555.

```{r hwise-convert}
hwise_pss <- hwise_pss %>%
  mutate(across(all_of(HWISE_ITEMS), ~ {
    x     <- str_trim(as.character(.))
    x_num <- suppressWarnings(as.numeric(x))
    case_when(
      is.na(x)                                   ~ NA_real_,
      x %in% as.character(MISSING_CODES)         ~ suppressWarnings(as.numeric(x)),
      !is.na(x_num) & x_num %in% 0:3            ~ x_num,
      !is.na(x_num)                              ~ 999,   # out-of-range numeric
      TRUE                                       ~ 555    # non-numeric text
    )
  }))
```

## 3.2b HWISE: cleaning change log

Rows where at least one HWISE item changed from its raw value (e.g., empty → NA,
non-numeric → 555, out-of-range → 999).

```{r hwise-changelog}
hwise_raw_long <- hwise_pss %>%
  select(ID, ends_with("_RAW")) %>%
  rename_with(~ str_remove(., "_RAW"), ends_with("_RAW")) %>%
  select(ID, all_of(HWISE_ITEMS)) %>%
  pivot_longer(-ID, names_to = "item", values_to = "raw")

hwise_clean_long <- hwise_pss %>%
  select(ID, all_of(HWISE_ITEMS)) %>%
  pivot_longer(-ID, names_to = "item", values_to = "cleaned")

hwise_changelog <- left_join(hwise_raw_long, hwise_clean_long, by = c("ID", "item")) %>%
  filter(as.character(raw) != as.character(cleaned) |
           (is.na(raw) != is.na(cleaned))) %>%
  mutate(change_type = case_when(
    is.na(cleaned)               ~ "empty → NA",
    cleaned %in% MISSING_CODES   ~ paste0("flagged as ", cleaned),
    TRUE                         ~ paste0(raw, " → ", cleaned)
  )) %>%
  arrange(ID, item)

if (nrow(hwise_changelog) == 0) {
  cat("No HWISE values changed during cleaning.\n")
} else {
  cat(nrow(hwise_changelog), "HWISE value(s) changed:\n")
  print(hwise_changelog, n = 30)
}
```

## 3.3 Check HWISE item distributions

```{r hwise-check}
cat("HWISE item value distributions:\n")
hwise_pss %>%
  select(all_of(HWISE_ITEMS)) %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(everything(), names_to = "item", values_to = "value") %>%
  count(item, value) %>%
  filter(!is.na(value)) %>%
  print(n = 30)

# Flag out-of-range or unexpected values
hwise_flags <- hwise_pss %>%
  select(ID, all_of(HWISE_ITEMS)) %>%
  pivot_longer(-ID, names_to = "item", values_to = "value") %>%
  filter(!is.na(value) & !(value %in% c(0:3))) %>%
  mutate(flag = case_when(
    value %in% MISSING_CODES ~ "sentinel",
    TRUE                     ~ "unexpected"
  ))

cat("\nFlagged HWISE values:\n")
print(hwise_flags, n = 30)
```

## 3.4 Compute HW_TOTAL

Sum of 12 items. Set to `NA` if any item is `NA` or a sentinel code.

```{r hwise-total}
hwise_pss <- hwise_pss %>%
  rowwise() %>%
  mutate(
    HW_TOTAL = {
      vals <- c_across(all_of(HWISE_ITEMS))
      if (any(is.na(vals)) | any(vals %in% MISSING_CODES, na.rm = TRUE)) {
        NA_real_
      } else {
        sum(vals)
      }
    }
  ) %>%
  ungroup()

cat("HW_TOTAL distribution:\n")
print(summary(hwise_pss$HW_TOTAL))
cat("Missing HW_TOTAL:", sum(is.na(hwise_pss$HW_TOTAL)), "\n")
hwise_pss %>% filter(is.na(HW_TOTAL)) %>% select(ID, all_of(HWISE_ITEMS), HW_INFO)


# Distribution plot
ggplot(hwise_pss %>% filter(!is.na(HW_TOTAL)),
       aes(x = HW_TOTAL)) +
  geom_histogram(binwidth = 1, fill = "#5bc0de", color = "white") +
  labs(title = "HW_TOTAL distribution",
       x = "HWISE total score (0–36)", y = "Count") +
  theme_minimal()
```

## 3.5 Compute HW_WORRY_SUB (water worry subscale)

The water worry subscale captures the **psychosocial burden** of household water
insecurity. It is computed as the sum of three HWISE items that reflect emotional
and social distress related to water access:

- `HW_WORRY` — worried the household would not have enough water
- `HW_ANGRY` — felt angry because of the household's water situation
- `HW_SHAME` — felt shame because the household did not have enough water

Each item is scored 0–3, giving a subscale range of **0–9**, where higher scores
indicate greater psychosocial distress related to water insecurity. This subscale
was proposed by Jepson et al. (2021), who used it to examine associations between
water insecurity and health outcomes in urban Mexico.

> Jepson, W. E., Stoler, J., Baek, J., Morán Martínez, J., Uribe Salas, F. J., &
> Carrillo, G. (2021). Cross-sectional study to measure household water insecurity
> and its health outcomes in urban Mexico. *BMJ Open*, 11(3), e040825.
> https://doi.org/10.1136/bmjopen-2020-040825

```{r hwise-worry-sub}
WORRY_ITEMS <- c("HW_WORRY", "HW_ANGRY", "HW_SHAME")

hwise_pss <- hwise_pss %>%
  rowwise() %>%
  mutate(
    HW_WORRY_SUB = {
      vals <- c_across(all_of(WORRY_ITEMS))
      if (any(is.na(vals)) | any(vals %in% MISSING_CODES, na.rm = TRUE)) {
        NA_real_
      } else {
        sum(vals)
      }
    }
  ) %>%
  ungroup()

cat("HW_WORRY_SUB distribution:\n")
print(summary(hwise_pss$HW_WORRY_SUB))
cat("Missing HW_WORRY_SUB:", sum(is.na(hwise_pss$HW_WORRY_SUB)), "\n")

ggplot(hwise_pss %>% filter(!is.na(HW_WORRY_SUB)),
       aes(x = HW_WORRY_SUB)) +
  geom_bar(fill = "#5bc0de", color = "white") +
  scale_x_continuous(breaks = 0:9) +
  labs(title = "HW_WORRY_SUB distribution",
       x = "Water worry subscale (0–9)", y = "Count") +
  theme_minimal()
```

---

# 3.6 Copmute HW_MEX_5SUB

The 5-item subscale was proposed by Huffman and Najera in 2024 as more reliable and cost effective than the complete 12-item HWISE. They reached these conclutions after analyzing the 2021 national survey on health and nutrition data (n=12619 households). They propose a scale including the items: Clothes, Plans, Food, Hands and Body. They propose a cutoff of 3 for low/high water insecurity using this scale. 

> Huffman C, Nájera H.
> The Household Water Insecurity Experiences Scale: answer patterns and cutoffs in Mexico. Salud Publica Mex 
> 8 de diciembre de 2023;66(1, ene-feb):59-66. 
> https://saludpublica.mx/index.php/spm/article/view/14735

```{r hwise-5items-sub}
HW_MEX_5SUB_ITEMS <- c("HW_CLOTHES", "HW_PLANS", "HW_FOOD", "HW_HANDS", "HW_BODY")

hwise_pss <- hwise_pss %>%
  rowwise() %>%
  mutate(
    HW_MEX_5SUB = {
      vals <- c_across(all_of(HW_MEX_5SUB_ITEMS))
      if (any(is.na(vals)) | any(vals %in% MISSING_CODES, na.rm = TRUE)) {
        NA_real_
      } else {
        sum(vals)
      }
    }
  ) %>%
  ungroup()

cat("HW_MEX_5SUB distribution:\n")
print(summary(hwise_pss$HW_MEX_5SUB))
cat("Missing HW_MEX_5SUB:", sum(is.na(hwise_pss$HW_MEX_5SUB)), "\n")

ggplot(hwise_pss %>% filter(!is.na(HW_MEX_5SUB)),
       aes(x = HW_MEX_5SUB)) +
  geom_bar(fill = "#5bc0de", color = "white") +
  scale_x_continuous(breaks = 0:9) +
  labs(title = "HW_MEX_5SUB distribution",
       x = "5-item subscale (0–9)", y = "Count") +
  theme_minimal()
```

---


# 4. Clean PSS data

## 4.1 Preserve raw PSS values

```{r pss-preserve-raw}
hwise_pss <- hwise_pss %>%
  mutate(across(all_of(PSS_ITEMS),
                ~ ., .names = "{.col}_RAW"))
```

## 4.2 Convert PSS items to numeric

Expected values: 0, 1, 2, 3, 4.

```{r pss-convert}
hwise_pss <- hwise_pss %>%
  mutate(across(all_of(PSS_ITEMS), ~ {
    x     <- str_trim(as.character(.))
    x_num <- suppressWarnings(as.numeric(x))
    case_when(
      is.na(x)                                   ~ NA_real_,
      x %in% as.character(MISSING_CODES)         ~ suppressWarnings(as.numeric(x)),
      !is.na(x_num) & x_num %in% 0:4            ~ x_num,
      !is.na(x_num)                              ~ 999,
      TRUE                                       ~ 555
    )
  }))
```

## 4.2b PSS: cleaning change log

```{r pss-changelog}
pss_raw_long <- hwise_pss %>%
  select(ID, ends_with("_RAW")) %>%
  rename_with(~ str_remove(., "_RAW"), ends_with("_RAW")) %>%
  select(ID, all_of(PSS_ITEMS)) %>%
  pivot_longer(-ID, names_to = "item", values_to = "raw")

pss_clean_long <- hwise_pss %>%
  select(ID, all_of(PSS_ITEMS)) %>%
  pivot_longer(-ID, names_to = "item", values_to = "cleaned")

pss_changelog <- left_join(pss_raw_long, pss_clean_long, by = c("ID", "item")) %>%
  filter(as.character(raw) != as.character(cleaned) |
           (is.na(raw) != is.na(cleaned))) %>%
  mutate(change_type = case_when(
    is.na(cleaned)               ~ "empty → NA",
    cleaned %in% MISSING_CODES   ~ paste0("flagged as ", cleaned),
    TRUE                         ~ paste0(raw, " → ", cleaned)
  )) %>%
  arrange(ID, item)

if (nrow(pss_changelog) == 0) {
  cat("No PSS values changed during cleaning.\n")
} else {
  cat(nrow(pss_changelog), "PSS value(s) changed:\n")
  print(pss_changelog, n = 20)
}
```

## 4.3 Check PSS item distributions

```{r pss-check}
cat("PSS item value distributions:\n")
hwise_pss %>%
  select(all_of(PSS_ITEMS)) %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(everything(), names_to = "item", values_to = "value") %>%
  count(item, value) %>%
  filter(!is.na(value)) %>%
  print(n = 20)

# Flag unexpected values
pss_flags <- hwise_pss %>%
  select(ID, all_of(PSS_ITEMS)) %>%
  pivot_longer(-ID, names_to = "item", values_to = "value") %>%
  filter(!is.na(value) & !(value %in% c(0:4))) %>%
  mutate(flag = case_when(
    value %in% MISSING_CODES ~ "sentinel",
    TRUE                     ~ "unexpected"
  ))

cat("\nFlagged PSS values:\n")
print(pss_flags, n = 30)
```

## 4.4 Reverse-code negatively-worded PSS items

PSS items 4, 5, 6, 7, 9, 10, 13 are negatively worded and must be reversed
before summing (0→4, 1→3, 2→2, 3→1, 4→0). Sentinel codes are left unchanged.

```{r pss-reverse}
reverse_pss <- function(x) {
  case_when(
    is.na(x)              ~ NA_real_,
    x %in% MISSING_CODES  ~ x,       # pass sentinel through unchanged
    x == 0 ~ 4,
    x == 1 ~ 3,
    x == 2 ~ 2,
    x == 3 ~ 1,
    x == 4 ~ 0,
    TRUE   ~ x
  )
}

hwise_pss <- hwise_pss %>%
  mutate(across(all_of(PSS_REVERSE),
                reverse_pss, .names = "{.col}_REV"))

# Renamed columns for scored total computation
PSS_SCORED_DIRECT  <- PSS_DIRECT
PSS_SCORED_REVERSE <- paste0(PSS_REVERSE, "_REV")
PSS_SCORED_ITEMS   <- c(PSS_SCORED_DIRECT, PSS_SCORED_REVERSE)
```

## 4.4b PSS reverse-coding: before/after

Mean item score before and after reverse coding for the 7 reversed items.
Direct items are shown for reference. If reverse-coding worked correctly, the
reversed items' means should shift (e.g., a mean of 2.8 becomes ~1.2).

```{r pss-reverse-check}
# Mean of original values for all 14 items
pss_means_before <- hwise_pss %>%
  select(all_of(PSS_ITEMS)) %>%
  summarise(across(everything(), ~ mean(.[!. %in% MISSING_CODES], na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "item", values_to = "mean_before")

# Mean of scored values (direct items unchanged, reverse items = _REV columns)
pss_means_after <- bind_rows(
  hwise_pss %>%
    select(all_of(PSS_DIRECT)) %>%
    summarise(across(everything(), ~ mean(.[!. %in% MISSING_CODES], na.rm = TRUE))) %>%
    pivot_longer(everything(), names_to = "item", values_to = "mean_after"),
  hwise_pss %>%
    select(all_of(paste0(PSS_REVERSE, "_REV"))) %>%
    rename_with(~ str_remove(., "_REV")) %>%
    summarise(across(everything(), ~ mean(.[!. %in% MISSING_CODES], na.rm = TRUE))) %>%
    pivot_longer(everything(), names_to = "item", values_to = "mean_after")
)

pss_rev_check <- left_join(pss_means_before, pss_means_after, by = "item") %>%
  mutate(
    direction = if_else(item %in% PSS_DIRECT, "Direct (unchanged)", "Reversed"),
    item      = factor(item, levels = c(PSS_DIRECT, PSS_REVERSE))
  )

# Print table
cat("PSS item means before and after reverse-coding:\n")
print(pss_rev_check %>% arrange(item) %>%
      mutate(across(c(mean_before, mean_after), ~ round(., 2))), n = 20)

# Plot
pss_rev_check %>%
  pivot_longer(c(mean_before, mean_after),
               names_to = "stage", values_to = "mean") %>%
  mutate(stage = factor(stage,
                        levels = c("mean_before", "mean_after"),
                        labels = c("Before", "After reverse-coding"))) %>%
  ggplot(aes(x = item, y = mean, color = direction, shape = stage, group = item)) +
  geom_point(size = 3) +
  geom_line(aes(group = item), color = "grey70", linewidth = 0.5) +
  scale_color_manual(values = c("Direct (unchanged)" = "#5bc0de",
                                "Reversed"           = "#d9534f")) +
  scale_shape_manual(values = c("Before" = 16, "After reverse-coding" = 17)) +
  labs(title    = "PSS item means: before vs after reverse-coding",
       subtitle = "Lines connect the same item; triangles = after reverse-coding",
       x = "PSS item", y = "Mean score (0–4)",
       color = "Item type", shape = "Stage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 4.5 Compute PSS_TOTAL

Sum of 7 direct items + 7 reverse-coded items. Set to `NA` if any item is
`NA` or a sentinel code. Participants with a `Revisar` note in `PSS_INFO` are
flagged for manual review.

```{r pss-total}
hwise_pss <- hwise_pss %>%
  rowwise() %>%
  mutate(
    PSS_TOTAL = {
      vals <- c_across(all_of(PSS_SCORED_ITEMS))
      if (any(is.na(vals)) | any(vals %in% MISSING_CODES, na.rm = TRUE)) {
        NA_real_
      } else {
        sum(vals)
      }
    }
  ) %>%
  ungroup()

# Flag participants with a "Revisar" note in PSS_INFO
hwise_pss <- hwise_pss %>%
  mutate(PSS_REVIEW_FLAG = if_else(
    str_detect(tolower(PSS_INFO), "revisar"), 1L, 0L
  ))

cat("PSS_TOTAL distribution:\n")
print(summary(hwise_pss$PSS_TOTAL))
cat("Missing PSS_TOTAL:", sum(is.na(hwise_pss$PSS_TOTAL)), "\n")

hwise_pss %>% filter(is.na(PSS_TOTAL)) %>% select(ID, PSS_INFO, all_of(PSS_ITEMS))

cat("\nParticipants flagged for PSS review (PSS_REVIEW_FLAG = 1):\n")
hwise_pss %>%
  filter(PSS_REVIEW_FLAG == 1) %>%
  select(ID,  PSS_INFO, all_of(PSS_ITEMS), PSS_TOTAL) %>%
  print()

# Distribution plot
ggplot(hwise_pss %>% filter(!is.na(PSS_TOTAL)),
       aes(x = PSS_TOTAL)) +
  geom_histogram(binwidth = 2, fill = "#d9534f", color = "white") +
  labs(title = "PSS_TOTAL distribution",
       x = "PSS-14 total score (0–56)", y = "Count") +
  theme_minimal()
```

---

# 5. Standardize SEASON in HWISE_PSS

The HWISE_PSS file encodes season as text ("Oct-Dec", "Apr-Jun"). Recode to
match the numeric convention in the screening file (0 = rainy, 1 = dry).

```{r season-recode}
hwise_pss <- hwise_pss %>%
  mutate(SEASON_CLEAN = case_when(
    str_detect(tolower(SEASON), "oct|rainy") ~ 0,
    str_detect(tolower(SEASON), "apr|dry")   ~ 1,
    TRUE ~ NA_real_
  ))

cat("SEASON_CLEAN distribution:\n")
print(table(hwise_pss$SEASON_CLEAN, useNA = "always"))
```

---

# 6. Select columns for merge

Keep only the columns needed from the HWISE_PSS file (raw items, scored totals,
flags, and notes). The `SEASON` variable will come from the screening file.

```{r select-hwise-cols}
hwise_to_merge <- hwise_pss %>%
  select(
    ID,
    W_WS_LOC,
    SEASON_CLEAN,                      # numeric season (for cross-check)
    # HWISE: raw items + computed total
    all_of(HWISE_ITEMS),
    HW_TOTAL,
    HW_WORRY_SUB,
    HW_MEX_5SUB,
    HW_INFO,                           # free-text notes
    # PSS: raw items, reversed items, computed total, flags
    all_of(PSS_ITEMS),
    all_of(PSS_SCORED_REVERSE),        # reversed versions of neg items
    PSS_TOTAL,
    PSS_REVIEW_FLAG,
    PSS_INFO                           # free-text notes
  )

cat("Columns in hwise_to_merge:", ncol(hwise_to_merge), "\n")
```

---

# 7. Merge with screening

Left join on `screening` to keep all screened participants. HWISE/PSS columns
will be `NA` for the handful of participants without HWISE_PSS data.

```{r merge}
df_merged <- screening %>%
  left_join(hwise_to_merge, by = "ID")

cat("Merged dimensions:", dim(df_merged), "\n")

# Cross-check SEASON from screening (0/1) vs HWISE_PSS text-derived value
season_check <- df_merged %>%
  filter(!is.na(SEASON) & !is.na(SEASON_CLEAN)) %>%
  mutate(season_match = SEASON == SEASON_CLEAN)

cat("\nSEASON cross-check (screening vs HWISE_PSS):\n")
print(table(season_check$season_match, useNA = "always"))

season_mismatch <- season_check %>% filter(!season_match)
if (nrow(season_mismatch) > 0) {
  cat("\nSEASON mismatches (review these IDs):\n")
  print(season_mismatch %>% select(ID, SEASON, SEASON_CLEAN))
}

```

---

# 8. Post-merge checks

## 8.1 Participants without HWISE_PSS data

```{r check-missing-hw}
missing_hw <- df_merged %>%
  filter(is.na(HW_TOTAL) | is.na(PSS_TOTAL)) %>%
  select(ID, SEASON, HW_TOTAL, PSS_TOTAL)

cat("Participants with no HWISE or PSS data after merge:\n")
print(missing_hw)
```

## 8.2 Missing data summary for key variables

```{r missing-summary}
key_vars <- c("HW_TOTAL", "PSS_TOTAL", HWISE_ITEMS, PSS_ITEMS)

missing_report <- df_merged %>%
  select(all_of(key_vars)) %>%
  summarise(across(everything(), list(
    n_na       = ~ sum(is.na(.)),
    n_sentinel = ~ sum(. %in% MISSING_CODES, na.rm = TRUE)
  ))) %>%
  pivot_longer(everything(),
               names_to  = c("variable", ".value"),
               names_sep = "_(?=n_na|n_sentinel)") %>%
  mutate(
    n_total  = nrow(df_merged),
    pct_na   = round(n_na / n_total * 100, 1),
    pct_flag = round(n_sentinel / n_total * 100, 1)
  ) %>%
  filter(n_na > 0 | n_sentinel > 0) %>%
  arrange(desc(pct_na))

# Collect up to 5 example IDs per variable that are NA
na_ids <- df_merged %>%
  select(ID, all_of(key_vars)) %>%
  pivot_longer(-ID, names_to = "variable", values_to = "value") %>%
  filter(is.na(value)) %>%
  group_by(variable) %>%
  summarise(sample_IDs = paste(head(ID, 10), collapse = ", "), .groups = "drop")

missing_report <- missing_report %>%
  left_join(na_ids, by = "variable")

cat("Missing data summary — HWISE & PSS variables:\n")
print(missing_report, n = 40)
```

## 8.3 Score distributions by season

```{r score-by-season}
df_merged %>%
  filter(!is.na(SEASON)) %>%
  mutate(Season = if_else(SEASON == 0, "Rainy (Oct-Dec)", "Dry (Apr-Jun)")) %>%
  select(Season, HW_TOTAL, PSS_TOTAL) %>%
  pivot_longer(-Season, names_to = "Scale", values_to = "Score") %>%
  filter(!is.na(Score)) %>%
  ggplot(aes(x = Score, fill = Season)) +
  geom_histogram(binwidth = 2, position = "identity", alpha = 0.6) +
  facet_wrap(~ Scale, scales = "free_x") +
  scale_fill_manual(values = c("Rainy (Oct-Dec)" = "#5bc0de",
                               "Dry (Apr-Jun)"   = "#d9534f")) +
  labs(title = "HWISE and PSS score distributions by season",
       x = "Score", y = "Count") +
  theme_minimal()
```

---

# 9. Remove uninformative columns

```{r drop-cols}
# Remove raw (un-reverse-coded) PSS items and the intermediate SEASON_CLEAN column
dim(df_merged)
df_merged <- df_merged %>% select(-SEASON_CLEAN, -PSS4, -PSS5, -PSS6, -PSS7, -PSS9, -PSS10, -PSS13)
dim(df_merged)
```

---

# 10. Save merged dataset

```{r save}
out_file <- file.path(data_path, "02.Merged_Screening_HWISE_PSS.csv")
write.csv(df_merged, out_file, row.names = FALSE)
cat("Saved:", out_file, "\n")
cat("Final dimensions:", dim(df_merged), "\n")
```


# Pending

- check samples that are only in screening file or in hwise_pss file
- HWISE have 8 samples with NAs (without final score)
- ID not in database: 71

