---
title: "Merge Datasets – Water Insecurity & Stress Study"
author: ""
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

## Overview

This script merges three supplemental files onto the base dataset
(`02.Merged_Screening_HWISE_PSS.csv`):

1. **03.Chronic_pain_illness.csv** – corrected `HLTH_CPAIN_CAT` and `HLTH_CDIS_CAT` values  
2. **03.Q2-3-4-5-6.csv** – supplemental water variables (Season 1, IDs 1–330)  
3. **03.Q1-2-3-4-5_season2.csv** – supplemental water variables (Season 2, IDs 331+)


Missing code system:

- **999** = Missing / lost / data entry error
- **333** = Pending (data exists but not yet entered)
- **888** = Don't know
- **777** = Refused / prefer not to answer
- **666** = Illegible
- **555** = Other (answer outside allowed categories)
- **444** = Ambiguous

---

## 1. Load base dataset

```{r load-base}
base <- read_csv("./data/02.Merged_Screening_HWISE_PSS.csv", show_col_types = FALSE)

cat("Base dimensions:", nrow(base), "rows ×", ncol(base), "cols\n")
cat("ID range:", min(base$ID), "–", max(base$ID), "\n")

```

---

## 2. Update chronic pain / illness categories

These replace the original `HLTH_CPAIN_CAT` and `HLTH_CDIS_CAT` columns with
corrected values.

```{r merge-chronic}
chronic <- read_csv(
  "./data/03.Chronic_pain_illness.csv",
  col_types = cols(
    ID             = col_double(),
    HLTH_CPAIN_CAT = col_double(),
    HLTH_CDIS_CAT  = col_double()
  )
)

cat("Chronic pain file:", nrow(chronic), "rows\n")
cat("IDs not in base:", sum(!chronic$ID %in% base$ID), "\n")

# Drop old columns and join corrected values
merged <- base %>%
  select(-HLTH_CPAIN_CAT, -HLTH_CDIS_CAT) %>%
  left_join(chronic, by = "ID")

# Verify counts
cat("After chronic merge:", nrow(merged), "rows\n")
cat("IDs not in base:", paste(chronic$ID[!chronic$ID %in% base$ID], collapse = ", "), "\n")
cat("IDs not in base:", paste(chronic$ID[!chronic$ID %in% base$ID], collapse = ", "), "\n")
```

---

## 3. Load and stack supplemental water files

### File 1 water variables (03.Q2-3-4-5-6.csv)

```{r load-s1}
water_s1 <- read_csv(
  "./data/03.Q2-3-4-5-6.csv",
  show_col_types = FALSE
) %>%
  rename_with(~ str_replace_all(., "/", "_PER_")) 

cat("Sample IDs:", water_s1$ID)
cat("IDs not in base:", paste(water_s1$ID[!water_s1$ID %in% base$ID], collapse = ", "), "\n")

```

### Season 2 water variables (03.Q1-2-3-4-5_season2.csv)

```{r load-s2}
water_s2 <- read_csv(
  "./data/03.Q1-2-3-4-5_season2.csv",
  show_col_types = FALSE
) %>%
  rename_with(~ str_replace_all(., "/", "_PER_")) 

cat("Season 2 water rows:", nrow(water_s2), "\n")
cat("Season 2 IDs range:", min(water_s2$ID), "–", max(water_s2$ID), "\n")
head(water_s2)
```

### Stack seasons (bind shared columns, retain all)

```{r stack-water}
# Full bind – fills NA for columns that differ between seasons
water_s1 <- water_s1 %>% mutate(ID = as.numeric(ID))
water_s2 <- water_s2 %>% mutate(ID = as.numeric(ID))

water_all <- bind_rows(water_s1, water_s2)
water_all <- bind_rows(water_s1, water_s2)

# Check for duplicate IDs across seasons
water_all <- bind_rows(water_s1, water_s2) %>%
  group_by(ID) %>%
  summarise(across(everything(), ~ {
    vals <- na.omit(unique(.x))
    if (length(vals) == 0) NA_character_
    else if (length(vals) == 1) as.character(vals)
    else paste(vals, collapse = " vs ")
  }), .groups = "drop") %>%
  mutate(ID = as.numeric(ID)) %>%
  arrange(ID)

cat("Rows after deduplication:", nrow(water_all), "\n")

dupes <- water_all %>% group_by(ID) %>% filter(n() > 1)

if (nrow(dupes) > 0) {
  conflicts <- dupes %>%
      group_by(ID) %>%
    summarise(across(everything(), ~ {
      vals <- na.omit(unique(.x))
      if (length(vals) > 1) paste(vals, collapse = " vs ") else NA_character_
    }), .groups = "drop") %>%
    arrange(ID) %>%
    select(ID, where(~ any(!is.na(.)))) 
  
  cat("Conflicting values by ID:\n")
  print(conflicts)
} else {
  cat("No duplicate IDs across seasons.\n")
}


cat("Combined water rows:", nrow(water_all), "\n")
head(water_all)
```

---

## 4. Merge water variables onto main dataset

```{r merge-water}
merged <- merged %>%
  left_join(water_all, by = "ID")

cat("Final dimensions:", nrow(merged), "rows ×", ncol(merged), "cols\n")

# IDs in base without a water record
no_water <- sum(!base$ID %in% water_all$ID)
cat("Base IDs with no water record:", no_water, "\n")
```

---

## 5. Reorder to match target column specification

Core columns are placed first in the required order; supplemental water columns
are appended after.

```{r reorder}
target_cols <- c(
  "ID", "D_YRBR", "D_AGE", "D_LOC", "D_LOC_CODE", "D_LOC_TIME", "D_LOC_INFO",
  "HLTH_SMK", "HLTH_TTM", "HLTH_MED", "HLTH_CPAIN", "HLTH_CPAIN_CAT",
  "HLTH_CDIS", "HLTH_CDIS_CAT", "D_HH_SIZE", "D_HH_SIZE_INFO", "D_CHLD",
  "D_EMP_TYPE", "D_EMP_DAYS", "W_HH_SIZE", "W_STORAGE", "D_INFO",
  "SES_EDU_CAT", "SES_EDU_SC", "SES_BTHR_CAT", "SES_BTHR_SC",
  "SES_CAR_CAT", "SES_CAR_SC", "SES_INT_CAT", "SES_INT_SC",
  "SES_WRK_CAT", "SES_WRK_SC", "SES_BEDR_CAT", "SES_BEDR_SC", "SES_INFO",
  "HLTH_TTM_RAW", "HLTH_MED_RAW", "HLTH_CPAIN_RAW", "HLTH_CDIS_RAW",
  "HLTH_SMK_RAW", "D_EMP_TYPE_RAW", "D_CHLD_RAW", "SEASON",
  "HLTH_TTM_CAT", "HLTH_MED_CAT", "HLTH_CPAIN_CAT", "HLTH_CDIS_CAT",
  "HLTH_SMK_CLEAN", "HLTH_SMK_CAT", "D_EMP_CLEAN", "D_EMP_FORM_CLEAN",
  "D_CHLD_CLEAN", "D_LOC_TIME_RAW", "D_LOC_TIME_CLEAN",
  "SES_EDU_SC_RAW", "SES_BTHR_SC_RAW", "SES_CAR_SC_RAW", "SES_INT_SC_RAW",
  "SES_WRK_SC_RAW", "SES_BEDR_SC_RAW", "SES_SC_Total", "SES_CAT_AMAI",
  "W_WS_LOC",
  "HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS", "HW_FOOD",
  "HW_HANDS", "HW_BODY", "HW_DRINK", "HW_ANGRY", "HW_SLEEP",
  "HW_NONE", "HW_SHAME", "HW_TOTAL", "HW_WORRY_SUB", "HW_MEX_5SUB", "HW_INFO",
  "PSS1", "PSS2", "PSS3", "PSS8", "PSS11", "PSS12", "PSS14",
  "PSS4", "PSS5", "PSS6", "PSS7", "PSS9", "PSS10", "PSS13",
  "PSS4_REV", "PSS5_REV", "PSS6_REV", "PSS7_REV",
  "PSS9_REV", "PSS10_REV", "PSS13_REV",
  "PSS_TOTAL", "PSS_REVIEW_FLAG", "PSS_INFO"
)

# Identify which target cols are present
present   <- target_cols[target_cols %in% names(merged)]
missing   <- target_cols[!target_cols %in% names(merged)]
extra     <- names(merged)[!names(merged) %in% target_cols]

if (length(missing) > 0) cat("Target cols NOT found in merged data:\n", paste(missing, collapse = ", "), "\n")
if (length(extra)   > 0) cat("Extra cols appended after target cols:\n", paste(extra,   collapse = ", "), "\n")

# Reorder: target first, then extra (supplemental water + source flag)
merged_final <- merged %>%
  select(all_of(present), any_of(extra))

cat("Final column count:", ncol(merged_final), "\n")
```

---

## 6. Quality checks

```{r qc}
# Missing values in key outcome columns
qc_cols <- c("HLTH_CPAIN","HLTH_CPAIN_CAT", "HLTH_CPAIN_RAW", "HLTH_CDIS_CAT", "HW_TOTAL", "PSS_TOTAL", "W_WS_LOC", "MX3_CONT_INT","MX2_WI_SEASON_TYPE", "MX2_WI_SEASON_INFO", "MX2_WI_SEASON", "MX5_WFREQ_HRS_PER_WEEK")

print(merged_final[, qc_cols], n=5)

merged_final %>%
  summarise(across(all_of(qc_cols), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
  arrange(desc(n_missing)) %>%
  print()

# SEASON distribution
cat("\nSEASON distribution:\n")
print(table(merged_final$SEASON, useNA = "always"))
```

---

## 7. Export

```{r export}
write_csv(merged_final, "./data/03.Merged_Q1-6.csv", na = "")
cat("Exported: 03.Merged_Q1-6.csv\n")
cat("Rows:", nrow(merged_final), "| Cols:", ncol(merged_final), "\n")
```

---

# 8. Load supplemental files –Q8/26/28, 19, 16, 14-19-20-21

## 8a. Q8-26-28 (water quality trust, household water emotions, water quality comparison)

```{r load-q826}
MISSING_CODES <- c(999L, 888L, 777L, 666L, 555L, 444L, 333L)

q826 <- read_csv(
  "./data/Q8-26-28.csv",
  col_types = cols(ID = col_character()),
  show_col_types = FALSE
) %>%
  mutate(ID = as.numeric(str_trim(ID))) %>%
  filter(!is.na(ID))

cat("Q8-26-28 rows:", nrow(q826), "| Unique IDs:", n_distinct(q826$ID), "\n")
cat("MX28_WQ_COMP unique values:", paste(sort(unique(q826$MX28_WQ_COMP)), collapse = ", "), "\n")
cat("MX26_EM_HHW_TYPE unique values:", paste(sort(unique(q826$MX26_EM_HHW_TYPE)), collapse = ", "), "\n")
```

## 8b. Q19 (consensus coding – all IDs)

```{r load-q19}
q19 <- read_csv(
  "./data/Q19.csv",
  col_types = cols(ID = col_double()),
  show_col_types = FALSE
)

cat("Q19.csv rows:", nrow(q19), "| Unique IDs:", n_distinct(q19$ID), "\n")
cat("Q19_CAT unique values:", paste(sort(unique(q19$Q19_CAT)), collapse = ", "), "\n")
```

## 8c. Q16 (water scenario preferences – Season 2 only)

```{r load-q16}
q16 <- read_csv(
  "./data/Q16.csv",
  col_types = cols(ID = col_double()),
  show_col_types = FALSE
)

cat("Q16 rows:", nrow(q16), "| ID range:", min(q16$ID, na.rm = TRUE), "–",
    max(q16$ID, na.rm = TRUE), "\n")
cat("MX16_SCEN_01 unique values:", paste(sort(unique(q16$MX16_SCEN_01)), collapse = ", "), "\n")
```

## 8d. Q14-19-20-21 (Jocelyne and Alice independent codings) and Q19_combined (three-rater adjudication)

```{r load-q14-21}
q_joc <- read_csv("./data/Q14-19-20-21_Jocelyne.csv", show_col_types = FALSE) %>%
  rename(ID = Folio) %>%
  select(ID, Q19_Jocelyne = `X19.CAT`)

q_alice <- read_csv("./data/Q14-19-20-21_Alice.csv", show_col_types = FALSE) %>%
  rename(ID = Folio) %>%
  select(ID, Q19_Alice = `X19.CAT`)

# Q19_combined: three-rater adjudication file (Season 1, IDs 1–398).
# CatA = Alice, CatJ = Jocelyne, CatL = lead rater (= Q19.csv Q19_CAT).
# Match == "match"     → all three raters agreed.
# Match == "different" → at least one rater disagreed; Consensus may be blank.
q19_combined <- read_csv("./data/Q19_combined.csv", show_col_types = FALSE) %>%
  rename(ID = SampleID) %>%
  select(ID,
         CatA_combined  = CatA,
         CatJ_combined  = CatJ,
         CatL_combined  = CatL,
         three_rater_match = Match,
         consensus_combined = Consensus)

cat("Jocelyne rows:", nrow(q_joc), "| Unique IDs:", n_distinct(q_joc$ID), "\n")
cat("Alice rows:", nrow(q_alice), "| Unique IDs:", n_distinct(q_alice$ID), "\n")
cat("Q19_combined rows:", nrow(q19_combined), "| three_rater_match:", "\n")
print(table(q19_combined$three_rater_match, useNA = "always"))
```

---

# 9. Process Q28 – Water Quality Comparison

Coding: **0 = Better**, **1 = Same**, **2 = Worse**.

Each ID has exactly one Q28 value. String "NA" and blank cells → 999 (missing).

```{r process-q28}
q28_clean <- q826 %>%
  # Keep the first non-missing Q28 per ID (row 1 always carries the value)
  filter(!is.na(MX28_WQ_COMP)) %>%
  distinct(ID, .keep_all = TRUE) %>%
  select(ID, MX28_WQ_COMP, MX28_WQ_COMP_INFO) %>%
  mutate(
    MX28_WQ_COMP_CAT = case_when(
      str_to_lower(str_trim(MX28_WQ_COMP)) == "better" ~ 0L,
      str_to_lower(str_trim(MX28_WQ_COMP)) == "same"   ~ 1L,
      str_to_lower(str_trim(MX28_WQ_COMP)) == "worse"  ~ 2L,
      str_to_lower(str_trim(MX28_WQ_COMP)) == "na"     ~ 999L,
      TRUE                                              ~ 999L
    )
  )

cat("Q28 distribution (MX28_WQ_COMP_CAT):\n")
cat("  0 = Better, 1 = Same, 2 = Worse, 999 = Missing\n")
print(table(q28_clean$MX28_WQ_COMP_CAT, useNA = "always"))
cat("IDs with Q28 not in base:", sum(!q28_clean$ID %in% base$ID), "\n")
```

---

# 10. Process Q26 – Household Water Emotions (Categorical)

Coding: **0 = positive**, **1 = negative**.
Some IDs report two emotions; if both share the same valence the code is unambiguous.
IDs where one emotion is positive and another is negative are coded **444 (Ambiguous)**
and exported for manual review.
Neutral / other valences → **555 (Other)**.

```{r process-q26}
# Normalise valence labels
q26_norm <- q826 %>%
  select(ID, MX26_EM_HHW, MX26_EM_HHW_TYPE, MX26_EM_HHW_INFO) %>%
  mutate(
    valence_norm = case_when(
      str_to_lower(str_trim(MX26_EM_HHW_TYPE)) == "positive" ~ "positive",
      str_to_lower(str_trim(MX26_EM_HHW_TYPE)) == "negative" ~ "negative",
      str_to_lower(str_trim(MX26_EM_HHW_TYPE)) %in%
        c("neutral", "otro")                                  ~ "other",
      str_to_lower(str_trim(MX26_EM_HHW_TYPE)) == "na"       ~ NA_character_,
      is.na(MX26_EM_HHW_TYPE)                                 ~ NA_character_,
      TRUE                                                     ~ NA_character_
    )
  ) %>%
  filter(!is.na(MX26_EM_HHW) | !is.na(MX26_EM_HHW_TYPE))

# Aggregate to one row per ID
q26_clean <- q26_norm %>%
  group_by(ID) %>%
  summarise(
    MX26_EM_HHW      = paste(na.omit(MX26_EM_HHW), collapse = "; "),
    MX26_EM_HHW_INFO = paste(na.omit(MX26_EM_HHW_INFO), collapse = "; "),
    valences_present = paste(sort(unique(na.omit(valence_norm))), collapse = "; "),
    .groups = "drop"
  ) %>%
  mutate(
    MX26_EM_HHW_CAT = case_when(
      valences_present == "positive"                                          ~ 0L,
      valences_present == "negative"                                          ~ 1L,
      valences_present == "other"                                             ~ 555L,
      str_detect(valences_present, "positive") &
        str_detect(valences_present, "negative")                              ~ 444L,
      valences_present == "" | is.na(valences_present)                        ~ 999L,
      TRUE                                                                    ~ 555L
    )
  ) %>%
  select(ID, MX26_EM_HHW, MX26_EM_HHW_INFO, MX26_EM_HHW_CAT, valences_present)

cat("Q26 distribution (MX26_EM_HHW_CAT):\n")
cat("  0=positive, 1=negative, 444=ambiguous, 555=other, 999=missing\n")
print(table(q26_clean$MX26_EM_HHW_CAT, useNA = "always"))
cat("IDs with conflicting valence (444):", sum(q26_clean$MX26_EM_HHW_CAT == 444L, na.rm = TRUE), "\n")
```

---

# 11. Process Q19 – Time Investment in Water Management

Coding: **0 = No**, **1 = Yes**.
**444** (Ambiguous) for responses coded "Somewhat" or "a bit" by any rater.
**555** (Other) for responses outside the No/Yes/Somewhat range.
**999** (Missing) for NA or blank.

**Source priority:** Q19.csv (consensus, covers all IDs) is the authoritative source.
Jocelyne and Alice codings are used solely to detect discrepancies.

```{r process-q19}
# Normalise Q19 values to a consistent vocabulary
norm_q19 <- function(x) {
  x <- str_to_lower(str_trim(x))
  case_when(
    x == "no"                           ~ "no",
    x %in% c("si", "yes")              ~ "yes",
    x %in% c("a bit", "somewhat",
              "a bit ", "un poco")      ~ "somewhat",
    x %in% c("other", "otro",
              "other ", "other?????",
              "other ??? ", "other??????",
              "other?????")             ~ "other",
    x %in% c("na", "")                 ~ NA_character_,
    TRUE                                ~ NA_character_   # ?, ????, no????, etc.
  )
}

# Normalise each source
q19_primary <- q19 %>%
  mutate(q19_norm = norm_q19(Q19_CAT)) %>%
  select(ID, MX19_TIME_RAW = Q19, MX19_TIME_ENG = Q19_ENG, q19_norm)

q19_joc <- q_joc %>%
  mutate(joc_norm = norm_q19(Q19_Jocelyne)) %>%
  select(ID, joc_norm)

q19_alice <- q_alice %>%
  mutate(alice_norm = norm_q19(Q19_Alice)) %>%
  select(ID, alice_norm)

# Combine all three
q19_all <- q19_primary %>%
  left_join(q19_joc,   by = "ID") %>%
  left_join(q19_alice, by = "ID")

# Final clean column (from primary/consensus source)
# Also attach the three-rater match flag from Q19_combined (Season 1 only; NA for Season 2)
q19_clean <- q19_all %>%
  left_join(
    q19_combined %>% select(ID, three_rater_match),
    by = "ID"
  ) %>%
  mutate(
    MX19_TIME_CAT = case_when(
      q19_norm == "no"       ~ 0L,
      q19_norm == "yes"      ~ 1L,
      q19_norm == "somewhat" ~ 444L,
      q19_norm == "other"    ~ 555L,
      is.na(q19_norm)        ~ 999L
    )
  ) %>%
  select(ID, MX19_TIME_RAW, MX19_TIME_ENG, MX19_TIME_CAT,
         q19_norm, joc_norm, alice_norm, three_rater_match)

cat("Q19 distribution (MX19_TIME_CAT):\n")
cat("  0=No, 1=Yes, 444=Ambiguous, 555=Other, 999=Missing\n")
print(table(q19_clean$MX19_TIME_CAT, useNA = "always"))
cat("Three-rater match (Season 1 only):\n")
print(table(q19_clean$three_rater_match, useNA = "always"))
```

```{r q19-conflict-check}
# Flag IDs where primary (Q19.csv) disagrees with Jocelyne or Alice
# (only compare clear yes/no disagreements – ignore NA and ambiguous)
q19_conflicts <- q19_clean %>%
  filter(
    q19_norm %in% c("no", "yes") &
    (
      (!is.na(joc_norm)   & joc_norm   %in% c("no", "yes") & q19_norm != joc_norm)   |
      (!is.na(alice_norm) & alice_norm %in% c("no", "yes") & q19_norm != alice_norm)
    )
  ) %>%
  mutate(
    conflict_with_jocelyne = case_when(
      !is.na(joc_norm) & joc_norm %in% c("no","yes") & q19_norm != joc_norm ~ "YES",
      TRUE ~ "no"
    ),
    conflict_with_alice = case_when(
      !is.na(alice_norm) & alice_norm %in% c("no","yes") & q19_norm != alice_norm ~ "YES",
      TRUE ~ "no"
    )
  ) %>%
  select(ID, MX19_TIME_RAW, MX19_TIME_ENG,
         Primary_Q19csv = q19_norm,
         Jocelyne = joc_norm,
         Alice = alice_norm,
         conflict_with_jocelyne, conflict_with_alice,
         three_rater_match)

cat("IDs with Q19 primary vs. Jocelyne or Alice conflict:", nrow(q19_conflicts), "\n")
if (nrow(q19_conflicts) > 0) print(head(q19_conflicts, 10))
```

---

# 12. Process Q16 – Water Scenario Preferences

Likert scale 1–4 across four vignette scenarios. Administered to Season 2 participants only (IDs 302+).
`Otro` / `Other` → **555**. Blank / NA → **999**.

```{r process-q16}
q16_clean <- q16 %>%
  mutate(
    across(
      c(MX16_SCEN_01, MX16_SCEN_03, MX16_SCEN_05, MX16_SCEN_99),
      ~ case_when(
          str_to_lower(str_trim(as.character(.x))) %in% c("otro", "other") ~ 555L,
          str_to_lower(str_trim(as.character(.x))) == "na"                  ~ 999L,
          is.na(.x)                                                          ~ 999L,
          TRUE                                                               ~ as.integer(.x)
        )
    )
  ) %>%
  select(ID, MX16_SCEN_01, MX16_SCEN_03, MX16_SCEN_05, MX16_SCEN_99, MX16_SCEN_INFO)

cat("Q16 rows:", nrow(q16_clean), "| ID range:",
    min(q16_clean$ID, na.rm = TRUE), "–", max(q16_clean$ID, na.rm = TRUE), "\n")
cat("Note: Q16 was administered to Season 2 participants only.\n")
for (col in c("MX16_SCEN_01","MX16_SCEN_03","MX16_SCEN_05","MX16_SCEN_99")) {
  cat(col, "distribution:", paste(sort(unique(q16_clean[[col]])), collapse = ", "), "\n")
}
```

---

# 13. Export Conflicts for Manual Review

Four issue types are exported to a single CSV, each with a blank `Suggested_code` column to fill in:

1. **Q26 – rater conflict** – IDs where recorded emotions have conflicting valence (positive vs. negative) → coded 444.
2. **Q26 – sentinel** – IDs coded 555 (neutral/other emotion).
3. **Q19 – rater conflict** – IDs where Q19.csv disagrees with Jocelyne's or Alice's yes/no coding. The `three_rater_match` column flags Season 1 IDs where the full three-rater panel also disagreed.
4. **Q19 – sentinel** – IDs coded 444 (Somewhat/ambiguous) or 555 (Other).
5. **Q16 – sentinel** – IDs with any scenario column coded 555 (Otro response).

```{r export-conflicts}
# ── Q26: rater conflict (coded 444) ──────────────────────────────────────────
q26_conflicts <- q26_clean %>%
  filter(MX26_EM_HHW_CAT == 444L) %>%
  transmute(
    ID,
    Question          = "Q26",
    Issue             = "Rater conflict: mixed positive and negative emotions",
    Coded_value       = as.character(MX26_EM_HHW_CAT),
    Value_source1     = str_extract(MX26_EM_HHW, "^[^;]+"),
    Value_source2     = str_extract(MX26_EM_HHW, "(?<=; ).+$"),
    three_rater_match = NA_character_,
    Notes             = MX26_EM_HHW_INFO,
    Suggested_code    = NA_character_
  )

# ── Q26: sentinel 555 (neutral/other emotion) ────────────────────────────────
q26_sentinel <- q26_clean %>%
  filter(MX26_EM_HHW_CAT == 555L) %>%
  transmute(
    ID,
    Question          = "Q26",
    Issue             = "Sentinel 555 (Other): emotion valence is neutral or 'otro'",
    Coded_value       = as.character(MX26_EM_HHW_CAT),
    Value_source1     = MX26_EM_HHW,
    Value_source2     = valences_present,
    three_rater_match = NA_character_,
    Notes             = MX26_EM_HHW_INFO,
    Suggested_code    = NA_character_
  )

# ── Q19: rater conflict (primary vs. Jocelyne / Alice) ───────────────────────
q19_conflicts_export <- q19_conflicts %>%
  transmute(
    ID,
    Question          = "Q19",
    Issue             = paste0(
      "Rater conflict: primary vs. ",
      if_else(conflict_with_jocelyne == "YES" & conflict_with_alice == "YES",
              "Jocelyne AND Alice",
              if_else(conflict_with_jocelyne == "YES", "Jocelyne", "Alice"))
    ),
    Coded_value       = NA_character_,
    Value_source1     = paste("Q19.csv (primary):", Primary_Q19csv),
    Value_source2     = paste("Jocelyne:", Jocelyne, "| Alice:", Alice),
    three_rater_match,
    Notes             = MX19_TIME_ENG,
    Suggested_code    = NA_character_
  )

# ── Q19: sentinel 444 (Somewhat) or 555 (Other) ──────────────────────────────
# Exclude IDs already in the rater-conflict block to avoid duplication
q19_sentinel <- q19_clean %>%
  filter(MX19_TIME_CAT %in% c(444L, 555L),
         !ID %in% q19_conflicts$ID) %>%
  transmute(
    ID,
    Question          = "Q19",
    Issue             = if_else(
      MX19_TIME_CAT == 444L,
      "Sentinel 444 (Ambiguous): response coded as 'Somewhat'",
      "Sentinel 555 (Other): response outside No/Yes/Somewhat"
    ),
    Coded_value       = as.character(MX19_TIME_CAT),
    Value_source1     = paste("Q19.csv (primary):", q19_norm),
    Value_source2     = paste("Jocelyne:", joc_norm, "| Alice:", alice_norm),
    three_rater_match,
    Notes             = MX19_TIME_ENG,
    Suggested_code    = NA_character_
  )

# ── Q16: sentinel 555 (Otro) in any scenario column ──────────────────────────
q16_sentinel <- q16_clean %>%
  filter(if_any(c(MX16_SCEN_01, MX16_SCEN_03, MX16_SCEN_05, MX16_SCEN_99),
                ~ .x == 555L)) %>%
  mutate(
    flagged_cols = paste(
      na.omit(c(
        if_else(MX16_SCEN_01 == 555L, "MX16_SCEN_01", NA_character_),
        if_else(MX16_SCEN_03 == 555L, "MX16_SCEN_03", NA_character_),
        if_else(MX16_SCEN_05 == 555L, "MX16_SCEN_05", NA_character_),
        if_else(MX16_SCEN_99 == 555L, "MX16_SCEN_99", NA_character_)
      )),
      collapse = "; "
    )
  ) %>%
  transmute(
    ID,
    Question          = "Q16",
    Issue             = "Sentinel 555 (Other): 'Otro' response in one or more scenario columns",
    Coded_value       = "555",
    Value_source1     = paste("Flagged cols:", flagged_cols),
    Value_source2     = paste0(
      "SCEN_01=", MX16_SCEN_01, " | SCEN_03=", MX16_SCEN_03,
      " | SCEN_05=", MX16_SCEN_05, " | SCEN_99=", MX16_SCEN_99
    ),
    three_rater_match = NA_character_,
    Notes             = MX16_SCEN_INFO,
    Suggested_code    = NA_character_
  )

# ── Combine and export ────────────────────────────────────────────────────────
conflicts_review <- bind_rows(
  q26_conflicts,
  q26_sentinel,
  q19_conflicts_export,
  q19_sentinel,
  q16_sentinel
) %>%
  arrange(Question, ID)

write_csv(conflicts_review, "./data/03.Conflicts_to_review.csv", na = "")
cat("Exported: 03.Conflicts_to_review.csv\n")
cat("  Q26 rater conflicts (444):", nrow(q26_conflicts), "IDs\n")
cat("  Q26 sentinel 555:         ", nrow(q26_sentinel),  "IDs\n")
cat("  Q19 rater conflicts:      ", nrow(q19_conflicts_export), "IDs\n")
cat("  Q19 sentinel 444/555:     ", nrow(q19_sentinel),  "IDs\n")
cat("  Q16 sentinel 555:         ", nrow(q16_sentinel),  "IDs\n")
cat("  Total rows:               ", nrow(conflicts_review), "\n")
```

---

# 13.5 Apply resolutions from reviewed conflicts file

Read `03.conflicts_reviewed.csv` — a copy of `03.Conflicts_to_review.csv` where you
have filled in the `Suggested_code` column — and patch `q19_clean`, `q26_clean`, and
`q16_clean` before the final merge.

**Workflow:**

1. Copy `data/03.Conflicts_to_review.csv` → `data/03.conflicts_reviewed.csv`.
2. Fill in `Suggested_code` for each row you have resolved. Leave blank for rows still pending.
3. Save and re-knit.

Valid codes by question:

- **Q19**: 0 (No), 1 (Yes), 444 (Ambiguous), 555 (Other), 999 (Missing)
- **Q26**: 0 (Positive), 1 (Negative), 555 (Other), 999 (Missing)
- **Q16**: 1–4, 555 (Other), 999 (Missing) — applied to every scenario column still coded 555 for that ID

```{r apply-resolutions}
reviewed_path <- "./data/03.conflicts_reviewed.csv"

if (!file.exists(reviewed_path)) {

  cat("No reviewed file found at", reviewed_path, "\n")
  cat("Create it by copying 03.Conflicts_to_review.csv, filling in Suggested_code, and re-knitting.\n")

} else {

  reviewed <- read_csv(reviewed_path, show_col_types = FALSE) %>%
    filter(!is.na(Suggested_code), str_trim(Suggested_code) != "")

  cat("Reviewed file loaded:", nrow(reviewed), "resolved rows (of",
      nrow(read_csv(reviewed_path, show_col_types = FALSE)), "total)\n\n")

  # Counters for the summary
  n_q19 <- 0L; n_q26 <- 0L; n_q16 <- 0L

  for (i in seq_len(nrow(reviewed))) {

    id_i   <- reviewed$ID[i]
    q_i    <- reviewed$Question[i]
    code_i <- as.integer(reviewed$Suggested_code[i])

    # ── Q19 ────────────────────────────────────────────────────────────────────
    if (q_i == "Q19") {
      idx <- which(q19_clean$ID == id_i)
      if (length(idx) == 1) {
        old <- q19_clean$MX19_TIME_CAT[idx]
        q19_clean$MX19_TIME_CAT[idx] <- code_i
        cat(sprintf("Q19 ID %d: %s → %d\n", id_i, old, code_i))
        n_q19 <- n_q19 + 1L
      } else {
        cat(sprintf("Q19 ID %d: not found in q19_clean – skipped\n", id_i))
      }
    }

    # ── Q26 ────────────────────────────────────────────────────────────────────
    else if (q_i == "Q26") {
      idx <- which(q26_clean$ID == id_i)
      if (length(idx) == 1) {
        old <- q26_clean$MX26_EM_HHW_CAT[idx]
        q26_clean$MX26_EM_HHW_CAT[idx] <- code_i
        cat(sprintf("Q26 ID %d: %s → %d\n", id_i, old, code_i))
        n_q26 <- n_q26 + 1L
      } else {
        cat(sprintf("Q26 ID %d: not found in q26_clean – skipped\n", id_i))
      }
    }

    # ── Q16 ────────────────────────────────────────────────────────────────────
    # Replace every scenario column still coded 555 for this ID
    else if (q_i == "Q16") {
      idx <- which(q16_clean$ID == id_i)
      if (length(idx) == 1) {
        q16_cols <- c("MX16_SCEN_01", "MX16_SCEN_03", "MX16_SCEN_05", "MX16_SCEN_99")
        updated_cols <- character(0)
        for (col in q16_cols) {
          if (!is.na(q16_clean[[col]][idx]) && q16_clean[[col]][idx] == 555L) {
            q16_clean[[col]][idx] <- code_i
            updated_cols <- c(updated_cols, col)
          }
        }
        if (length(updated_cols) > 0) {
          cat(sprintf("Q16 ID %d: 555 → %d in %s\n",
                      id_i, code_i, paste(updated_cols, collapse = ", ")))
          n_q16 <- n_q16 + 1L
        } else {
          cat(sprintf("Q16 ID %d: no 555 values found in scenario columns – skipped\n", id_i))
        }
      } else {
        cat(sprintf("Q16 ID %d: not found in q16_clean – skipped\n", id_i))
      }
    }
  }

  # Summary
  cat(sprintf("\n── Resolution summary ──────────────────────────────\n"))
  cat(sprintf("  Q19 updated: %d\n", n_q19))
  cat(sprintf("  Q26 updated: %d\n", n_q26))
  cat(sprintf("  Q16 updated: %d\n", n_q16))

  # Check how many conflicts remain unresolved (still 444 or 555 in coded vars)
  still_open_q19 <- sum(q19_clean$MX19_TIME_CAT %in% c(444L, 555L), na.rm = TRUE)
  still_open_q26 <- sum(q26_clean$MX26_EM_HHW_CAT %in% c(444L, 555L), na.rm = TRUE)
  still_open_q16 <- sum(
    q16_clean$MX16_SCEN_01 == 555L | q16_clean$MX16_SCEN_03 == 555L |
    q16_clean$MX16_SCEN_05 == 555L | q16_clean$MX16_SCEN_99 == 555L,
    na.rm = TRUE
  )
  cat(sprintf("  Still unresolved – Q19: %d | Q26: %d | Q16: %d\n",
              still_open_q19, still_open_q26, still_open_q16))
}
```

---

# 13.6. Review file – HRS_PER_WEEK, D_HH_SIZE, and Cont_int

Exports `03.continuous_to_review.csv`: one row per (ID × variable) where
`MX5_WFREQ_HRS_PER_WEEK`, `D_HH_SIZE`, or `MX3_CONT_INT` is NA or carries a
sentinel code. The student fills in the `consensus` column, saves it as
`03.continuous_consensus.csv`, and re-runs section 13.7 to patch the dataset
before the final merge in section 14.

```{r export-continuous-review}
MISSING_CODES_INT <- c(999L, 888L, 777L, 666L, 555L, 444L, 333L)
MISSING_CODES_CHR <- as.character(MISSING_CODES_INT)

review_vars <- c("MX5_WFREQ_HRS_PER_WEEK", "D_HH_SIZE", "MX3_CONT_INT")

sentinel_labels <- c(
  "999" = "Lost / data entry error",
  "888" = "Don't know",
  "777" = "Refused",
  "666" = "Illegible",
  "555" = "Other — outside allowed categories",
  "444" = "Ambiguous — needs review",
  "333" = "Pending — data exists but not entered"
)

hints <- c(
  MX5_WFREQ_HRS_PER_WEEK = "Enter hours per week as a positive number (e.g. 3.5). Use 0 if none.",
  D_HH_SIZE              = "Enter number of household members as a positive integer.",
  MX3_CONT_INT           = "Enter 0 = no interruption, 1 = interrupted (continuous supply)."
)

# Long-format: one row per problematic (ID × variable)
cont_long <- merged_final %>%
  select(ID, all_of(review_vars)) %>%
  mutate(across(-ID, as.character)) %>%
  pivot_longer(cols = -ID, names_to = "variable", values_to = "current_value") %>%
  mutate(
    issue = case_when(
      is.na(current_value)                  ~ "Missing / not entered",
      current_value %in% MISSING_CODES_CHR  ~ paste0(
        "Sentinel ", current_value, " — ", sentinel_labels[current_value]
      ),
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(issue))

# Pull D_HH_SIZE_INFO as a notes column where available
notes_df <- merged_final %>%
  select(ID, notes = any_of("D_HH_SIZE_INFO")) %>%
  mutate(variable = "D_HH_SIZE")

cont_review <- cont_long %>%
  left_join(notes_df, by = c("ID", "variable")) %>%
  mutate(
    hint      = hints[variable],
    consensus = NA_character_
  ) %>%
  select(ID, variable, current_value,
         any_of("notes"),
         issue, hint, consensus)

write_csv(cont_review, "./data/03.continuous_to_review.csv", na = "")

cat("Exported: 03.continuous_to_review.csv\n")
for (v in review_vars) {
  cat(sprintf("  %-28s %d rows\n", paste0(v, ":"), sum(cont_review$variable == v)))
}
cat("  Total:", nrow(cont_review), "rows\n")
```

---

# 13.7. Apply continuous variable consensus

Fill in `consensus` in `03.continuous_to_review.csv`, save as
`03.continuous_consensus.csv`, then re-run this chunk **before** proceeding to
section 14. The corrected values will flow automatically into `merged_final2`
during the merge.

```{r apply-continuous-consensus}
cont_consensus_path <- "./data/03.continuous_consensus.csv"

if (!file.exists(cont_consensus_path)) {

  cat("Consensus file not found:", cont_consensus_path, "\n",
      "Fill in the 'consensus' column in 03.continuous_to_review.csv,\n",
      "save as 03.continuous_consensus.csv, and re-run this chunk.\n")

} else {

  cont_consensus <- read_csv(cont_consensus_path, show_col_types = FALSE) %>%
    filter(!is.na(consensus), str_trim(as.character(consensus)) != "")

  cat("Applying", nrow(cont_consensus), "consensus values...\n")

  n_applied <- 0L
  for (i in seq_len(nrow(cont_consensus))) {
    id_i  <- cont_consensus$ID[i]
    v_i   <- cont_consensus$variable[i]
    val_i <- suppressWarnings(as.numeric(cont_consensus$consensus[i]))

    if (is.na(val_i)) {
      cat(sprintf("Non-numeric value skipped — ID: %d, variable: %s, value: %s\n",
                  id_i, v_i, cont_consensus$consensus[i]))
      next
    }

    idx <- which(merged_final$ID == id_i)
    if (length(idx) == 1 && v_i %in% names(merged_final)) {
      merged_final[[v_i]][idx] <- val_i
      n_applied <- n_applied + 1L
    } else {
      cat(sprintf("ID %d or variable '%s' not found — skipped\n", id_i, v_i))
    }
  }

  cat("Applied:", n_applied, "values\n\n")

  # Updated missingness summary
  cat("Missingness after consensus update:\n")
  merged_final %>%
    select(all_of(review_vars)) %>%
    mutate(across(everything(), as.character)) %>%
    summarise(across(everything(),
                     ~ sum(is.na(.) | . %in% MISSING_CODES_CHR))) %>%
    pivot_longer(everything(),
                 names_to  = "variable",
                 values_to = "n_missing_or_flagged") %>%
    print()
}
```

---

# 14. Merge Q16, Q19, Q26, Q28 onto main dataset

```{r merge-supplemental}
merged_final2 <- merged_final %>%
  # Q28
  left_join(
    q28_clean %>% select(ID, MX28_WQ_COMP, MX28_WQ_COMP_CAT, MX28_WQ_COMP_INFO),
    by = "ID"
  ) %>%
  # Q26
  left_join(
    q26_clean %>% select(ID, MX26_EM_HHW, MX26_EM_HHW_INFO, MX26_EM_HHW_CAT),
    by = "ID"
  ) %>%
  # Q19
  left_join(
    q19_clean %>% select(ID, MX19_TIME_RAW, MX19_TIME_ENG, MX19_TIME_CAT),
    by = "ID"
  ) %>%
  # Q16
  left_join(
    q16_clean,
    by = "ID"
  )

cat("Final dimensions:", nrow(merged_final2), "rows ×", ncol(merged_final2), "cols\n")
cat("New columns added:", paste(setdiff(names(merged_final2), names(merged_final)), collapse = ", "), "\n")

# Missingness summary for new variables
new_vars <- c("MX28_WQ_COMP_CAT", "MX26_EM_HHW_CAT", "MX19_TIME_CAT",
              "MX16_SCEN_01", "MX16_SCEN_03", "MX16_SCEN_05", "MX16_SCEN_99")

cat("\nMissing-value counts for new variables (NA = ID not in supplemental file):\n")
merged_final2 %>%
  summarise(across(all_of(new_vars), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "n_NA") %>%
  arrange(desc(n_NA)) %>%
  print()
```

---

# 15. Export updated dataset

```{r export-v2}
write_csv(merged_final2, "./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv", na = "")
cat("Exported: 03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv\n")
cat("Rows:", nrow(merged_final2), "| Cols:", ncol(merged_final2), "\n")
```

---

# 16. Summary Table – Questions of Interest

Distributions for Q28, Q26, Q19, and Q16.
Sentinel/missing codes are shown separately from valid responses.

```{r summary-table}
# Helper: tabulate one variable, split valid vs. sentinel
summarise_q <- function(df, col, label, value_labels) {
  x <- df[[col]]

  valid  <- x[!is.na(x) & !(x %in% MISSING_CODES)]
  sentinels <- x[!is.na(x) &   x %in% MISSING_CODES]
  n_na   <- sum(is.na(x))

  sent_labels <- c(
    "999" = "Missing", "888" = "Don't know", "777" = "Refused",
    "666" = "Illegible", "555" = "Other", "444" = "Ambiguous", "333" = "Pending"
  )

  valid_rows <- if (length(valid) > 0) {
    tbl <- table(valid)
    tibble(
      Variable = label,
      Code     = names(tbl),
      Label    = value_labels[names(tbl)],
      n        = as.integer(tbl),
      pct_valid = round(100 * as.integer(tbl) / length(valid), 1),
      Type     = "Valid"
    )
  } else tibble()

  sent_rows <- if (length(sentinels) > 0) {
    tbl <- table(sentinels)
    tibble(
      Variable  = label,
      Code      = names(tbl),
      Label     = sent_labels[names(tbl)],
      n         = as.integer(tbl),
      pct_valid = NA_real_,
      Type      = "Sentinel/Missing"
    )
  } else tibble()

  na_rows <- if (n_na > 0) {
    tibble(Variable = label, Code = "NA", Label = "Not in file (no record)",
           n = n_na, pct_valid = NA_real_, Type = "Sentinel/Missing")
  } else tibble()

  bind_rows(valid_rows, sent_rows, na_rows)
}

lbl_q28  <- c("0" = "Better", "1" = "Same",     "2" = "Worse")
lbl_q26  <- c("0" = "Positive", "1" = "Negative")
lbl_q19  <- c("0" = "No",     "1" = "Yes")
lbl_q16  <- c("1" = "Option 1", "2" = "Option 2", "3" = "Option 3", "4" = "Option 4")

summary_tbl <- bind_rows(
  summarise_q(merged_final2, "MX28_WQ_COMP_CAT",
              "Q28 – Water Quality Comparison (0=Better, 1=Same, 2=Worse)", lbl_q28),
  summarise_q(merged_final2, "MX26_EM_HHW_CAT",
              "Q26 – Household Water Emotion (0=Positive, 1=Negative)",     lbl_q26),
  summarise_q(merged_final2, "MX19_TIME_CAT",
              "Q19 – Time Investment in Water Mgmt (0=No, 1=Yes)",          lbl_q19),
  summarise_q(merged_final2, "MX16_SCEN_01",
              "Q16 – Scenario 01 (Season 2 only)",                          lbl_q16),
  summarise_q(merged_final2, "MX16_SCEN_03",
              "Q16 – Scenario 03 (Season 2 only)",                          lbl_q16),
  summarise_q(merged_final2, "MX16_SCEN_05",
              "Q16 – Scenario 05 (Season 2 only)",                          lbl_q16),
  summarise_q(merged_final2, "MX16_SCEN_99",
              "Q16 – Scenario 99 (Season 2 only)",                          lbl_q16)
) %>%
  arrange(Variable, Type, Code)

cat("\n=== SUMMARY TABLE: Questions of Interest ===\n\n")
print(summary_tbl, n = 100)

knitr::kable(
  summary_tbl,
  caption = paste(
    "Distribution of Q16 (scenario preferences), Q19 (time investment),",
    "Q26 (household water emotions), and Q28 (water quality comparison).",
    "Percentages (pct_valid) are calculated over valid responses only.",
    "Missing codes: 444=Ambiguous, 555=Other, 999=Missing/not in file."
  ),
  col.names = c("Variable", "Code", "Label", "n", "% (valid only)", "Type"),
  digits = 1
)
```
