---
title: "Merge Datasets – Water Insecurity & Stress Study"
author: ""
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

## Overview

This script merges three supplemental files onto the base dataset
(`Merged_Screening_HWISE_PSS.csv`):

1. **Chronic_pain_illness.csv** – corrected `HLTH_CPAIN_CAT` and `HLTH_CDIS_CAT` values  
2. **Q2-3-4-5-6.csv** – supplemental water variables (Season 1, IDs 1–330)  
3. **Q1-2-3-4-5_season2.csv** – supplemental water variables (Season 2, IDs 331+)

---

## 1. Load base dataset

```{r load-base}
base <- read_csv("./data/Merged_Screening_HWISE_PSS.csv", show_col_types = FALSE)

cat("Base dimensions:", nrow(base), "rows ×", ncol(base), "cols\n")
cat("ID range:", min(base$ID), "–", max(base$ID), "\n")

```

---

## 2. Update chronic pain / illness categories

These replace the original `HLTH_CPAIN_CAT` and `HLTH_CDIS_CAT` columns with
corrected values.

```{r merge-chronic}
chronic <- read_csv(
  "./data/Chronic_pain_illness.csv",
  col_types = cols(
    ID             = col_double(),
    HLTH_CPAIN_CAT = col_double(),
    HLTH_CDIS_CAT  = col_double()
  )
)

cat("Chronic pain file:", nrow(chronic), "rows\n")
cat("IDs not in base:", sum(!chronic$ID %in% base$ID), "\n")

# Drop old columns and join corrected values
merged <- base %>%
  select(-HLTH_CPAIN_CAT, -HLTH_CDIS_CAT) %>%
  left_join(chronic, by = "ID")

# Verify counts
cat("After chronic merge:", nrow(merged), "rows\n")
cat("IDs not in base:", paste(chronic$ID[!chronic$ID %in% base$ID], collapse = ", "), "\n")
cat("IDs not in base:", paste(chronic$ID[!chronic$ID %in% base$ID], collapse = ", "), "\n")
```

---

## 3. Load and stack supplemental water files

### Season 1 water variables (Q2-3-4-5-6.csv)

```{r load-s1}
water_s1 <- read_csv(
  "./data/Q2-3-4-5-6.csv",
  show_col_types = FALSE
) %>%
  rename_with(~ str_replace_all(., "/", "_PER_")) %>%   # fix slash in col names
  mutate(WATER_FILE_SOURCE = "S1_Q2-3-4-5-6")

cat("Season 1 water rows:", nrow(water_s1), "\n")
cat("Season 1 IDs range:", min(water_s1$ID), "–", max(water_s1$ID), "\n")

cat("Sample IDs:", water_s1$ID)
cat("IDs not in base:", paste(chronic$ID[!chronic$ID %in% base$ID], collapse = ", "), "\n")
cat("IDs not in base:", paste(chronic$ID[!chronic$ID %in% base$ID], collapse = ", "), "\n")
```

### Season 2 water variables (Q1-2-3-4-5_season2.csv)

```{r load-s2}
water_s2 <- read_csv(
  "./data/Q1-2-3-4-5_season2.csv",
  show_col_types = FALSE
) %>%
  rename_with(~ str_replace_all(., "/", "_PER_")) %>%
  mutate(WATER_FILE_SOURCE = "S2_Q1-2-3-4-5")

cat("Season 2 water rows:", nrow(water_s2), "\n")
cat("Season 2 IDs range:", min(water_s2$ID), "–", max(water_s2$ID), "\n")
head(water_s2)
```

### Stack seasons (bind shared columns, retain all)

```{r stack-water}
# Full bind – fills NA for columns that differ between seasons
water_s1 <- water_s1 %>% mutate(ID = as.numeric(ID))
water_s2 <- water_s2 %>% mutate(ID = as.numeric(ID))

water_all <- bind_rows(water_s1, water_s2)
water_all <- bind_rows(water_s1, water_s2)

# Check for duplicate IDs across seasons
water_all <- bind_rows(water_s1, water_s2) %>%
  group_by(ID) %>%
  summarise(across(everything(), ~ {
    vals <- na.omit(unique(.x))
    if (length(vals) == 0) NA_character_
    else if (length(vals) == 1) as.character(vals)
    else paste(vals, collapse = " vs ")
  }), .groups = "drop") %>%
  mutate(ID = as.numeric(ID)) %>%
  arrange(ID)

cat("Rows after deduplication:", nrow(water_all), "\n")

dupes <- water_all %>% group_by(ID) %>% filter(n() > 1)

if (nrow(dupes) > 0) {
  conflicts <- dupes %>%
      group_by(ID) %>%
    summarise(across(everything(), ~ {
      vals <- na.omit(unique(.x))
      if (length(vals) > 1) paste(vals, collapse = " vs ") else NA_character_
    }), .groups = "drop") %>%
    arrange(ID) %>%
    select(ID, where(~ any(!is.na(.)))) %>%
    select(-any_of("WATER_FILE_SOURCE"))
  
  cat("Conflicting values by ID:\n")
  print(conflicts)
} else {
  cat("No duplicate IDs across seasons.\n")
}


cat("Combined water rows:", nrow(water_all), "\n")
head(water_all)
```

---

## 4. Merge water variables onto main dataset

```{r merge-water}
merged <- merged %>%
  left_join(water_all, by = "ID")

cat("Final dimensions:", nrow(merged), "rows ×", ncol(merged), "cols\n")

# IDs in base without a water record
no_water <- sum(!base$ID %in% water_all$ID)
cat("Base IDs with no water record:", no_water, "\n")
```

---

## 5. Reorder to match target column specification

Core columns are placed first in the required order; supplemental water columns
are appended after.

```{r reorder}
target_cols <- c(
  "ID", "D_YRBR", "D_AGE", "D_LOC", "D_LOC_CODE", "D_LOC_TIME", "D_LOC_INFO",
  "HLTH_SMK", "HLTH_TTM", "HLTH_MED", "HLTH_CPAIN", "HLTH_CPAIN_CAT",
  "HLTH_CDIS", "HLTH_CDIS_CAT", "D_HH_SIZE", "D_HH_SIZE_INFO", "D_CHLD",
  "D_EMP_TYPE", "D_EMP_DAYS", "W_HH_SIZE", "W_STORAGE", "D_INFO",
  "SES_EDU_CAT", "SES_EDU_SC", "SES_BTHR_CAT", "SES_BTHR_SC",
  "SES_CAR_CAT", "SES_CAR_SC", "SES_INT_CAT", "SES_INT_SC",
  "SES_WRK_CAT", "SES_WRK_SC", "SES_BEDR_CAT", "SES_BEDR_SC", "SES_INFO",
  "HLTH_TTM_RAW", "HLTH_MED_RAW", "HLTH_CPAIN_RAW", "HLTH_CDIS_RAW",
  "HLTH_SMK_RAW", "D_EMP_TYPE_RAW", "D_CHLD_RAW", "SEASON",
  "HLTH_TTM_CLEAN", "HLTH_MED_CLEAN", "HLTH_CPAIN_CLEAN", "HLTH_CDIS_CLEAN",
  "HLTH_SMK_CLEAN", "HLTH_SMK_CAT", "D_EMP_CLEAN", "D_EMP_FORM_CLEAN",
  "D_CHLD_CLEAN", "D_LOC_TIME_RAW", "D_LOC_TIME_CLEAN",
  "SES_EDU_SC_RAW", "SES_BTHR_SC_RAW", "SES_CAR_SC_RAW", "SES_INT_SC_RAW",
  "SES_WRK_SC_RAW", "SES_BEDR_SC_RAW", "SES_SC_Total", "SES_CAT_AMAI",
  "W_WS_LOC",
  "HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS", "HW_FOOD",
  "HW_HANDS", "HW_BODY", "HW_DRINK", "HW_ANGRY", "HW_SLEEP",
  "HW_NONE", "HW_SHAME", "HW_TOTAL", "HW_WORRY_SUB", "HW_INFO",
  "PSS1", "PSS2", "PSS3", "PSS8", "PSS11", "PSS12", "PSS14",
  "PSS4", "PSS5", "PSS6", "PSS7", "PSS9", "PSS10", "PSS13",
  "PSS4_REV", "PSS5_REV", "PSS6_REV", "PSS7_REV",
  "PSS9_REV", "PSS10_REV", "PSS13_REV",
  "PSS_TOTAL", "PSS_REVIEW_FLAG", "PSS_INFO"
)

# Identify which target cols are present
present   <- target_cols[target_cols %in% names(merged)]
missing   <- target_cols[!target_cols %in% names(merged)]
extra     <- names(merged)[!names(merged) %in% target_cols]

if (length(missing) > 0) cat("Target cols NOT found in merged data:\n", paste(missing, collapse = ", "), "\n")
if (length(extra)   > 0) cat("Extra cols appended after target cols:\n", paste(extra,   collapse = ", "), "\n")

# Reorder: target first, then extra (supplemental water + source flag)
merged_final <- merged %>%
  select(all_of(present), any_of(extra))

cat("Final column count:", ncol(merged_final), "\n")
```

---

## 6. Quality checks

```{r qc}
# Missing values in key outcome columns
qc_cols <- c("HLTH_CPAIN","HLTH_CPAIN_CAT", "HLTH_CPAIN_RAW", "HLTH_CPAIN_CLEAN", "HLTH_CDIS_CAT", "HW_TOTAL", "PSS_TOTAL", "W_WS_LOC", "MX3_CONT_INT","MX2_WI_SEASON_TYPE", "MX2_WI_SEASON_INFO", "MX2_WI_SEASON", "MX5_WFREQ_HRS_PER_WEEK")

print(merged_final[, qc_cols], n=5)

merged_final %>%
  summarise(across(all_of(qc_cols), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "column", values_to = "n_missing") %>%
  print()

# SEASON distribution
cat("\nSEASON distribution:\n")
print(table(merged_final$SEASON, useNA = "always"))
```

---

## 7. Export

```{r export}
write_csv(merged_final, "./data/Merged_Final.csv", na = "")
cat("Exported: Merged_Final.csv\n")
cat("Rows:", nrow(merged_final), "| Cols:", ncol(merged_final), "\n")
```
