---
title: "Descriptive Statistics"
author: "Paloma"
date: "2026-03-08"
output: workflowr::wflow_html
editor_options: 
  chunk_output_type: console
---

## Introduction

This document produces descriptive statistics for the merged dataset
(`Merged_Final.csv`, produced by `merge_datasets.Rmd`).
It includes:

1. A comprehensive descriptive statistics table (all analytic variables)

2. A filtered table with only the variables relevant to the manuscript

3. HWISE and PSS descriptive statistics (overall and by season)

4. Sample characteristics stratified by water insecurity level (HWISE categories)

5. A missing data summary table

6. Objective water supply variables (new): source, continuity, frequency, hours of supply, pressure, and seasonality

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(gtsummary)
library(gt)
library(ggplot2)

data_path <- "./data"

MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

# Helper: convert sentinel codes and recode to NA for descriptive stats
# Sentinel-coded values are treated as NA in tables (they are not true responses)
clean_for_table <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES | is.na(x_num), NA, x_num)
}
```

---

# 1. Load cleaned data

```{r load}
df <- read.csv(file.path(data_path, "Merged_Final.csv"),
               stringsAsFactors = FALSE,
               na.strings = c(""))

cat("Dimensions:", dim(df), "\n")
cat("Participants:", nrow(df), "\n")
```

---

# 2. Prepare analytic dataset

Select and recode variables for the descriptive table.
All sentinel codes are converted to NA here since they represent non-informative responses for descriptive purposes.

```{r prepare}
df_table <- df %>%
  mutate(
    # --- Demographics ---
    Age          = clean_for_table(D_AGE),
    Yrs_in_nbhd  = clean_for_table(D_LOC_TIME_CLEAN),
    HH_size      = clean_for_table(D_HH_SIZE),
    Parity       = clean_for_table(D_CHLD_CLEAN),

    # --- Season ---
    Season = factor(SEASON,
                    levels = c(0, 1),
                    labels = c("Rainy (Oct-Dec 2022)", "Dry (Apr-Jun 2023)")),

    # --- SES ---
    SES_Total    = clean_for_table(SES_SC_Total),
    SES_Category = factor(SES_CAT_AMAI,
                          levels = c("E", "D-", "D+", "C-", "C", "C+", "AB"),
                          ordered = TRUE),

    # --- Employment ---
    Employed = factor(clean_for_table(D_EMP_CLEAN),
                      levels = c(0, 1),
                      labels = c("No", "Yes")),
    Employment_type = factor(clean_for_table(D_EMP_FORM_CLEAN),
                             levels = c(0, 1),
                             labels = c("Informal", "Formal")),

    # --- Health ---
    Smoking = factor(clean_for_table(HLTH_SMK_CLEAN),
                     levels = c(0, 1, 2),
                     labels = c("No", "Yes", "Occasionally")),
    Hormonal_tx   = factor(clean_for_table(HLTH_TTM_CLEAN),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Medication    = factor(clean_for_table(HLTH_MED_CLEAN),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_pain  = factor(clean_for_table(HLTH_CPAIN_CLEAN),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_dis   = factor(clean_for_table(HLTH_CDIS_CLEAN),
                           levels = c(0, 1), labels = c("No", "Yes")),

    # --- HWISE ---
    HWISE_Total   = clean_for_table(HW_TOTAL),
    # Standard HWISE categories (Young et al. 2019, BMJ Global Health)
    # 0 = secure; 1-11 = marginal; 12-23 = moderate; 24-36 = severe
    HWISE_Cat     = factor(
      case_when(
        is.na(clean_for_table(HW_TOTAL))             ~ NA_character_,
        clean_for_table(HW_TOTAL) == 0               ~ "Secure (0)",
        clean_for_table(HW_TOTAL) <= 11              ~ "Marginal (1-11)",
        clean_for_table(HW_TOTAL) <= 23              ~ "Moderate (12-23)",
        TRUE                                         ~ "Severe (24-36)"
      ),
      levels = c("Secure (0)", "Marginal (1-11)",
                 "Moderate (12-23)", "Severe (24-36)")
    ),
    Water_worry   = clean_for_table(HW_WORRY_SUB),

    # --- PSS-14 ---
    PSS_Total     = clean_for_table(PSS_TOTAL),
    # Tertile-based PSS categories
    PSS_Cat       = factor(
      case_when(
        is.na(clean_for_table(PSS_TOTAL))            ~ NA_character_,
        clean_for_table(PSS_TOTAL) <= 22             ~ "Low stress (≤22)",
        clean_for_table(PSS_TOTAL) <= 31             ~ "Moderate stress (23-31)",
        TRUE                                         ~ "High stress (≥32)"
      ),
      levels = c("Low stress (≤22)", "Moderate stress (23-31)",
                 "High stress (≥32)")
    )
  )
head(df_table)
```

---

# 3. Comprehensive descriptive statistics table

Includes all analytic variables to inspect the full sample before deciding what to include in the manuscript table.

```{r table-full}
tbl_full <- df_table %>%
  select(
    # Demographics
    Season, Age, Yrs_in_nbhd, HH_size, Parity,
    # SES
    SES_Total, SES_Category,
    # Employment
    Employed, Employment_type,
    # Health
    Smoking, Hormonal_tx, Medication, Chronic_pain, Chronic_dis,
    # Water insecurity
    HWISE_Total, HWISE_Cat, Water_worry,
    # Perceived stress
    PSS_Total, PSS_Cat
  ) %>%
  tbl_summary(
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) / {median} [{p25}, {p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits    = list(all_continuous() ~ 1),
    missing   = "no",   # missing handled in separate table
    label     = list(
      Season          ~ "Season of data collection",
      Age             ~ "Age (years)",
      Yrs_in_nbhd     ~ "Years in neighborhood",
      HH_size         ~ "Household size (n persons)",
      Parity          ~ "Number of children",
      SES_Total       ~ "SES composite score (NSE-AMAI)",
      SES_Category    ~ "SES category (NSE-AMAI)",
      Employed        ~ "Currently employed",
      Employment_type ~ "Employment type (among employed)",
      Smoking         ~ "Smoking status",
      Hormonal_tx     ~ "Hormonal treatment",
      Medication      ~ "Regular medication use",
      Chronic_pain    ~ "Chronic pain",
      Chronic_dis     ~ "Chronic disease",
      HWISE_Total     ~ "HWISE total score (0–36)",
      HWISE_Cat       ~ "Water insecurity level",
      Water_worry     ~ "Water worry subscale (0–9)",
      PSS_Total       ~ "PSS-14 total score (0–56)",
      PSS_Cat         ~ "Perceived stress level"
    )
  ) %>%
  bold_labels() %>%
  modify_caption("**Table S1. Comprehensive descriptive statistics (n = {N})**")

tbl_full
```

---

# 4. Manuscript table (filtered)

Contains only variables directly relevant to the paper (for now).
Stratified by season to show comparability across data collection waves.

```{r table-manuscript}
tbl_ms <- df_table %>%
  select(
    Season, Age, Yrs_in_nbhd, HH_size, Parity,
    SES_Total, SES_Category,
    Employed, Employment_type,
    Smoking, Chronic_pain, Chronic_dis, W_WS_LOC
  ) %>%
  tbl_summary(
    by        = W_WS_LOC,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits    = list(all_continuous() ~ 1),
    missing   = "no",
    label     = list(
      Age             ~ "Age (years)",
      Yrs_in_nbhd     ~ "Years in neighborhood",
      HH_size         ~ "Household size",
      Parity          ~ "Parity (n children)",
      SES_Total       ~ "SES score (NSE-AMAI)",
      SES_Category    ~ "SES category",
      Employed        ~ "Employed",
      Employment_type ~ "Employment type",
      Smoking         ~ "Smoking",
      Chronic_pain    ~ "Chronic pain" #,
    #  Chronic_dis     ~ "Chronic disease"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season of data collection**") %>%
  modify_caption("**Table 1. Sample characteristics by season (n = {N})**")

tbl_ms
```

---

# 5. Missing data summary table

Reports true NAs and sentinel-coded values separately for all analytic
variables. Include as supplementary table in dissertation; summarize
in-text for the journal manuscript.

```{r table-missing}
# Variables to include in missing data report
analytic_cols <- c(
  "D_AGE", "D_LOC_TIME_CLEAN", "D_HH_SIZE", "D_CHLD_CLEAN",
  "SEASON", "SES_SC_Total", "SES_CAT_AMAI",
  "D_EMP_CLEAN", "D_EMP_FORM_CLEAN",
  "HLTH_SMK_CLEAN", "HLTH_TTM_CLEAN", "HLTH_MED_CLEAN",
  "HLTH_CPAIN_CLEAN", "HLTH_CDIS_CLEAN"
)

# Readable labels
var_labels <- c(
  D_AGE             = "Age",
  D_LOC_TIME_CLEAN  = "Years in neighborhood",
  D_HH_SIZE         = "Household size",
  D_CHLD_CLEAN      = "Parity",
  SEASON            = "Season",
  SES_SC_Total      = "SES composite score",
  SES_CAT_AMAI      = "SES category",
  D_EMP_CLEAN       = "Employed (yes/no)",
  D_EMP_FORM_CLEAN  = "Employment type",
  HLTH_SMK_CLEAN    = "Smoking",
  HLTH_TTM_CLEAN    = "Hormonal treatment",
  HLTH_MED_CLEAN    = "Medication use",
  HLTH_CPAIN_CLEAN  = "Chronic pain",
  HLTH_CDIS_CLEAN   = "Chronic disease"
)

N <- nrow(df)

missing_tbl <- df %>%
  select(all_of(analytic_cols)) %>%
  summarise(across(everything(), list(
    n_na      = ~ sum(is.na(.)),
    n_999     = ~ sum(as.character(.) == "999", na.rm = TRUE),
    n_888     = ~ sum(as.character(.) == "888", na.rm = TRUE),
    n_777     = ~ sum(as.character(.) == "777", na.rm = TRUE),
    n_666     = ~ sum(as.character(.) == "666", na.rm = TRUE),
    n_555     = ~ sum(as.character(.) == "555", na.rm = TRUE),
    n_444     = ~ sum(as.character(.) == "444", na.rm = TRUE),
    n_333     = ~ sum(as.character(.) == "333", na.rm = TRUE)
  ))) %>%
  pivot_longer(everything(),
               names_to  = c("variable", ".value"),
               names_sep = "_(?=n_)") %>%
  mutate(
    label        = var_labels[variable],
    n_total      = N,
    pct_na       = round(n_na   / N * 100, 1),
    n_flagged    = n_999 + n_888 + n_777 + n_666 + n_555 + n_444 + n_333,
    pct_flagged  = round(n_flagged / N * 100, 1),
    n_complete   = N - n_na - n_flagged,
    pct_complete = round(n_complete / N * 100, 1)
  ) %>%
  select(label, n_complete, pct_complete,
         n_na, pct_na,
         n_flagged, pct_flagged,
         n_999, n_888, n_777, n_666, n_555, n_444, n_333) %>%
  arrange(desc(n_na))

# Print full missing table
missing_tbl %>%
  gt() %>%
  tab_header(title = "Missing Data Summary") %>%
  tab_spanner(label = "Complete",   columns = c(n_complete, pct_complete)) %>%
  tab_spanner(label = "True NA",    columns = c(n_na, pct_na)) %>%
  tab_spanner(label = "Flagged",    columns = c(n_flagged, pct_flagged)) %>%
  tab_spanner(label = "Sentinel code breakdown",
              columns = c(n_999, n_888, n_777, n_666, n_555, n_444, n_333)) %>%
  cols_label(
    label        = "Variable",
    n_complete   = "n", pct_complete = "%",
    n_na         = "n", pct_na       = "%",
    n_flagged    = "n", pct_flagged  = "%",
    n_999 = "999", n_888 = "888", n_777 = "777", n_666 = "666",
    n_555 = "555", n_444 = "444", n_333 = "333"
  ) %>%
  fmt_number(columns = where(is.numeric), decimals = 0) %>%
  tab_footnote(
    footnote = "999=Missing/lost, 888=Don't know, 777=Refused, 666=Illegible, 555=Other, 444=Ambiguous, 333=Pending"
  )
```

---

# 6. Visualization: missing data overview

```{r plot-missing, fig.width=8, fig.height=5}
missing_tbl %>%
  select(label, pct_complete, pct_na, pct_flagged) %>%
  pivot_longer(-label, names_to = "type", values_to = "pct") %>%
  mutate(type = factor(type,
                       levels = c("pct_complete", "pct_na", "pct_flagged"),
                       labels = c("Complete", "True NA", "Flagged (sentinel)")),
         label = factor(label, levels = rev(missing_tbl$label))) %>%
  ggplot(aes(x = label, y = pct, fill = type)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = c("Complete"            = "#5bc0de",
                               "True NA"             = "#d9534f",
                               "Flagged (sentinel)"  = "#f0ad4e")) +
  labs(title = "Missing data by variable",
       x = NULL, y = "Percentage (%)", fill = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")
```


---

# 7. HWISE and PSS descriptive statistics

## 7.1 Overall HWISE and PSS summary

```{r hwise-pss-overall}
tbl_scores <- df_table %>%
  select(HWISE_Total, HWISE_Cat, Water_worry, PSS_Total, PSS_Cat) %>%
  tbl_summary(
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) / {median} [{p25}, {p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      HWISE_Total  ~ "HWISE total score (0-36)",
      HWISE_Cat    ~ "Water insecurity level",
      Water_worry  ~ "Water worry subscale (0-9)",
      PSS_Total    ~ "PSS-14 total score (0-56)",
      PSS_Cat      ~ "Perceived stress level"
    )
  ) %>%
  bold_labels() %>%
  modify_caption("**Table. HWISE and PSS scores — full sample (n = {N})**")

tbl_scores
```

## 7.2 HWISE and PSS by season

```{r hwise-pss-by-season}
tbl_scores_season <- df_table %>%
  select(Season, HWISE_Total, HWISE_Cat, Water_worry, PSS_Total, PSS_Cat) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      HWISE_Total  ~ "HWISE total score (0-36)",
      HWISE_Cat    ~ "Water insecurity level",
      Water_worry  ~ "Water worry subscale (0-9)",
      PSS_Total    ~ "PSS-14 total score (0-56)",
      PSS_Cat      ~ "Perceived stress level"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season**") %>%
  modify_caption("**Table. HWISE and PSS scores by season (n = {N})**")

tbl_scores_season
```

## 7.3 Distribution plots

```{r hwise-pss-plots, fig.width=9, fig.height=4, message=FALSE, warning=FALSE}
library(gridExtra)

p_hwise <- ggplot(df_table %>% filter(!is.na(HWISE_Total)),
                  aes(x = HWISE_Total, fill = Season)) +
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.6) +
  scale_fill_manual(values = c("Rainy (Oct-Dec 2022)" = "#5bc0de",
                               "Dry (Apr-Jun 2023)"   = "#d9534f")) +
  labs(title = "HWISE total score",
       subtitle = paste0("n = ", sum(!is.na(df_table$HWISE_Total))),
       x = "Score (0-36)", y = "Count", fill = NULL) +
  theme_minimal() + theme(legend.position = "bottom")

p_pss <- ggplot(df_table %>% filter(!is.na(PSS_Total)),
                aes(x = PSS_Total, fill = Season)) +
  geom_histogram(binwidth = 2, position = "identity", alpha = 0.6) +
  scale_fill_manual(values = c("Rainy (Oct-Dec 2022)" = "#5bc0de",
                               "Dry (Apr-Jun 2023)"   = "#d9534f")) +
  labs(title = "PSS-14 total score",
       subtitle = paste0("n = ", sum(!is.na(df_table$PSS_Total))),
       x = "Score (0-56)", y = "Count", fill = NULL) +
  theme_minimal() + theme(legend.position = "bottom")

grid.arrange(p_hwise, p_pss, ncol = 2)
```

---

# 8. Sample characteristics by water insecurity level

Demographics and health variables stratified by HWISE category, using the
standard cut-points from Young et al. (2019, *BMJ Global Health*):
0 = water secure; 1-11 = marginal; 12-23 = moderate; 24-36 = severe.
One-way ANOVA used for continuous variables; chi-square for categorical.

```{r table-by-hwise}
tbl_by_hwise <- df_table %>%
  filter(!is.na(HWISE_Cat)) %>%
  select(
    HWISE_Cat,
    Age, Yrs_in_nbhd, HH_size, Parity,
    SES_Total, SES_Category,
    Employed, Smoking, Chronic_pain, Chronic_dis,
    PSS_Total, PSS_Cat, Water_worry
  ) %>%
  tbl_summary(
    by        = HWISE_Cat,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      Age          ~ "Age (years)",
      Yrs_in_nbhd  ~ "Years in neighborhood",
      HH_size      ~ "Household size",
      Parity       ~ "Parity (n children)",
      SES_Total    ~ "SES score",
      SES_Category ~ "SES category",
      Employed     ~ "Employed",
      Smoking      ~ "Smoking",
      Chronic_pain ~ "Chronic pain",
      Chronic_dis  ~ "Chronic disease",
      PSS_Total    ~ "PSS-14 total score",
      PSS_Cat      ~ "Perceived stress level",
      Water_worry  ~ "Water worry subscale"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "aov",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4") ~
                           "**Water insecurity level (HWISE)**") %>%
  modify_caption("**Table. Sample characteristics by water insecurity level (n = {N})**")

tbl_by_hwise
```

---

# 9. PSS and water worry by water insecurity level

```{r pss-by-hwise-plot, fig.width=9, fig.height=4}
p_pss_hwise <- df_table %>%
  filter(!is.na(HWISE_Cat) & !is.na(PSS_Total)) %>%
  ggplot(aes(x = HWISE_Cat, y = PSS_Total, fill = HWISE_Cat)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  stat_summary(fun.data = function(x) data.frame(y = -2, label = paste0("n=", length(x))),
               geom = "text", size = 3.2) +
  scale_fill_manual(values = c(
    "Secure (0)"       = "#2ecc71",
    "Marginal (1-11)"  = "#f1c40f",
    "Moderate (12-23)" = "#e67e22",
    "Severe (24-36)"   = "#e74c3c"
  )) +
  labs(title = "PSS-14 by water insecurity level",
       x = NULL, y = "PSS-14 score (0-56)") +
  theme_minimal() + theme(legend.position = "none")

p_worry_hwise <- df_table %>%
  filter(!is.na(HWISE_Cat) & !is.na(Water_worry)) %>%
  ggplot(aes(x = HWISE_Cat, y = Water_worry, fill = HWISE_Cat)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  stat_summary(fun.data = function(x) data.frame(y = -0.5, label = paste0("n=", length(x))),
               geom = "text", size = 3.2) +
  scale_fill_manual(values = c(
    "Secure (0)"       = "#2ecc71",
    "Marginal (1-11)"  = "#f1c40f",
    "Moderate (12-23)" = "#e67e22",
    "Severe (24-36)"   = "#e74c3c"
  )) +
  scale_y_continuous(breaks = 0:9) +
  labs(title = "Water worry subscale by water insecurity level",
       x = NULL, y = "Water worry score (0-9)") +
  theme_minimal() + theme(legend.position = "none")

grid.arrange(p_pss_hwise, p_worry_hwise, ncol = 2)
```

---

# 10. Objective water supply variables (new)

These variables were added from the supplemental water questionnaires (Q2-3-4-5-6 for Season 1;
Q1-2-3-4-5 for Season 2). They capture objective, interviewer-recorded characteristics of each
participant's household water supply: source, continuity, frequency, hours per week, and pressure.

Sentinel codes used in these variables:

- **999** = Missing / lost / data entry error
- **333** = Pending (data exists but not yet entered)
- **888** = Don't know
- **777** = Refused / prefer not to answer
- **666** = Illegible
- **555** = Other (answer outside allowed categories)
- **444** = Ambiguous

All sentinel-coded values are excluded from analytic summaries and plots but are reported
separately in the missing data table (Section 10.5).

```{r water-vars-prepare}
# Helper: flag sentinel codes in free-text water fields
# These fields are character, so sentinels may appear as strings
SENTINEL_STRINGS <- as.character(c(999, 888, 777, 666, 555, 444, 333))

is_sentinel_str <- function(x) {
  str_trim(as.character(x)) %in% SENTINEL_STRINGS
}

# Harmonise and recode the raw water supply variables
# Sentinel-coded entries are set to NA for analysis; tracked separately below
df_water <- df %>%
  mutate(
    # Water source (Season 2 only: MX1_W_SRC)
    W_Source = case_when(
      is_sentinel_str(MX1_W_SRC)                              ~ NA_character_,
      str_detect(tolower(MX1_W_SRC), "servicio|publ")         ~ "Public network",
      str_detect(tolower(MX1_W_SRC), "pipa|truck")            ~ "Water truck (pipa)",
      str_detect(tolower(MX1_W_SRC), "pozo|well")             ~ "Well",
      MX1_W_SRC %in% c("", "N/A") | is.na(MX1_W_SRC)        ~ NA_character_,
      TRUE                                                     ~ "Other"
    ),

    # Continuity of supply (MX3_CONT_INT — both seasons)
    W_Continuity = case_when(
      is_sentinel_str(MX3_CONT_INT)                                  ~ NA_character_,
      str_detect(tolower(MX3_CONT_INT), "^contin")                   ~ "Continuous",
      str_detect(tolower(MX3_CONT_INT), "^intermi|cortan|deja")      ~ "Intermittent",
      str_detect(tolower(MX3_CONT_INT), "disminu")                   ~ "Reduced pressure/flow",
      MX3_CONT_INT %in% c("", "N/A") | is.na(MX3_CONT_INT)         ~ NA_character_,
      TRUE                                                            ~ "Other"
    ),
    W_Continuity = factor(W_Continuity,
                          levels = c("Continuous", "Intermittent",
                                     "Reduced pressure/flow", "Other")),

    # Frequency of supply (MX4_WFREQ / MX4_W_FREQ — harmonised across seasons)
    W_Freq_raw = coalesce(MX4_WFREQ, MX4_W_FREQ),
    W_Frequency = case_when(
      is_sentinel_str(W_Freq_raw)                                    ~ NA_character_,
      str_detect(tolower(W_Freq_raw), "^diari|every day")            ~ "Daily",
      str_detect(tolower(W_Freq_raw), "^intermi")                    ~ "Intermittent",
      str_detect(tolower(W_Freq_raw), "1 a 3|1-3")                   ~ "1–3 days/week",
      str_detect(tolower(W_Freq_raw), "4 a 6|4-6")                   ~ "4–6 days/week",
      str_detect(tolower(W_Freq_raw), "^contin")                     ~ "Continuous",
      W_Freq_raw %in% c("", "N/A") | is.na(W_Freq_raw)             ~ NA_character_,
      TRUE                                                            ~ "Other"
    ),
    W_Frequency = factor(W_Frequency,
                         levels = c("Continuous", "Daily", "4–6 days/week",
                                    "1–3 days/week", "Intermittent", "Other")),

    # Hours per week (MX5_WFREQ_HRS_PER_WEEK)
    # First strip sentinel strings, then parse first numeric value
    W_Hrs_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_HRS_PER_WEEK), NA,
             as.numeric(str_extract(MX5_WFREQ_HRS_PER_WEEK, "^[0-9\\.]+")))
    ),
    # Also catch sentinel values that parsed as numbers
    W_Hrs_week = ifelse(W_Hrs_week %in% MISSING_CODES, NA, W_Hrs_week),
    # Cap implausible values (>168 hrs/week is impossible)
    W_Hrs_week = ifelse(W_Hrs_week > 168, NA, W_Hrs_week),

    # Days per week (MX5_WFREQ_NDAYS)
    W_Days_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_NDAYS), NA,
             as.numeric(str_extract(MX5_WFREQ_NDAYS, "^[0-9]+")))
    ),
    W_Days_week = ifelse(W_Days_week %in% MISSING_CODES, NA, W_Days_week),
    W_Days_week = ifelse(W_Days_week > 7, NA, W_Days_week),

    # Water pressure (MX6_WPRESS)
    W_Pressure = case_when(
      is_sentinel_str(MX6_WPRESS)                                    ~ NA_character_,
      str_detect(tolower(MX6_WPRESS), "^s[ií]|^yes|^buena|^bien")   ~ "Yes",
      str_detect(tolower(MX6_WPRESS), "^no|^baja|^poca")            ~ "No/Low",
      MX6_WPRESS %in% c("", "N/A") | is.na(MX6_WPRESS)            ~ NA_character_,
      TRUE                                                            ~ "Variable"
    ),
    W_Pressure = factor(W_Pressure, levels = c("Yes", "No/Low", "Variable")),

    # Dry-season scarcity (MX2_WI_SEASON)
    W_Season_scarcity = case_when(
      is_sentinel_str(MX2_WI_SEASON)                                 ~ NA_character_,
      str_detect(tolower(MX2_WI_SEASON), "n/a|no hay|no se|^no$")   ~ "No scarcity reported",
      MX2_WI_SEASON %in% c("", "N/A") | is.na(MX2_WI_SEASON)      ~ NA_character_,
      TRUE                                                            ~ "Seasonal scarcity reported"
    ),
    W_Season_scarcity = factor(W_Season_scarcity,
                                levels = c("No scarcity reported",
                                           "Seasonal scarcity reported")),

    # Neighborhood factor for plots
    W_WS_LOC_F = factor(W_WS_LOC, levels = c("WS", "WI"),
                         labels = c("Water Secure", "Water Insecure"))
  )
```

## 10.1 Summary table: objective water supply

```{r water-vars-table}
tbl_water <- df_water %>%
  select(
    W_WS_LOC_F,
    W_Source, W_Continuity, W_Frequency,
    W_Days_week, W_Hrs_week,
    W_Pressure, W_Season_scarcity
  ) %>%
  tbl_summary(
    by        = W_WS_LOC_F,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) / {median} [{p25}, {p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "ifany",
    label   = list(
      W_Source          ~ "Water source (Season 2 only)",
      W_Continuity      ~ "Supply continuity",
      W_Frequency       ~ "Supply frequency",
      W_Days_week       ~ "Days of supply per week",
      W_Hrs_week        ~ "Hours of supply per week",
      W_Pressure        ~ "Adequate water pressure",
      W_Season_scarcity ~ "Seasonal scarcity reported"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Neighborhood type**") %>%
  modify_caption("**Table. Objective water supply characteristics by neighborhood (n = {N})**")

tbl_water
```

## 10.2 Distribution of hours of supply per week

```{r water-hrs-plot, fig.width=8, fig.height=4, message=FALSE}
df_water %>%
  filter(!is.na(W_Hrs_week) & !is.na(W_WS_LOC_F)) %>%
  ggplot(aes(x = W_Hrs_week, fill = W_WS_LOC_F)) +
  geom_histogram(binwidth = 6, position = "identity", alpha = 0.6, color = "white") +
  scale_fill_manual(values = c("Water Secure" = "#4E8098", "Water Insecure" = "#E76F51")) +
  labs(
    title    = "Hours of water supply per week by neighborhood",
    subtitle = paste0("n = ", sum(!is.na(df_water$W_Hrs_week) & !is.na(df_water$W_WS_LOC_F)),
                      " (excludes sentinel codes and missing)"),
    x = "Hours per week", y = "Count", fill = NULL
  ) +
  theme_minimal() + theme(legend.position = "bottom")
```

## 10.3 Supply continuity and frequency by neighborhood

```{r water-cat-plots, fig.width=10, fig.height=5}
# Pre-compute counts for label annotations
cont_counts <- df_water %>%
  filter(!is.na(W_Continuity) & !is.na(W_WS_LOC_F)) %>%
  count(W_WS_LOC_F, W_Continuity)

freq_counts <- df_water %>%
  filter(!is.na(W_Frequency) & !is.na(W_WS_LOC_F)) %>%
  count(W_WS_LOC_F, W_Frequency)

p_cont <- ggplot(cont_counts,
                 aes(x = W_Continuity, y = n, fill = W_WS_LOC_F)) +
  geom_col(position = "dodge", alpha = 0.85) +
  geom_text(aes(label = n),
            position = position_dodge(width = 0.9),
            vjust = -0.4, size = 3.2) +
  scale_fill_manual(values = c("Water Secure" = "#4E8098", "Water Insecure" = "#E76F51")) +
  labs(title = "Supply continuity", x = NULL, y = "Count", fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1), legend.position = "bottom")

p_freq <- ggplot(freq_counts,
                 aes(x = W_Frequency, y = n, fill = W_WS_LOC_F)) +
  geom_col(position = "dodge", alpha = 0.85) +
  geom_text(aes(label = n),
            position = position_dodge(width = 0.9),
            vjust = -0.4, size = 3.2) +
  scale_fill_manual(values = c("Water Secure" = "#4E8098", "Water Insecure" = "#E76F51")) +
  labs(title = "Supply frequency", x = NULL, y = "Count", fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1), legend.position = "bottom")

grid.arrange(p_cont, p_freq, ncol = 2)
```

## 10.4 Objective water supply vs HWISE and PSS

```{r water-supply-vs-outcomes, fig.width=10, fig.height=4, message=FALSE}
df_water_scores <- df_water %>%
  mutate(
    HWISE_Total = clean_for_table(HW_TOTAL),
    PSS_Total   = clean_for_table(PSS_TOTAL)
  ) %>%
  filter(!is.na(W_Hrs_week) & !is.na(HWISE_Total) & !is.na(PSS_Total))

n_label <- nrow(df_water_scores)

p_hw <- ggplot(df_water_scores, aes(x = W_Hrs_week, y = HWISE_Total)) +
  geom_point(alpha = 0.4, size = 1.5, color = "#90A955") +
  geom_smooth(method = "lm", se = TRUE, color = "#386641") +
  labs(title = "Hours of supply vs HWISE",
       subtitle = paste0("n = ", n_label),
       x = "Hours of supply per week", y = "HWISE total score") +
  theme_minimal()

p_pss <- ggplot(df_water_scores, aes(x = W_Hrs_week, y = PSS_Total)) +
  geom_point(alpha = 0.4, size = 1.5, color = "#4E8098") +
  geom_smooth(method = "lm", se = TRUE, color = "#1d3557") +
  labs(title = "Hours of supply vs PSS",
       subtitle = paste0("n = ", n_label),
       x = "Hours of supply per week", y = "PSS-14 total score") +
  theme_minimal()

grid.arrange(p_hw, p_pss, ncol = 2)
```

## 10.5 Sentinel code summary for water supply variables

Reports how many entries per variable were sentinel-coded (excluded from analysis above).

```{r water-sentinel-summary}
water_raw_cols <- c(
  "MX1_W_SRC", "MX2_WI_SEASON", "MX3_CONT_INT",
  "MX4_WFREQ", "MX4_W_FREQ",
  "MX5_WFREQ_NDAYS", "MX5_WFREQ_NHRS", "MX5_WFREQ_HRS_PER_WEEK",
  "MX6_WPRESS"
)

sentinel_labels <- c(
  "999" = "Missing/lost/entry error",
  "333" = "Pending",
  "888" = "Don't know",
  "777" = "Refused",
  "666" = "Illegible",
  "555" = "Other (outside categories)",
  "444" = "Ambiguous"
)

N_total <- nrow(df_water)

# Build long table: one row per variable × sentinel code
sentinel_tbl <- purrr::map_dfr(water_raw_cols, function(col) {
  if (!col %in% names(df_water)) return(NULL)
  x <- str_trim(as.character(df_water[[col]]))
  purrr::map_dfr(names(sentinel_labels), function(code) {
    n <- sum(x == code, na.rm = TRUE)
    if (n > 0) {
      tibble(
        variable      = col,
        sentinel_code = code,
        meaning       = sentinel_labels[[code]],
        n             = n,
        pct           = round(n / N_total * 100, 1)
      )
    }
  })
})

if (nrow(sentinel_tbl) == 0) {
  cat("No sentinel codes found in water supply variables.\n")
} else {
  sentinel_tbl %>%
    arrange(variable, sentinel_code) %>%
    gt() %>%
    tab_header(title = "Sentinel codes in water supply variables") %>%
    cols_label(
      variable      = "Variable",
      sentinel_code = "Code",
      meaning       = "Meaning",
      n             = "n",
      pct           = "%"
    ) %>%
    tab_footnote(paste0("% of total sample (N = ", N_total, ")"))
}


```
