---
title: "Descriptive Statistics"
author: "Paloma"
date: "2026-03-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

This document produces descriptive statistics tables for the merged dataset
(`03.Merged_Q1-6.csv`, produced by `03.Data_merge_Q1-6.Rmd`).
It includes:

1. A comprehensive descriptive statistics table (all analytic variables)

2. A manuscript table with sample characteristics stratified by season

3. A missing data summary table

4. HWISE and PSS descriptive statistics (overall and by season)

5. Sample characteristics stratified by water insecurity level (HWISE categories)

6. Objective water supply characteristics (source, continuity, frequency, hours of supply, pressure, and seasonality)

Visualizations for all of the above are in `05.Descriptive_plots.Rmd`.

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(gtsummary)
library(gt)
library(flextable)
library(purrr)
library(officer)
library(ggplot2)

data_path <- "./data"

MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

# Helper: convert sentinel codes to NA for descriptive stats
clean_for_table <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES | is.na(x_num), NA, x_num)
}

# Helper: flag sentinel codes in free-text / character fields
SENTINEL_STRINGS <- as.character(c(999, 888, 777, 666, 555, 444, 333))
is_sentinel_str <- function(x) {
  str_trim(as.character(x)) %in% SENTINEL_STRINGS
}
```

---

# 1. Load cleaned data

```{r load}
df <- read.csv(file.path(data_path, "03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv"),
               stringsAsFactors = FALSE,
               na.strings = c(""))

cat("Dimensions:", dim(df), "\n")
cat("Participants:", nrow(df), "\n")
```

---

# 2. Prepare analytic dataset

Select and recode variables for the descriptive tables.
All sentinel codes are converted to NA since they represent non-informative responses.

```{r prepare}

edu_levels <- c(
  "No formal education",
  "Incomplete primary (1–5 yrs)",
  "Complete primary",
  "Incomplete secondary (7–8 yrs)",
  "Complete secondary",
  "Incomplete high school (10–11 yrs)",
  "Complete high school",
  "Incomplete university (13–15 yrs)",
  "Complete university",
  "Postgraduate"
)

df_table <- df %>%
  mutate(
    # --- Demographics ---
    Age          = clean_for_table(D_AGE),
    Yrs_in_nbhd  = clean_for_table(D_LOC_TIME_CLEAN),
    HH_size      = clean_for_table(D_HH_SIZE),
    Parity       = clean_for_table(D_CHLD_CLEAN),

    # --- Season ---
    Season = factor(SEASON,
                    levels = c(0, 1),
                    labels = c("Rainy", "Dry")),

    # --- SES ---
    Education    = factor(
      dplyr::recode(as.character(SES_EDU_CAT_CLEAN),
        "Incomplete primary"    = "Incomplete primary (1–5 yrs)",
        "Incomplete secondary"  = "Incomplete secondary (7–8 yrs)",
        "Incomplete high school" = "Incomplete high school (10–11 yrs)",
        "Incomplete university"  = "Incomplete university (13–15 yrs)"
      ),
      levels = edu_levels, ordered = TRUE
    ),
    SES          = factor(SES_CAT_AMAI,
                          levels = c("E", "D-", "D+", "C-", "C", "C+", "AB"),
                          ordered = TRUE),

    # --- Employment ---
    Employed = factor(clean_for_table(D_EMP_CLEAN),
                      levels = c(0, 1),
                      labels = c("No", "Yes")),
    Employment_type = factor(clean_for_table(D_EMP_FORM_CLEAN),
                             levels = c(0, 1),
                             labels = c("Informal", "Formal")),

    # --- Health ---
    Smoking = factor(clean_for_table(HLTH_SMK_CLEAN),
                     levels = c(0, 1, 2),
                     labels = c("No", "Yes", "Occasionally")),
    Hormonal_tx   = factor(clean_for_table(HLTH_TTM_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Medication    = factor(clean_for_table(HLTH_MED_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_pain  = factor(clean_for_table(HLTH_CPAIN_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_dis   = factor(clean_for_table(HLTH_CDIS_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),

    # --- HWISE ---
    HWISE_Total   = clean_for_table(HW_TOTAL),
    # Standard HWISE categories (Young et al. 2019, BMJ Global Health)
    # 0 = secure; 1-11 = marginal; 12-23 = moderate; 24-36 = severe
    HWISE_Cat     = factor(
      case_when(
        is.na(clean_for_table(HW_TOTAL))             ~ NA_character_,
        clean_for_table(HW_TOTAL) == 0               ~ "Secure (0)",
        clean_for_table(HW_TOTAL) <= 11              ~ "Marginal (1-11)",
        clean_for_table(HW_TOTAL) <= 23              ~ "Moderate (12-23)",
        TRUE                                         ~ "Severe (24-36)"
      ),
      levels = c("Secure (0)", "Marginal (1-11)",
                 "Moderate (12-23)", "Severe (24-36)")
    ),
    Water_worry   = clean_for_table(HW_WORRY_SUB),

    # --- Individual HWISE items ---
HW_Worry    = clean_for_table(HW_WORRY),
HW_Interr   = clean_for_table(HW_INTERR),
HW_Clothes  = clean_for_table(HW_CLOTHES),
HW_Plans    = clean_for_table(HW_PLANS),
HW_Food     = clean_for_table(HW_FOOD),
HW_Hands    = clean_for_table(HW_HANDS),
HW_Body     = clean_for_table(HW_BODY),
HW_Drink    = clean_for_table(HW_DRINK),
HW_Angry    = clean_for_table(HW_ANGRY),
HW_Sleep    = clean_for_table(HW_SLEEP),
HW_None     = clean_for_table(HW_NONE),
HW_Shame    = clean_for_table(HW_SHAME),
    # --- PSS-14 ---
    PSS_Total     = clean_for_table(PSS_TOTAL),
    # Tertile-based PSS categories
    PSS_Cat       = factor(
      case_when(
        is.na(clean_for_table(PSS_TOTAL))            ~ NA_character_,
        clean_for_table(PSS_TOTAL) <= 22             ~ "Low stress (≤22)",
        clean_for_table(PSS_TOTAL) <= 31             ~ "Moderate stress (23-31)",
        TRUE                                         ~ "High stress (≥32)"
      ),
      levels = c("Low stress (≤22)", "Moderate stress (23-31)",
                 "High stress (≥32)")
    )
  )
```

---

# 3. Supplementary descriptive statistics table

Includes all analytic variables. Intended as a supplementary table for the dissertation.

```{r table-full}
tbl_full <- df_table %>%
  select(Age, Yrs_in_nbhd, HH_size, Parity,
         SES, Education, Employed, Employment_type,
         Smoking, Hormonal_tx, Medication, Chronic_pain, Chronic_dis,
         Season) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    type    = list(Parity ~ "continuous"),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      Age             ~ "Age (years)",
      Yrs_in_nbhd     ~ "Years in neighborhood",
      HH_size         ~ "Household size",
      Parity          ~ "Parity",
      SES             ~ "SES (NSE-AMAI)",
      Education       ~ "Education",
      Employed        ~ "Currently employed",
      Employment_type ~ "Employment type",
      Smoking         ~ "Smoking status",
      Hormonal_tx     ~ "Hormonal treatment",
      Medication      ~ "Regular medication use",
      Chronic_pain    ~ "Chronic pain",
      Chronic_dis     ~ "Chronic disease"
    )
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  add_n(statistic = "{p_miss}%",
        col_label = "**Missing (%)**") %>%
  modify_table_body(~ .x %>% relocate(n, .after = last_col())) %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**By season**") %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Table S1. Sample characteristics — full variable set (n = {N})**")

tbl_full
```

---

# 4. Manuscript table

Sample characteristics stratified by season to show comparability across data collection waves.

```{r table-manuscript}
tbl_ms <- df_table %>%
  select(
    Season, Age, Yrs_in_nbhd, HH_size, Parity,
    SES, Education,
    Employed, Employment_type,
    Smoking, Chronic_pain, Chronic_dis
  ) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    type      = list(Parity ~ "continuous"),
    digits    = list(all_continuous() ~ 1),
    missing   = "no",
    label     = list(
      Age             ~ "Age",
      Yrs_in_nbhd     ~ "Years in neighborhood",
      HH_size         ~ "Household size",
      Parity          ~ "Parity",
      SES             ~ "SES (NSE-AMAI)",
      Education       ~ "Education",
      Employed        ~ "Employed",
      Employment_type ~ "Employment type",
      Smoking         ~ "Smoking",
      Chronic_pain    ~ "Chronic pain",
      Chronic_dis     ~ "Chronic disease"
    )
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season of data collection**") %>%
  modify_header(label = "**Variable**") %>%
  modify_caption("**Table 1. Sample characteristics by season**")

tbl_ms
```

---

# 5. Missing data summary table

Reports true NAs and sentinel-coded values separately for all analytic variables.
Include as supplementary table in dissertation; summarize in-text for the journal manuscript.
See `05.Descriptive_plots.Rmd` for the corresponding visualization.

```{r table-missing}
analytic_cols <- c(
  "D_AGE", "D_LOC_TIME_CLEAN", "D_HH_SIZE", "D_CHLD_CLEAN",
  "SEASON", "SES_SC_Total", "SES_CAT_AMAI",
  "D_EMP_CLEAN", "D_EMP_FORM_CLEAN",
  "HLTH_SMK_CLEAN", "HLTH_TTM_CAT", "HLTH_MED_CAT",
  "HLTH_CPAIN_CAT", "HLTH_CDIS_CAT"
)

var_labels <- c(
  D_AGE             = "Age",
  D_LOC_TIME_CLEAN  = "Years in neighborhood",
  D_HH_SIZE         = "Household size",
  D_CHLD_CLEAN      = "Parity",
  SEASON            = "Season",
  SES_SC_Total      = "SES composite score",
  SES_CAT_AMAI      = "SES category",
  D_EMP_CLEAN       = "Employed (yes/no)",
  D_EMP_FORM_CLEAN  = "Employment type",
  HLTH_SMK_CLEAN    = "Smoking",
  HLTH_TTM_CAT      = "Hormonal treatment",
  HLTH_MED_CAT      = "Medication use",
  HLTH_CPAIN_CAT    = "Chronic pain",
  HLTH_CDIS_CAT     = "Chronic disease"
)

N <- nrow(df)

missing_tbl <- df %>%
  select(all_of(analytic_cols)) %>%
  summarise(across(everything(), list(
    n_na  = ~ sum(is.na(.)),
    n_999 = ~ sum(as.character(.) == "999", na.rm = TRUE),
    n_888 = ~ sum(as.character(.) == "888", na.rm = TRUE),
    n_777 = ~ sum(as.character(.) == "777", na.rm = TRUE),
    n_666 = ~ sum(as.character(.) == "666", na.rm = TRUE),
    n_555 = ~ sum(as.character(.) == "555", na.rm = TRUE),
    n_444 = ~ sum(as.character(.) == "444", na.rm = TRUE),
    n_333 = ~ sum(as.character(.) == "333", na.rm = TRUE)
  ))) %>%
  pivot_longer(everything(),
               names_to  = c("variable", ".value"),
               names_sep = "_(?=n_)") %>%
  mutate(
    label        = var_labels[variable],
    n_total      = N,
    pct_na       = round(n_na   / N * 100, 1),
    n_flagged    = n_999 + n_888 + n_777 + n_666 + n_555 + n_444 + n_333,
    pct_flagged  = round(n_flagged / N * 100, 1),
    n_complete   = N - n_na - n_flagged,
    pct_complete = round(n_complete / N * 100, 1)
  ) %>%
  select(label, n_complete, pct_complete,
         n_na, pct_na,
         n_flagged, pct_flagged,
         n_999, n_888, n_777, n_666, n_555, n_444, n_333) %>%
  arrange(desc(n_na))

missing_tbl_print <- missing_tbl %>%
  gt() %>%
  tab_header(title = "Missing Data Summary") %>%
  tab_spanner(label = "Complete",  columns = c(n_complete, pct_complete)) %>%
  tab_spanner(label = "True NA",   columns = c(n_na, pct_na)) %>%
  tab_spanner(label = "Flagged",   columns = c(n_flagged, pct_flagged)) %>%
  tab_spanner(label = "Sentinel code breakdown",
              columns = c(n_999, n_888, n_777, n_666, n_555, n_444, n_333)) %>%
  cols_label(
    label        = "Variable",
    n_complete   = "n", pct_complete = "%",
    n_na         = "n", pct_na       = "%",
    n_flagged    = "n", pct_flagged  = "%",
    n_999 = "999", n_888 = "888", n_777 = "777", n_666 = "666",
    n_555 = "555", n_444 = "444", n_333 = "333"
  ) %>%
  fmt_number(columns = where(is.numeric), decimals = 0) %>%
  tab_footnote(
    footnote = "999=Missing/lost, 888=Don't know, 777=Refused, 666=Illegible, 555=Other, 444=Ambiguous, 333=Pending"
  )

missing_tbl_print
```

---

# 6. HWISE and PSS descriptive statistics

## 6.1 Overall summary

```{r hwise-pss-overall}
tbl_scores <- df_table %>%
  select(HWISE_Total, HWISE_Cat, Water_worry, PSS_Total, PSS_Cat) %>%
  tbl_summary(
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) / {median} [{p25}, {p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      HWISE_Total  ~ "HWISE total score (0-36)",
      HWISE_Cat    ~ "Water insecurity level",
      Water_worry  ~ "Water worry subscale (0-9)",
      PSS_Total    ~ "PSS-14 total score (0-56)",
      PSS_Cat      ~ "Perceived stress level"
    )
  ) %>%
  bold_labels() %>%
  modify_caption("**Table. HWISE and PSS scores — full sample (n = {N})**")

tbl_scores
```

## 6.2 By season

```{r hwise-pss-by-season}
tbl_scores_season <- df_table %>%
  select(Season, HWISE_Total, HWISE_Cat, Water_worry, PSS_Total, PSS_Cat) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      HWISE_Total  ~ "HWISE total score (0-36)",
      HWISE_Cat    ~ "Water insecurity level",
      Water_worry  ~ "Water worry subscale (0-9)",
      PSS_Total    ~ "PSS-14 total score (0-56)",
      PSS_Cat      ~ "Perceived stress level"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season**") %>%
  modify_caption("**Table. HWISE and PSS scores by season (n = {N})**")

tbl_scores_season
```

---

# 7. Sample characteristics by water insecurity level

Demographics and health variables stratified by HWISE category, using the
standard cut-points from Young et al. (2019, *BMJ Global Health*):
0 = water secure; 1-11 = marginal; 12-23 = moderate; 24-36 = severe.
One-way ANOVA used for continuous variables; chi-square for categorical.

```{r table-by-hwise}
tbl_by_hwise <- df_table %>%
  filter(!is.na(HWISE_Cat)) %>%
  select(
    HWISE_Cat,
    Age, Yrs_in_nbhd, HH_size, Parity,
    SES, Education,
    Employed, Smoking, Chronic_pain, Chronic_dis,
    PSS_Total, PSS_Cat, Water_worry
  ) %>%
  tbl_summary(
    by        = HWISE_Cat,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      Age          ~ "Age (years)",
      Yrs_in_nbhd  ~ "Years in neighborhood",
      HH_size      ~ "Household size",
      Parity       ~ "Parity (n children)",
      SES          ~ "SES (NSE-AMAI)",
      Education    ~ "Education",
      Employed     ~ "Employed",
      Smoking      ~ "Smoking",
      Chronic_pain ~ "Chronic pain",
      Chronic_dis  ~ "Chronic disease",
      PSS_Total    ~ "PSS-14 total score",
      PSS_Cat      ~ "Perceived stress level",
      Water_worry  ~ "Water worry subscale"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "aov",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4") ~
                           "**Water insecurity level (HWISE)**") %>%
  modify_caption("**Table. Sample characteristics by water insecurity level (n = {N})**")

tbl_by_hwise
```

## 7. HWISE item-level table

This table shows the distribution of each HWISE item (scored 0–3: never, rarely,
sometimes, always), the water worry subscale (HW1), the HWISE total score, and
the overall HWISE category. Presented overall, stratified by neighborhood
(water secure vs. water insecure), and by season. Kruskal-Wallis test used for
item scores and totals (ordinal); chi-square for categories.

## 7.1 Overall and by neighborhood
```{r hwise-items-by-neighborhood}
hwise_item_labels <- list(
  HW_Worry   ~ "1. Worried about water",
  HW_Interr  ~ "2. Water interrupted",
  HW_Clothes ~ "3. Unable to wash clothes",
  HW_Plans   ~ "4. Changed plans due to water",
  HW_Food    ~ "5. Unable to prepare food",
  HW_Hands   ~ "6. Unable to wash hands",
  HW_Body    ~ "7. Unable to wash body",
  HW_Drink   ~ "8. No water to drink",
  HW_Angry   ~ "9. Felt angry about water",
  HW_Sleep   ~ "10. Lost sleep over water",
  HW_None    ~ "11. Had no water at all",
  HW_Shame   ~ "12. Felt shame about water",
  Water_worry  ~ "Water worry subscale (items 1+9)",
  HWISE_Total  ~ "HWISE total score (0–36)",
  HWISE_Cat    ~ "Water insecurity category"
)

tbl_hwise_nbhd <- df_table %>%
  filter(!is.na(W_WS_LOC)) %>%
  select(W_WS_LOC,
         HW_Worry, HW_Interr, HW_Clothes, HW_Plans,
         HW_Food, HW_Hands, HW_Body, HW_Drink,
         HW_Angry, HW_Sleep, HW_None, HW_Shame,
         Water_worry, HWISE_Total, HWISE_Cat) %>%
  tbl_summary(
    by        = W_WS_LOC,
    statistic = list(
      all_continuous()  ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = hwise_item_labels
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "kruskal.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  add_q(method = "BH") %>%
  bold_p(t = 0.05) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Neighborhood**") %>%
  modify_caption("**Table. HWISE item scores by neighborhood (n = {N})**") %>%
  modify_footnote(
    all_stat_cols() ~ "Median (Q1, Q3) for continuous; n (%) for categorical.",
    p.value ~ "Kruskal-Wallis for item scores/totals; chi-square for categories.",
    q.value ~ "FDR-adjusted p-value (Benjamini-Hochberg)."
  )

tbl_hwise_nbhd
```

## 7.2 By season
```{r hwise-items-by-season}
tbl_hwise_season <- df_table %>%
  filter(!is.na(Season)) %>%
  select(Season,
         HW_Worry, HW_Interr, HW_Clothes, HW_Plans,
         HW_Food, HW_Hands, HW_Body, HW_Drink,
         HW_Angry, HW_Sleep, HW_None, HW_Shame,
         Water_worry, HWISE_Total, HWISE_Cat) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = hwise_item_labels
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "wilcox.test",   # 2 groups only
    all_categorical() ~ "chisq.test"
  )) %>%
  add_q(method = "BH") %>%
  bold_p(t = 0.05) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season**") %>%
  modify_caption("**Table. HWISE item scores by season (n = {N})**") %>%
  modify_footnote(
    all_stat_cols() ~ "Median (Q1, Q3) for continuous; n (%) for categorical.",
    p.value ~ "Wilcoxon rank-sum for item scores/totals; chi-square for categories.",
    q.value ~ "FDR-adjusted p-value (Benjamini-Hochberg)."
  )

tbl_hwise_season
```

## 7.3 HWISE item response distributions

```{r hwise-response-dist, fig.width = 9, fig.height = 6, message = FALSE, warning = FALSE}
library(tidyr)
library(forcats)

hwise_item_names <- c(
  "HW_Worry"   = "1. Worried about water",
  "HW_Interr"  = "2. Water interrupted",
  "HW_Clothes" = "3. Unable to wash clothes",
  "HW_Plans"   = "4. Changed plans due to water",
  "HW_Food"    = "5. Unable to prepare food",
  "HW_Hands"   = "6. Unable to wash hands",
  "HW_Body"    = "7. Unable to wash body",
  "HW_Drink"   = "8. No water to drink",
  "HW_Angry"   = "9. Felt angry about water",
  "HW_Sleep"   = "10. Lost sleep over water",
  "HW_None"    = "11. Had no water at all",
  "HW_Shame"   = "12. Felt shame about water"
)

response_labels <- c(
  "0" = "Never",
  "1" = "Rarely",
  "2" = "Sometimes",
  "3" = "Often/Always"
)

hwise_long <- df_table %>%
  select(all_of(names(hwise_item_names))) %>%
  pivot_longer(everything(), names_to = "item", values_to = "response") %>%
  filter(!is.na(response)) %>%
  mutate(
    item     = factor(hwise_item_names[item], levels = rev(hwise_item_names)),
    response = factor(response, levels = c(0, 1, 2, 3),
                      labels = response_labels)
  ) %>%
  count(item, response) %>%
  group_by(item) %>%
  mutate(pct = n / sum(n) * 100) %>%
  ungroup()

ggplot(hwise_long, aes(x = pct, y = item, fill = response)) +
  geom_col(position = "stack", width = 0.7) +
  geom_text(
    aes(label = ifelse(pct >= 5, paste0(round(pct), "%"), "")),
    position = position_stack(vjust = 0.5),
    size = 3, color = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Never"        = "#2166ac",
      "Rarely"       = "#74add1",
      "Sometimes"    = "#f4a582",
      "Often/Always" = "#d6604d"
    ),
    name = "Response"
  ) +
  scale_x_continuous(labels = function(x) paste0(x, "%"), expand = c(0, 0)) +
  labs(
    title    = "HWISE item response distributions",
    subtitle = "Percentage of respondents per response category (n excludes missing)",
    x        = "Percentage of respondents",
    y        = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position   = "bottom",
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.text.y        = element_text(size = 10)
  )
```

---

# 8. Objective water supply variables

These variables capture objective, interviewer-recorded characteristics of each
participant's household water supply. Sentinel codes are excluded from summaries
and tracked in the audit table at the end of this section.

Sentinel code key:

- **999** = Missing / lost / data entry error
- **333** = Pending
- **888** = Don't know
- **777** = Refused
- **666** = Illegible
- **555** = Other
- **444** = Ambiguous

## 8.1 Prepare water supply variables

```{r water-vars-prepare}
df_water <- df %>%
  mutate(
    # Water source (Season 2 only: MX1_W_SRC)
    W_Source = case_when(
      is_sentinel_str(MX1_W_SRC)                              ~ NA_character_,
      str_detect(tolower(MX1_W_SRC), "servicio|publ")         ~ "Public network",
      str_detect(tolower(MX1_W_SRC), "pipa|truck")            ~ "Water truck (pipa)",
      str_detect(tolower(MX1_W_SRC), "pozo|well")             ~ "Well",
      MX1_W_SRC %in% c("", "N/A") | is.na(MX1_W_SRC)        ~ NA_character_,
      TRUE                                                     ~ "Other"
    ),

    # Continuity of supply (MX3_CONT_INT — both seasons)
    W_Continuity = case_when(
      is_sentinel_str(MX3_CONT_INT)                                  ~ NA_character_,
      str_detect(tolower(MX3_CONT_INT), "^contin")                   ~ "Continuous",
      str_detect(tolower(MX3_CONT_INT), "^intermi|cortan|deja")      ~ "Intermittent",
      str_detect(tolower(MX3_CONT_INT), "disminu")                   ~ "Reduced pressure/flow",
      MX3_CONT_INT %in% c("", "N/A") | is.na(MX3_CONT_INT)         ~ NA_character_,
      TRUE                                                            ~ "Other"
    ),
    W_Continuity = factor(W_Continuity,
                          levels = c("Continuous", "Intermittent",
                                     "Reduced pressure/flow", "Other")),

    # Frequency of supply (MX4_WFREQ / MX4_W_FREQ — harmonised across seasons)
    W_Freq_raw = coalesce(MX4_WFREQ, MX4_W_FREQ),
    W_Frequency = case_when(
      is_sentinel_str(W_Freq_raw)                                    ~ NA_character_,
      str_detect(tolower(W_Freq_raw), "^diari|every day")            ~ "Daily",
      str_detect(tolower(W_Freq_raw), "^intermi")                    ~ "Intermittent",
      str_detect(tolower(W_Freq_raw), "1 a 3|1-3")                   ~ "1–3 days/week",
      str_detect(tolower(W_Freq_raw), "4 a 6|4-6")                   ~ "4–6 days/week",
      str_detect(tolower(W_Freq_raw), "^contin")                     ~ "Continuous",
      W_Freq_raw %in% c("", "N/A") | is.na(W_Freq_raw)             ~ NA_character_,
      TRUE                                                            ~ "Other"
    ),
    W_Frequency = factor(W_Frequency,
                         levels = c("Continuous", "Daily", "4–6 days/week",
                                    "1–3 days/week", "Intermittent", "Other")),

    # Hours per week (MX5_WFREQ_HRS_PER_WEEK)
    W_Hrs_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_HRS_PER_WEEK), NA,
             as.numeric(str_extract(MX5_WFREQ_HRS_PER_WEEK, "^[0-9\\.]+")))
    ),
    W_Hrs_week = ifelse(W_Hrs_week %in% MISSING_CODES, NA, W_Hrs_week),
    W_Hrs_week = ifelse(W_Hrs_week > 168, NA, W_Hrs_week),

    # Days per week (MX5_WFREQ_NDAYS)
    W_Days_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_NDAYS), NA,
             as.numeric(str_extract(MX5_WFREQ_NDAYS, "^[0-9]+")))
    ),
    W_Days_week = ifelse(W_Days_week %in% MISSING_CODES, NA, W_Days_week),
    W_Days_week = ifelse(W_Days_week > 7, NA, W_Days_week),

    # Water pressure (MX6_WPRESS)
    W_Pressure = case_when(
      is_sentinel_str(MX6_WPRESS)                                    ~ NA_character_,
      str_detect(tolower(MX6_WPRESS), "^s[ií]|^yes|^buena|^bien")   ~ "Yes",
      str_detect(tolower(MX6_WPRESS), "^no|^baja|^poca")            ~ "No/Low",
      MX6_WPRESS %in% c("", "N/A") | is.na(MX6_WPRESS)            ~ NA_character_,
      TRUE                                                            ~ "Variable"
    ),
    W_Pressure = factor(W_Pressure, levels = c("Yes", "No/Low", "Variable")),

    # Neighborhood factor
    W_WS_LOC_F = factor(W_WS_LOC, levels = c("WS", "WI"),
                         labels = c("Water Secure", "Water Insecure"))
  )
```

## 8.2 Summary table by neighborhood

```{r water-vars-table}
tbl_water <- df_water %>%
  select(
    W_WS_LOC_F,
    W_Source, W_Continuity, W_Frequency,
    W_Days_week, W_Hrs_week,
    W_Pressure
  ) %>%
  tbl_summary(
    by        = W_WS_LOC_F,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) / {median} [{p25}, {p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "ifany",
    label   = list(
      W_Source          ~ "Water source (Season 2 only)",
      W_Continuity      ~ "Supply continuity",
      W_Frequency       ~ "Supply frequency",
      W_Days_week       ~ "Days of supply per week",
      W_Hrs_week        ~ "Hours of supply per week",
      W_Pressure        ~ "Adequate water pressure"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Neighborhood type**") %>%
  modify_caption("**Table. Objective water supply characteristics by neighborhood (n = {N})**")

tbl_water
```

## 8.3 Sentinel code audit for water supply variables

```{r water-sentinel-summary}
water_raw_cols <- c(
  "MX1_W_SRC", "MX3_CONT_INT",
  "MX4_WFREQ", "MX4_W_FREQ",
  "MX5_WFREQ_NDAYS", "MX5_WFREQ_NHRS", "MX5_WFREQ_HRS_PER_WEEK",
  "MX6_WPRESS"
)

sentinel_labels <- c(
  "999" = "Missing/lost/entry error",
  "333" = "Pending",
  "888" = "Don't know",
  "777" = "Refused",
  "666" = "Illegible",
  "555" = "Other (outside categories)",
  "444" = "Ambiguous"
)

N_total <- nrow(df_water)

sentinel_tbl <- map_dfr(water_raw_cols, function(col) {
  if (!col %in% names(df_water)) return(NULL)
  x <- str_trim(as.character(df_water[[col]]))
  map_dfr(names(sentinel_labels), function(code) {
    n <- sum(x == code, na.rm = TRUE)
    if (n > 0) {
      tibble(
        variable      = col,
        sentinel_code = code,
        meaning       = sentinel_labels[[code]],
        n             = n,
        pct           = round(n / N_total * 100, 1)
      )
    }
  })
})

if (nrow(sentinel_tbl) == 0) {
  cat("No sentinel codes found in water supply variables.\n")
} else {
  sentinel_tbl %>%
    arrange(variable, sentinel_code) %>%
    gt() %>%
    tab_header(title = "Sentinel codes in water supply variables") %>%
    cols_label(
      variable      = "Variable",
      sentinel_code = "Code",
      meaning       = "Meaning",
      n             = "n",
      pct           = "%"
    ) %>%
    tab_footnote(paste0("% of total sample (N = ", N_total, ")"))
}
```

---

# 9. Export tables to Word

```{r export-tables, message=FALSE, warning=FALSE}
output_dir <- "./output"
if (!dir.exists(output_dir)) dir.create(output_dir)

tables <- list(
  "Table1_sample_characteristics.docx"     = tbl_ms,
  "TableS1_full_descriptives.docx"         = tbl_full,
  "TableS2_HWISE_PSS_overall.docx"         = tbl_scores,
  "TableS3_HWISE_PSS_by_season.docx"       = tbl_scores_season,
  "TableS4_characteristics_by_HWISE.docx"  = tbl_by_hwise,
  "TableS5_water_supply.docx"              = tbl_water,
  "TableS6_water_by_neighborhood.docx"     = tbl_hwise_nbhd,
  "TableS7_water_by_season.docx"          = tbl_hwise_season
)

for (fname in names(tables)) {
  tables[[fname]] %>%
    as_flex_table() %>%
    save_as_docx(path = file.path(output_dir, fname))
  cat("Exported:", fname, "\n")
}

missing_tbl_print %>%
  gtsave(file.path(output_dir, "TableS1.2_missing_data.docx"))

```


