---
title: "Descriptive Statistics"
author: "Paloma"
date: "2026-03-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

This document produces descriptive statistics tables for the merged dataset
(`03.Merged_Q1-6.csv`, produced by `03.Data_merge_Q1-6.Rmd`).
It includes:

1. A comprehensive descriptive statistics table (all analytic variables)

2. A manuscript table with sample characteristics stratified by season

3. A missing data summary table

4. HWISE and PSS descriptive statistics (overall and by season)

5. Sample characteristics stratified by water insecurity level (HWISE categories)

6. Objective water supply characteristics (source, continuity, frequency, hours of supply, pressure, and seasonality)

Visualizations for all of the above are in `05.Descriptive_plots.Rmd`.

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(gtsummary)
library(gt)
library(purrr)

data_path <- "./data"

MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

# Helper: convert sentinel codes to NA for descriptive stats
clean_for_table <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES | is.na(x_num), NA, x_num)
}

# Helper: flag sentinel codes in free-text / character fields
SENTINEL_STRINGS <- as.character(c(999, 888, 777, 666, 555, 444, 333))
is_sentinel_str <- function(x) {
  str_trim(as.character(x)) %in% SENTINEL_STRINGS
}
```

---

# 1. Load cleaned data

```{r load}
df <- read.csv(file.path(data_path, "03.Merged_Q1-6.csv"),
               stringsAsFactors = FALSE,
               na.strings = c(""))

cat("Dimensions:", dim(df), "\n")
cat("Participants:", nrow(df), "\n")
```

---

# 2. Prepare analytic dataset

Select and recode variables for the descriptive tables.
All sentinel codes are converted to NA since they represent non-informative responses.

```{r prepare}
df_table <- df %>%
  mutate(
    # --- Demographics ---
    Age          = clean_for_table(D_AGE),
    Yrs_in_nbhd  = clean_for_table(D_LOC_TIME_CLEAN),
    HH_size      = clean_for_table(D_HH_SIZE),
    Parity       = clean_for_table(D_CHLD_CLEAN),

    # --- Season ---
    Season = factor(SEASON,
                    levels = c(0, 1),
                    labels = c("Rainy (Oct-Dec 2022)", "Dry (Apr-Jun 2023)")),

    # --- SES ---
    SES_Total    = clean_for_table(SES_SC_Total),
    SES_Category = factor(SES_CAT_AMAI,
                          levels = c("E", "D-", "D+", "C-", "C", "C+", "AB"),
                          ordered = TRUE),

    # --- Employment ---
    Employed = factor(clean_for_table(D_EMP_CLEAN),
                      levels = c(0, 1),
                      labels = c("No", "Yes")),
    Employment_type = factor(clean_for_table(D_EMP_FORM_CLEAN),
                             levels = c(0, 1),
                             labels = c("Informal", "Formal")),

    # --- Health ---
    Smoking = factor(clean_for_table(HLTH_SMK_CLEAN),
                     levels = c(0, 1, 2),
                     labels = c("No", "Yes", "Occasionally")),
    Hormonal_tx   = factor(clean_for_table(HLTH_TTM_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Medication    = factor(clean_for_table(HLTH_MED_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_pain  = factor(clean_for_table(HLTH_CPAIN_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_dis   = factor(clean_for_table(HLTH_CDIS_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),

    # --- HWISE ---
    HWISE_Total   = clean_for_table(HW_TOTAL),
    # Standard HWISE categories (Young et al. 2019, BMJ Global Health)
    # 0 = secure; 1-11 = marginal; 12-23 = moderate; 24-36 = severe
    HWISE_Cat     = factor(
      case_when(
        is.na(clean_for_table(HW_TOTAL))             ~ NA_character_,
        clean_for_table(HW_TOTAL) == 0               ~ "Secure (0)",
        clean_for_table(HW_TOTAL) <= 11              ~ "Marginal (1-11)",
        clean_for_table(HW_TOTAL) <= 23              ~ "Moderate (12-23)",
        TRUE                                         ~ "Severe (24-36)"
      ),
      levels = c("Secure (0)", "Marginal (1-11)",
                 "Moderate (12-23)", "Severe (24-36)")
    ),
    Water_worry   = clean_for_table(HW_WORRY_SUB),

    # --- PSS-14 ---
    PSS_Total     = clean_for_table(PSS_TOTAL),
    # Tertile-based PSS categories
    PSS_Cat       = factor(
      case_when(
        is.na(clean_for_table(PSS_TOTAL))            ~ NA_character_,
        clean_for_table(PSS_TOTAL) <= 22             ~ "Low stress (≤22)",
        clean_for_table(PSS_TOTAL) <= 31             ~ "Moderate stress (23-31)",
        TRUE                                         ~ "High stress (≥32)"
      ),
      levels = c("Low stress (≤22)", "Moderate stress (23-31)",
                 "High stress (≥32)")
    )
  )
```

---

# 3. Comprehensive descriptive statistics table

Includes all analytic variables to inspect the full sample.

```{r table-full}
tbl_full <- df_table %>%
  select(
    # Demographics
    Season, Age, Yrs_in_nbhd, HH_size, Parity,
    # SES
    SES_Total, SES_Category,
    # Employment
    Employed, Employment_type,
    # Health
    Smoking, Hormonal_tx, Medication, Chronic_pain, Chronic_dis,
    # Water insecurity
    HWISE_Total, HWISE_Cat, Water_worry,
    # Perceived stress
    PSS_Total, PSS_Cat
  ) %>%
  tbl_summary(
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) / {median} [{p25}, {p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits    = list(all_continuous() ~ 1),
    missing   = "no",
    label     = list(
      Season          ~ "Season of data collection",
      Age             ~ "Age (years)",
      Yrs_in_nbhd     ~ "Years in neighborhood",
      HH_size         ~ "Household size (n persons)",
      Parity          ~ "Number of children",
      SES_Total       ~ "SES composite score (NSE-AMAI)",
      SES_Category    ~ "SES category (NSE-AMAI)",
      Employed        ~ "Currently employed",
      Employment_type ~ "Employment type (among employed)",
      Smoking         ~ "Smoking status",
      Hormonal_tx     ~ "Hormonal treatment",
      Medication      ~ "Regular medication use",
      Chronic_pain    ~ "Chronic pain",
      Chronic_dis     ~ "Chronic disease",
      HWISE_Total     ~ "HWISE total score (0–36)",
      HWISE_Cat       ~ "Water insecurity level",
      Water_worry     ~ "Water worry subscale (0–9)",
      PSS_Total       ~ "PSS-14 total score (0–56)",
      PSS_Cat         ~ "Perceived stress level"
    )
  ) %>%
  bold_labels() %>%
  modify_caption("**Table S1. Comprehensive descriptive statistics (n = {N})**")

tbl_full
```

---

# 4. Manuscript table

Sample characteristics stratified by season to show comparability across data collection waves.

```{r table-manuscript}
tbl_ms <- df_table %>%
  select(
    Season, Age, Yrs_in_nbhd, HH_size, Parity,
    SES_Total, SES_Category,
    Employed, Employment_type,
    Smoking, Chronic_pain, Chronic_dis
  ) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits    = list(all_continuous() ~ 1),
    missing   = "no",
    label     = list(
      Age             ~ "Age (years)",
      Yrs_in_nbhd     ~ "Years in neighborhood",
      HH_size         ~ "Household size",
      Parity          ~ "Parity (n children)",
      SES_Total       ~ "SES score (NSE-AMAI)",
      SES_Category    ~ "SES category",
      Employed        ~ "Employed",
      Employment_type ~ "Employment type",
      Smoking         ~ "Smoking",
      Chronic_pain    ~ "Chronic pain",
      Chronic_dis     ~ "Chronic disease"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season of data collection**") %>%
  modify_caption("**Table 1. Sample characteristics by season (n = {N})**")

tbl_ms
```

---

# 5. Missing data summary table

Reports true NAs and sentinel-coded values separately for all analytic variables.
Include as supplementary table in dissertation; summarize in-text for the journal manuscript.
See `05.Descriptive_plots.Rmd` for the corresponding visualization.

```{r table-missing}
analytic_cols <- c(
  "D_AGE", "D_LOC_TIME_CLEAN", "D_HH_SIZE", "D_CHLD_CLEAN",
  "SEASON", "SES_SC_Total", "SES_CAT_AMAI",
  "D_EMP_CLEAN", "D_EMP_FORM_CLEAN",
  "HLTH_SMK_CLEAN", "HLTH_TTM_CAT", "HLTH_MED_CAT",
  "HLTH_CPAIN_CAT", "HLTH_CDIS_CAT"
)

var_labels <- c(
  D_AGE             = "Age",
  D_LOC_TIME_CLEAN  = "Years in neighborhood",
  D_HH_SIZE         = "Household size",
  D_CHLD_CLEAN      = "Parity",
  SEASON            = "Season",
  SES_SC_Total      = "SES composite score",
  SES_CAT_AMAI      = "SES category",
  D_EMP_CLEAN       = "Employed (yes/no)",
  D_EMP_FORM_CLEAN  = "Employment type",
  HLTH_SMK_CLEAN    = "Smoking",
  HLTH_TTM_CAT      = "Hormonal treatment",
  HLTH_MED_CAT      = "Medication use",
  HLTH_CPAIN_CAT    = "Chronic pain",
  HLTH_CDIS_CAT     = "Chronic disease"
)

N <- nrow(df)

missing_tbl <- df %>%
  select(all_of(analytic_cols)) %>%
  summarise(across(everything(), list(
    n_na  = ~ sum(is.na(.)),
    n_999 = ~ sum(as.character(.) == "999", na.rm = TRUE),
    n_888 = ~ sum(as.character(.) == "888", na.rm = TRUE),
    n_777 = ~ sum(as.character(.) == "777", na.rm = TRUE),
    n_666 = ~ sum(as.character(.) == "666", na.rm = TRUE),
    n_555 = ~ sum(as.character(.) == "555", na.rm = TRUE),
    n_444 = ~ sum(as.character(.) == "444", na.rm = TRUE),
    n_333 = ~ sum(as.character(.) == "333", na.rm = TRUE)
  ))) %>%
  pivot_longer(everything(),
               names_to  = c("variable", ".value"),
               names_sep = "_(?=n_)") %>%
  mutate(
    label        = var_labels[variable],
    n_total      = N,
    pct_na       = round(n_na   / N * 100, 1),
    n_flagged    = n_999 + n_888 + n_777 + n_666 + n_555 + n_444 + n_333,
    pct_flagged  = round(n_flagged / N * 100, 1),
    n_complete   = N - n_na - n_flagged,
    pct_complete = round(n_complete / N * 100, 1)
  ) %>%
  select(label, n_complete, pct_complete,
         n_na, pct_na,
         n_flagged, pct_flagged,
         n_999, n_888, n_777, n_666, n_555, n_444, n_333) %>%
  arrange(desc(n_na))

missing_tbl %>%
  gt() %>%
  tab_header(title = "Missing Data Summary") %>%
  tab_spanner(label = "Complete",  columns = c(n_complete, pct_complete)) %>%
  tab_spanner(label = "True NA",   columns = c(n_na, pct_na)) %>%
  tab_spanner(label = "Flagged",   columns = c(n_flagged, pct_flagged)) %>%
  tab_spanner(label = "Sentinel code breakdown",
              columns = c(n_999, n_888, n_777, n_666, n_555, n_444, n_333)) %>%
  cols_label(
    label        = "Variable",
    n_complete   = "n", pct_complete = "%",
    n_na         = "n", pct_na       = "%",
    n_flagged    = "n", pct_flagged  = "%",
    n_999 = "999", n_888 = "888", n_777 = "777", n_666 = "666",
    n_555 = "555", n_444 = "444", n_333 = "333"
  ) %>%
  fmt_number(columns = where(is.numeric), decimals = 0) %>%
  tab_footnote(
    footnote = "999=Missing/lost, 888=Don't know, 777=Refused, 666=Illegible, 555=Other, 444=Ambiguous, 333=Pending"
  )
```

---

# 6. HWISE and PSS descriptive statistics

## 6.1 Overall summary

```{r hwise-pss-overall}
tbl_scores <- df_table %>%
  select(HWISE_Total, HWISE_Cat, Water_worry, PSS_Total, PSS_Cat) %>%
  tbl_summary(
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) / {median} [{p25}, {p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      HWISE_Total  ~ "HWISE total score (0-36)",
      HWISE_Cat    ~ "Water insecurity level",
      Water_worry  ~ "Water worry subscale (0-9)",
      PSS_Total    ~ "PSS-14 total score (0-56)",
      PSS_Cat      ~ "Perceived stress level"
    )
  ) %>%
  bold_labels() %>%
  modify_caption("**Table. HWISE and PSS scores — full sample (n = {N})**")

tbl_scores
```

## 6.2 By season

```{r hwise-pss-by-season}
tbl_scores_season <- df_table %>%
  select(Season, HWISE_Total, HWISE_Cat, Water_worry, PSS_Total, PSS_Cat) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      HWISE_Total  ~ "HWISE total score (0-36)",
      HWISE_Cat    ~ "Water insecurity level",
      Water_worry  ~ "Water worry subscale (0-9)",
      PSS_Total    ~ "PSS-14 total score (0-56)",
      PSS_Cat      ~ "Perceived stress level"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season**") %>%
  modify_caption("**Table. HWISE and PSS scores by season (n = {N})**")

tbl_scores_season
```

---

# 7. Sample characteristics by water insecurity level

Demographics and health variables stratified by HWISE category, using the
standard cut-points from Young et al. (2019, *BMJ Global Health*):
0 = water secure; 1-11 = marginal; 12-23 = moderate; 24-36 = severe.
One-way ANOVA used for continuous variables; chi-square for categorical.

```{r table-by-hwise}
tbl_by_hwise <- df_table %>%
  filter(!is.na(HWISE_Cat)) %>%
  select(
    HWISE_Cat,
    Age, Yrs_in_nbhd, HH_size, Parity,
    SES_Total, SES_Category,
    Employed, Smoking, Chronic_pain, Chronic_dis,
    PSS_Total, PSS_Cat, Water_worry
  ) %>%
  tbl_summary(
    by        = HWISE_Cat,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      Age          ~ "Age (years)",
      Yrs_in_nbhd  ~ "Years in neighborhood",
      HH_size      ~ "Household size",
      Parity       ~ "Parity (n children)",
      SES_Total    ~ "SES score",
      SES_Category ~ "SES category",
      Employed     ~ "Employed",
      Smoking      ~ "Smoking",
      Chronic_pain ~ "Chronic pain",
      Chronic_dis  ~ "Chronic disease",
      PSS_Total    ~ "PSS-14 total score",
      PSS_Cat      ~ "Perceived stress level",
      Water_worry  ~ "Water worry subscale"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "aov",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4") ~
                           "**Water insecurity level (HWISE)**") %>%
  modify_caption("**Table. Sample characteristics by water insecurity level (n = {N})**")

tbl_by_hwise
```

---

# 8. Objective water supply variables

These variables capture objective, interviewer-recorded characteristics of each
participant's household water supply. Sentinel codes are excluded from summaries
and tracked in the audit table at the end of this section.

Sentinel code key:

- **999** = Missing / lost / data entry error
- **333** = Pending
- **888** = Don't know
- **777** = Refused
- **666** = Illegible
- **555** = Other
- **444** = Ambiguous

## 8.1 Prepare water supply variables

```{r water-vars-prepare}
df_water <- df %>%
  mutate(
    # Water source (Season 2 only: MX1_W_SRC)
    W_Source = case_when(
      is_sentinel_str(MX1_W_SRC)                              ~ NA_character_,
      str_detect(tolower(MX1_W_SRC), "servicio|publ")         ~ "Public network",
      str_detect(tolower(MX1_W_SRC), "pipa|truck")            ~ "Water truck (pipa)",
      str_detect(tolower(MX1_W_SRC), "pozo|well")             ~ "Well",
      MX1_W_SRC %in% c("", "N/A") | is.na(MX1_W_SRC)        ~ NA_character_,
      TRUE                                                     ~ "Other"
    ),

    # Continuity of supply (MX3_CONT_INT — both seasons)
    W_Continuity = case_when(
      is_sentinel_str(MX3_CONT_INT)                                  ~ NA_character_,
      str_detect(tolower(MX3_CONT_INT), "^contin")                   ~ "Continuous",
      str_detect(tolower(MX3_CONT_INT), "^intermi|cortan|deja")      ~ "Intermittent",
      str_detect(tolower(MX3_CONT_INT), "disminu")                   ~ "Reduced pressure/flow",
      MX3_CONT_INT %in% c("", "N/A") | is.na(MX3_CONT_INT)         ~ NA_character_,
      TRUE                                                            ~ "Other"
    ),
    W_Continuity = factor(W_Continuity,
                          levels = c("Continuous", "Intermittent",
                                     "Reduced pressure/flow", "Other")),

    # Frequency of supply (MX4_WFREQ / MX4_W_FREQ — harmonised across seasons)
    W_Freq_raw = coalesce(MX4_WFREQ, MX4_W_FREQ),
    W_Frequency = case_when(
      is_sentinel_str(W_Freq_raw)                                    ~ NA_character_,
      str_detect(tolower(W_Freq_raw), "^diari|every day")            ~ "Daily",
      str_detect(tolower(W_Freq_raw), "^intermi")                    ~ "Intermittent",
      str_detect(tolower(W_Freq_raw), "1 a 3|1-3")                   ~ "1–3 days/week",
      str_detect(tolower(W_Freq_raw), "4 a 6|4-6")                   ~ "4–6 days/week",
      str_detect(tolower(W_Freq_raw), "^contin")                     ~ "Continuous",
      W_Freq_raw %in% c("", "N/A") | is.na(W_Freq_raw)             ~ NA_character_,
      TRUE                                                            ~ "Other"
    ),
    W_Frequency = factor(W_Frequency,
                         levels = c("Continuous", "Daily", "4–6 days/week",
                                    "1–3 days/week", "Intermittent", "Other")),

    # Hours per week (MX5_WFREQ_HRS_PER_WEEK)
    W_Hrs_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_HRS_PER_WEEK), NA,
             as.numeric(str_extract(MX5_WFREQ_HRS_PER_WEEK, "^[0-9\\.]+")))
    ),
    W_Hrs_week = ifelse(W_Hrs_week %in% MISSING_CODES, NA, W_Hrs_week),
    W_Hrs_week = ifelse(W_Hrs_week > 168, NA, W_Hrs_week),

    # Days per week (MX5_WFREQ_NDAYS)
    W_Days_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_NDAYS), NA,
             as.numeric(str_extract(MX5_WFREQ_NDAYS, "^[0-9]+")))
    ),
    W_Days_week = ifelse(W_Days_week %in% MISSING_CODES, NA, W_Days_week),
    W_Days_week = ifelse(W_Days_week > 7, NA, W_Days_week),

    # Water pressure (MX6_WPRESS)
    W_Pressure = case_when(
      is_sentinel_str(MX6_WPRESS)                                    ~ NA_character_,
      str_detect(tolower(MX6_WPRESS), "^s[ií]|^yes|^buena|^bien")   ~ "Yes",
      str_detect(tolower(MX6_WPRESS), "^no|^baja|^poca")            ~ "No/Low",
      MX6_WPRESS %in% c("", "N/A") | is.na(MX6_WPRESS)            ~ NA_character_,
      TRUE                                                            ~ "Variable"
    ),
    W_Pressure = factor(W_Pressure, levels = c("Yes", "No/Low", "Variable")),

    # Dry-season scarcity (MX2_WI_SEASON)
    W_Season_scarcity = case_when(
      is_sentinel_str(MX2_WI_SEASON)                                 ~ NA_character_,
      str_detect(tolower(MX2_WI_SEASON), "n/a|no hay|no se|^no$")   ~ "No scarcity reported",
      MX2_WI_SEASON %in% c("", "N/A") | is.na(MX2_WI_SEASON)      ~ NA_character_,
      TRUE                                                            ~ "Seasonal scarcity reported"
    ),
    W_Season_scarcity = factor(W_Season_scarcity,
                                levels = c("No scarcity reported",
                                           "Seasonal scarcity reported")),

    # Neighborhood factor
    W_WS_LOC_F = factor(W_WS_LOC, levels = c("WS", "WI"),
                         labels = c("Water Secure", "Water Insecure"))
  )
```

## 8.2 Summary table by neighborhood

```{r water-vars-table}
tbl_water <- df_water %>%
  select(
    W_WS_LOC_F,
    W_Source, W_Continuity, W_Frequency,
    W_Days_week, W_Hrs_week,
    W_Pressure, W_Season_scarcity
  ) %>%
  tbl_summary(
    by        = W_WS_LOC_F,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd}) / {median} [{p25}, {p75}]",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "ifany",
    label   = list(
      W_Source          ~ "Water source (Season 2 only)",
      W_Continuity      ~ "Supply continuity",
      W_Frequency       ~ "Supply frequency",
      W_Days_week       ~ "Days of supply per week",
      W_Hrs_week        ~ "Hours of supply per week",
      W_Pressure        ~ "Adequate water pressure",
      W_Season_scarcity ~ "Seasonal scarcity reported"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "t.test",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Neighborhood type**") %>%
  modify_caption("**Table. Objective water supply characteristics by neighborhood (n = {N})**")

tbl_water
```

## 8.3 Sentinel code audit for water supply variables

```{r water-sentinel-summary}
water_raw_cols <- c(
  "MX1_W_SRC", "MX2_WI_SEASON", "MX3_CONT_INT",
  "MX4_WFREQ", "MX4_W_FREQ",
  "MX5_WFREQ_NDAYS", "MX5_WFREQ_NHRS", "MX5_WFREQ_HRS_PER_WEEK",
  "MX6_WPRESS"
)

sentinel_labels <- c(
  "999" = "Missing/lost/entry error",
  "333" = "Pending",
  "888" = "Don't know",
  "777" = "Refused",
  "666" = "Illegible",
  "555" = "Other (outside categories)",
  "444" = "Ambiguous"
)

N_total <- nrow(df_water)

sentinel_tbl <- map_dfr(water_raw_cols, function(col) {
  if (!col %in% names(df_water)) return(NULL)
  x <- str_trim(as.character(df_water[[col]]))
  map_dfr(names(sentinel_labels), function(code) {
    n <- sum(x == code, na.rm = TRUE)
    if (n > 0) {
      tibble(
        variable      = col,
        sentinel_code = code,
        meaning       = sentinel_labels[[code]],
        n             = n,
        pct           = round(n / N_total * 100, 1)
      )
    }
  })
})

if (nrow(sentinel_tbl) == 0) {
  cat("No sentinel codes found in water supply variables.\n")
} else {
  sentinel_tbl %>%
    arrange(variable, sentinel_code) %>%
    gt() %>%
    tab_header(title = "Sentinel codes in water supply variables") %>%
    cols_label(
      variable      = "Variable",
      sentinel_code = "Code",
      meaning       = "Meaning",
      n             = "n",
      pct           = "%"
    ) %>%
    tab_footnote(paste0("% of total sample (N = ", N_total, ")"))
}
```
