---
title: "Descriptive Statistics"
author: "Paloma"
date: "2026-03-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

This document produces descriptive statistics tables for the merged dataset
(`3.Merged_Q1-6_Q16_Q19_Q26_Q28.csv`, produced by `03.Data_merge_Q1-6.Rmd`).
It includes:

1. A comprehensive descriptive statistics table (all analytic variables)

2. A manuscript table with sample characteristics stratified by season

3. A missing data summary table

4. HWISE and PSS descriptive statistics (overall and by season)

5. Sample characteristics stratified by water insecurity level (HWISE categories)

6. Objective water supply characteristics (source, continuity, frequency, hours of supply, pressure, and seasonality)

Visualizations for all of the above are in `05.Descriptive_plots.Rmd`.

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(gtsummary)
library(gt)
library(flextable)
library(purrr)
library(officer)
library(ggplot2)

# --- Pretty tables theme ---
# Apply JAMA theme for overall styling, then immediately reset to wipe out its
# automatic add_stat_label() post-processing (which appends "Mean (SD)" etc.
# to every variable name). Re-apply only the settings we actually want.
theme_gtsummary_journal(journal = "jama")
reset_gtsummary_theme()
set_gtsummary_theme(list(
  "pkgwide-fn:pvalue_fun"      = label_style_pvalue(digits = 3),
  "tbl_summary-fn:percent_fun" = label_style_percent(digits = 1)
))

data_path <- "./data"

MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

# Helper: convert sentinel codes to NA for descriptive stats
clean_for_table <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES | is.na(x_num), NA, x_num)
}

# Helper: flag sentinel codes in free-text / character fields
SENTINEL_STRINGS <- as.character(c(999, 888, 777, 666, 555, 444, 333))
is_sentinel_str <- function(x) {
  str_trim(as.character(x)) %in% SENTINEL_STRINGS
}
```

---

# 1. Load cleaned data

```{r load}
df <- read.csv(file.path(data_path, "03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv"),
               stringsAsFactors = FALSE,
               na.strings = c(""))

cat("Dimensions:", dim(df), "\n")
cat("Participants:", nrow(df), "\n")
```

---

# 2. Prepare analytic dataset

Select and recode variables for the descriptive tables.
All sentinel codes are converted to NA since they represent non-informative responses.

```{r prepare}

edu_levels <- c(
  "No formal / incomplete primary",
  "Complete primary",
  "Incomplete secondary",
  "Complete secondary",
  "Incomplete high school",
  "Complete high school",
  "University or higher"
)

df_table <- df %>%
  mutate(
    # --- Demographics ---
    Age          = clean_for_table(D_AGE),
    Yrs_in_nbhd  = clean_for_table(D_LOC_TIME_CLEAN),
    HH_size      = clean_for_table(D_HH_SIZE),
    Parity       = clean_for_table(D_CHLD_CLEAN),

    # --- Season ---
    Season = factor(SEASON,
                    levels = c(0, 1),
                    labels = c("Rainy", "Dry")),

    # --- SES ---
    Education    = factor(
      case_when(
        is.na(SES_EDU_CAT_CLEAN) | SES_EDU_CAT_CLEAN == ""          ~ NA_character_,
        SES_EDU_CAT_CLEAN %in% c("No formal education",
                                  "Incomplete primary")              ~ "No formal / incomplete primary",
        SES_EDU_CAT_CLEAN == "Complete primary"                      ~ "Complete primary",
        SES_EDU_CAT_CLEAN == "Incomplete secondary"                  ~ "Incomplete secondary",
        SES_EDU_CAT_CLEAN == "Complete secondary"                    ~ "Complete secondary",
        SES_EDU_CAT_CLEAN == "Incomplete high school"                ~ "Incomplete high school",
        SES_EDU_CAT_CLEAN == "Complete high school"                  ~ "Complete high school",
        SES_EDU_CAT_CLEAN %in% c("Incomplete university",
                                  "Complete university",
                                  "Postgraduate")                    ~ "University or higher",
        TRUE                                                         ~ NA_character_
      ),
      levels = edu_levels, ordered = TRUE
    ),
    SES          = factor(SES_CAT_AMAI,
                          levels = c("E", "D-", "D+", "C-", "C", "C+", "AB"),
                          ordered = TRUE),
    # SES items — derived from item scores (sentinel 666 → NA)
    SES_Bthr_cat = factor(
      case_when(
        clean_for_table(SES_BTHR_SC) ==  0 ~ "0 bathrooms",
        clean_for_table(SES_BTHR_SC) == 24 ~ "1 bathroom",
        clean_for_table(SES_BTHR_SC) == 47 ~ "2+ bathrooms",
        TRUE                               ~ NA_character_
      ),
      levels = c("0 bathrooms", "1 bathroom", "2+ bathrooms"),
      ordered = TRUE
    ),

    SES_Car_cat = factor(
      case_when(
        clean_for_table(SES_CAR_SC) ==  0 ~ "0 cars",
        clean_for_table(SES_CAR_SC) == 18 ~ "1 car",
        clean_for_table(SES_CAR_SC) == 37 ~ "2+ cars",
        TRUE                              ~ NA_character_
      ),
      levels = c("0 cars", "1 car", "2+ cars"),
      ordered = TRUE
    ),

    SES_Int_cat = factor(
      case_when(
        clean_for_table(SES_INT_SC) ==  0 ~ "No",
        clean_for_table(SES_INT_SC) == 31 ~ "Yes",
        TRUE                              ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    SES_Wrk_cat = factor(
      case_when(
        clean_for_table(SES_WRK_SC) ==  0 ~ "0 workers",
        clean_for_table(SES_WRK_SC) == 15 ~ "1 worker",
        clean_for_table(SES_WRK_SC) == 31 ~ "2 workers",
        clean_for_table(SES_WRK_SC) == 46 ~ "3 workers",
        clean_for_table(SES_WRK_SC) == 61 ~ "4+ workers",
        TRUE                              ~ NA_character_
      ),
      levels = c("0 workers", "1 worker", "2 workers", "3 workers", "4+ workers"),
      ordered = TRUE
    ),

    SES_Bedr_cat = factor(
      case_when(
        clean_for_table(SES_BEDR_SC) ==  0 ~ "0 bedrooms",
        clean_for_table(SES_BEDR_SC) ==  6 ~ "1 bedroom",
        clean_for_table(SES_BEDR_SC) == 12 ~ "2 bedrooms",
        clean_for_table(SES_BEDR_SC) == 17 ~ "3 bedrooms",
        clean_for_table(SES_BEDR_SC) == 23 ~ "4+ bedrooms",
        TRUE                               ~ NA_character_
      ),
      levels = c("0 bedrooms", "1 bedroom", "2 bedrooms", "3 bedrooms", "4+ bedrooms"),
      ordered = TRUE
    ),

    # --- Employment ---
    Employed = factor(clean_for_table(D_EMP_CLEAN),
                      levels = c(0, 1),
                      labels = c("No", "Yes")),
    Employment_type = factor(clean_for_table(D_EMP_FORM_CLEAN),
                             levels = c(0, 1),
                             labels = c("Informal", "Formal")),

    # --- Health ---
    Smoking = factor(
      case_when(
        clean_for_table(HLTH_SMK_CLEAN) == 0 ~ "No",
        clean_for_table(HLTH_SMK_CLEAN) %in% c(1, 2) ~ "Yes",
        TRUE ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),
    Hormonal_tx   = factor(clean_for_table(HLTH_TTM_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Medication    = factor(clean_for_table(HLTH_MED_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_pain  = factor(clean_for_table(HLTH_CPAIN_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_dis   = factor(clean_for_table(HLTH_CDIS_CAT),
                           levels = c(0, 1), labels = c("No", "Yes")),

    # --- HWISE ---
    HWISE_Total   = clean_for_table(HW_TOTAL),
    Water_worry   = clean_for_table(as.numeric(HW_WORRY_SUB)),
    Water_5sub   = clean_for_table(HW_MEX_5SUB),
    
    HWISE_Cat     = factor(
      case_when(
        is.na(clean_for_table(HW_TOTAL))  ~ NA_character_,
        clean_for_table(HW_TOTAL) <= 2    ~ "Marginal (0\u20132)",
        clean_for_table(HW_TOTAL) <= 11   ~ "Low (3\u201311)",
        clean_for_table(HW_TOTAL) <= 23   ~ "Moderate (12\u201323)",
        TRUE                              ~ "High (\u226524)"
      ),
      levels = c("Marginal (0\u20132)", "Low (3\u201311)",
                 "Moderate (12\u201323)", "High (\u226524)")
    ),

    # --- Individual HWISE items ---
HW_Worry    = clean_for_table(HW_WORRY),
HW_Interr   = clean_for_table(HW_INTERR),
HW_Clothes  = clean_for_table(HW_CLOTHES),
HW_Plans    = clean_for_table(HW_PLANS),
HW_Food     = clean_for_table(HW_FOOD),
HW_Hands    = clean_for_table(HW_HANDS),
HW_Body     = clean_for_table(HW_BODY),
HW_Drink    = clean_for_table(HW_DRINK),
HW_Angry    = clean_for_table(HW_ANGRY),
HW_Sleep    = clean_for_table(HW_SLEEP),
HW_None     = clean_for_table(HW_NONE),
HW_Shame    = clean_for_table(HW_SHAME),
    # --- PSS-14 ---
    PSS_Total     = clean_for_table(PSS_TOTAL),
    # Tertile-based PSS categories
    PSS_Cat       = factor(
      case_when(
        is.na(clean_for_table(PSS_TOTAL))            ~ NA_character_,
        clean_for_table(PSS_TOTAL) <= 18             ~ "Low stress (0–18)",
        clean_for_table(PSS_TOTAL) <= 37             ~ "Moderate stress (19–37)",
        TRUE                                         ~ "High stress (38–56)"
      ),
      levels = c("Low stress (0–18)", "Moderate stress (19–37)",
                 "High stress (38–56)")
    )
  )

# Shared type list: force HWISE items and Water_worry to continuous
hwise_items_type <- list(
  HW_Worry    ~ "continuous", HW_Interr  ~ "continuous",
  HW_Clothes  ~ "continuous", HW_Plans   ~ "continuous",
  HW_Food     ~ "continuous", HW_Hands   ~ "continuous",
  HW_Body     ~ "continuous", HW_Drink   ~ "continuous",
  HW_Angry    ~ "continuous", HW_Sleep   ~ "continuous",
  HW_None     ~ "continuous", HW_Shame   ~ "continuous",
  Water_worry ~ "continuous"
)

# Helper: build water supply variables from a raw df
make_water_vars <- function(df) {
  df %>%
    mutate(
      W_Continuity = case_when(
        is_sentinel_str(MX3_CONT_INT)                             ~ NA_character_,
        str_detect(tolower(MX3_CONT_INT), "^contin")              ~ "Continuous",
        str_detect(tolower(MX3_CONT_INT), "^intermi|cortan|deja") ~ "Intermittent",
        MX3_CONT_INT %in% c("", "N/A") | is.na(MX3_CONT_INT)    ~ NA_character_,
        TRUE                                                       ~ "Other"
      ),
      W_Continuity = factor(W_Continuity,
                            levels = c("Continuous", "Intermittent", "Other")),
      W_Freq_raw  = coalesce(MX4_WFREQ, MX4_W_FREQ),
      W_Frequency = case_when(
        is_sentinel_str(W_Freq_raw)                               ~ NA_character_,
        str_detect(tolower(W_Freq_raw), "^diari|every day")       ~ "Daily",
        str_detect(tolower(W_Freq_raw), "1 a 3|1-3")              ~ "1–3 days/week",
        str_detect(tolower(W_Freq_raw), "4 a 6|4-6")              ~ "4–6 days/week",
        W_Freq_raw %in% c("", "N/A") | is.na(W_Freq_raw)        ~ NA_character_,
        TRUE                                                       ~ NA_character_
      ),
      W_Frequency = factor(W_Frequency,
                           levels = c("Daily", "4–6 days/week", "1–3 days/week")),
      W_Days_week_num = suppressWarnings(
        ifelse(is_sentinel_str(MX5_WFREQ_NDAYS), NA,
               as.numeric(str_extract(MX5_WFREQ_NDAYS, "^[0-9]+")))
      ),
      W_Days_week_num = ifelse(W_Days_week_num %in% MISSING_CODES | W_Days_week_num > 7,
                               NA, W_Days_week_num),
      W_Days_week = factor(
        case_when(
          is.na(W_Days_week_num) ~ NA_character_,
          W_Days_week_num <= 2   ~ "1–2 days/week",
          W_Days_week_num <= 5   ~ "3–5 days/week",
          TRUE                   ~ "6–7 days/week"
        ),
        levels = c("1–2 days/week", "3–5 days/week", "6–7 days/week")
      ),
      W_Hrs_week  = suppressWarnings(
        ifelse(is_sentinel_str(MX5_WFREQ_HRS_PER_WEEK), NA,
               as.numeric(str_extract(MX5_WFREQ_HRS_PER_WEEK, "^[0-9\\.]+")))
      ),
      W_Hrs_week  = ifelse(W_Hrs_week %in% MISSING_CODES | W_Hrs_week > 168,
                           NA, W_Hrs_week),
      W_Pressure  = case_when(
        is_sentinel_str(MX6_WPRESS)                                  ~ NA_character_,
        str_detect(tolower(MX6_WPRESS), "^s[ií]|^yes|^buena|^bien") ~ "Yes",
        str_detect(tolower(MX6_WPRESS), "^no|^baja|^poca")          ~ "No/Low",
        MX6_WPRESS %in% c("", "N/A") | is.na(MX6_WPRESS)          ~ NA_character_,
        TRUE                                                          ~ "Variable"
      ),
      W_Pressure  = factor(W_Pressure, levels = c("Yes", "No/Low", "Variable"))
    )
}

analytic_cols <- c(
  "D_AGE", "D_LOC_TIME_CLEAN", "D_HH_SIZE", "D_CHLD_CLEAN",
  "SEASON", "SES_SC_Total", "SES_CAT_AMAI",
  "D_EMP_CLEAN", "D_EMP_FORM_CLEAN",
  "HLTH_SMK_CLEAN", "HLTH_TTM_CAT", "HLTH_MED_CAT",
  "HLTH_CPAIN_CAT", "HLTH_CDIS_CAT"
)
```

---

# 3. Supplementary descriptive statistics table

Includes all analytic variables.

```{r table-full}
tbl_full <- df_table %>%
  select(Age, Yrs_in_nbhd, HH_size, Parity,
         SES, Education, Employed, Employment_type,
         Smoking, Hormonal_tx, Medication, Chronic_pain, Chronic_dis,
         Season) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    type    = list(Parity ~ "continuous"),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      Age             ~ "Age (years)",
      Yrs_in_nbhd     ~ "Years in neighborhood",
      HH_size         ~ "Household size",
      Parity          ~ "Parity",
      SES             ~ "SES (NSE-AMAI)",
      Education       ~ "Education",
      Employed        ~ "Employed",
      Employment_type ~ "Employment type",
      Smoking         ~ "Smoking status",
      Hormonal_tx     ~ "Hormonal treatment",
      Medication      ~ "Regular medication use",
      Chronic_pain    ~ "Chronic pain",
      Chronic_dis     ~ "Chronic disease"
    )
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(test = list(
    all_continuous()  ~ "wilcox.test",
    all_categorical() ~ "fisher.test",
    where(is.ordered) ~ "kruskal.test"
  )) %>%
  bold_labels() %>%
  add_n(statistic = "{p_miss}%",
        col_label = "**Missing (%)**") %>%
  modify_table_body(~ .x %>% relocate(n, .after = last_col())) %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**By season**") %>%
  modify_header(label = "**Variable**") %>%
  modify_footnote(all_stat_cols() ~ "Mean (SD) for continuous variables; n (%) for categorical variables.") %>%
  modify_caption("**Table S1. Sample characteristics — supplementary full variable set**")

tbl_full |> as_gt() |> opt_stylize(style = 6, color = "gray")
```

---

# 4. Manuscript table

Sample characteristics stratified by season to show comparability across data collection waves.

```{r table-manuscript}
tbl_ms <- df_table %>%
  select(
    Season, Age, Yrs_in_nbhd, HH_size, Parity,
    SES,
    SES_Bthr_cat, SES_Car_cat,
    SES_Int_cat, SES_Wrk_cat, SES_Bedr_cat,
    Education,
    Employed, Employment_type,
    Smoking, Chronic_pain, Chronic_dis
  ) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    type      = list(Parity ~ "continuous"),
    digits    = list(all_continuous() ~ 1),
    missing   = "no",
    label     = list(
      Age             ~ "Age",
      Yrs_in_nbhd     ~ "Years in neighborhood",
      HH_size         ~ "Household size",
      Parity          ~ "Parity",
      SES             ~ "SES (NSE-AMAI)",
      SES_Bthr_cat    ~ "  Bathrooms",
      SES_Car_cat     ~ "  Car ownership",
      SES_Int_cat     ~ "  Internet access",
      SES_Wrk_cat     ~ "  Workers in household",
      SES_Bedr_cat    ~ "  Bedrooms",
      Education       ~ "Education",
      Employed        ~ "Employed",
      Employment_type ~ "Employment type",
      Smoking         ~ "Smoking",
      Chronic_pain    ~ "Chronic pain",
      Chronic_dis     ~ "Chronic disease"
    )
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(test = list(
    all_continuous()  ~ "wilcox.test",
    all_categorical() ~ "fisher.test",
    where(is.ordered) ~ "kruskal.test"
  )) %>%
  bold_labels() %>%
  add_n(statistic = "{p_miss}%", col_label = "**Missing (%)**") %>%
  modify_table_body(~ .x %>% relocate(n, .after = last_col())) %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season of data collection**") %>%
  modify_header(label = "**Variable**") %>%
  modify_footnote(all_stat_cols() ~ "Mean (SD) for continuous variables; n (%) for categorical variables.") %>%
  modify_caption("Table 1. Descriptive statistics")

tbl_ms |> as_gt() |> opt_stylize(style = 6, color = "gray")
```

---

# 5. Missing data summary table

Reports true NAs and sentinel-coded values separately for all analytic variables.
Include as supplementary table in dissertation; summarize in-text for the journal manuscript.
See `05.Descriptive_plots.Rmd` for the corresponding visualization.

```{r table-missing}
var_labels <- c(
  # Demographics & SES
  Age             = "Age",
  Yrs_in_nbhd     = "Years in neighborhood",
  HH_size         = "Household size",
  Parity          = "Parity",
  SES             = "SES (NSE-AMAI)",
  Education       = "Education",
  Employed        = "Employed",
  Employment_type = "Employment type",
  # Health
  Smoking         = "Smoking",
  Hormonal_tx     = "Hormonal treatment",
  Medication      = "Medication use",
  Chronic_pain    = "Chronic pain",
  Chronic_dis     = "Chronic disease",
  # PSS & HWISE
  PSS_Total       = "PSS-14 total score",
  HWISE_Total     = "HWISE total score",
  Water_worry     = "Water worry subscale",
  # Water supply
  W_Continuity    = "Supply continuity",
  W_Frequency     = "Supply frequency",
  W_Days_week     = "Days of supply per week",
  W_Hrs_week      = "Hours of supply per week",
  W_Pressure      = "Water pressure"
)

miss_cols <- names(var_labels)

N       <- nrow(df)
N_rainy <- sum(df$SEASON == 0, na.rm = TRUE)
N_dry   <- sum(df$SEASON == 1, na.rm = TRUE)

# ── Build comprehensive dataframe for missingness assessment ─────────────────
df_miss <- make_water_vars(df) %>%
  select(ID, W_Continuity, W_Frequency, W_Days_week, W_Hrs_week, W_Pressure) %>%
  right_join(
    df_table %>% select(ID, SEASON = Season,
                        Age, Yrs_in_nbhd, HH_size, Parity,
                        SES, Education, Employed, Employment_type,
                        Smoking, Hormonal_tx, Medication,
                        Chronic_pain, Chronic_dis,
                        PSS_Total, HWISE_Total, Water_worry),
    by = "ID"
  ) %>%
  mutate(SEASON = as.integer(SEASON) - 1L)  # restore 0=Rainy, 1=Dry

# ── Overall missing counts ──────────────────────────────────────────────────
overall_miss <- df_miss %>%
  select(all_of(miss_cols)) %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_na") %>%
  mutate(
    label        = dplyr::coalesce(var_labels[variable], variable),
    pct_na       = round(n_na / N * 100, 1),
    n_complete   = N - n_na,
    pct_complete = round(n_complete / N * 100, 1)
  )

# ── By-season NA % ──────────────────────────────────────────────────────────
season_miss <- df_miss %>%
  select(SEASON, all_of(miss_cols)) %>%
  group_by(SEASON) %>%
  summarise(across(all_of(miss_cols), ~ round(mean(is.na(.)) * 100, 1)),
            .groups = "drop") %>%
  pivot_longer(-SEASON, names_to = "variable", values_to = "pct") %>%
  pivot_wider(names_from = SEASON, values_from = pct,
              names_glue = "pct_na_{SEASON}") %>%
  rename(pct_na_rainy = pct_na_0, pct_na_dry = pct_na_1)

# ── Chi-square / Fisher p-values (% missing ~ season) ───────────────────────
miss_pvals_tbl <- sapply(miss_cols, function(v) {
  tbl <- table(is.na(df_miss[[v]]), df_miss$SEASON)
  if (any(dim(tbl) < 2) || nrow(tbl) < 2) return(NA_real_)
  tryCatch(
    if (any(tbl < 5)) fisher.test(tbl)$p.value
    else              chisq.test(tbl, correct = FALSE)$p.value,
    error = function(e) NA_real_
  )
})

miss_pval_tbl_df <- tibble(
  variable = names(miss_pvals_tbl),
  p_value  = miss_pvals_tbl,
  p_fmt    = case_when(
    is.na(p_value)  ~ "\u2014",
    p_value < 0.001 ~ "<0.001",
    TRUE            ~ sprintf("%.3f", p_value)
  )
)

# ── Combined table ───────────────────────────────────────────────────────────
missing_tbl_trimmed <- overall_miss %>%
  left_join(season_miss,       by = "variable") %>%
  left_join(miss_pval_tbl_df,  by = "variable") %>%
  select(label, n_complete, pct_complete, n_na, pct_na,
         pct_na_rainy, pct_na_dry, p_fmt) %>%
  arrange(desc(n_na))

# ── HTML preview ─────────────────────────────────────────────────────────────
missing_tbl_trimmed %>%
  gt() %>%
  tab_header(title = "Missing Data Summary") %>%
  tab_spanner(label = "Complete",         columns = c(n_complete, pct_complete)) %>%
  tab_spanner(label = "Overall NA",       columns = c(n_na, pct_na)) %>%
  tab_spanner(label = "By season (% NA)", columns = c(pct_na_rainy, pct_na_dry)) %>%
  cols_label(
    label        = "Variable",
    n_complete   = "n",     pct_complete  = "%",
    n_na         = "n",     pct_na        = "%",
    pct_na_rainy = paste0("Rainy (n=", N_rainy, ")"),
    pct_na_dry   = paste0("Dry (n=",   N_dry,   ")"),
    p_fmt        = "p-value"
  ) %>%
  fmt_number(columns = c(n_complete, n_na), decimals = 0) %>%
  fmt_number(columns = c(pct_complete, pct_na, pct_na_rainy, pct_na_dry), decimals = 1) %>%
  opt_stylize(style = 6, color = "gray")
```

---

# 6. HWISE and PSS descriptive statistics

## 6.1 Overall summary

```{r hwise-pss-overall}
# Build water supply variables (same logic as section 8) and join to df_table
df_s2 <- make_water_vars(df) %>%
  mutate(
    W_Nbhd = factor(W_WS_LOC, levels = c("WS", "WI"),
                    labels = c("Water Secure", "Water Insecure"))
  ) %>%
  # Add all df_table variables via a join
  select(ID, W_Continuity, W_Frequency, W_Days_week, W_Hrs_week,
         W_Pressure, W_Nbhd) %>%   # W_Source excluded (incomplete data entry)
  right_join(
    df_table %>%
      mutate(
        HWISE_Binary = factor(
          case_when(
            is.na(HWISE_Total)      ~ NA_character_,
            HWISE_Total <= 11       ~ "Water secure (0–11)",
            TRUE                    ~ "Water insecure (≥12)"
          ),
          levels = c("Water secure (0–11)", "Water insecure (≥12)")
        ),
        HWISE_Cat4 = factor(
          case_when(
            is.na(HWISE_Total)      ~ NA_character_,
            HWISE_Total == 0        ~ "Secure (0)",
            HWISE_Total <= 3        ~ "Marginal (1–3)",
            HWISE_Total <= 11       ~ "Low (4–11)",
            TRUE                    ~ "Moderate/Severe (≥12)"
          ),
          levels = c("Secure (0)", "Marginal (1–3)",
                     "Low (4–11)", "Moderate/Severe (≥12)")
        ),
        Frongillo_2024 = factor(
          case_when(
            is.na(HWISE_Total)      ~ NA_character_,
            HWISE_Total <= 2        ~ "No to marginal (0–2)",
            HWISE_Total <= 11       ~ "Low (3–11)",
            HWISE_Total <= 23       ~ "Moderate (12–23)",
            TRUE                    ~ "High (24–36)"
          ),
          levels = c("No to marginal (0–2)", "Low (3–11)",
                     "Moderate (12–23)", "High (24–36)")
        ),
        Jepson_2021 = factor(
          case_when(
            is.na(HWISE_Total)      ~ NA_character_,
            HWISE_Total <= 3        ~ "Marginal (0–3)",
            HWISE_Total <= 11       ~ "Low (4–11)",
            HWISE_Total <= 20       ~ "Moderate (12–20)",
            HWISE_Total <= 29       ~ "High (21–29)",
            TRUE                    ~ "Severe (30–36)"
          ),
          levels = c("Marginal (0–3)", "Low (4–11)",
                     "Moderate (12–20)", "High (21–29)", "Severe (30–36)")
        ),
        Water_5sub_binary = factor(
          case_when(
            is.na(Water_5sub)  ~ NA_character_,
            Water_5sub >= 3    ~ "High (≥3)",
            TRUE               ~ "Low (0–2)"
          ),
          levels = c("Low (0–2)", "High (≥3)")
        )
      ),
    by = "ID"
  )

s2_labels <- list(
  PSS_Total      ~ "PSS-14 total score (0–56)",
  PSS_Cat        ~ "Perceived stress level",
  HWISE_Total    ~ "HWISE total score (0–36)",
  Water_worry    ~ "Water worry subscale (0–9)",
  Water_5sub        ~ "Water 5-item subscale (0–15)",
  Water_5sub_binary ~ "Water 5-item subscale — binary (high ≥3)",
  HWISE_Binary      ~ "HWISE binary — Young 2019 (≥12 = insecure)",
  Frongillo_2024    ~ "HWISE — Frongillo et al. 2024 (4 cat.)",
  Jepson_2021       ~ "HWISE — Jepson et al. 2021 (5 cat.)",
  HW_Worry       ~ "1. Worried about water",
  HW_Interr      ~ "2. Water interrupted",
  HW_Clothes     ~ "3. Unable to wash clothes",
  HW_Plans       ~ "4. Changed plans due to water",
  HW_Food        ~ "5. Unable to prepare food",
  HW_Hands       ~ "6. Unable to wash hands",
  HW_Body        ~ "7. Unable to wash body",
  HW_Drink       ~ "8. No drinking water",
  HW_Angry       ~ "9. Felt angry about water",
  HW_Sleep       ~ "10. Went to sleep thirsty",
  HW_None        ~ "11. Had no water at all",
  HW_Shame       ~ "12. Felt shame about water",
  W_Continuity   ~ "Supply continuity",
  W_Frequency    ~ "Supply frequency",
  W_Hrs_week     ~ "Hours of supply per week",
  W_Days_week    ~ "Days of supply per week",
  W_Pressure     ~ "Adequate water pressure"
)

hwise_item_vars <- c("HW_Worry","HW_Interr","HW_Clothes","HW_Plans",
                     "HW_Food","HW_Hands","HW_Body","HW_Drink",
                     "HW_Angry","HW_Sleep","HW_None","HW_Shame")

tbl_hwise_pss_overall <- df_s2 %>%
  select(
    Season,
    PSS_Total, PSS_Cat,
    HWISE_Total,
    HWISE_Binary, Frongillo_2024, Jepson_2021,
    Water_5sub, Water_5sub_binary,
    Water_worry,
    HW_Worry, HW_Interr, HW_Clothes, HW_Plans,
    HW_Food, HW_Hands, HW_Body, HW_Drink,
    HW_Angry, HW_Sleep, HW_None, HW_Shame,
    W_Continuity, W_Frequency,
    W_Hrs_week, W_Days_week, W_Pressure
  ) %>%
  tbl_summary(
    by        = Season,
    type      = hwise_items_type,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = s2_labels
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(test = list(
    all_continuous()  ~ "wilcox.test",
    all_categorical() ~ "fisher.test"
  )) %>%
  bold_labels() %>%
  bold_p(t = 0.05) %>%
  modify_table_body(~{
    # Unbold the 12 HWISE item label rows
    .x <- .x %>% mutate(label = ifelse(variable %in% hwise_item_vars,
                                       gsub("\\*\\*(.+)\\*\\*", "\\1", label), label))
    # Insert a bold group header row just before HW_Worry
    first_idx <- which(.x$variable == "HW_Worry" & .x$row_type == "label")[1]
    header <- .x[first_idx, ]
    header$variable  <- "hwise_items_group"
    header$var_label <- "**HWISE individual items**"
    header$label     <- "**HWISE individual items**"
    header$row_type  <- "label"
    stat_cols <- grep("^stat_", names(header), value = TRUE)
    header[stat_cols] <- NA_character_
    if ("p.value" %in% names(header)) header$p.value <- NA_real_
    dplyr::bind_rows(.x[seq_len(first_idx - 1), ], header, .x[first_idx:nrow(.x), ])
  }) %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season**") %>%
  modify_caption("**Table S2. Water-related variables and PSS by season)**")

tbl_hwise_pss_overall |> as_gt() |> opt_stylize(style = 6, color = "gray")

# --- Table 2 (short): key HWISE, PSS and water supply variables by season ---
tbl_hwise_pss_short <- df_s2 %>%
  select(
    Season,
    PSS_Total, PSS_Cat,
    HWISE_Total, Frongillo_2024,
    W_Continuity, W_Frequency, W_Pressure
  ) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      PSS_Total      ~ "PSS-14 total score (0\u201356)",
      PSS_Cat        ~ "Perceived stress level",
      HWISE_Total    ~ "HWISE total score (0\u201336)",
      Frongillo_2024 ~ "HWISE \u2014 Frongillo et al. 2024",
      W_Continuity   ~ "Supply continuity",
      W_Frequency    ~ "Supply frequency",
      W_Pressure     ~ "Adequate water pressure"
    )
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(test = list(
    all_continuous()  ~ "wilcox.test",
    all_categorical() ~ "fisher.test"
  )) %>%
  bold_labels() %>%
  bold_p(t = 0.05) %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season**") %>%
  modify_caption("**Table 2. HWISE, PSS and water supply characteristics by season**")

tbl_hwise_pss_short |> as_gt() |> opt_stylize(style = 6, color = "gray")

# --- Table S3: same variables, by neighborhood ---
tbl_s3_by_nbhd <- df_s2 %>%
  filter(!is.na(W_Nbhd)) %>%
  select(
    W_Nbhd,
    PSS_Total, PSS_Cat,
    HWISE_Total,
    HWISE_Binary, Frongillo_2024, Jepson_2021,
    Water_5sub, Water_5sub_binary,
    Water_worry,
    HW_Worry, HW_Interr, HW_Clothes, HW_Plans,
    HW_Food, HW_Hands, HW_Body, HW_Drink,
    HW_Angry, HW_Sleep, HW_None, HW_Shame,
    W_Continuity, W_Frequency,
    W_Hrs_week, W_Days_week, W_Pressure
  ) %>%
  tbl_summary(
    by        = W_Nbhd,
    type      = hwise_items_type,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = s2_labels
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(test = list(
    all_continuous()  ~ "wilcox.test",
    all_categorical() ~ "fisher.test"
  )) %>%
  bold_labels() %>%
  bold_p(t = 0.05) %>%
  modify_table_body(~{
    # Unbold the 12 HWISE item label rows
    .x <- .x %>% mutate(label = ifelse(variable %in% hwise_item_vars,
                                       gsub("\\*\\*(.+)\\*\\*", "\\1", label), label))
    # Insert a bold group header row just before HW_Worry
    first_idx <- which(.x$variable == "HW_Worry" & .x$row_type == "label")[1]
    header <- .x[first_idx, ]
    header$variable  <- "hwise_items_group"
    header$var_label <- "**HWISE individual items**"
    header$label     <- "**HWISE individual items**"
    header$row_type  <- "label"
    stat_cols <- grep("^stat_", names(header), value = TRUE)
    header[stat_cols] <- NA_character_
    if ("p.value" %in% names(header)) header$p.value <- NA_real_
    dplyr::bind_rows(.x[seq_len(first_idx - 1), ], header, .x[first_idx:nrow(.x), ])
  }) %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Neighborhood**") %>%
  modify_caption("**Table S3. Water-related variables and PSS by neighborhood (n = {N})**")

tbl_s3_by_nbhd |> as_gt() |> opt_stylize(style = 6, color = "gray")
```


## 6.2 By season

```{r hwise-pss-by-season}
tbl_scores_season <- df_table %>%
  select(Season, HWISE_Total, HWISE_Cat, Water_worry, Water_5sub, PSS_Total, PSS_Cat) %>%
  tbl_summary(
    by        = Season,
    type = list(Water_worry ~ "continuous"),
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      HWISE_Total  ~ "HWISE total score (0–36)",
      HWISE_Cat    ~ "HWISE 4 categories",
      Water_worry  ~ "Water worry subscale (0–9)",
      Water_5sub   ~ "Water 5-item subscale (0–15)",
      PSS_Total    ~ "PSS-14 total score (0–56)",
      PSS_Cat      ~ "Perceived stress level"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "wilcox.test",
    all_categorical() ~ "fisher.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season**") %>%
  modify_caption("**Table. HWISE and PSS scores by season (n = {N})**")

tbl_scores_season |> as_gt() |> opt_stylize(style = 6, color = "gray")
```

---


# 7. Sample characteristics by water insecurity level

Demographics and health variables stratified by HWISE category, using the
standard cut-points from Young et al. (2019, *BMJ Global Health*):
0 = water secure; 1-11 = marginal; 12-23 = moderate; 24-36 = severe.
One-way ANOVA used for continuous variables; chi-square for categorical.

```{r table-by-hwise}
tbl_by_hwise <- df_table %>%
  filter(!is.na(HWISE_Cat)) %>%
  select(
    HWISE_Cat,
    Age, Yrs_in_nbhd, HH_size, Parity,
    SES, Education,
    Employed, Smoking, Chronic_pain, Chronic_dis,
    PSS_Total, PSS_Cat, Water_worry, Water_5sub
  ) %>%
  tbl_summary(
    by        = HWISE_Cat,
    type      = list(Water_worry ~ "continuous"),
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits  = list(all_continuous() ~ 1),
    missing = "no",
    label   = list(
      Age          ~ "Age (years)",
      Yrs_in_nbhd  ~ "Years in neighborhood",
      HH_size      ~ "Household size",
      Parity       ~ "Parity (n children)",
      SES          ~ "SES (NSE-AMAI)",
      Education    ~ "Education",
      Employed     ~ "Employed",
      Smoking      ~ "Smoking",
      Chronic_pain ~ "Chronic pain",
      Chronic_dis  ~ "Chronic disease",
      PSS_Total    ~ "PSS-14 total score",
      PSS_Cat      ~ "Perceived stress level",
      Water_worry  ~ "Water worry subscale (0–9)"
    )
  ) %>%
  add_overall() %>%
  add_p(test = list(
    all_continuous()  ~ "aov",
    all_categorical() ~ "chisq.test"
  )) %>%
  bold_labels() %>%
  modify_spanning_header(c("stat_1", "stat_2", "stat_3") ~
                           "**HWISE 3 categories**") %>%
  modify_caption("**Table. Sample characteristics by water insecurity level (n = {N})**")

tbl_by_hwise |> as_gt() |> opt_stylize(style = 6, color = "gray")
```

## 7.3 HWISE item response distributions

```{r hwise-response-dist, fig.width = 9, fig.height = 6, message = FALSE, warning = FALSE}
library(tidyr)
library(forcats)

hwise_item_names <- c(
  "HW_Worry"   = "1. Worry",
  "HW_Interr"  = "2. Interruption",
  "HW_Clothes" = "3. Clothes",
  "HW_Plans"   = "4. Plans",
  "HW_Food"    = "5. Food",
  "HW_Hands"   = "6. Hands",
  "HW_Body"    = "7. Body",
  "HW_Drink"   = "8. Drink",
  "HW_Angry"   = "9. Angry",
  "HW_Sleep"   = "10. Sleep",
  "HW_None"    = "11. No water",
  "HW_Shame"   = "12. Shame"
)

response_labels <- c(
  "0" = "Never",
  "1" = "Rarely",
  "2" = "Sometimes",
  "3" = "Often/Always"
)

hwise_long <- df_table %>%
  select(all_of(names(hwise_item_names))) %>%
  pivot_longer(everything(), names_to = "item", values_to = "response") %>%
  filter(!is.na(response)) %>%
  mutate(
    item     = factor(hwise_item_names[item], levels = (hwise_item_names)),
    response = factor(response, levels = c(0, 1, 2, 3),
                      labels = response_labels)
  ) %>%
  count(item, response) %>%
  group_by(item) %>%
  mutate(pct = n / sum(n) * 100) %>%
  ungroup()

ggplot(hwise_long, aes(x = pct, y = item, fill = response)) +
  geom_col(position = "stack", width = 0.7) +
  geom_text(
    aes(label = ifelse(pct >= 5, paste0(round(pct), "%"), "")),
    position = position_stack(vjust = 0.5),
    size = 3, color = "white", fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Never"        = "#2166ac",
      "Rarely"       = "#74add1",
      "Sometimes"    = "#f4a582",
      "Often/Always" = "#d6604d"
    ),
    name = "Response"
  ) +
  scale_x_continuous(labels = function(x) paste0(x, "%"), expand = c(0, 0)) +
  labs(
    title    = "HWISE item response distributions",
    subtitle = "Percentage of respondents per response category",
    x        = "Percentage of respondents",
    y        = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position   = "bottom",
    panel.grid.major.y = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.text.y        = element_text(size = 10),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)
  ) + coord_flip()
```

---

# 8. Objective water supply variables

## 8.1 Prepare water supply variables

```{r water-vars-prepare}
df_water <- make_water_vars(df) %>%
  mutate(
    # Neighborhood factor
    W_WS_LOC_F = factor(W_WS_LOC, levels = c("WS", "WI"),
                         labels = c("Water Secure", "Water Insecure"))
  )
```



---

# 9. Internal consistency (Cronbach's α)

Cronbach's α measures how well the items in a scale correlate with each other —
i.e., whether they are all capturing the same underlying construct. Values ≥ 0.7
are generally considered acceptable. α is computed here on complete cases only
(rows with any `NA` or sentinel code in the relevant items are excluded).

```{r cronbach, message=FALSE, warning=FALSE}
library(psych)

HWISE_ITEMS <- c("HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
                 "HW_FOOD", "HW_HANDS", "HW_BODY", "HW_DRINK",
                 "HW_ANGRY", "HW_SLEEP", "HW_NONE", "HW_SHAME")

PSS_SCORED_DIRECT  <- c("PSS1","PSS2","PSS3","PSS8","PSS11","PSS12","PSS14")
PSS_SCORED_REVERSE <- c("PSS4_REV","PSS5_REV","PSS6_REV","PSS7_REV",
                        "PSS9_REV","PSS10_REV","PSS13_REV")
PSS_SCORED_ITEMS   <- c(PSS_SCORED_DIRECT, PSS_SCORED_REVERSE)

WORRY_ITEMS <- c("HW_WORRY", "HW_ANGRY", "HW_SHAME")

# Helper to clean a set of items: exclude NAs and sentinel codes row-wise
clean_items <- function(data, items) {
  data %>%
    select(all_of(items)) %>%
    mutate(across(everything(), ~ if_else(. %in% MISSING_CODES, NA_real_,
                                          as.numeric(.)))) %>%
    filter(if_all(everything(), ~ !is.na(.)))
}

# HWISE — 12 items, scored 0–3
alpha_hwise <- clean_items(df, HWISE_ITEMS) %>% psych::alpha()

# PSS-14 — use reverse-coded versions of the negative items
alpha_pss <- clean_items(df, PSS_SCORED_ITEMS) %>% psych::alpha()

# Water worry subscale — 3 items
alpha_worry <- clean_items(df, WORRY_ITEMS) %>% psych::alpha()

# Summary table
alpha_summary <- tibble(
  Scale            = c("HWISE (12 items)", "PSS-14 (14 items)",
                       "Water worry subscale (3 items)"),
  n_items          = c(12, 14, 3),
  n_complete       = c(nrow(clean_items(df, HWISE_ITEMS)),
                       nrow(clean_items(df, PSS_SCORED_ITEMS)),
                       nrow(clean_items(df, WORRY_ITEMS))),
  alpha            = round(c(alpha_hwise$total$raw_alpha,
                             alpha_pss$total$raw_alpha,
                             alpha_worry$total$raw_alpha), 3),
  alpha_95CI_lower = round(c(alpha_hwise$total$raw_alpha -
                               1.96 * alpha_hwise$total$ase,
                             alpha_pss$total$raw_alpha -
                               1.96 * alpha_pss$total$ase,
                             alpha_worry$total$raw_alpha -
                               1.96 * alpha_worry$total$ase), 3),
  alpha_95CI_upper = round(c(alpha_hwise$total$raw_alpha +
                               1.96 * alpha_hwise$total$ase,
                             alpha_pss$total$raw_alpha +
                               1.96 * alpha_pss$total$ase,
                             alpha_worry$total$raw_alpha +
                               1.96 * alpha_worry$total$ase), 3)
)

cat("Cronbach's α summary:\n")
print(alpha_summary)

ggplot(alpha_summary,
       aes(x = reorder(Scale, alpha), y = alpha)) +
  geom_pointrange(aes(ymin = alpha_95CI_lower, ymax = alpha_95CI_upper),
                  size = 0.8, color = "#5bc0de") +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "grey50") +
  geom_hline(yintercept = 0.9, linetype = "dotted", color = "grey50") +
  annotate("text", x = 0.6, y = 0.71, label = "acceptable (0.7)",
           hjust = 0, size = 3, color = "grey40") +
  annotate("text", x = 0.6, y = 0.91, label = "excellent (0.9)",
           hjust = 0, size = 3, color = "grey40") +
  coord_flip() +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
  labs(title = "Cronbach's α by scale (95% CI)",
       x = NULL, y = "Cronbach's α") +
  theme_minimal()
```

### Interpretation

**HWISE (α = 0.848, 95% CI: 0.827–0.869):** Good internal consistency, comfortably
above the 0.8 threshold. All 12 items contribute positively, with item-total
correlations ranging from 0.36 to 0.66. The weakest items are `HW_SLEEP`
(r.drop = 0.296) and `HW_DRINK` (r.drop = 0.381), though neither is low enough
to warrant removal — they likely capture a slightly different facet of water
insecurity (disrupted sleep and drinking water access) that is less correlated
with the broader scale but still conceptually relevant.

**PSS-14 (α = 0.813, 95% CI: 0.787–0.840):** Good internal consistency. The
weakest items are `PSS12` (r.drop = 0.274) and `PSS13_REV` (r.drop = 0.292),
both of which are among the reverse-coded items. This pattern is common in PSS
data and has been noted in validation studies. The overall α is consistent with
values reported in the literature for Spanish-language administrations of the PSS.

**Water worry subscale (α = 0.539, 95% CI: 0.468–0.610):** Below the conventional
0.7 threshold, which warrants caution in interpretation. However, this is partly
expected — Cronbach's α is sensitive to the number of items, and a 3-item scale
will almost always yield lower α than the full 12-item HWISE even if the items are
genuinely related. Among the three items, `HW_SHAME` has the lowest item-total
correlation (r.drop = 0.261), suggesting it is the least consistent with the other
two. `HW_ANGRY` and `HW_WORRY` are more closely related to each other. These
results suggest the subscale has limited reliability as a standalone measure and
findings based on it should be interpreted carefully.

### Item-level detail

Item-total correlations and α-if-item-deleted for each scale — useful for
spotting items that weaken internal consistency.

```{r cronbach-itemlevel}
cat("--- HWISE item-level stats ---\n")
print(alpha_hwise$item.stats %>%
        select(r.cor, r.drop) %>%
        round(3))

cat("\n--- PSS-14 item-level stats ---\n")
print(alpha_pss$item.stats %>%
        select(r.cor, r.drop) %>%
        round(3))

cat("\n--- Water worry subscale item-level stats ---\n")
print(alpha_worry$item.stats %>%
        select(r.cor, r.drop) %>%
        round(3))
```

---

# 10. Export tables to Word

```{r export-tables, message=FALSE, warning=FALSE}
output_dir <- "./output/tables"
if (!dir.exists(output_dir)) dir.create(output_dir, recursive = TRUE)

# Helper: convert a gtsummary table to a formatted flextable ready for Word
fmt_docx <- function(tbl) {
  ft <- tbl %>% as_flex_table()
  n_body <- nrow(ft$body$dataset)
  ft %>%
    font(fontname = "Calibri", part = "all") %>%
    fontsize(size = 10, part = "body") %>%
    fontsize(size = 10, part = "header") %>%
    fontsize(size = 9,  part = "footer") %>%
    # Header: dark navy background, white bold text
    bg(bg = "#1A3A5C", part = "header") %>%
    color(color = "white", part = "header") %>%
    bold(part = "header") %>%
    # Alternating row shading
    bg(i = seq(2, n_body, by = 2), bg = "#F9F9F9", part = "body") %>%
    # Horizontal-only borders
    border_remove() %>%
    hline_top(part = "header",
              border = fp_border(color = "#1A3A5C", width = 1.5)) %>%
    hline_bottom(part = "header",
                 border = fp_border(color = "#1A3A5C", width = 1.5)) %>%
    hline_bottom(part = "body",
                 border = fp_border(color = "#1A3A5C", width = 1.5)) %>%
    hline(border = fp_border(color = "#bdc3c7", width = 0.5), part = "body") %>%
    # Align: label column left, stats/p-value columns centre
    align(j = 1, align = "left",   part = "all") %>%
    align(j = -1, align = "center", part = "all") %>%
    # Compact row height and padding
    padding(padding.top = 2, padding.bottom = 2, part = "all") %>%
    # Footer: italic, grey
    italic(part = "footer") %>%
    color(color = "#555555", part = "footer") %>%
    set_table_properties(layout = "autofit", width = 1) %>%
    fix_border_issues()
}

tables <- list(
  "04.Table1_sample_full.docx"                   = tbl_ms,
  "04.Table2_HWISE_PSS_short.docx"               = tbl_hwise_pss_short,
  "04.Table2_HWISE_PSS_by_season_full.docx"      = tbl_hwise_pss_overall,
  "04.Table3_HWISE_PSS_by_neighborhood.docx"     = tbl_s3_by_nbhd
)

for (fname in names(tables)) {
  tables[[fname]] %>%
    fmt_docx() %>%
    save_as_docx(path = file.path(output_dir, fname))
  cat("Exported:", fname, "\n")
}

# Main manuscript table — selected variables only
tbl_main <- df_table %>%
  select(Season, Age, HH_size, Parity, SES, Employed, Chronic_pain, Chronic_dis) %>%
  tbl_summary(
    by        = Season,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    type      = list(Parity ~ "continuous"),
    digits    = list(all_continuous() ~ 1),
    missing   = "no",
    label     = list(
      Age          ~ "Age",
      HH_size      ~ "Household size",
      Parity       ~ "Parity",
      SES          ~ "SES (NSE-AMAI)",
      Employed     ~ "Employed",
      Chronic_pain ~ "Chronic pain",
      Chronic_dis  ~ "Chronic disease"
    )
  ) %>%
  add_overall(last = FALSE) %>%
  add_p(test = list(
    all_continuous()  ~ "wilcox.test",
    all_categorical() ~ "fisher.test",
    where(is.ordered) ~ "kruskal.test"
  )) %>%
  bold_labels() %>%
  add_n(statistic = "{p_miss}%", col_label = "**Missing (%)**") %>%
  modify_table_body(~ .x %>% relocate(n, .after = last_col())) %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Season of data collection**") %>%
  modify_header(label = "**Variable**") %>%
  modify_footnote(all_stat_cols() ~ "Mean (SD) for continuous variables; n (%) for categorical variables.") %>%
  modify_caption("**Table 1. Sample characteristics by season**")

tbl_main %>%
  fmt_docx() %>%
  save_as_docx(path = file.path(output_dir, "04.Table1_sample_main.docx"))
cat("Exported: 04.Table1_sample_main.docx\n")

# Missing data table — build flextable directly from data (gt_tbl can't use as_flex_table)
COL_HDR_MISS  <- "#1A3A5C"
COL_SUB_MISS  <- "#2E6DA4"
COL_ALT_MISS  <- "#F9F9F9"
COL_HIGH_MISS <- "#D5F5E3"

n_total_cols   <- 8L  # label + n_complete + pct_complete + n_na + pct_na + rainy + dry + p_fmt
cols_complete  <- c("n_complete", "pct_complete")
cols_na        <- c("n_na", "pct_na")
rows_sig_p     <- which(miss_pval_tbl_df$p_value < 0.05)

rainy_lbl <- paste0("Rainy\n(n=", N_rainy, ")")
dry_lbl   <- paste0("Dry\n(n=",   N_dry,   ")")

missing_table <- flextable(missing_tbl_trimmed) %>%
  # ── Column labels ──────────────────────────────────────────────────────────
  set_header_labels(
    label        = "Variable",
    n_complete   = "n",       pct_complete = "%",
    n_na         = "n",       pct_na       = "%",
    pct_na_rainy = rainy_lbl, pct_na_dry   = dry_lbl,
    p_fmt        = "p-value"
  ) %>%
  # ── Spanner row ────────────────────────────────────────────────────────────
  add_header_row(
    values    = c("", "Complete", "Overall NA", "By season (% NA)", ""),
    colwidths = c(1, 2, 2, 2, 1)
  ) %>%
  # ── Title row ──────────────────────────────────────────────────────────────
  add_header_row(
    values    = "Table S1.2. Missing data summary",
    colwidths = n_total_cols
  ) %>%
  # ── Typography ─────────────────────────────────────────────────────────────
  font(fontname = "Calibri", part = "all") %>%
  fontsize(size = 10, part = "body") %>%
  fontsize(size = 10, part = "header") %>%
  fontsize(size = 9,  part = "footer") %>%
  bold(part = "header") %>%
  color(color = "white", part = "header") %>%
  # ── Header colours (uniform #1A3A5C across all header rows) ────────────────
  bg(bg = COL_HDR_MISS, part = "header") %>%
  # ── Body ───────────────────────────────────────────────────────────────────
  bg(i = seq(2, nrow(missing_tbl_trimmed), by = 2), bg = COL_ALT_MISS, part = "body") %>%
  { if (length(rows_sig_p) > 0)
      bold(., i = rows_sig_p, j = "p_fmt", part = "body")
    else . } %>%
  # ── Borders (horizontal-only, matching fmt_docx) ───────────────────────────
  border_remove() %>%
  hline_top(part = "header",
            border = fp_border(color = COL_HDR_MISS, width = 1.5)) %>%
  hline_bottom(part = "header",
               border = fp_border(color = COL_HDR_MISS, width = 1.5)) %>%
  hline_bottom(part = "body",
               border = fp_border(color = COL_HDR_MISS, width = 1.5)) %>%
  hline(border = fp_border(color = "#bdc3c7", width = 0.5), part = "body") %>%
  # ── Alignment ──────────────────────────────────────────────────────────────
  align(j = 1,  align = "left",   part = "all") %>%
  align(j = -1, align = "center", part = "all") %>%
  # ── Number formatting ──────────────────────────────────────────────────────
  colformat_num(j = c("n_complete", "n_na"), big.mark = ",", digits = 0) %>%
  colformat_num(j = c("pct_complete", "pct_na", "pct_na_rainy", "pct_na_dry"), digits = 1) %>%
  # ── Footnote ───────────────────────────────────────────────────────────────
  add_footer_lines(c(
    "Rows sorted by number of true NAs (descending).",
    "p-value from chi-square test (Fisher\u2019s exact if any cell < 5) comparing proportion missing between seasons."
  )) %>%
  italic(part = "footer") %>%
  color(color = "#555555", part = "footer") %>%
  # ── Layout ─────────────────────────────────────────────────────────────────
  padding(padding.top = 2, padding.bottom = 2, part = "all") %>%
  set_table_properties(layout = "autofit", width = 1) %>%
  fix_border_issues()

missing_table
save_as_docx(missing_table, path = file.path(output_dir, "04.Table1.2_missing.docx"))

cat("Exported: 04.Table1.2_missing.docx\n")
```

```

