---
title: "Descriptive Plots"
author: "Paloma"
date: "2026-03-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

This document produces descriptive visualizations for the merged dataset
(`03.Merged_Q1-6.csv`, produced by `03.Data_merge_Q1-6.Rmd`).
It complements the tables in `04.Descriptive_stats.Rmd` and includes:

1. Missing data overview (stacked bar chart)

2. HWISE and PSS score distributions by season (histograms)

3. PSS and water worry by water insecurity level (boxplots)

4. Objective water supply distributions by neighborhood (histograms and bar charts)

5. Hours of water supply vs HWISE and PSS (scatter plots)

---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(gridExtra)
library(purrr)

data_path <- "./data"

MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

clean_for_table <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES | is.na(x_num), NA, x_num)
}

SENTINEL_STRINGS <- as.character(c(999, 888, 777, 666, 555, 444, 333))
is_sentinel_str <- function(x) {
  str_trim(as.character(x)) %in% SENTINEL_STRINGS
}
```

---

# 1. Load and prepare data

```{r load}
df <- read.csv(file.path(data_path, "03.Merged_Q1-6.csv"),
               stringsAsFactors = FALSE,
               na.strings = c(""))

cat("Dimensions:", dim(df), "\n")
cat("Participants:", nrow(df), "\n")
```

```{r prepare-df-table}
df_table <- df %>%
  mutate(
    Age          = clean_for_table(D_AGE),
    Yrs_in_nbhd  = clean_for_table(D_LOC_TIME_CLEAN),
    HH_size      = clean_for_table(D_HH_SIZE),
    Parity       = clean_for_table(D_CHLD_CLEAN),

    Season = factor(SEASON,
                    levels = c(0, 1),
                    labels = c("Rainy (Oct-Dec 2022)", "Dry (Apr-Jun 2023)")),

    SES_Total    = clean_for_table(SES_SC_Total),
    SES_Category = factor(SES_CAT_AMAI,
                          levels = c("E", "D-", "D+", "C-", "C", "C+", "AB"),
                          ordered = TRUE),

    Employed = factor(clean_for_table(D_EMP_CLEAN),
                      levels = c(0, 1), labels = c("No", "Yes")),
    Employment_type = factor(clean_for_table(D_EMP_FORM_CLEAN),
                             levels = c(0, 1), labels = c("Informal", "Formal")),

    Smoking = factor(clean_for_table(HLTH_SMK_CLEAN),
                     levels = c(0, 1, 2), labels = c("No", "Yes", "Occasionally")),
    Hormonal_tx  = factor(clean_for_table(HLTH_TTM_CAT),
                          levels = c(0, 1), labels = c("No", "Yes")),
    Medication   = factor(clean_for_table(HLTH_MED_CAT),
                          levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_pain = factor(clean_for_table(HLTH_CPAIN_CAT),
                          levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_dis  = factor(clean_for_table(HLTH_CDIS_CAT),
                          levels = c(0, 1), labels = c("No", "Yes")),

    HWISE_Total = clean_for_table(HW_TOTAL),
    HWISE_Cat   = factor(
      case_when(
        is.na(clean_for_table(HW_TOTAL))  ~ NA_character_,
        clean_for_table(HW_TOTAL) == 0    ~ "Secure (0)",
        clean_for_table(HW_TOTAL) <= 11   ~ "Marginal (1-11)",
        clean_for_table(HW_TOTAL) <= 23   ~ "Moderate (12-23)",
        TRUE                              ~ "Severe (24-36)"
      ),
      levels = c("Secure (0)", "Marginal (1-11)", "Moderate (12-23)", "Severe (24-36)")
    ),
    Water_worry = clean_for_table(HW_WORRY_SUB),

    PSS_Total = clean_for_table(PSS_TOTAL),
    PSS_Cat   = factor(
      case_when(
        is.na(clean_for_table(PSS_TOTAL))  ~ NA_character_,
        clean_for_table(PSS_TOTAL) <= 22   ~ "Low stress (≤22)",
        clean_for_table(PSS_TOTAL) <= 31   ~ "Moderate stress (23-31)",
        TRUE                               ~ "High stress (≥32)"
      ),
      levels = c("Low stress (≤22)", "Moderate stress (23-31)", "High stress (≥32)")
    )
  )
```

```{r prepare-df-water}
df_water <- df %>%
  mutate(
    W_Source = case_when(
      is_sentinel_str(MX1_W_SRC)                              ~ NA_character_,
      str_detect(tolower(MX1_W_SRC), "servicio|publ")         ~ "Public network",
      str_detect(tolower(MX1_W_SRC), "pipa|truck")            ~ "Water truck (pipa)",
      str_detect(tolower(MX1_W_SRC), "pozo|well")             ~ "Well",
      MX1_W_SRC %in% c("", "N/A") | is.na(MX1_W_SRC)        ~ NA_character_,
      TRUE                                                     ~ "Other"
    ),

    W_Continuity = case_when(
      is_sentinel_str(MX3_CONT_INT)                                  ~ NA_character_,
      str_detect(tolower(MX3_CONT_INT), "^contin")                   ~ "Continuous",
      str_detect(tolower(MX3_CONT_INT), "^intermi|cortan|deja")      ~ "Intermittent",
      str_detect(tolower(MX3_CONT_INT), "disminu")                   ~ "Reduced pressure/flow",
      MX3_CONT_INT %in% c("", "N/A") | is.na(MX3_CONT_INT)         ~ NA_character_,
      TRUE                                                            ~ "Other"
    ),
    W_Continuity = factor(W_Continuity,
                          levels = c("Continuous", "Intermittent",
                                     "Reduced pressure/flow", "Other")),

    W_Freq_raw = coalesce(MX4_WFREQ, MX4_W_FREQ),
    W_Frequency = case_when(
      is_sentinel_str(W_Freq_raw)                                    ~ NA_character_,
      str_detect(tolower(W_Freq_raw), "^diari|every day")            ~ "Daily",
      str_detect(tolower(W_Freq_raw), "^intermi")                    ~ "Intermittent",
      str_detect(tolower(W_Freq_raw), "1 a 3|1-3")                   ~ "1–3 days/week",
      str_detect(tolower(W_Freq_raw), "4 a 6|4-6")                   ~ "4–6 days/week",
      str_detect(tolower(W_Freq_raw), "^contin")                     ~ "Continuous",
      W_Freq_raw %in% c("", "N/A") | is.na(W_Freq_raw)             ~ NA_character_,
      TRUE                                                            ~ "Other"
    ),
    W_Frequency = factor(W_Frequency,
                         levels = c("Continuous", "Daily", "4–6 days/week",
                                    "1–3 days/week", "Intermittent", "Other")),

    W_Hrs_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_HRS_PER_WEEK), NA,
             as.numeric(str_extract(MX5_WFREQ_HRS_PER_WEEK, "^[0-9\\.]+")))
    ),
    W_Hrs_week = ifelse(W_Hrs_week %in% MISSING_CODES, NA, W_Hrs_week),
    W_Hrs_week = ifelse(W_Hrs_week > 168, NA, W_Hrs_week),

    W_Days_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_NDAYS), NA,
             as.numeric(str_extract(MX5_WFREQ_NDAYS, "^[0-9]+")))
    ),
    W_Days_week = ifelse(W_Days_week %in% MISSING_CODES, NA, W_Days_week),
    W_Days_week = ifelse(W_Days_week > 7, NA, W_Days_week),

    W_Pressure = case_when(
      is_sentinel_str(MX6_WPRESS)                                    ~ NA_character_,
      str_detect(tolower(MX6_WPRESS), "^s[ií]|^yes|^buena|^bien")   ~ "Yes",
      str_detect(tolower(MX6_WPRESS), "^no|^baja|^poca")            ~ "No/Low",
      MX6_WPRESS %in% c("", "N/A") | is.na(MX6_WPRESS)            ~ NA_character_,
      TRUE                                                            ~ "Variable"
    ),
    W_Pressure = factor(W_Pressure, levels = c("Yes", "No/Low", "Variable")),

    W_Season_scarcity = case_when(
      is_sentinel_str(MX2_WI_SEASON)                                 ~ NA_character_,
      str_detect(tolower(MX2_WI_SEASON), "n/a|no hay|no se|^no$")   ~ "No scarcity reported",
      MX2_WI_SEASON %in% c("", "N/A") | is.na(MX2_WI_SEASON)      ~ NA_character_,
      TRUE                                                            ~ "Seasonal scarcity reported"
    ),
    W_Season_scarcity = factor(W_Season_scarcity,
                                levels = c("No scarcity reported",
                                           "Seasonal scarcity reported")),

    W_WS_LOC_F = factor(W_WS_LOC, levels = c("WS", "WI"),
                         labels = c("Water Secure", "Water Insecure"))
  )
```

---

# 2. Missing data overview

Stacked bar chart showing the proportion of complete, true-NA, and sentinel-flagged
observations per analytic variable. See `04.Descriptive_stats.Rmd` for the full
numerical table.

```{r plot-missing, fig.width=8, fig.height=5}
analytic_cols <- c(
  "D_AGE", "D_LOC_TIME_CLEAN", "D_HH_SIZE", "D_CHLD_CLEAN",
  "SEASON", "SES_SC_Total", "SES_CAT_AMAI",
  "D_EMP_CLEAN", "D_EMP_FORM_CLEAN",
  "HLTH_SMK_CLEAN", "HLTH_TTM_CAT", "HLTH_MED_CAT",
  "HLTH_CPAIN_CAT", "HLTH_CDIS_CAT"
)

var_labels <- c(
  D_AGE             = "Age",
  D_LOC_TIME_CLEAN  = "Years in neighborhood",
  D_HH_SIZE         = "Household size",
  D_CHLD_CLEAN      = "Parity",
  SEASON            = "Season",
  SES_SC_Total      = "SES composite score",
  SES_CAT_AMAI      = "SES category",
  D_EMP_CLEAN       = "Employed (yes/no)",
  D_EMP_FORM_CLEAN  = "Employment type",
  HLTH_SMK_CLEAN    = "Smoking",
  HLTH_TTM_CAT      = "Hormonal treatment",
  HLTH_MED_CAT      = "Medication use",
  HLTH_CPAIN_CAT    = "Chronic pain",
  HLTH_CDIS_CAT     = "Chronic disease"
)

N <- nrow(df)

missing_tbl <- df %>%
  select(all_of(analytic_cols)) %>%
  summarise(across(everything(), list(
    n_na  = ~ sum(is.na(.)),
    n_999 = ~ sum(as.character(.) == "999", na.rm = TRUE),
    n_888 = ~ sum(as.character(.) == "888", na.rm = TRUE),
    n_777 = ~ sum(as.character(.) == "777", na.rm = TRUE),
    n_666 = ~ sum(as.character(.) == "666", na.rm = TRUE),
    n_555 = ~ sum(as.character(.) == "555", na.rm = TRUE),
    n_444 = ~ sum(as.character(.) == "444", na.rm = TRUE),
    n_333 = ~ sum(as.character(.) == "333", na.rm = TRUE)
  ))) %>%
  pivot_longer(everything(),
               names_to  = c("variable", ".value"),
               names_sep = "_(?=n_)") %>%
  mutate(
    label       = var_labels[variable],
    pct_na      = round(n_na / N * 100, 1),
    n_flagged   = n_999 + n_888 + n_777 + n_666 + n_555 + n_444 + n_333,
    pct_flagged = round(n_flagged / N * 100, 1),
    n_complete  = N - n_na - n_flagged,
    pct_complete = round(n_complete / N * 100, 1)
  ) %>%
  arrange(desc(n_na))

missing_tbl %>%
  select(label, pct_complete, pct_na, pct_flagged) %>%
  pivot_longer(-label, names_to = "type", values_to = "pct") %>%
  mutate(
    type  = factor(type,
                   levels = c("pct_complete", "pct_na", "pct_flagged"),
                   labels = c("Complete", "True NA", "Flagged (sentinel)")),
    label = factor(label, levels = rev(missing_tbl$label))
  ) %>%
  ggplot(aes(x = label, y = pct, fill = type)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = c("Complete"           = "#5bc0de",
                               "True NA"            = "#d9534f",
                               "Flagged (sentinel)" = "#f0ad4e")) +
  labs(title = "Missing data by variable",
       x = NULL, y = "Percentage (%)", fill = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

---

# 3. HWISE and PSS score distributions by season

```{r hwise-pss-histograms, fig.width=9, fig.height=4, message=FALSE, warning=FALSE}
p_hwise <- ggplot(df_table %>% filter(!is.na(HWISE_Total)),
                  aes(x = HWISE_Total, fill = Season)) +
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.6) +
  scale_fill_manual(values = c("Rainy (Oct-Dec 2022)" = "#5bc0de",
                               "Dry (Apr-Jun 2023)"   = "#d9534f")) +
  labs(title = "HWISE total score",
       subtitle = paste0("n = ", sum(!is.na(df_table$HWISE_Total))),
       x = "Score (0-36)", y = "Count", fill = NULL) +
  theme_minimal() + theme(legend.position = "bottom")

p_pss <- ggplot(df_table %>% filter(!is.na(PSS_Total)),
                aes(x = PSS_Total, fill = Season)) +
  geom_histogram(binwidth = 2, position = "identity", alpha = 0.6) +
  scale_fill_manual(values = c("Rainy (Oct-Dec 2022)" = "#5bc0de",
                               "Dry (Apr-Jun 2023)"   = "#d9534f")) +
  labs(title = "PSS-14 total score",
       subtitle = paste0("n = ", sum(!is.na(df_table$PSS_Total))),
       x = "Score (0-56)", y = "Count", fill = NULL) +
  theme_minimal() + theme(legend.position = "bottom")

grid.arrange(p_hwise, p_pss, ncol = 2)
```

---

# 4. PSS and water worry by water insecurity level

Boxplots stratified by HWISE category (Young et al. 2019 cut-points).
N per group is annotated below each box.

```{r pss-by-hwise-boxplots, fig.width=9, fig.height=4}
hwise_colors <- c(
  "Secure (0)"       = "#2ecc71",
  "Marginal (1-11)"  = "#f1c40f",
  "Moderate (12-23)" = "#e67e22",
  "Severe (24-36)"   = "#e74c3c"
)

p_pss_hwise <- df_table %>%
  filter(!is.na(HWISE_Cat) & !is.na(PSS_Total)) %>%
  ggplot(aes(x = HWISE_Cat, y = PSS_Total, fill = HWISE_Cat)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  stat_summary(fun.data = function(x) data.frame(y = -2, label = paste0("n=", length(x))),
               geom = "text", size = 3.2) +
  scale_fill_manual(values = hwise_colors) +
  labs(title = "PSS-14 by water insecurity level",
       x = NULL, y = "PSS-14 score (0-56)") +
  theme_minimal() + theme(legend.position = "none")

p_worry_hwise <- df_table %>%
  filter(!is.na(HWISE_Cat) & !is.na(Water_worry)) %>%
  ggplot(aes(x = HWISE_Cat, y = Water_worry, fill = HWISE_Cat)) +
  geom_boxplot(alpha = 0.7, outlier.shape = 21) +
  stat_summary(fun.data = function(x) data.frame(y = -0.5, label = paste0("n=", length(x))),
               geom = "text", size = 3.2) +
  scale_fill_manual(values = hwise_colors) +
  scale_y_continuous(breaks = 0:9) +
  labs(title = "Water worry subscale by water insecurity level",
       x = NULL, y = "Water worry score (0-9)") +
  theme_minimal() + theme(legend.position = "none")

grid.arrange(p_pss_hwise, p_worry_hwise, ncol = 2)
```

---

# 5. Objective water supply distributions

## 5.1 Hours of supply per week

```{r water-hrs-histogram, fig.width=8, fig.height=4, message=FALSE}
df_water %>%
  filter(!is.na(W_Hrs_week) & !is.na(W_WS_LOC_F)) %>%
  ggplot(aes(x = W_Hrs_week, fill = W_WS_LOC_F)) +
  geom_histogram(binwidth = 6, position = "identity", alpha = 0.6, color = "white") +
  scale_fill_manual(values = c("Water Secure" = "#4E8098", "Water Insecure" = "#E76F51")) +
  labs(
    title    = "Hours of water supply per week by neighborhood",
    subtitle = paste0("n = ", sum(!is.na(df_water$W_Hrs_week) & !is.na(df_water$W_WS_LOC_F)),
                      " (excludes sentinel codes and missing)"),
    x = "Hours per week", y = "Count", fill = NULL
  ) +
  theme_minimal() + theme(legend.position = "bottom")
```

## 5.2 Supply continuity and frequency by neighborhood

```{r water-cat-barplots, fig.width=10, fig.height=5}
nbhd_colors <- c("Water Secure" = "#4E8098", "Water Insecure" = "#E76F51")

cont_counts <- df_water %>%
  filter(!is.na(W_Continuity) & !is.na(W_WS_LOC_F)) %>%
  count(W_WS_LOC_F, W_Continuity)

freq_counts <- df_water %>%
  filter(!is.na(W_Frequency) & !is.na(W_WS_LOC_F)) %>%
  count(W_WS_LOC_F, W_Frequency)

p_cont <- ggplot(cont_counts,
                 aes(x = W_Continuity, y = n, fill = W_WS_LOC_F)) +
  geom_col(position = "dodge", alpha = 0.85) +
  geom_text(aes(label = n),
            position = position_dodge(width = 0.9),
            vjust = -0.4, size = 3.2) +
  scale_fill_manual(values = nbhd_colors) +
  labs(title = "Supply continuity", x = NULL, y = "Count", fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1), legend.position = "bottom")

p_freq <- ggplot(freq_counts,
                 aes(x = W_Frequency, y = n, fill = W_WS_LOC_F)) +
  geom_col(position = "dodge", alpha = 0.85) +
  geom_text(aes(label = n),
            position = position_dodge(width = 0.9),
            vjust = -0.4, size = 3.2) +
  scale_fill_manual(values = nbhd_colors) +
  labs(title = "Supply frequency", x = NULL, y = "Count", fill = NULL) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1), legend.position = "bottom")

grid.arrange(p_cont, p_freq, ncol = 2)
```

---

# 6. Hours of water supply vs HWISE and PSS

Scatter plots with linear regression line (± 95% CI).

```{r water-supply-vs-outcomes, fig.width=10, fig.height=4, message=FALSE}
df_water_scores <- df_water %>%
  mutate(
    HWISE_Total = clean_for_table(HW_TOTAL),
    PSS_Total   = clean_for_table(PSS_TOTAL)
  ) %>%
  filter(!is.na(W_Hrs_week) & !is.na(HWISE_Total) & !is.na(PSS_Total))

n_label <- nrow(df_water_scores)

p_hw <- ggplot(df_water_scores, aes(x = W_Hrs_week, y = HWISE_Total)) +
  geom_point(alpha = 0.4, size = 1.5, color = "#90A955") +
  geom_smooth(method = "lm", se = TRUE, color = "#386641") +
  labs(title = "Hours of supply vs HWISE",
       subtitle = paste0("n = ", n_label),
       x = "Hours of supply per week", y = "HWISE total score") +
  theme_minimal()

p_pss2 <- ggplot(df_water_scores, aes(x = W_Hrs_week, y = PSS_Total)) +
  geom_point(alpha = 0.4, size = 1.5, color = "#4E8098") +
  geom_smooth(method = "lm", se = TRUE, color = "#1d3557") +
  labs(title = "Hours of supply vs PSS-14",
       subtitle = paste0("n = ", n_label),
       x = "Hours of supply per week", y = "PSS-14 total score") +
  theme_minimal()

grid.arrange(p_hw, p_pss2, ncol = 2)
```
