---
title: "06.1 PSS Models – Perceived Stress as Outcome"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(gtsummary)
library(lmtest)
library(flextable)
library(ggplot2)
library(officer)
library(ggstance)
```

## Overview

SHOULD consider removing data point with household size of 40 (vecindad), as it may be distorting results. 

This script fits a series of **linear regression models** predicting **Perceived
Stress Scale total score (`PSS_TOTAL`)**. Each model shares the same set of core
control variables; one water-related predictor is added per model. Models are
defined in a named list (`model_specs`).

| Role | Variables |
|------|-----------|
| Outcome | `PSS_TOTAL` (continuous, 0–56) |
| Core (every model) | Age (`D_AGE`), SES composite (`SES_SC_Total`), Household size (`D_HH_SIZE`) |
| Augmenting (one per model) | HWISE total · individual HWISE items · hours water supply/week · continuous/intermittent supply · WS/WI neighborhood · neighborhood comparison · time handling water |

---

# Load and Prepare Data

```{r load-data}
MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

# Replace sentinel codes with NA for numeric variables
clean_num <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES, NA_real_, x_num)
}

df_raw <- read_csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
                   show_col_types = FALSE)

cat("Dataset:", nrow(df_raw), "rows ×", ncol(df_raw), "cols\n")
```

```{r prepare-variables}
hwise_items <- c("HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
                 "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
                 "HW_ANGRY", "HW_SLEEP",  "HW_NONE",    "HW_SHAME")

df <- df_raw %>%
  mutate(
    # ── Outcome ────────────────────────────────────────────────────────────────
    PSS_TOTAL    = clean_num(PSS_TOTAL),

    # ── Core predictors ────────────────────────────────────────────────────────
    D_AGE        = clean_num(D_AGE),
    SES_SC_Total = clean_num(SES_SC_Total),
    D_HH_SIZE    = clean_num(D_HH_SIZE),

    # ── HWISE ──────────────────────────────────────────────────────────────────
    HW_TOTAL     = clean_num(HW_TOTAL),
    across(all_of(hwise_items), clean_num),

    # ── Hours of water supply per week ─────────────────────────────────────────
    HRS_PER_WEEK = clean_num(MX5_WFREQ_HRS_PER_WEEK),

    # ── Continuous vs. intermittent supply ─────────────────────────────────────
    # Accepts numeric (0/1) or text ("Continuo"/"Intermitente")
    CONT_INT = {
      raw <- as.character(MX3_CONT_INT)
      raw_num <- suppressWarnings(as.numeric(raw))
      factor(case_when(
        raw_num %in% MISSING_CODES                         ~ NA_character_,
        raw_num == 0                                       ~ "Continuous",
        raw_num == 1                                       ~ "Intermittent",
        str_detect(str_to_lower(raw), "contin")            ~ "Continuous",
        str_detect(str_to_lower(raw), "inter")             ~ "Intermittent",
        is.na(raw) | str_trim(raw) == "" | raw == "NA"    ~ NA_character_,
        TRUE                                               ~ raw
      ), levels = c("Continuous", "Intermittent"))  # Continuous = reference
    },

    # ── WS/WI neighborhood classification ─────────────────────────────────────
    WS_LOC = factor(W_WS_LOC),

    # ── Neighborhood water quality comparison (Q28) ────────────────────────────
    # 0 = Better (reference), 1 = Same, 2 = Worse
    NB_COMP = factor(
      case_when(
        clean_num(MX28_WQ_COMP_CAT) == 0 ~ "Better",
        clean_num(MX28_WQ_COMP_CAT) == 1 ~ "Same",
        clean_num(MX28_WQ_COMP_CAT) == 2 ~ "Worse",
        TRUE                             ~ NA_character_
      ),
      levels = c("Better", "Same", "Worse")
    ),

    # ── Time handling water (Q19) ──────────────────────────────────────────────
    # 0 = No (reference), 1 = Yes
    TIME_WATER = factor(
      case_when(
        clean_num(MX19_TIME_CAT) == 0 ~ "No",
        clean_num(MX19_TIME_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── Season ─────────────────────────────────────────────────────────────────
    # 1 = Season 1, 2 = Season 2; used in CORE2 and stratified sensitivity
    SEASON = factor(SEASON, levels = c(0, 1), labels = c("Season 1", "Season 2"))
  )
```

```{r check-missingness}
analytic_vars <- c("PSS_TOTAL", "D_AGE", "SES_SC_Total", "D_HH_SIZE", "SEASON",
                   "HW_TOTAL", hwise_items,
                   "HRS_PER_WEEK", "CONT_INT", "WS_LOC", "NB_COMP", "TIME_WATER")

df %>%
  summarise(across(all_of(analytic_vars), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_missing") %>%
  mutate(pct_missing = round(100 * n_missing / nrow(df), 1)) %>%
  arrange(desc(n_missing)) %>%
  print(n = Inf)
```

---

# Model Specifications {#model-specs}

Models are defined in a **named list**. To add a model later, append a new entry
following the same structure: `list(label = "...", group = "...", formula = ...)`.

```{r model-specs}
core_formula  <- PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE
core2_formula <- PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + SEASON

# ── Individual HWISE item models (one per item, built on CORE1) ────────────────
hwise_item_specs <- setNames(
  lapply(hwise_items, function(item) {
    list(
      label   = paste0("Core + ", item),
      group   = "Individual HWISE items",
      formula = as.formula(
        paste("PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE +", item)
      )
    )
  }),
  paste0("hwise_", tolower(hwise_items))
)

# ── Main registry ──────────────────────────────────────────────────────────────
model_specs <- c(
  list(
    # ── Base models ─────────────────────────────────────────────────────────────
    core = list(
      label   = "M0: Core only (base)",
      group   = "Base",
      formula = core_formula
    ),
    core2 = list(
      label   = "M0b: Core2 (+ Season)",
      group   = "Base",
      formula = core2_formula
    ),
    # ── HWISE on Core1 ──────────────────────────────────────────────────────────
    hwise_total = list(
      label   = "M1: Core + HWISE Total",
      group   = "HWISE",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + HW_TOTAL
    ),
    hwise_log = list(
      label   = "M1: Core + LogHWISE Total",
      group   = "HWISE",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + log1p(HW_TOTAL)
    ),
    # ── HWISE on Core2 (controls for Season) ────────────────────────────────────
    core2_hwise_total = list(
      label   = "M1b: Core2 + HWISE Total",
      group   = "HWISE",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + SEASON + HW_TOTAL
    ),
    core2_hwise_log = list(
      label   = "M1b: Core2 + LogHWISE Total",
      group   = "HWISE",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + SEASON + log1p(HW_TOTAL)
    ),
    # ── HWISE × Season interaction ───────────────────────────────────────────
    core3_hwise_total = list(
      label   = "M1c: Core2 × HWISE Total (interaction)",
      group   = "HWISE",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + SEASON + HW_TOTAL + SEASON:HW_TOTAL
    ),
    core3_hwise_log = list(
      label   = "M1c: Core2 × LogHWISE Total (interaction)",
      group   = "HWISE",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + SEASON + log1p(HW_TOTAL) + SEASON:log1p(HW_TOTAL)
    )
  ),
  hwise_item_specs,
  list(
    hrs_supply = list(
      label   = "M2: Core + Hours water supply/week",
      group   = "QUAIL-Mex",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + HRS_PER_WEEK
    ),
    cont_int = list(
      label   = "M3: Core + Continuous/intermittent supply",
      group   = "QUAIL-Mex",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + CONT_INT
    ),
    ws_loc = list(
      label   = "M4: Core + WS/WI neighborhood",
      group   = "QUAIL-Mex",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + WS_LOC
    ),
    nb_comp = list(
      label   = "M5: Core + Neighborhood comparison (Q28)",
      group   = "QUAIL-Mex",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + NB_COMP
    ),
    time_water = list(
      label   = "M6: Core + Time handling water (Q19)",
      group   = "QUAIL-Mex",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + TIME_WATER
    )
  )
)

cat("Total models registered:", length(model_specs), "\n")
```

---

# Fit Models

Each model uses listwise deletion (`na.action = na.omit`) for the variables it
includes. For valid F-tests comparing an augmented model to M0, the core model
is **refit on the same rows** as the augmented model.

```{r fit-models}
fitted_models <- lapply(model_specs, function(spec) {
  lm(spec$formula, data = df, na.action = na.omit)
})

# Sample sizes per model
tibble(
  model_id = names(model_specs),
  label    = map_chr(model_specs, "label"),
  n        = map_int(fitted_models, nobs)
) %>% print(n = Inf)
```

---

# Per-Model Coefficient Tables

```{r results-tables, results="asis"}
core_label_list <- list(
  D_AGE        ~ "Age (years)",
  SES_SC_Total ~ "SES composite score",
  D_HH_SIZE    ~ "Household size"
)

for (nm in names(fitted_models)) {
  spec <- model_specs[[nm]]
  mod  <- fitted_models[[nm]]
  fit  <- glance(mod)

  cat("\n\n## ", spec$label, " {.unnumbered}\n\n", sep = "")
  cat("**N =**", nobs(mod),
      "| **R² =**", round(fit$r.squared, 3),
      "| **Adj. R² =**", round(fit$adj.r.squared, 3),
      "| **AIC =**", round(fit$AIC, 1),
      "| **BIC =**", round(fit$BIC, 1), "\n\n")

  print(
    tbl_regression(
      mod,
      label    = core_label_list,
      conf.int = TRUE
    ) %>%
      bold_p(t = 0.05) %>%
      italicize_levels() %>%
      modify_caption(paste("**", spec$label, "**"))
  )
}
```

---

# Model Comparison

Each augmented model is compared to M0 re-fit on the same complete cases,
using an **F-test for added predictors**. Models with different N from M0
(due to missing data in the augmenting predictor) are flagged.

```{r model-comparison}
comparison <- map_dfr(names(fitted_models), function(nm) {
  mod  <- fitted_models[[nm]]
  spec <- model_specs[[nm]]
  fit  <- glance(mod)
  n_mod <- nobs(mod)

  base_row <- tibble(
    model_id     = nm,
    label        = spec$label,
    group        = spec$group,
    n            = n_mod,
    R2           = round(fit$r.squared,     3),
    adj_R2       = round(fit$adj.r.squared, 3),
    delta_adj_R2 = NA_real_,
    AIC          = round(fit$AIC, 1),
    BIC          = round(fit$BIC, 1),
    F_stat       = NA_real_,
    F_df1        = NA_integer_,
    F_df2        = NA_integer_,
    F_p          = NA_real_,
    note         = ""
  )

  if (nm == "core") return(base_row)

  # Refit M0 on same rows as augmented model
  rows_used <- as.integer(rownames(model.frame(mod)))
  m0_sub    <- lm(core_formula, data = df[rows_used, ])
  lrt_tab   <- anova(m0_sub, mod)
  f_stat    <- lrt_tab$F[2]
  f_df1     <- lrt_tab$Df[2]
  f_df2     <- lrt_tab$Res.Df[2]
  f_p       <- lrt_tab$`Pr(>F)`[2]

  base_row %>% mutate(
    delta_adj_R2 = round(fit$adj.r.squared - glance(m0_sub)$adj.r.squared, 3),
    F_stat       = round(f_stat, 2),
    F_df1        = as.integer(f_df1),
    F_df2        = as.integer(f_df2),
    F_p          = round(f_p, 4),
    note         = if_else(n_mod != nobs(fitted_models[["core"]]),
                           "N differs from full M0", "")
  )
})

comparison %>%
  select(label, group, n, R2, adj_R2, delta_adj_R2, AIC, BIC,
         F_stat, F_df1, F_df2, F_p, note) %>%
  flextable() %>%
  set_header_labels(
    label        = "Model",
    group        = "Group",
    n            = "N",
    R2           = "R²",
    adj_R2       = "Adj. R²",
    delta_adj_R2 = "Δ Adj. R²",
    AIC          = "AIC",
    BIC          = "BIC",
    F_stat       = "F",
    F_df1        = "df₁",
    F_df2        = "df₂",
    F_p          = "p (F-test)",
    note         = "Note"
  ) %>%
  color(~ !is.na(F_p) & F_p < 0.05, j = "F_p", color = "orange2") %>%
  color(~ !is.na(F_p) & F_p >= 0.05, j = "F_p", color = "grey50") %>%
  autofit()
```

## On interactions

```{r}
summary(lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + HW_TOTAL, data = filter(df, SEASON == "Season 2")))
```

```{r interaction-f2}
# Cohen's f² for the SEASON × HW_TOTAL interaction term
# f² = (R²_M1c - R²_M1b) / (1 - R²_M1c)
r2_m1b <- summary(fitted_models[["core2_hwise_total"]])$r.squared
r2_m1c <- summary(fitted_models[["core3_hwise_total"]])$r.squared
f2_interaction <- (r2_m1c - r2_m1b) / (1 - r2_m1c)
cat(sprintf(
  "R² M1b (Core2 + HWISE):            %.4f\n
   R² M1c (Core2 + HWISE × Season):   %.4f\n
   ΔR²  (interaction added):           %.4f\n
   Cohen's f² (interaction term):      %.4f\n
   Interpretation: %s\n",
  r2_m1b, r2_m1c, r2_m1c - r2_m1b, f2_interaction,
  dplyr::case_when(
    f2_interaction >= 0.35 ~ "large",
    f2_interaction >= 0.15 ~ "medium",
    f2_interaction >= 0.02 ~ "small",
    TRUE                   ~ "negligible"
  )
))
```

### Individual HWISE Items – Summary

```{r hwise-item-comparison}
# Subset comparison table to individual items for a tidy summary
comparison %>%
  filter(group == "Individual HWISE items") %>%
  select(label, n, R2, adj_R2, delta_adj_R2, F_stat, F_p) %>%
  arrange(F_p) %>%
  flextable() %>%
  set_header_labels(
    label        = "Item model",
    n            = "N",
    R2           = "R²",
    adj_R2       = "Adj. R²",
    delta_adj_R2 = "Δ Adj. R²",
    F_stat       = "F",
    F_p          = "p (F-test)"
  ) %>%
  color(~ !is.na(F_p) & F_p < 0.05, j = "F_p", color = "orange2") %>%
  autofit()
```

---

# Coefficient Plot (Augmenting Predictors)

```{r coef-plot, fig.width = 10, fig.height = 7}
core_terms <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE")

coef_data <- map_dfr(names(fitted_models), function(nm) {
  tidy(fitted_models[[nm]], conf.int = TRUE) %>%
    filter(!term %in% core_terms) %>%
    mutate(
      model_id = nm,
      label    = model_specs[[nm]]$label,
      group    = model_specs[[nm]]$group
    )
})

if (nrow(coef_data) > 0) {
  ggplot(coef_data,
         aes(x = estimate, y = reorder(label, estimate),
             xmin = conf.low, xmax = conf.high, color = group)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
    geom_errorbarh(height = 0.2, linewidth = 0.6) +
    geom_point(size = 2.5) +
    scale_color_brewer(palette = "Dark2") +
    labs(
      title    = "Augmenting predictor coefficients — PSS Total as outcome",
      subtitle = "Core variables (Age, SES, HH size) omitted; bars = 95% CI",
      x        = "β coefficient",
      y        = NULL,
      color    = "Predictor group"
    ) +
    theme_bw(base_size = 11) +
    theme(legend.position = "bottom")
}
```

```{r hwise-coef-plot, fig.width = 8, fig.height = 6}
# Readable labels for HWISE items
hwise_labels <- c(
  HW_WORRY   = "Worry about water",
  HW_INTERR  = "Interrupted supply",
  HW_CLOTHES = "Couldn't wash clothes",
  HW_PLANS   = "Plans disrupted",
  HW_FOOD    = "Food preparation affected",
  HW_HANDS   = "Couldn't wash hands",
  HW_BODY    = "Couldn't bathe",
  HW_DRINK   = "Not enough to drink",
  HW_ANGRY   = "Felt angry",
  HW_SLEEP   = "Sleep disrupted",
  HW_NONE    = "No water at all",
  HW_SHAME   = "Felt shame"
)

hwise_ids <- c("hwise_total", paste0("hwise_", tolower(hwise_items)))

hwise_coef <- map_dfr(hwise_ids, function(nm) {
  tidy(fitted_models[[nm]], conf.int = TRUE) %>%
    filter(!term %in% c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE")) %>%
    mutate(
      model_id  = nm,
      item_label = if_else(
        term == "HW_TOTAL",
        "HWISE Total Score",
        hwise_labels[term]
      ),
      significant = p.value < 0.05
    )
})

ggplot(hwise_coef,
       aes(x = estimate, y = reorder(item_label, estimate),
           xmin = conf.low, xmax = conf.high,
           color = significant, shape = (term == "HW_TOTAL"))) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(height = 0.25, linewidth = 0.7) +
  geom_point(size = 3) +
  scale_color_manual(
    values = c("TRUE" = "#1a7abf", "FALSE" = "grey60"),
    labels = c("TRUE" = "p < 0.05", "FALSE" = "p ≥ 0.05")
  ) +
  scale_shape_manual(
    values = c("TRUE" = 18, "FALSE" = 16),
    labels = c("TRUE" = "Total score", "FALSE" = "Individual item"),
    guide  = guide_legend(title = NULL)
  ) +
  labs(
    title    = "HWISE items as predictors of PSS Total",
    subtitle = "Each item in a separate model; core variables (Age, SES, HH size) controlled; bars = 95% CI",
    x        = "β coefficient",
    y        = NULL,
    color    = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom")
```

---

# Interpreting the Results

## What each coefficient means

For each augmented model, the **β coefficient** for the water-related predictor
is the expected change in PSS_TOTAL for a one-unit increase in that predictor,
*holding Age, SES, and Household size constant*. A positive β indicates higher
water burden associates with higher perceived stress.

**95% CI**: if the interval does not cross zero, the association is significant
at α = 0.05. Wide intervals signal imprecision — check your N in that model.

## Model fit criteria

| Criterion | Prefer |
|-----------|--------|
| **Adj. R²** | Higher (more variance explained, penalized for parameters) |
| **Δ Adj. R²** | Positive values mean the water predictor explains additional variance |
| **AIC / BIC** | Lower. If AIC/BIC of an augmented model > M0's AIC/BIC, the predictor adds noise |
| **F-test p** | < 0.05 suggests the augmenting predictor(s) jointly improve fit |

## Core variable benchmarks (M0)

Check M0 first:

- **SES_SC_Total**: expect a **negative** β (higher SES → lower stress). If
  positive or near-zero, investigate SES operationalization.
- **D_AGE**: typically a modest association (positive or U-shaped).
- **D_HH_SIZE**: larger households may share water burden — direction is theoretically ambiguous.

## HWISE Total vs. Individual Items

- HWISE total (HW_TOTAL) provides a summary test: *does overall water insecurity
  predict stress beyond demographics?*
- Individual item models reveal **which dimensions** drive stress (e.g., worry
  vs. physical access). Interpret items one at a time — they are intercorrelated
  components of the same scale. **Do not** enter all 12 items in a single model
  without addressing multicollinearity (VIF > 5 is a concern; use `car::vif()`).



## Categorical predictors

For CONT_INT, WS_LOC, NB_COMP, TIME_WATER, the F-test p-value tests the
**overall** effect of the variable. Inspect individual level contrasts from the
coefficient table. Reference levels: Continuous supply, "Better" neighborhood
comparison, "No" time investment.

---

# Sensitivity Analysis: Season-Stratified Models

Models are re-run separately for each season on a user-defined subset of
`model_specs`. This checks whether associations hold across both data collection
periods, or whether a particular season drives the pooled results.

To add or remove models from the stratified analysis, edit `strat_ids` below.

```{r season-strat, results="asis"}
# ── Models to stratify ─────────────────────────────────────────────────────────
strat_ids <- names(model_specs)  # all registered models

seasons <- levels(df$SEASON)
cat("Seasons in data:", paste(seasons, collapse = ", "), "\n")
cat("N per season:\n")
print(table(df$SEASON, useNA = "always"))

# ── Fit stratified models ──────────────────────────────────────────────────────
strat_models <- lapply(seasons, function(s) {
  df_s <- filter(df, SEASON == s)
  lapply(model_specs[strat_ids], function(spec) {
    # Drop SEASON (main effect + any interactions) using R formula tools —
    # robust to function calls (e.g. log1p()) that break regex approaches
    tt          <- terms(spec$formula)
    all_terms   <- attr(tt, "term.labels")
    keep_terms  <- all_terms[!grepl("SEASON", all_terms)]
    response    <- as.character(attr(tt, "variables")[[2]])
    f_clean     <- reformulate(keep_terms, response = response)
    lm(f_clean, data = df_s, na.action = na.omit)
  })
})
names(strat_models) <- seasons

# ── Comparison table: focal predictor β per season ────────────────────────────
core_terms <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE")

strat_comparison <- map_dfr(strat_ids, function(nm) {
  map_dfr(seasons, function(s) {
    mod <- strat_models[[s]][[nm]]
    if (is.null(mod)) return(NULL)
    fit <- glance(mod)
    coefs <- tidy(mod, conf.int = TRUE) %>%
      filter(!term %in% core_terms)
    if (nrow(coefs) == 0) {
      coefs <- tibble(term = "—", estimate = NA_real_,
                      conf.low = NA_real_, conf.high = NA_real_, p.value = NA_real_)
    }
    coefs %>%
      mutate(
        model_id  = nm,
        label     = model_specs[[nm]]$label,
        season    = s,
        n         = nobs(mod),
        adj_R2    = round(fit$adj.r.squared, 3)
      )
  })
})

# Wide format: Season 1 and Season 2 side by side
strat_wide <- strat_comparison %>%
  select(label, term, season, n, adj_R2, estimate, conf.low, conf.high, p.value) %>%
  mutate(
    ci   = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_fmt = case_when(
      is.na(p.value)      ~ "—",
      p.value < 0.001     ~ "<0.001",
      TRUE                ~ sprintf("%.3f", p.value)
    )
  ) %>%
  select(label, term, season, n, adj_R2, ci, p_fmt) %>%
  pivot_wider(
    names_from  = season,
    values_from = c(n, adj_R2, ci, p_fmt),
    names_glue  = "{season}_{.value}"
  )

strat_wide %>%
  flextable() %>%
  set_header_labels(
    label  = "Model",
    term   = "Predictor"
  ) %>%
  autofit() %>%
  print()
```

```{r season-coef-plot, fig.width = 10, fig.height = 10}
# Coefficient plot by season for focal predictors (core variables excluded)
strat_comparison %>%
  filter(!is.na(estimate)) %>%
  ggplot(aes(x = estimate, y = reorder(label, estimate),
             xmin = conf.low, xmax = conf.high, color = season)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(height = 0.2, linewidth = 0.6,
                 position = position_dodgev(height = 0.5)) +
  geom_point(size = 2.5, position = position_dodgev(height = 0.5)) +
  scale_color_manual(values = c("Season 1" = "#2166ac", "Season 2" = "#d6604d")) +
  labs(
    title    = "Season-stratified coefficients — PSS Total as outcome",
    subtitle = "Core variables omitted; bars = 95% CI",
    x        = "β coefficient",
    y        = NULL,
    color    = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom")
```

```{r combined-coef-plot, fig.width = 11, fig.height = 11}
# ── Terms to exclude from all sources ─────────────────────────────────────────
exclude_terms <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE",
                   "SEASONSeason 2", "SEASONSeason 2:HW_TOTAL",
                   "HW_TOTAL:SEASONSeason 2")

# ── Overall (pooled) estimates ─────────────────────────────────────────────────
overall_coef <- map_dfr(strat_ids, function(nm) {
  tidy(fitted_models[[nm]], conf.int = TRUE) %>%
    filter(!term %in% exclude_terms) %>%
    mutate(
      source   = "Overall",
      model_id = nm,
      label    = model_specs[[nm]]$label
    )
})

# ── Season-stratified estimates ────────────────────────────────────────────────
strat_coef <- map_dfr(strat_ids, function(nm) {
  map_dfr(seasons, function(s) {
    mod <- strat_models[[s]][[nm]]
    if (is.null(mod)) return(NULL)
    tidy(mod, conf.int = TRUE) %>%
      filter(!term %in% exclude_terms) %>%
      mutate(
        source   = s,
        model_id = nm,
        label    = model_specs[[nm]]$label
      )
  })
})

# ── Combine and plot ───────────────────────────────────────────────────────────
combined_coef <- bind_rows(overall_coef, strat_coef) %>%
  filter(!is.na(estimate)) %>%
  mutate(source = factor(source, levels = c("Overall", seasons)))

ggplot(combined_coef,
       aes(x = estimate, y = reorder(label, estimate),
           xmin = conf.low, xmax = conf.high,
           color = source)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(height = 0.25, linewidth = 0.6,
                 position = position_dodgev(height = 0.7)) +
  geom_point(size = 2.5, position = position_dodgev(height = 0.7)) +
  scale_color_manual(
    values = c("Overall"  = "#e87fae",
               "Season 1" = "#2166ac",
               "Season 2" = "#d6604d"),
    name = NULL
  ) +
  labs(
    title    = "Overall and season-stratified coefficients — PSS Total as outcome",
    subtitle = "Core variables (Age, SES, HH size) and SEASON main effect omitted; bars = 95% CI",
    x        = "β coefficient",
    y        = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom")
```


---

# Strengthening the Analysis

1. **Check regression assumptions** after fitting each model:
```{r check-regression}
plot(fitted_models[["hwise_total"]])   # base R diagnostic plots
# or: performance::check_model(fitted_models[["hwise_total"]])


plot(fitted_models[["core3_hwise_total"]], which = 4)

```
   Look for: non-linearity (residuals vs. fitted), heteroscedasticity
   (scale–location), non-normality (Q–Q), and influential points (Cook's D > 4/n).

2. **Standardize continuous predictors** before fitting to make β coefficients
   directly comparable across different scales:
   
```{r}
   df <- df %>% mutate(across(c(D_AGE, SES_SC_Total, D_HH_SIZE, HW_TOTAL),
                               ~ scale(.)[,1], .names = "{.col}_z"))


augment(fitted_models[["core3_hwise_total"]]) %>%
  filter(.cooksd > 4/339) %>%
  select(.cooksd, PSS_TOTAL, HW_TOTAL, SEASON, everything())

```

3. **Multiple testing correction**: With ~18 models on the same sample, the
   familywise error rate is elevated. Apply Benjamini–Hochberg correction to
   the vector of F-test p-values:
```{r}
   comparison %>%
     filter(!is.na(F_p)) %>%
     mutate(F_p_adj = p.adjust(F_p, method = "BH"))
```

4. **Missing data**: If any key predictor has >5% missing, consider multiple imputation with `mice`:


5. **Additional confounders** to consider adding to the core model:
   - `HLTH_CPAIN_CAT` / `HLTH_CDIS_CAT` (chronic illness/pain)
   - `D_EMP_FORM_CLEAN` (formal vs. informal employment)
   - `D_CHLD_CLEAN` (number of children)
   - `SEASON` (dry vs. wet; especially important if testing water supply
     continuity)

6. **Sensitivity analyses**:
   - Stratify by `SEASON` to assess whether associations differ seasonally.
   - Exclude participants flagged in `PSS_REVIEW_FLAG`.
   - Test interaction terms: `HW_TOTAL * SES_SC_Total`, `HW_TOTAL * D_AGE`.

7. **Effect size**: report Cohen's *f*² = R²_change / (1 − R²_augmented) as a
   scale-independent effect size alongside β:

```{r}

   comparison %>%
     filter(!is.na(delta_adj_R2)) %>%
     mutate(cohens_f2 = delta_adj_R2 / (1 - adj_R2))
```

---

# Limitations

1. **Cross-sectional design**: Predictors and outcome were measured at the same
   time point. Temporal precedence cannot be established — stress could
   precede and bias water-related perceptions (reverse causality).

2. **Common method bias**: Both PSS and most water variables are self-reported.
   Shared measurement method can inflate associations regardless of any true
   relationship.

3. **Listwise deletion**: Each model drops participants missing any included
   variable. Effective N varies across models, complicating direct comparisons.
   If missingness is informative (e.g., more common in extreme SES groups),
   results may be biased.

4. **Multiple testing**: Testing ~18 separate models on the same outcome and
   sample substantially increases the familywise error rate. All findings should
   be treated as exploratory and replicated in independent samples.

5. **HWISE item multicollinearity**: Individual items are components of the
   total score and are highly intercorrelated. Running 12 item models is
   exploratory; enter items individually, not jointly, without a specific
   dimensionality hypothesis.

6. **HWISE response scale treated as interval**: Individual HWISE items use
   4-point response options (0 = never, 1 = rarely, 2 = sometimes,
   3 = often/always) and are entered as numeric predictors, assuming equal
   spacing between categories. This is consistent with prior HWISE validation
   literature and standard practice for Likert-type items in regression, but
   the interval assumption is unverified. Sensitivity analyses using dummy-coded
   categories could assess whether results are robust to this assumption.

7. **Unmeasured confounders**: Violence exposure, family conflict, financial
   shocks, mental health history, and healthcare access are not captured and may
   confound the water–stress relationship.

8. **Generalizability**: Findings reflect a specific sample from colonias in one
   urban context. Caution is warranted when extrapolating to other settings,
   water systems, or populations.

9. **PSS as a state measure**: PSS captures perceived stress over the past month.
   Acute events unrelated to water may dominate PSS scores, diluting
   water-specific associations.

10. **Seasonal conflation**: Data from two seasons may blend seasonal within-person
    change with stable between-person differences. Include `SEASON` as a covariate
    or stratify in confirmatory analyses.

11. **Linear assumption**: Linear regression assumes a constant marginal
    relationship between predictors and PSS. Non-linear or threshold effects
    (e.g., stress only rising above a certain HWISE score) would be missed.
    Check with splines or polynomial terms if diagnostics suggest non-linearity.
