---
title: "06.2 Emotional Response Models – Negative Emotion as Outcome"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(gtsummary)
library(lmtest)
library(flextable)
library(ggplot2)
library(officer)
library(ggstance)
```

## Overview

This script fits a series of **logistic regression models** predicting
**negative emotional response to household water (`MX26_EM_HHW_CAT`)**,
coded 0 = positive, 1 = negative. Each model shares the same core predictors;
one water-related predictor is added per model. Models are defined in a named
list (`model_specs`) that is easy to extend.

Results are reported as **odds ratios (OR)** with 95% CI.

| Role | Variables |
|------|-----------|
| Outcome | `MX26_EM_HHW_CAT` — 0 = positive emotion, 1 = negative emotion (Q26) |
| Core (every model) | Age (`D_AGE`), SES composite (`SES_SC_Total`), Household size (`D_HH_SIZE`), Perceived stress (`PSS_TOTAL`) |
| Augmenting (one per model) | HWISE total · individual HWISE items · hours water supply/week · continuous/intermittent supply · WS/WI neighborhood · neighborhood comparison · time handling water |

> **Note on PSS_TOTAL in the core**: PSS_TOTAL is included as a core covariate
> to test whether water variables predict emotional responses *over and above*
> general stress — i.e., whether the water–emotion link is not fully mediated
> by overall stress.

---

# Load and Prepare Data

```{r load-data}
MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

clean_num <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES, NA_real_, x_num)
}

df_raw <- read_csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
                   show_col_types = FALSE)

cat("Dataset:", nrow(df_raw), "rows ×", ncol(df_raw), "cols\n")
```

```{r prepare-variables}
hwise_items <- c("HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
                 "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
                 "HW_ANGRY", "HW_SLEEP",  "HW_NONE",    "HW_SHAME")

df <- df_raw %>%
  mutate(
    # ── Outcome ────────────────────────────────────────────────────────────────
    # Keep only valid 0/1 values; exclude sentinels (444, 555, 999)
    EMO_NEG = {
      raw <- clean_num(MX26_EM_HHW_CAT)
      case_when(
        raw == 0 ~ 0L,   # positive
        raw == 1 ~ 1L,   # negative
        TRUE     ~ NA_integer_
      )
    },

    # ── Core predictors ────────────────────────────────────────────────────────
    D_AGE        = clean_num(D_AGE),
    SES_SC_Total = clean_num(SES_SC_Total),
    D_HH_SIZE    = clean_num(D_HH_SIZE),
    PSS_TOTAL    = clean_num(PSS_TOTAL),

    # ── HWISE ──────────────────────────────────────────────────────────────────
    HW_TOTAL     = clean_num(HW_TOTAL),
    across(all_of(hwise_items), clean_num),

    # ── Hours of water supply per week ─────────────────────────────────────────
    HRS_PER_WEEK = clean_num(MX5_WFREQ_HRS_PER_WEEK),

    # ── Continuous vs. intermittent supply ─────────────────────────────────────
    CONT_INT = {
      raw     <- as.character(MX3_CONT_INT)
      raw_num <- suppressWarnings(as.numeric(raw))
      factor(case_when(
        raw_num %in% MISSING_CODES                      ~ NA_character_,
        raw_num == 0                                    ~ "Continuous",
        raw_num == 1                                    ~ "Intermittent",
        str_detect(str_to_lower(raw), "contin")         ~ "Continuous",
        str_detect(str_to_lower(raw), "inter")          ~ "Intermittent",
        is.na(raw) | str_trim(raw) == "" | raw == "NA" ~ NA_character_,
        TRUE                                            ~ raw
      ), levels = c("Continuous", "Intermittent"))
    },

    # ── WS/WI neighborhood ─────────────────────────────────────────────────────
    WS_LOC = factor(W_WS_LOC),

    # ── Neighborhood water quality comparison (Q28) ────────────────────────────
    NB_COMP = factor(
      case_when(
        clean_num(MX28_WQ_COMP_CAT) == 0 ~ "Better",
        clean_num(MX28_WQ_COMP_CAT) == 1 ~ "Same",
        clean_num(MX28_WQ_COMP_CAT) == 2 ~ "Worse",
        TRUE                             ~ NA_character_
      ),
      levels = c("Better", "Same", "Worse")
    ),

    # ── Time handling water (Q19) ──────────────────────────────────────────────
    TIME_WATER = factor(
      case_when(
        clean_num(MX19_TIME_CAT) == 0 ~ "No",
        clean_num(MX19_TIME_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── Season ─────────────────────────────────────────────────────────────────
    SEASON = factor(SEASON, levels = c(0, 1), labels = c("Season 1", "Season 2"))
  )
```

```{r check-outcome}
cat("EMO_NEG distribution:\n")
print(table(df$EMO_NEG, useNA = "always"))
cat("\nProportion negative (valid responses):",
    round(mean(df$EMO_NEG == 1, na.rm = TRUE), 3), "\n")
```

```{r check-missingness}
analytic_vars <- c("EMO_NEG", "D_AGE", "SES_SC_Total", "D_HH_SIZE", "PSS_TOTAL",
                   "SEASON", "HW_TOTAL", hwise_items,
                   "HRS_PER_WEEK", "CONT_INT", "WS_LOC", "NB_COMP", "TIME_WATER")

df %>%
  summarise(across(all_of(analytic_vars), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_missing") %>%
  mutate(pct_missing = round(100 * n_missing / nrow(df), 1)) %>%
  arrange(desc(n_missing)) %>%
  print(n = Inf)
```

---

# Model Specifications {#model-specs}

As in `06.1.pss_models.Rmd`, models live in a named list. Add new models by
appending entries with `label`, `group`, and `formula`.

```{r model-specs}
core_formula  <- EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL
core2_formula <- EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + SEASON

# ── Individual HWISE item models (built on Core1) ──────────────────────────────
hwise_item_specs <- setNames(
  lapply(hwise_items, function(item) {
    list(
      label   = paste0("Core + ", item),
      group   = "Individual HWISE items",
      formula = as.formula(
        paste("EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL +", item)
      )
    )
  }),
  paste0("hwise_", tolower(hwise_items))
)

# ── Main registry ──────────────────────────────────────────────────────────────
model_specs <- c(
  list(
    core = list(
      label   = "M0: Core only (base)",
      group   = "Base",
      formula = core_formula
    ),
    core2 = list(
      label   = "M0b: Core2 (+ Season)",
      group   = "Base",
      formula = core2_formula
    ),
    hwise_total = list(
      label   = "M1: Core + HWISE Total",
      group   = "HWISE",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + HW_TOTAL
    ),
    core2_hwise_total = list(
      label   = "M1b: Core2 + HWISE Total",
      group   = "HWISE",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + SEASON + HW_TOTAL
    ),
    # ── HWISE × Season interaction ───────────────────────────────────────────
    core3_hwise_total = list(
      label   = "M1c: Core2 × HWISE Total (interaction)",
      group   = "HWISE",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + SEASON + HW_TOTAL + SEASON:HW_TOTAL
    )
  ),
  hwise_item_specs,
  list(
    hrs_supply = list(
      label   = "M2: Core + Hours water supply/week",
      group   = "Water supply",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + HRS_PER_WEEK
    ),
    cont_int = list(
      label   = "M3: Core + Continuous/intermittent supply",
      group   = "Water supply",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + CONT_INT
    ),
    ws_loc = list(
      label   = "M4: Core + WS/WI neighborhood",
      group   = "Water supply",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + WS_LOC
    ),
    nb_comp = list(
      label   = "M5: Core + Neighborhood comparison (Q28)",
      group   = "Perception",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + NB_COMP
    ),
    time_water = list(
      label   = "M6: Core + Time handling water (Q19)",
      group   = "Water behavior",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + TIME_WATER
    )
  )
)

cat("Total models registered:", length(model_specs), "\n")
```

---

# Fit Models

```{r fit-models}
fitted_models <- lapply(model_specs, function(spec) {
  glm(spec$formula, data = df, family = binomial, na.action = na.omit)
})

# Sample sizes and basic convergence check
tibble(
  model_id = names(model_specs),
  label    = map_chr(model_specs, "label"),
  n        = map_int(fitted_models, nobs),
  converged = map_lgl(fitted_models, ~ .x$converged)
) %>% print(n = Inf)
```

---

# Per-Model Coefficient Tables (Odds Ratios)

```{r results-tables, results="asis"}
# McFadden pseudo-R²
mcfadden_r2 <- function(mod) {
  round(1 - mod$deviance / mod$null.deviance, 3)
}

core_label_list <- list(
  D_AGE        ~ "Age (years)",
  SES_SC_Total ~ "SES composite score",
  D_HH_SIZE    ~ "Household size",
  PSS_TOTAL    ~ "PSS Total (perceived stress)"
)

for (nm in names(fitted_models)) {
  spec <- model_specs[[nm]]
  mod  <- fitted_models[[nm]]
  fit  <- glance(mod)

  cat("\n\n## ", spec$label, " {.unnumbered}\n\n", sep = "")
  cat("**N =**", nobs(mod),
      "| **McFadden R² =**", mcfadden_r2(mod),
      "| **AIC =**", round(fit$AIC, 1),
      "| **BIC =**", round(fit$BIC, 1), "\n\n")

  print(
    tbl_regression(
      mod,
      exponentiate = TRUE,          # report as OR
      label        = core_label_list,
      conf.int     = TRUE
    ) %>%
      bold_p(t = 0.05) %>%
      italicize_levels() %>%
      modify_caption(paste("**", spec$label, "**  (OR, 95% CI)"))
  )
}
```

---

# Model Comparison

Each augmented model is compared to M0 re-fit on the same complete cases using
a **likelihood ratio test (LRT)**. The LRT chi-square statistic tests whether
the augmenting predictor(s) significantly improve model fit.

```{r model-comparison}
# McFadden R² = 1 - (residual deviance / null deviance)
# Using stored deviances avoids refitting the null model in a new environment
mcfadden_r2 <- function(mod) {
  round(1 - mod$deviance / mod$null.deviance, 3)
}

comparison <- map_dfr(names(fitted_models), function(nm) {
  mod   <- fitted_models[[nm]]
  spec  <- model_specs[[nm]]
  fit   <- glance(mod)
  n_mod <- nobs(mod)

  base_row <- tibble(
    model_id    = nm,
    label       = spec$label,
    group       = spec$group,
    n           = n_mod,
    mcfadden_r2 = mcfadden_r2(mod),
    delta_r2    = NA_real_,
    AIC         = round(fit$AIC, 1),
    BIC         = round(fit$BIC, 1),
    LRT_chi2    = NA_real_,
    LRT_df      = NA_integer_,
    LRT_p       = NA_real_,
    note        = ""
  )

  if (nm == "core") return(base_row)

  # Refit M0 on the same rows used by the augmented model
  rows_used  <- as.integer(rownames(model.frame(mod)))
  m0_sub     <- glm(core_formula, data = df[rows_used, ],
                    family = binomial, na.action = na.omit)
  lrt_result <- lrtest(m0_sub, mod)

  base_row %>% mutate(
    delta_r2  = round(mcfadden_r2(mod) - mcfadden_r2(m0_sub), 3),
    LRT_chi2  = round(lrt_result$Chisq[2], 2),
    LRT_df    = as.integer(lrt_result$Df[2]),
    LRT_p     = round(lrt_result$`Pr(>Chisq)`[2], 4),
    note      = if_else(n_mod != nobs(fitted_models[["core"]]),
                        "N differs from full M0", "")
  )
})

comparison %>%
  select(label, group, n, mcfadden_r2, delta_r2, AIC, BIC,
         LRT_chi2, LRT_df, LRT_p, note) %>%
  flextable() %>%
  set_header_labels(
    label       = "Model",
    group       = "Group",
    n           = "N",
    mcfadden_r2 = "McFadden R²",
    delta_r2    = "Δ McFadden R²",
    AIC         = "AIC",
    BIC         = "BIC",
    LRT_chi2    = "LRT χ²",
    LRT_df      = "df",
    LRT_p       = "p (LRT)",
    note        = "Note"
  ) %>%
  color(~ !is.na(LRT_p) & LRT_p < 0.05, j = "LRT_p", color = "darkgreen") %>%
  color(~ !is.na(LRT_p) & LRT_p >= 0.05, j = "LRT_p", color = "grey50") %>%
  autofit()
```

## On interactions

```{r interaction-season-slope}
# Simple slope: HWISE → EMO_NEG within Season 2 only
summary(glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + HW_TOTAL,
            data = filter(df, SEASON == "Season 2"),
            family = binomial))
```

```{r interaction-f2}
# McFadden pseudo-R² increment for the SEASON × HW_TOTAL interaction
r2_m1b <- mcfadden_r2(fitted_models[["core2_hwise_total"]])
r2_m1c <- mcfadden_r2(fitted_models[["core3_hwise_total"]])
cat(sprintf(
  "McFadden R² M1b (Core2 + HWISE):           %.4f\n
   McFadden R² M1c (Core2 + HWISE × Season):  %.4f\n
   ΔMcFadden R² (interaction added):           %.4f\n",
  r2_m1b, r2_m1c, r2_m1c - r2_m1b
))
```

---

### Individual HWISE Items – Summary

```{r hwise-item-comparison}
comparison %>%
  filter(group == "Individual HWISE items") %>%
  select(label, n, mcfadden_r2, delta_r2, LRT_chi2, LRT_p) %>%
  arrange(LRT_p) %>%
  flextable() %>%
  set_header_labels(
    label       = "Item model",
    n           = "N",
    mcfadden_r2 = "McFadden R²",
    delta_r2    = "Δ McFadden R²",
    LRT_chi2    = "LRT χ²",
    LRT_p       = "p (LRT)"
  ) %>%
  color(~ !is.na(LRT_p) & LRT_p < 0.05, j = "LRT_p", color = "darkgreen") %>%
  autofit()
```

---

# Odds Ratio Plot (Augmenting Predictors)

```{r or-plot, fig.width = 10, fig.height = 7}
core_terms <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE", "PSS_TOTAL")

or_data <- map_dfr(names(fitted_models), function(nm) {
  tidy(fitted_models[[nm]], exponentiate = TRUE, conf.int = TRUE) %>%
    filter(!term %in% core_terms) %>%
    mutate(
      model_id = nm,
      label    = model_specs[[nm]]$label,
      group    = model_specs[[nm]]$group
    )
})

if (nrow(or_data) > 0) {
  ggplot(or_data,
         aes(x = estimate, y = reorder(label, estimate),
             xmin = conf.low, xmax = conf.high, color = group)) +
    geom_vline(xintercept = 1, linetype = "dashed", color = "grey40") +
    geom_errorbarh(height = 0.2, linewidth = 0.6) +
    geom_point(size = 2.5) +
    scale_x_log10() +
    scale_color_brewer(palette = "Dark2") +
    labs(
      title    = "Odds ratios for augmenting predictors — Negative emotion as outcome",
      subtitle = "Core variables (Age, SES, HH size, PSS) omitted; bars = 95% CI; x-axis log-scaled",
      x        = "Odds Ratio (log scale)",
      y        = NULL,
      color    = "Predictor group"
    ) +
    theme_bw(base_size = 11) +
    theme(legend.position = "bottom")
}
```

```{r hwise-or-plot, fig.width = 8, fig.height = 6}
# Readable labels for HWISE items
hwise_labels <- c(
  HW_WORRY   = "Worry about water",
  HW_INTERR  = "Interrupted supply",
  HW_CLOTHES = "Couldn't wash clothes",
  HW_PLANS   = "Plans disrupted",
  HW_FOOD    = "Food preparation affected",
  HW_HANDS   = "Couldn't wash hands",
  HW_BODY    = "Couldn't bathe",
  HW_DRINK   = "Not enough to drink",
  HW_ANGRY   = "Felt angry",
  HW_SLEEP   = "Sleep disrupted",
  HW_NONE    = "No water at all",
  HW_SHAME   = "Felt shame"
)

hwise_ids <- c("hwise_total", paste0("hwise_", tolower(hwise_items)))

hwise_or <- map_dfr(hwise_ids, function(nm) {
  tidy(fitted_models[[nm]], exponentiate = TRUE, conf.int = TRUE) %>%
    filter(!term %in% c("(Intercept)", "D_AGE", "SES_SC_Total",
                        "D_HH_SIZE", "PSS_TOTAL")) %>%
    mutate(
      model_id   = nm,
      item_label = if_else(
        term == "HW_TOTAL",
        "HWISE Total Score",
        hwise_labels[term]
      ),
      significant = p.value < 0.05
    )
})

ggplot(hwise_or,
       aes(x = estimate, y = reorder(item_label, estimate),
           xmin = conf.low, xmax = conf.high,
           color = significant, shape = (term == "HW_TOTAL"))) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey40") +
  geom_errorbarh(height = 0.25, linewidth = 0.7) +
  geom_point(size = 3) +
  scale_x_log10() +
  scale_color_manual(
    values = c("TRUE" = "#1a7abf", "FALSE" = "grey60"),
    labels = c("TRUE" = "p < 0.05", "FALSE" = "p ≥ 0.05")
  ) +
  scale_shape_manual(
    values = c("TRUE" = 18, "FALSE" = 16),
    labels = c("TRUE" = "Total score", "FALSE" = "Individual item"),
    guide  = guide_legend(title = NULL)
  ) +
  labs(
    title    = "HWISE items as predictors of negative emotional response",
    subtitle = "Each item in a separate model; core variables (Age, SES, HH size, PSS) controlled; bars = 95% CI",
    x        = "Odds Ratio (log scale)",
    y        = NULL,
    color    = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom")
```

---

# Interpreting the Results

## Reading logistic regression output

This is a **binary logistic regression**. The outcome is the odds of reporting
a **negative** (vs. positive) emotional response to household water.

**Odds Ratio (OR)**:

- OR = 1.0 → no difference from reference / one-unit change has no effect
- OR > 1.0 → higher odds of negative emotion
- OR < 1.0 → lower odds of negative emotion (more likely positive)

**Example**: if OR for HW_TOTAL = 1.12, each one-point increase in HWISE score
is associated with a 12% higher odds of negative emotion, *adjusting for Age,
SES, Household size, and PSS*.

**95% CI**: if the interval excludes 1.0, the association is significant at α = 0.05.

## The role of PSS_TOTAL in the core model

PSS_TOTAL is a covariate, not a mediator in this specification. Including it
tests: *does water insecurity predict emotional responses independently of how
stressed someone already feels?*

- If a water variable is significant WITH PSS in the model, water has an effect
  that is not fully explained by general stress.
- If a water variable drops to non-significance when PSS is added (compared to
  a model without PSS), the effect may be mediated by stress. Formal mediation
  analysis would be needed to confirm this (e.g., `mediation` package).

## HWISE Total vs. Individual Items

Same logic as in `06.1.pss_models.Rmd` applies:

- HW_TOTAL tests the overall insecurity–emotion link.
- Individual items reveal *which* water insecurity experiences (physical vs.
  psychological vs. social) most strongly predict negative affect.
- Items related to **shame** (HW_SHAME), **anger** (HW_ANGRY), and **worry**
  (HW_WORRY) are the most psychologically proximal to emotional responses and
  may show the strongest associations.

## McFadden R²

McFadden's pseudo-R² is not directly comparable to OLS R². General benchmarks:
0.2–0.4 is considered a good fit; values < 0.1 are weak. Focus on **Δ McFadden
R²** (improvement over M0) and the LRT p-value rather than absolute values.

## Categorical predictors

For CONT_INT, WS_LOC, NB_COMP, TIME_WATER: the LRT p-value tests the overall
variable. Inspect specific level ORs in the coefficient table. Reference levels:
Continuous supply, "Better" neighborhood comparison, "No" time investment.

---

# Sensitivity Analysis: Season-Stratified Models

Logistic regression models are re-run separately for each season. This checks
whether the water–emotion association holds across both data collection periods.
Edit `strat_ids` to add or remove models from the stratified analysis.

```{r season-strat, results="asis"}
# ── Models to stratify ─────────────────────────────────────────────────────────
strat_ids <- names(model_specs)  # all registered models

seasons <- levels(df$SEASON)
cat("Seasons in data:", paste(seasons, collapse = ", "), "\n")
cat("N per season:\n")
print(table(df$SEASON, useNA = "always"))

# ── Fit stratified models ──────────────────────────────────────────────────────
strat_models <- lapply(seasons, function(s) {
  df_s <- filter(df, SEASON == s)
  lapply(model_specs[strat_ids], function(spec) {
    # Drop SEASON (main effect + any interactions) using R formula tools —
    # robust to function calls (e.g. log1p()) that break regex approaches
    tt          <- terms(spec$formula)
    all_terms   <- attr(tt, "term.labels")
    keep_terms  <- all_terms[!grepl("SEASON", all_terms)]
    response    <- as.character(attr(tt, "variables")[[2]])
    f_clean     <- reformulate(keep_terms, response = response)
    glm(f_clean, data = df_s, family = binomial, na.action = na.omit)
  })
})
names(strat_models) <- seasons

mcfadden_r2 <- function(mod) round(1 - mod$deviance / mod$null.deviance, 3)

# ── Comparison table: focal OR per season ─────────────────────────────────────
core_terms <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE", "PSS_TOTAL")

strat_comparison <- map_dfr(strat_ids, function(nm) {
  map_dfr(seasons, function(s) {
    mod <- strat_models[[s]][[nm]]
    if (is.null(mod)) return(NULL)
    coefs <- tidy(mod, exponentiate = TRUE, conf.int = TRUE) %>%
      filter(!term %in% core_terms)
    if (nrow(coefs) == 0) {
      coefs <- tibble(term = "—", estimate = NA_real_,
                      conf.low = NA_real_, conf.high = NA_real_, p.value = NA_real_)
    }
    coefs %>%
      mutate(
        model_id    = nm,
        label       = model_specs[[nm]]$label,
        season      = s,
        n           = nobs(mod),
        mcfadden_r2 = mcfadden_r2(mod)
      )
  })
})

strat_wide <- strat_comparison %>%
  mutate(
    or_fmt = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_fmt  = case_when(
      is.na(p.value)  ~ "—",
      p.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", p.value)
    )
  ) %>%
  select(label, term, season, n, mcfadden_r2, or_fmt, p_fmt) %>%
  pivot_wider(
    names_from  = season,
    values_from = c(n, mcfadden_r2, or_fmt, p_fmt),
    names_glue  = "{season}_{.value}"
  )

strat_wide %>%
  flextable() %>%
  set_header_labels(label = "Model", term = "Predictor") %>%
  autofit() %>%
  print()
```

```{r season-or-plot, fig.width = 10, fig.height = 6}
strat_comparison %>%
  filter(!is.na(estimate)) %>%
  ggplot(aes(x = estimate, y = reorder(label, estimate),
             xmin = conf.low, xmax = conf.high, color = season)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey40") +
  geom_errorbarh(height = 0.2, linewidth = 0.6,
                 position = position_dodgev(height = 0.5)) +
  geom_point(size = 2.5, position = position_dodgev(height = 0.5)) +
  scale_x_log10() +
  scale_color_manual(values = c("Season 1" = "#2166ac", "Season 2" = "#d6604d")) +
  labs(
    title    = "Season-stratified odds ratios — Negative emotion as outcome",
    subtitle = "Core variables omitted; bars = 95% CI; x-axis log-scaled",
    x        = "Odds Ratio (log scale)",
    y        = NULL,
    color    = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom")
```

```{r combined-or-plot, fig.width = 11, fig.height = 11}
# ── Terms to exclude from all sources ─────────────────────────────────────────
exclude_terms <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE",
                   "PSS_TOTAL", "SEASONSeason 2",
                   "SEASONSeason 2:HW_TOTAL", "HW_TOTAL:SEASONSeason 2")

# ── Overall (pooled) ORs ───────────────────────────────────────────────────────
overall_or <- map_dfr(strat_ids, function(nm) {
  tidy(fitted_models[[nm]], conf.int = TRUE, exponentiate = TRUE) %>%
    filter(!term %in% exclude_terms) %>%
    mutate(
      source   = "Overall",
      model_id = nm,
      label    = model_specs[[nm]]$label
    )
})

# ── Season-stratified ORs ──────────────────────────────────────────────────────
strat_or <- map_dfr(strat_ids, function(nm) {
  map_dfr(seasons, function(s) {
    mod <- strat_models[[s]][[nm]]
    if (is.null(mod)) return(NULL)
    tidy(mod, conf.int = TRUE, exponentiate = TRUE) %>%
      filter(!term %in% exclude_terms) %>%
      mutate(
        source   = s,
        model_id = nm,
        label    = model_specs[[nm]]$label
      )
  })
})

# ── Combine and plot ───────────────────────────────────────────────────────────
combined_or <- bind_rows(overall_or, strat_or) %>%
  filter(!is.na(estimate)) %>%
  mutate(source = factor(source, levels = c("Overall", seasons)))

ggplot(combined_or,
       aes(x = estimate, y = reorder(label, estimate),
           xmin = conf.low, xmax = conf.high,
           color = source)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey40") +
  geom_errorbarh(height = 0.25, linewidth = 0.6,
                 position = position_dodgev(height = 0.7)) +
  geom_point(size = 2.5, position = position_dodgev(height = 0.7)) +
  scale_x_log10() +
  scale_color_manual(
    values = c("Overall"  = "#e87fae",
               "Season 1" = "#2166ac",
               "Season 2" = "#d6604d"),
    name = NULL
  ) +
  labs(
    title    = "Overall and season-stratified odds ratios — Negative emotion as outcome",
    subtitle = "Core variables (Age, SES, HH size, PSS) and SEASON main effect omitted; bars = 95% CI; x-axis log-scaled",
    x        = "Odds Ratio (log scale)",
    y        = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom")
```

---

# Strengthening the Analysis

1. **Separation check**: with a binary outcome and small N, some predictor–outcome
   combinations may exhibit complete or quasi-complete separation (a predictor
   perfectly predicts the outcome in one direction), causing unstable estimates.
   Check with:
   ```r
   # install.packages("brglm2")
   library(brglm2)
   glm(formula, data = df, family = binomial, method = "brglmFit")
   ```

2. **Check influential observations**: Logistic regression is sensitive to
   outliers. Use Cook's distance and leverage:
   ```r
   plot(fitted_models[["hwise_total"]], which = 4)  # Cook's D
   # or: performance::check_model(fitted_models[["hwise_total"]])
   
   plot(fitted_models[["core3_hwise_total"]], which = 4)
   
   ```

3. **Standardize continuous predictors** for comparability across ORs:
   ```r
   df <- df %>% mutate(
     across(c(D_AGE, SES_SC_Total, D_HH_SIZE, PSS_TOTAL, HW_TOTAL),
            ~ scale(.)[,1], .names = "{.col}_z")
   )
   ```
   Standardized ORs can be compared in magnitude across predictors measured on
   different scales.

4. **Multiple testing correction**: Apply BH correction to LRT p-values:
   ```r
   comparison %>%
     filter(!is.na(LRT_p)) %>%
     mutate(LRT_p_adj = p.adjust(LRT_p, method = "BH"))
   ```

5. **Mediation analysis**: If PSS_TOTAL attenuates the water–emotion association,
   formally test whether PSS mediates the relationship:
   ```r
   # install.packages("mediation")
   library(mediation)
   ```

6. **Missing outcome data**: Q26 (EMO_NEG) may have substantial missingness,
   especially for "positive" emotions being coded 0. Verify that missingness is
   not systematically related to water insecurity (i.e., test whether HWISE
   predicts missingness in EMO_NEG using a logistic regression on the missingness
   indicator).

7. **Additional covariates** to consider adding to the core model:
   - `HLTH_CPAIN_CAT` / `HLTH_CDIS_CAT` (chronic illness — affects baseline
     negative affect)
   - `D_EMP_FORM_CLEAN` (employment precarity)
   - `SEASON` (emotions may differ seasonally with water availability)
   - `D_CHLD_CLEAN` (children in household increase water management burden)

8. **Explore the outcome in more detail**: EMO_NEG collapses specific emotions
   (fear, sadness, anger, etc.) into a binary valence. Consider whether specific
   negative emotion types show different patterns with water predictors.

---

# Limitations

1. **Cross-sectional design**: Temporal precedence cannot be established.
   Negative emotions could lead to more negative water perceptions rather than
   the reverse.

2. **Binary outcome compression**: Collapsing all negative emotions (anger,
   sadness, fear, shame) into one binary variable loses important information
   about the type and intensity of emotional responses.

3. **Common method bias**: Both EMO_NEG and all water predictors are self-reported
   in the same interview context, inflating associations through shared variance.

4. **Listwise deletion and varying N**: Each model excludes participants missing
   any included variable. EMO_NEG itself may be missing for a non-trivial
   proportion, and missing emotion data may not be random.

5. **Multiple testing**: ~18 models are tested on the same outcome. Familywise
   error rate is elevated. Findings are exploratory.

6. **PSS_TOTAL as covariate vs. mediator**: Including PSS_TOTAL as a covariate
   does not fully disentangle the stress–emotion pathway. Mediation analysis
   is required to assess indirect effects.

7. **Rare outcome concern**: If the proportion of negative emotions is very high
   or very low (< 10% or > 90%), logistic regression can produce biased estimates.
   Report the outcome prevalence and consider Firth's penalized likelihood if
   needed.

8. **HWISE item intercorrelations**: See `06.1.pss_models.Rmd`. Enter items
   individually, not jointly, without a specific dimensionality hypothesis.

9. **HWISE response scale treated as interval**: Individual HWISE items use
   4-point response options (0 = never, 1 = rarely, 2 = sometimes,
   3 = often/always) and are entered as numeric predictors, assuming equal
   spacing between categories. This is consistent with prior HWISE validation
   literature and standard practice for Likert-type items in regression, but
   the interval assumption is unverified. Sensitivity analyses using dummy-coded
   categories could assess whether results are robust to this assumption.

10. **Unmeasured confounders**: Trait negative affect, neuroticism, or prior
    experiences with water scarcity could confound the association between
    current water conditions and emotional responses.

11. **Generalizability**: Sample is specific to urban colonias in one city.
    Emotional responses to water insecurity may differ across cultural, climatic,
    and infrastructural contexts.

12. **Seasonal variation**: Emotion valence about water may vary with actual
    water availability (dry vs. wet season). Include `SEASON` as a covariate
    or stratify in confirmatory analyses.
