---
title: "06. Unadjusted Regression Models"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(broom)
library(ggplot2)
library(gt)
library(flextable)
library(officer)

output_dir <- "./output/tables"
fig_dir    <- "./output/figures"
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(fig_dir,    recursive = TRUE, showWarnings = FALSE)

# Group labels for predictor grouping in tables / plots
group_map <- c(
  Age = "Demographics", Parity = "Demographics",
  HH_size = "Demographics", Yrs_in_nbhd = "Demographics",
  SES_Score = "Socioeconomic", Edu_Score = "Socioeconomic",
  Employed = "Socioeconomic", Employment_type = "Socioeconomic",
  Chronic_pain = "Health", Chronic_dis = "Health",
  PSS_Total = "Health",
  HWISE_Total = "HWISE", Water_5sub = "HWISE",
  HWISE_Hi = "HWISE", HWISE_Cat = "HWISE", Water_worry = "HWISE",
  HW_Worry = "HWISE items", HW_Interr = "HWISE items",
  HW_Clothes = "HWISE items", HW_Plans = "HWISE items",
  HW_Food = "HWISE items", HW_Hands = "HWISE items",
  HW_Body = "HWISE items", HW_Drink = "HWISE items",
  HW_Angry = "HWISE items", HW_Sleep = "HWISE items",
  HW_None = "HWISE items", HW_Shame = "HWISE items",
  Season = "Water supply",
  W_WS_LOC_F = "Water supply", W_Continuity = "Water supply",
  W_Frequency = "Water supply", W_Days_week = "Water supply",
  W_Hrs_week = "Water supply", W_Pressure = "Water supply"
)

group_order <- c("Demographics", "Socioeconomic", "Health",
                 "HWISE", "HWISE items", "Water supply")

# Significance symbols
sig_sym <- function(p) {
  case_when(p < 0.001 ~ "***", p < 0.01 ~ "**", p < 0.05 ~ "*",
            p < 0.1 ~ "\u2020", TRUE ~ "")
}

# Publication theme
pub_theme <- theme_classic(base_size = 11) +
  theme(
    plot.title      = element_text(face = "bold", size = 12),
    axis.text.y     = element_text(size = 9),
    axis.text.x     = element_text(size = 9),
    legend.position = "none",
    strip.text      = element_text(face = "bold", size = 9),
    strip.background = element_rect(fill = "grey92", colour = NA),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    plot.caption       = element_text(size = 9, colour = "grey40", hjust = 0)
  )

MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

clean_for_table <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES | is.na(x_num), NA, x_num)
}

SENTINEL_STRINGS <- as.character(c(999, 888, 777, 666, 555, 444, 333))
is_sentinel_str <- function(x) {
  str_trim(as.character(x)) %in% SENTINEL_STRINGS
}
```

---

# Load Data

```{r load-data}
df <- read.csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
               stringsAsFactors = FALSE,
               na.strings = c(""))

cat("Dimensions:", dim(df), "\n")
cat("Participants:", nrow(df), "\n")
```

---

# Prepare Analytic Dataset

```{r prepare}
edu_levels <- c(
  "No formal education",
  "Incomplete primary (1–5 yrs)",
  "Complete primary",
  "Incomplete secondary (7–8 yrs)",
  "Complete secondary",
  "Incomplete high school (10–11 yrs)",
  "Complete high school",
  "Incomplete university (13–15 yrs)",
  "Complete university",
  "Postgraduate"
)

df_table <- df %>%
  mutate(
    Age          = clean_for_table(D_AGE),
    Yrs_in_nbhd  = clean_for_table(D_LOC_TIME_CLEAN),
    HH_size      = clean_for_table(D_HH_SIZE),
    Parity       = clean_for_table(D_CHLD_CLEAN),

    Season = factor(SEASON, levels = c(0, 1), labels = c("Rainy", "Dry")),

    Education = factor(
      dplyr::recode(as.character(SES_EDU_CAT_CLEAN),
        "Incomplete primary"     = "Incomplete primary (1–5 yrs)",
        "Incomplete secondary"   = "Incomplete secondary (7–8 yrs)",
        "Incomplete high school" = "Incomplete high school (10–11 yrs)",
        "Incomplete university"  = "Incomplete university (13–15 yrs)"
      ),
      levels = edu_levels, ordered = TRUE
    ),
    SES_Score       = clean_for_table(SES_SC_Total),
    Edu_Score       = as.numeric(Education),

    Employed = factor(clean_for_table(D_EMP_CLEAN),
                      levels = c(0, 1), labels = c("No", "Yes")),
    Employment_type = factor(clean_for_table(D_EMP_FORM_CLEAN),
                             levels = c(0, 1), labels = c("Informal", "Formal")),

    Hormonal_tx  = factor(clean_for_table(HLTH_TTM_CAT),
                          levels = c(0, 1), labels = c("No", "Yes")),
    Medication   = factor(clean_for_table(HLTH_MED_CAT),
                          levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_pain = factor(clean_for_table(HLTH_CPAIN_CAT),
                          levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_dis  = factor(clean_for_table(HLTH_CDIS_CAT),
                          levels = c(0, 1), labels = c("No", "Yes")),

    HWISE_Total  = clean_for_table(HW_TOTAL),
    Water_worry  = clean_for_table(as.numeric(HW_WORRY_SUB)),
    Water_5sub   = clean_for_table(HW_MEX_5SUB),
    HWISE_Cat    = factor(
      case_when(
        is.na(clean_for_table(HW_TOTAL))   ~ NA_character_,
        clean_for_table(HW_TOTAL) <= 2     ~ "Marginal (0\u20132)",
        clean_for_table(HW_TOTAL) <= 11    ~ "Low (3\u201311)",
        clean_for_table(HW_TOTAL) <= 23    ~ "Moderate (12\u201323)",
        TRUE                               ~ "High (\u226524)"
      ),
      levels = c("Marginal (0\u20132)", "Low (3\u201311)",
                 "Moderate (12\u201323)", "High (\u226524)")
    ),
    HWISE_Hi = factor(
      case_when(
        is.na(HWISE_Total) ~ NA_character_,
        HWISE_Total <= 11  ~ "Low (\u226411)",
        TRUE               ~ "High (\u226512)"
      ),
      levels = c("Low (\u226411)", "High (\u226512)")
    ),

    HW_Worry   = clean_for_table(HW_WORRY),
    HW_Interr  = clean_for_table(HW_INTERR),
    HW_Clothes = clean_for_table(HW_CLOTHES),
    HW_Plans   = clean_for_table(HW_PLANS),
    HW_Food    = clean_for_table(HW_FOOD),
    HW_Hands   = clean_for_table(HW_HANDS),
    HW_Body    = clean_for_table(HW_BODY),
    HW_Drink   = clean_for_table(HW_DRINK),
    HW_Angry   = clean_for_table(HW_ANGRY),
    HW_Sleep   = clean_for_table(HW_SLEEP),
    HW_None    = clean_for_table(HW_NONE),
    HW_Shame   = clean_for_table(HW_SHAME),

    PSS_Total  = clean_for_table(PSS_TOTAL)
  )
```

---

# Prepare Water Supply Variables

```{r water-vars}
df_water <- df %>%
  mutate(
    W_Continuity = case_when(
      is_sentinel_str(MX3_CONT_INT)                             ~ NA_character_,
      str_detect(tolower(MX3_CONT_INT), "^contin")              ~ "Continuous",
      str_detect(tolower(MX3_CONT_INT), "^intermi|cortan|deja") ~ "Intermittent",
      str_detect(tolower(MX3_CONT_INT), "disminu")              ~ "Reduced pressure/flow",
      MX3_CONT_INT %in% c("", "N/A") | is.na(MX3_CONT_INT)    ~ NA_character_,
      TRUE                                                       ~ "Other"
    ),
    W_Continuity = factor(W_Continuity,
                          levels = c("Continuous", "Intermittent",
                                     "Reduced pressure/flow", "Other")),

    W_Freq_raw = coalesce(MX4_WFREQ, MX4_W_FREQ),
    W_Frequency = case_when(
      is_sentinel_str(W_Freq_raw)                              ~ NA_character_,
      str_detect(tolower(W_Freq_raw), "^diari|every day")      ~ "Daily",
      str_detect(tolower(W_Freq_raw), "^intermi")              ~ "Intermittent",
      str_detect(tolower(W_Freq_raw), "1 a 3|1-3")            ~ "1–3 days/week",
      str_detect(tolower(W_Freq_raw), "4 a 6|4-6")            ~ "4–6 days/week",
      str_detect(tolower(W_Freq_raw), "^contin")               ~ "Continuous",
      W_Freq_raw %in% c("", "N/A") | is.na(W_Freq_raw)       ~ NA_character_,
      TRUE                                                      ~ "Other"
    ),
    W_Frequency = factor(W_Frequency,
                         levels = c("Continuous", "Daily", "4–6 days/week",
                                    "1–3 days/week", "Intermittent", "Other")),

    W_Hrs_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_HRS_PER_WEEK), NA,
             as.numeric(str_extract(MX5_WFREQ_HRS_PER_WEEK, "^[0-9\\.]+")))
    ),
    W_Hrs_week = ifelse(W_Hrs_week %in% MISSING_CODES | W_Hrs_week > 168, NA, W_Hrs_week),

    W_Days_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_NDAYS), NA,
             as.numeric(str_extract(MX5_WFREQ_NDAYS, "^[0-9]+")))
    ),
    W_Days_week = ifelse(W_Days_week %in% MISSING_CODES | W_Days_week > 7, NA, W_Days_week),

    W_Pressure = case_when(
      is_sentinel_str(MX6_WPRESS)                                ~ NA_character_,
      str_detect(tolower(MX6_WPRESS), "^s[ií]|^yes|^buena|^bien") ~ "Yes",
      str_detect(tolower(MX6_WPRESS), "^no|^baja|^poca")         ~ "No/Low",
      MX6_WPRESS %in% c("", "N/A") | is.na(MX6_WPRESS)         ~ NA_character_,
      TRUE                                                         ~ "Variable"
    ),
    W_Pressure = factor(W_Pressure, levels = c("Yes", "No", "Variable")),

    W_WS_LOC_F = factor(W_WS_LOC, levels = c("WS", "WI"),
                        labels = c("Water Secure", "Water Insecure"))
  )
```

---

# Prepare Regression Dataset

```{r uvreg-prepare}
EMO_raw <- suppressWarnings(as.numeric(as.character(df$MX26_EM_HHW_CAT)))
EMO_raw <- ifelse(EMO_raw %in% MISSING_CODES, NA, EMO_raw)

df_reg <- df_table %>%
  mutate(
    EMO_NEG = case_when(
      EMO_raw == 0 ~ 0L,
      EMO_raw == 1 ~ 1L,
      TRUE         ~ NA_integer_
    ),
    W_WS_LOC_F  = df_water$W_WS_LOC_F,
    W_Continuity = df_water$W_Continuity,
    W_Frequency  = df_water$W_Frequency,
    W_Days_week  = df_water$W_Days_week,
    W_Hrs_week   = df_water$W_Hrs_week,
    W_Pressure   = df_water$W_Pressure
  )

# Desired within-group display order (lower number = appears first in plot/table)
predictor_order <- c(
  # Demographics
  Age = 1, Parity = 2, HH_size = 3, Yrs_in_nbhd = 4,
  # Socioeconomic
  SES_Score = 1, Edu_Score = 2, Employed = 3, Employment_type = 4,
  # Health
  Chronic_pain = 1, Chronic_dis = 2, PSS_Total = 3,
  # HWISE (Water_worry moved to end)
  HWISE_Total = 1, Water_5sub = 2, HWISE_Hi = 3, HWISE_Cat = 4, Water_worry = 5,
  # HWISE items
  HW_Worry = 1, HW_Interr = 2, HW_Clothes = 3, HW_Plans = 4,
  HW_Food = 5, HW_Hands = 6, HW_Body = 7, HW_Drink = 8,
  HW_Angry = 9, HW_Sleep = 10, HW_None = 11, HW_Shame = 12,
  # Water supply
  Season = 1, W_WS_LOC_F = 2, W_Continuity = 3, W_Frequency = 4,
  W_Days_week = 5, W_Hrs_week = 6, W_Pressure = 7
)

# Shared predictor labels
pred_labels <- c(
  Age             = "Age (years)",
  Parity          = "Parity",
  HH_size         = "Household size",
  Yrs_in_nbhd    = "Years in neighborhood",
  Season          = "Season",
  SES_Score       = "SES composite score",
  Edu_Score       = "Education (ordinal score)",
  Employed        = "Employed",
  Employment_type = "Employment type",
  Chronic_pain    = "Chronic pain",
  Chronic_dis     = "Chronic disease",
  PSS_Total       = "PSS-14 total score",
  HWISE_Total     = "HWISE total score",
  Water_5sub      = "5-item subscale (continuous)",
  HWISE_Hi        = "HWISE binary",
  HWISE_Cat       = "HWISE",
  Water_worry     = "Water worry subscale",
  HW_Worry        = "HW1: Worried about water",
  HW_Interr       = "HW2: Water interrupted",
  HW_Clothes      = "HW3: Unable to wash clothes",
  HW_Plans        = "HW4: Changed plans",
  HW_Food         = "HW5: Unable to prepare food",
  HW_Hands        = "HW6: Unable to wash hands",
  HW_Body         = "HW7: Unable to wash body",
  HW_Drink        = "HW8: No drinking water",
  HW_Angry        = "HW9: Felt angry",
  HW_Sleep        = "HW10: Went to sleep thirsty",
  HW_None         = "HW11: Had no water at all",
  HW_Shame        = "HW12: Felt shame",
  W_WS_LOC_F      = "Neighborhood",
  W_Continuity    = "Supply continuity",
  W_Frequency     = "Supply frequency",
  W_Days_week     = "Days of supply per week",
  W_Hrs_week      = "Hours of supply per week",
  W_Pressure      = "Adequate water pressure"
)

# Reference levels for every factor predictor in df_reg
ref_levels <- sapply(names(df_reg), function(v) {
  x <- df_reg[[v]]
  if (is.factor(x)) levels(x)[1] else NA_character_
})

# Build a display label that appends level & ref for categorical terms
make_display_label <- function(predictor, term, label) {
  if (predictor == term) return(label)          # continuous
  level <- gsub(predictor, "", term, fixed = TRUE)
  ref   <- ref_levels[predictor]
  if (!is.na(ref)) {
    paste0(label, ": ", level, " (ref: ", ref, ")")
  } else {
    paste0(label, ": ", level)
  }
}

# HWISE individual items (shared between short and full predictor lists)
hwise_items <- c(
  "HW_Worry", "HW_Interr", "HW_Clothes", "HW_Plans", "HW_Food",
  "HW_Hands", "HW_Body", "HW_Drink", "HW_Angry", "HW_Sleep",
  "HW_None", "HW_Shame"
)

# Helper: drop sparse levels and run unadjusted linear regression
run_lm_uv <- function(outcome, predictors, data, labels) {
  map_dfr(predictors, function(v) {
    dat <- data[, c(outcome, v)]
    for (drop_lvl in c("Variable", "Other")) {
      if (v %in% c("W_Pressure", "W_Frequency", "W_Continuity") &&
          drop_lvl %in% levels(dat[[v]])) {
        dat <- dat[is.na(dat[[v]]) | dat[[v]] != drop_lvl, ]
        dat[[v]] <- droplevels(dat[[v]])
      }
    }
    dat <- dat[complete.cases(dat), ]
    fit <- tryCatch(
      lm(as.formula(paste(outcome, "~ .")), data = dat),
      error = function(e) NULL
    )
    if (is.null(fit)) return(NULL)
    tidy(fit, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(predictor = v, label = labels[v], AIC = AIC(fit), n = nrow(dat))
  })
}

# Helper: drop sparse levels and run unadjusted logistic regression
run_glm_uv <- function(outcome, predictors, data, labels) {
  map_dfr(predictors, function(v) {
    dat <- data[, c(outcome, v)]
    for (drop_lvl in c("Variable", "Other")) {
      if (v %in% c("W_Pressure", "W_Frequency", "W_Continuity") &&
          drop_lvl %in% levels(dat[[v]])) {
        dat <- dat[is.na(dat[[v]]) | dat[[v]] != drop_lvl, ]
        dat[[v]] <- droplevels(dat[[v]])
      }
    }
    dat <- dat[complete.cases(dat), ]
    fit <- tryCatch(
      glm(as.formula(paste(outcome, "~ .")), data = dat, family = binomial),
      error = function(e) NULL
    )
    if (is.null(fit)) return(NULL)
    tidy(fit, conf.int = TRUE, exponentiate = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(predictor = v, label = labels[v], AIC = AIC(fit), n = nrow(dat))
  })
}
```

---

# Unadjusted Models: PSS Total Score (linear)

β (95% CI), p-value, BH-adjusted q-value, and AIC per model. Sorted by predictor group then variable order.

```{r uvreg-pss}
pss_predictors <- c(
  # Demographics
  "Age", "Parity", "HH_size", "Yrs_in_nbhd",
  # Socioeconomic
  "SES_Score", "Edu_Score", "Employed", "Employment_type",
  # Health (no hormonal_tx, no medication)
  "Chronic_pain", "Chronic_dis",
  # HWISE (Water_worry at end)
  "HWISE_Total", "Water_5sub", "HWISE_Hi", "HWISE_Cat", "Water_worry",
  # Water supply
  "Season", "W_WS_LOC_F", "W_Continuity", "W_Frequency",
  "W_Days_week", "W_Hrs_week", "W_Pressure"
)
pss_predictors_fp <- append(pss_predictors, hwise_items,
                             after = which(pss_predictors == "Water_worry"))
pss_predictors_fp <- pss_predictors_fp[pss_predictors_fp != "HWISE_Cat"]

pss_results <- run_lm_uv("PSS_Total", pss_predictors, df_reg, pred_labels) %>%
  mutate(
    q.value       = p.adjust(p.value, method = "BH"),
    group         = group_map[predictor],
    group         = factor(group, levels = group_order),
    var_order     = predictor_order[predictor],
    display_label = mapply(make_display_label, predictor, term, label)
  ) %>%
  arrange(group, var_order)

pss_tbl <- pss_results %>%
  mutate(
    `β (95% CI)` = sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
    sig          = sig_sym(q.value),
    p_fmt        = case_when(
      p.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", p.value)
    ),
    q_fmt = case_when(
      q.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", q.value)
    ),
    AIC_r = round(AIC, 1)
  ) %>%
  select(display_label, n, `β (95% CI)`, p_fmt, q_fmt, sig, AIC_r, p.value, q.value) %>%
  rename(
    Variable       = display_label,
    N              = n,
    `p-value`      = p_fmt,
    `q-value (BH)` = q_fmt,
    Sig            = sig,
    AIC            = AIC_r
  )
```

## Table: PSS (publication-ready)

```{r pss-table-pub}
pss_sig_rows <- which(pss_tbl$q.value < 0.05)

pss_gt <- pss_tbl %>%
  select(-p.value, -q.value) %>%
  gt() %>%
  tab_header(
    title    = md("**Table Unadj1. Unadjusted Linear Regression: Predictors of Perceived Stress Score**"),
    subtitle = md("Each predictor entered separately as the sole regressor. Grouped by predictor domain.")
  ) %>%
  cols_label(
    Variable       = md("**Predictor**"),
    N              = md("***n***"),
    `β (95% CI)`   = md("**β (95% CI)**"),
    `p-value`      = md("***p***"),
    `q-value (BH)` = md("***q***"),
    Sig            = md("**Sig.**"),
    AIC            = md("**AIC**")
  ) %>%
  # Header background
  tab_style(
    style     = list(cell_fill(color = "#2c3e50"), cell_text(color = "white", weight = "bold")),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style     = list(cell_fill(color = "#2c3e50"), cell_text(color = "white")),
    locations = cells_title()
  ) %>%
  # Significant rows
  tab_style(
    style     = list(cell_fill(color = "#fff3cd"), cell_text(weight = "bold")),
    locations = cells_body(rows = pss_sig_rows)
  ) %>%
  # Centred columns
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_body(columns = c(N, `p-value`, `q-value (BH)`, Sig, AIC))
  ) %>%
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = c(N, `p-value`, `q-value (BH)`, Sig, AIC))
  ) %>%
  cols_width(
    Variable       ~ px(210),
    N              ~ px(50),
    `β (95% CI)`   ~ px(155),
    `p-value`      ~ px(70),
    `q-value (BH)` ~ px(70),
    Sig            ~ px(48),
    AIC            ~ px(68)
  ) %>%
  tab_footnote(
    footnote = md(
      "**Highlighted rows:** *q* < 0.05 (BH-adjusted).
      **Sig.:** \u2020 *q* < 0.10 &nbsp; \\* *q* < 0.05 &nbsp; \\*\\* *q* < 0.01 &nbsp; \\*\\*\\* *q* < 0.001.
      ***p*-value:** unadjusted two-sided Wald test.
      ***q*-value (BH):** *p*-value adjusted for multiple comparisons using the Benjamini–Hochberg false discovery rate procedure across all predictors in this table; values below 0.05 indicate that the finding is expected to remain significant after controlling the false discovery rate at 5%.
      **AIC:** Akaike Information Criterion — lower values indicate better model fit for comparisons within the same outcome."
    )
  ) %>%
  opt_row_striping(row_striping = TRUE) %>%
  opt_table_font(font = list(google_font("Source Sans Pro"), default_fonts())) %>%
  tab_options(
    table.border.top.color          = "#2c3e50",
    table.border.top.width          = px(2),
    table.border.bottom.color       = "#2c3e50",
    table.border.bottom.width       = px(2),
    column_labels.border.bottom.color = "#2c3e50",
    column_labels.border.bottom.width = px(1),
    row.striping.background_color   = "#f7f9fc",
    data_row.padding                = px(5),
    footnotes.font.size             = px(11),
    heading.title.font.size         = px(14),
    heading.subtitle.font.size      = px(11),
    heading.padding                 = px(6)
  )

pss_gt
```

## Coefficient Plot: PSS

```{r coef-plot-pss, fig.width=9, fig.height=8}
plot_df_pss <- pss_results %>%
  mutate(
    label_full = factor(display_label,
                        levels = rev(unique(display_label))),
    sig        = sig_sym(p.value),
    n_label    = paste0("n=", n)
  )

p_pss <- ggplot(plot_df_pss, aes(x = estimate, y = label_full,
                                  colour = q.value < 0.05)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.2) +
  geom_text(aes(label = sig), nudge_y = 0.25, size = 3.5,
            fontface = "bold", show.legend = FALSE) +
  geom_text(aes(x = max(conf.high, na.rm = TRUE) * 1.08,
                label = n_label),
            size = 2.8, colour = "grey40", hjust = 0) +
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  scale_colour_manual(values = c("FALSE" = "grey55", "TRUE" = "#c0392b")) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.18))) +
  scale_y_discrete(expand = expansion(add = 0.4)) +
  labs(
    title   = "Unadjusted Linear Regression — Perceived Stress Score (β)",
    caption = "Red = q < 0.05 (Benjamini–Hochberg false discovery rate adjustment)  |  Symbols: \u2020p<0.10  *p<0.05  **p<0.01  ***p<0.001",
    x       = "β coefficient (95% CI)",
    y       = NULL
  ) +
  pub_theme +
  theme(panel.spacing = unit(0.8, "lines"))

p_pss

ggsave(file.path(fig_dir, "06.Unadj_coef_PSS.png"), p_pss,
       width = 9, height = 8, dpi = 300, bg = "white")
```

---

# Unadjusted Models: Emotion Type (logistic)

OR (95% CI), p-value, BH-adjusted q-value, and AIC per model.
Reference: positive emotion (0). Sorted by predictor group then variable order.

```{r uvreg-emo}
emo_predictors <- c(
  # Demographics
  "Age", "Parity", "HH_size", "Yrs_in_nbhd",
  # Socioeconomic
  "SES_Score", "Edu_Score", "Employed", "Employment_type",
  # Health (no hormonal_tx, no medication)
  "Chronic_pain", "Chronic_dis", "PSS_Total",
  # HWISE (Water_worry at end)
  "HWISE_Total", "Water_5sub", "HWISE_Hi", "HWISE_Cat", "Water_worry",
  # Water supply
  "Season", "W_WS_LOC_F", "W_Continuity", "W_Frequency",
  "W_Days_week", "W_Hrs_week", "W_Pressure"
)
emo_predictors_fp <- append(emo_predictors, hwise_items,
                             after = which(emo_predictors == "Water_worry"))
emo_predictors_fp <- emo_predictors_fp[emo_predictors_fp != "HWISE_Cat"]

emo_results <- run_glm_uv("EMO_NEG", emo_predictors, df_reg, pred_labels) %>%
  mutate(
    q.value       = p.adjust(p.value, method = "BH"),
    group         = group_map[predictor],
    group         = factor(group, levels = group_order),
    var_order     = predictor_order[predictor],
    display_label = mapply(make_display_label, predictor, term, label)
  ) %>%
  arrange(group, var_order)

emo_tbl <- emo_results %>%
  mutate(
    `OR (95% CI)` = sprintf("%.2f (%.2f, %.2f)", estimate, conf.low, conf.high),
    sig           = sig_sym(q.value),
    p_fmt = case_when(
      p.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", p.value)
    ),
    q_fmt = case_when(
      q.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", q.value)
    ),
    AIC_r = round(AIC, 1)
  ) %>%
  select(display_label, n, `OR (95% CI)`, p_fmt, q_fmt, sig, AIC_r, p.value, q.value) %>%
  rename(
    Variable       = display_label,
    N              = n,
    `p-value`      = p_fmt,
    `q-value (BH)` = q_fmt,
    Sig            = sig,
    AIC            = AIC_r
  )
```

## Table: Emotion Type (publication-ready)

```{r emo-table-pub}
emo_sig_rows <- which(emo_tbl$q.value < 0.05)

emo_gt <- emo_tbl %>%
  select(-p.value, -q.value) %>%
  gt() %>%
  tab_header(
    title    = md("**Table Unadj2. Unadjusted Logistic Regression: Predictors of Negative Emotional Response to Water**"),
    subtitle = md("Each predictor entered separately as the sole regressor. Grouped by predictor domain. Reference category: positive emotional response.")
  ) %>%
  cols_label(
    Variable       = md("**Predictor**"),
    N              = md("***n***"),
    `OR (95% CI)`  = md("**OR (95% CI)**"),
    `p-value`      = md("***p***"),
    `q-value (BH)` = md("***q***"),
    Sig            = md("**Sig.**"),
    AIC            = md("**AIC**")
  ) %>%
  tab_style(
    style     = list(cell_fill(color = "#1a3a5c"), cell_text(color = "white", weight = "bold")),
    locations = cells_column_labels()
  ) %>%
  tab_style(
    style     = list(cell_fill(color = "#1a3a5c"), cell_text(color = "white")),
    locations = cells_title()
  ) %>%
  tab_style(
    style     = list(cell_fill(color = "#d6eaf8"), cell_text(weight = "bold")),
    locations = cells_body(rows = emo_sig_rows)
  ) %>%
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_body(columns = c(N, `p-value`, `q-value (BH)`, Sig, AIC))
  ) %>%
  tab_style(
    style     = cell_text(align = "center"),
    locations = cells_column_labels(columns = c(N, `p-value`, `q-value (BH)`, Sig, AIC))
  ) %>%
  cols_width(
    Variable       ~ px(210),
    N              ~ px(50),
    `OR (95% CI)`  ~ px(155),
    `p-value`      ~ px(70),
    `q-value (BH)` ~ px(70),
    Sig            ~ px(48),
    AIC            ~ px(68)
  ) %>%
  tab_footnote(
    footnote = md(
      "**Highlighted rows:** *q* < 0.05 (BH-adjusted).
      **Sig.:** \u2020 *q* < 0.10 &nbsp; \\* *q* < 0.05 &nbsp; \\*\\* *q* < 0.01 &nbsp; \\*\\*\\* *q* < 0.001.
      **OR:** Odds ratio from binary logistic regression (outcome = negative emotion = 1 vs. positive = 0); OR > 1 indicates higher odds of negative response.
      ***p*-value:** unadjusted two-sided Wald test.
      ***q*-value (BH):** *p*-value adjusted for multiple comparisons using the Benjamini–Hochberg false discovery rate procedure across all predictors in this table; values below 0.05 indicate that the finding is expected to remain significant after controlling the false discovery rate at 5%.
      **AIC:** Akaike Information Criterion — lower values indicate better model fit for comparisons within the same outcome."
    )
  ) %>%
  opt_row_striping(row_striping = TRUE) %>%
  opt_table_font(font = list(google_font("Source Sans Pro"), default_fonts())) %>%
  tab_options(
    table.border.top.color          = "#1a3a5c",
    table.border.top.width          = px(2),
    table.border.bottom.color       = "#1a3a5c",
    table.border.bottom.width       = px(2),
    column_labels.border.bottom.color = "#1a3a5c",
    column_labels.border.bottom.width = px(1),
    row.striping.background_color   = "#f4f8fb",
    data_row.padding                = px(5),
    footnotes.font.size             = px(11),
    heading.title.font.size         = px(14),
    heading.subtitle.font.size      = px(11),
    heading.padding                 = px(6)
  )

emo_gt
```

## Coefficient Plot: Emotion Type (OR scale)

```{r coef-plot-emo, fig.width=9, fig.height=8}
plot_df_emo <- emo_results %>%
  mutate(
    label_full = factor(display_label,
                        levels = rev(unique(display_label))),
    sig        = sig_sym(p.value),
    n_label    = paste0("n=", n)
  )

p_emo <- ggplot(plot_df_emo, aes(x = estimate, y = label_full,
                                   colour = q.value < 0.05)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.2) +
  geom_text(aes(label = sig), nudge_y = 0.25, size = 3.5,
            fontface = "bold", show.legend = FALSE) +
  geom_text(aes(x = max(conf.high, na.rm = TRUE) * 1.12,
                label = n_label),
            size = 2.8, colour = "grey40", hjust = 0) +
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  scale_colour_manual(values = c("FALSE" = "grey55", "TRUE" = "#2471a3")) +
  scale_x_log10(expand = expansion(mult = c(0.05, 0.22))) +
  scale_y_discrete(expand = expansion(add = 0.4)) +
  labs(
    title   = stringr::str_wrap("Unadjusted Logistic Regression — Negative Emotional Response (OR)", width = 55),
    caption = "Blue = q < 0.05 (Benjamini–Hochberg false discovery rate adjustment)  |  Log scale  |  Symbols: \u2020p<0.10  *p<0.05  **p<0.01  ***p<0.001",
    x       = "Odds Ratio (95% CI, log scale)",
    y       = NULL
  ) +
  pub_theme +
  theme(panel.spacing = unit(0.8, "lines"))

p_emo

ggsave(file.path(fig_dir, "06.Unadj_coef_EMO.png"), p_emo,
       width = 9, height = 8, dpi = 300, bg = "white")
```

---

# Export to Word

```{r uvreg-export}

make_uv_flextable <- function(tbl, sig_rows, title_str, footnote_str,
                               hi_color = "#fff3cd") {
  tbl %>%
    select(-p.value, -q.value) %>%
    flextable() %>%
    set_header_labels(
      Variable       = "Predictor",
      N              = "n",
      Sig            = "Sig.",
      `p-value`      = "p-value",
      `q-value (BH)` = "q-value (BH)",
      AIC            = "AIC"
    ) %>%
    # Title row above header
    add_header_lines(title_str) %>%
    # Fonts
    font(fontname = "Calibri", part = "all") %>%
    fontsize(size = 10, part = "body") %>%
    fontsize(size = 10, part = "header") %>%
    fontsize(size = 9,  part = "footer") %>%
    # Header styling
    bold(part = "header") %>%
    color(part = "header", color = "white") %>%
    bg(part = "header", bg = "#1A3A5C") %>%
    # Title row styling (first header row = title)
    fontsize(i = 1, part = "header", size = 11) %>%
    bold(i = 1, part = "header") %>%
    # Alternating row shading
    bg(i = seq(2, nrow(tbl %>% select(-p.value)), by = 2),
       bg = "#F9F9F9", part = "body") %>%
    # Significant row highlight
    bold(i = sig_rows) %>%
    bg(i = sig_rows, bg = hi_color) %>%
    # Borders
    border_remove() %>%
    hline_top(part = "header",
              border = fp_border(color = "#1A3A5C", width = 1.5)) %>%
    hline_bottom(part = "header",
                 border = fp_border(color = "#1A3A5C", width = 1.5)) %>%
    hline_bottom(part = "body",
                 border = fp_border(color = "#1A3A5C", width = 1.5)) %>%
    hline(border = fp_border(color = "#bdc3c7", width = 0.5), part = "body") %>%
    # Alignment
    align(j = 1,  align = "left",   part = "all") %>%
    align(j = -1, align = "center", part = "all") %>%
    # Footer
    add_footer_lines(footnote_str) %>%
    italic(part = "footer") %>%
    color(part = "footer", color = "#555555") %>%
    # Layout
    padding(padding.top = 2, padding.bottom = 2, part = "all") %>%
    set_table_properties(layout = "autofit", width = 1) %>%
    fix_border_issues()
}

pss_footnote <- paste0(
  "Highlighted rows: q < 0.05 (BH-adjusted, shown in amber). ",
  "Sig.: \u2020 q<0.10, * q<0.05, ** q<0.01, *** q<0.001 (based on BH-adjusted q-values). ",
  "p-value: unadjusted two-sided Wald test. ",
  "q-value (BH): p-value corrected for multiple comparisons using the ",
  "Benjamini-Hochberg false discovery rate (FDR) procedure across all predictors ",
  "in this table. A q-value < 0.05 indicates the result is expected to remain ",
  "significant while controlling the FDR at 5%. ",
  "AIC: Akaike Information Criterion; lower = better fit."
)

emo_footnote <- paste0(
  "Highlighted rows: q < 0.05 (BH-adjusted, shown in blue). ",
  "OR: odds ratio from binary logistic regression (outcome: negative emotion = 1 ",
  "vs. positive = 0); OR > 1 indicates higher odds of negative emotional response. ",
  "Sig.: \u2020 q<0.10, * q<0.05, ** q<0.01, *** q<0.001 (based on BH-adjusted q-values). ",
  "p-value: unadjusted two-sided Wald test. ",
  "q-value (BH): p-value corrected for multiple comparisons using the ",
  "Benjamini-Hochberg false discovery rate (FDR) procedure across all predictors ",
  "in this table. A q-value < 0.05 indicates the result is expected to remain ",
  "significant while controlling the FDR at 5%. ",
  "AIC: Akaike Information Criterion; lower = better fit."
)

pss_ft <- make_uv_flextable(
  pss_tbl, pss_sig_rows,
  title_str   = "Table Unadj1. Unadjusted Linear Regression: Predictors of Perceived Stress Score",
  footnote_str = pss_footnote,
  hi_color    = "#fff3cd"
)

emo_ft <- make_uv_flextable(
  emo_tbl, emo_sig_rows,
  title_str   = "Table Unadj2. Unadjusted Logistic Regression: Predictors of Negative Emotional Response to Water",
  footnote_str = emo_footnote,
  hi_color    = "#d6eaf8"
)

combined_outpath <- file.path(output_dir, "06.Tables_Unadjusted_Models.docx")
```

---

# Forest Plot Tables (Publication-Ready)

```{r unadj-forest-helpers}
# ── Colour palette ────────────────────────────────────────────────────────────
COL_HEADER  <- "#1A3A5C"
COL_SUBHEAD <- "#2E6DA4"
COL_SIG     <- "#D5F5E3"   # BH-significant rows
COL_NS      <- "#F9F9F9"
COL_GROUP   <- "#EAF2FB"   # group header rows

# ── Helper functions ──────────────────────────────────────────────────────────
p_fmt_uv <- function(p) {
  case_when(is.na(p) ~ "—", p < 0.001 ~ "<0.001", TRUE ~ sprintf("%.3f", p))
}

is_sig_uv <- function(p_str) {
  suppressWarnings(as.numeric(gsub("<0.001", "0.0005", p_str))) < 0.05
}

```


```{r unadj-forest-models}
pss_results_fp <- run_lm_uv("PSS_Total", pss_predictors_fp, df_reg, pred_labels) %>%
  mutate(
    q.value       = p.adjust(p.value, method = "BH"),
    group         = group_map[predictor],
    group         = factor(group, levels = group_order),
    var_order     = predictor_order[predictor],
    display_label = mapply(make_display_label, predictor, term, label),
    est_ci        = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_fmt         = p_fmt_uv(p.value),
    q_fmt         = p_fmt_uv(q.value)
  ) %>%
  arrange(group, var_order)

emo_results_fp <- run_glm_uv("EMO_NEG", emo_predictors_fp, df_reg, pred_labels) %>%
  mutate(
    q.value       = p.adjust(p.value, method = "BH"),
    group         = group_map[predictor],
    group         = factor(group, levels = group_order),
    var_order     = predictor_order[predictor],
    display_label = mapply(make_display_label, predictor, term, label),
    est_ci        = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_fmt         = p_fmt_uv(p.value),
    q_fmt         = p_fmt_uv(q.value)
  ) %>%
  arrange(group, var_order)
```

```{r unadj-forest-table-builder}
insert_group_headers_uv <- function(df, group_col = "Group") {
  groups <- unique(df[[group_col]])
  purrr::map_dfr(groups, function(g) {
    rows       <- df[df[[group_col]] == g, ]
    header_row <- rows[1, ]
    for (col in names(header_row)) header_row[[col]] <- NA
    header_row[[group_col]]   <- g
    header_row[["Predictor"]] <- g
    bind_rows(header_row, rows)
  })
}

build_uv_ft <- function(results, outcome_label, est_col_label,
                         footnote_txt) {
  tbl_data <- results %>%
    transmute(
      Group     = as.character(group),
      Predictor = display_label,
      N         = n,
      est_ci    = est_ci,
      AIC       = round(AIC, 1),
      p_fmt     = p_fmt,
      q_fmt     = q_fmt,
      p_raw     = p.value,
      q_raw     = q.value
    )

  tbl_with_headers <- insert_group_headers_uv(tbl_data) %>%
    select(-Group)

  data <- tbl_with_headers

  grp_rows  <- which(is.na(data$est_ci) | data$est_ci == "")
  sig_rows  <- which(!data$q_fmt %in% c("—", NA) & is_sig_uv(data$q_fmt))
  nsig_rows <- setdiff(seq_len(nrow(data)), sig_rows)
  sig_coef  <- which(!is.na(data$p_raw) & data$p_raw < 0.05)

  ft <- data %>%
    select(Predictor, N, est_ci, AIC, p_fmt, q_fmt) %>%
    flextable() %>%
    set_header_labels(
      Predictor = "Predictor",
      N         = "N",
      est_ci    = est_col_label,
      AIC       = "AIC",
      p_fmt     = "p-value",
      q_fmt     = "q-value (BH)"
    ) %>%
    add_header_lines(outcome_label) %>%
    # Fonts
    font(fontname = "Calibri", part = "all") %>%
    fontsize(size = 10, part = "body") %>%
    fontsize(size = 10, part = "header") %>%
    fontsize(size = 9,  part = "footer") %>%
    fontsize(i = 1, part = "header", size = 11) %>%
    # Header styling (uniform #1A3A5C)
    bg(bg = "#1A3A5C", part = "header") %>%
    color(color = "white", part = "header") %>%
    bold(part = "header") %>%
    # Body row shading
    bg(i = sig_rows,  bg = COL_SIG,   part = "body") %>%
    bg(i = nsig_rows, bg = COL_NS,    part = "body") %>%
    bg(i = grp_rows,  bg = COL_GROUP, part = "body") %>%
    bold(i = grp_rows, part = "body") %>%
    color(i = sig_coef, j = "p_fmt", color = "#1A5276") %>%
    bold( i = sig_coef, j = "p_fmt") %>%
    color(i = sig_rows,  j = "q_fmt", color = "#1E8449") %>%
    bold( i = sig_rows,  j = "q_fmt") %>%
    color(i = nsig_rows, j = "q_fmt", color = "grey55") %>%
    # Borders (horizontal-only)
    border_remove() %>%
    hline_top(part = "header",
              border = fp_border(color = "#1A3A5C", width = 1.5)) %>%
    hline_bottom(part = "header",
                 border = fp_border(color = "#1A3A5C", width = 1.5)) %>%
    hline_bottom(part = "body",
                 border = fp_border(color = "#1A3A5C", width = 1.5)) %>%
    hline(border = fp_border(color = "#bdc3c7", width = 0.5), part = "body") %>%
    # Alignment
    align(j = 1,  align = "left",   part = "all") %>%
    align(j = -1, align = "center", part = "all") %>%
    # Footer
    add_footer_lines(footnote_txt) %>%
    italic(part = "footer") %>%
    color(color = "#555555", part = "footer") %>%
    # Layout
    padding(padding.top = 2, padding.bottom = 2, part = "all") %>%
    set_table_properties(layout = "autofit", width = 1) %>%
    fix_border_issues()

  ft
}
```

```{r unadj-forest-tables-export}
pss_footnote_fp <- paste0(
  "Each predictor entered in a separate unadjusted linear regression model. ",
  "Predictor groups: Demographics (Age, Parity, Household size, Years in neighborhood); ",
  "Socioeconomic (SES composite score, Education, Employed, Employment type); ",
  "Health (Chronic pain, Chronic disease); ",
  "HWISE (total score, 5-item subscale continuous, HWISE high [\u226512 vs. low \u226411], ",
  "HWISE category [marginal/low/moderate/high], Water worry subscale); ",
  "HWISE items (each item separately); Water supply variables. ",
  "\u03b2 = unstandardised regression coefficient; 95% CI = confidence interval. ",
  "AIC = Akaike Information Criterion (lower = better fit). ",
  "p-value: unadjusted two-sided Wald test. ",
  "q-value (BH): p-value adjusted for multiple comparisons using the Benjamini\u2013Hochberg ",
  "false discovery rate procedure across all predictors for this outcome. ",
  "Green rows: BH-adjusted q < 0.05. Bold blue: unadjusted p < 0.05."
)

emo_footnote_fp <- paste0(
  "Each predictor entered in a separate unadjusted logistic regression model. ",
  "Outcome: negative water-related emotion (1) vs. positive (0). ",
  "Predictor groups: Demographics (Age, Parity, Household size, Years in neighborhood); ",
  "Socioeconomic (SES composite score, Education, Employed, Employment type); ",
  "Health (Chronic pain, Chronic disease, PSS total); ",
  "HWISE (total score, 5-item subscale continuous, HWISE high [\u226512 vs. low \u226411], ",
  "HWISE category [marginal/low/moderate/high], Water worry subscale); ",
  "HWISE items (each item separately); Water supply variables. ",
  "OR = odds ratio; 95% CI = confidence interval. OR > 1 = higher odds of negative response. ",
  "AIC = Akaike Information Criterion (lower = better fit). ",
  "p-value: unadjusted two-sided Wald test. ",
  "q-value (BH): p-value adjusted for multiple comparisons using the Benjamini\u2013Hochberg ",
  "false discovery rate procedure across all predictors for this outcome. ",
  "Green rows: BH-adjusted q < 0.05. Bold blue: unadjusted p < 0.05."
)

ft_pss_fp <- build_uv_ft(
  results       = pss_results_fp,
  outcome_label = "Table Unadj1. Unadjusted Linear Regression \u2014 Perceived Stress Score (PSS)",
  est_col_label = "\u03b2 [95% CI]",
  footnote_txt  = pss_footnote_fp
)

ft_emo_fp <- build_uv_ft(
  results       = emo_results_fp,
  outcome_label = "Table Unadj2. Unadjusted Logistic Regression \u2014 Negative Emotional Response to Water",
  est_col_label = "OR [95% CI]",
  footnote_txt  = emo_footnote_fp
)

portrait_sec_fp <- prop_section(
  page_size    = page_size(orient = "portrait"),
  page_margins = page_mar(top = 0.8, bottom = 0.8, left = 0.9, right = 0.9)
)

doc_combined <- read_docx() %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(pss_ft) %>%
  body_add_break(pos = "after") %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(emo_ft) %>%
  body_add_break(pos = "after") %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_pss_fp) %>%
  body_add_break(pos = "after") %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_emo_fp) %>%
  body_set_default_section(portrait_sec_fp)

print(doc_combined, target = combined_outpath)
cat("Saved:", combined_outpath, "\n")
```

---

```{r session-info}
sessionInfo()
```
