---
title: "06z. Unadjusted Forest Plot + Sparklines — Prototype (PSS)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    code_folding: hide
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(broom)
library(ggplot2)

MISSING_CODES   <- c(999, 888, 777, 666, 555, 444, 333)
SENTINEL_STRINGS <- as.character(MISSING_CODES)

clean_for_table <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES | is.na(x_num), NA, x_num)
}
is_sentinel_str <- function(x) str_trim(as.character(x)) %in% SENTINEL_STRINGS

sig_sym <- function(p) {
  case_when(p < 0.001 ~ "***", p < 0.01 ~ "**", p < 0.05 ~ "*",
            p < 0.1 ~ "\u2020", TRUE ~ "")
}

group_map <- c(
  Age = "Demographics", Parity = "Demographics",
  HH_size = "Demographics", Yrs_in_nbhd = "Demographics",
  SES_Score = "Socioeconomic", Edu_Score = "Socioeconomic",
  Employed = "Socioeconomic", Employment_type = "Socioeconomic",
  Chronic_pain = "Health", Chronic_dis = "Health",
  HWISE_Total = "HWISE", Water_5sub = "HWISE",
  HWISE_Hi = "HWISE", Water_worry = "HWISE",
  HW_Worry = "HWISE items", HW_Interr = "HWISE items",
  HW_Clothes = "HWISE items", HW_Plans = "HWISE items",
  HW_Food = "HWISE items", HW_Hands = "HWISE items",
  HW_Body = "HWISE items", HW_Drink = "HWISE items",
  HW_Angry = "HWISE items", HW_Sleep = "HWISE items",
  HW_None = "HWISE items", HW_Shame = "HWISE items",
  Season = "Water supply",
  W_WS_LOC_F = "Water supply", W_Continuity = "Water supply",
  W_Frequency = "Water supply", W_Days_week = "Water supply",
  W_Hrs_week = "Water supply", W_Pressure = "Water supply",
  W_Trust = "Water infrastructure", W_Garrafon_brand = "Water infrastructure",
  W_Has_tinaco = "Water infrastructure", W_Has_cisterna = "Water infrastructure",
  W_Bill_waived = "Water infrastructure", W_Cost_2mnths = "Water infrastructure"
)

group_order <- c("Demographics", "Socioeconomic", "Health",
                 "HWISE", "HWISE items", "Water supply", "Water infrastructure")

predictor_order <- c(
  Age = 1, Parity = 2, HH_size = 3, Yrs_in_nbhd = 4,
  SES_Score = 1, Edu_Score = 2, Employed = 3, Employment_type = 4,
  Chronic_pain = 1, Chronic_dis = 2,
  HWISE_Total = 1, Water_5sub = 2, HWISE_Hi = 3, Water_worry = 5,
  HW_Worry = 1, HW_Interr = 2, HW_Clothes = 3, HW_Plans = 4,
  HW_Food = 5, HW_Hands = 6, HW_Body = 7, HW_Drink = 8,
  HW_Angry = 9, HW_Sleep = 10, HW_None = 11, HW_Shame = 12,
  Season = 1, W_WS_LOC_F = 2, W_Continuity = 3, W_Frequency = 4,
  W_Days_week = 5, W_Hrs_week = 6, W_Pressure = 7,
  W_Trust = 1, W_Garrafon_brand = 2, W_Has_tinaco = 3, W_Has_cisterna = 4,
  W_Bill_waived = 5, W_Cost_2mnths = 6
)

pred_labels <- c(
  Age             = "Age (years)",
  Parity          = "Parity",
  HH_size         = "Household size",
  Yrs_in_nbhd    = "Years in neighborhood",
  Season          = "Season",
  SES_Score       = "SES composite score",
  Edu_Score       = "Education (ordinal score)",
  Employed        = "Employed",
  Employment_type = "Employment type",
  Chronic_pain    = "Chronic pain",
  Chronic_dis     = "Chronic disease",
  HWISE_Total     = "HWISE total score",
  Water_5sub      = "5-item subscale (continuous)",
  HWISE_Hi        = "HWISE binary",
  Water_worry     = "Water worry subscale",
  HW_Worry        = "HW1: Worried about water",
  HW_Interr       = "HW2: Water interrupted",
  HW_Clothes      = "HW3: Unable to wash clothes",
  HW_Plans        = "HW4: Changed plans",
  HW_Food         = "HW5: Unable to prepare food",
  HW_Hands        = "HW6: Unable to wash hands",
  HW_Body         = "HW7: Unable to wash body",
  HW_Drink        = "HW8: No drinking water",
  HW_Angry        = "HW9: Felt angry",
  HW_Sleep        = "HW10: Went to sleep thirsty",
  HW_None         = "HW11: Had no water at all",
  HW_Shame        = "HW12: Felt shame",
  W_WS_LOC_F       = "Neighborhood",
  W_Continuity     = "Supply continuity",
  W_Frequency      = "Supply frequency",
  W_Days_week      = "Days of supply per week",
  W_Hrs_week       = "Hours of supply per week",
  W_Pressure       = "Adequate water pressure",
  W_Trust          = "Trusts tap water quality",
  W_Garrafon_brand = "Uses brand-name garrafón",
  W_Has_tinaco     = "Has tinaco (rooftop tank)",
  W_Has_cisterna   = "Has cisterna (underground tank)",
  W_Bill_waived    = "Water bill waived (condonado)",
  W_Cost_2mnths    = "Bimonthly water cost (pesos)"
)

make_display_label <- function(predictor, term, label) {
  if (predictor == term) return(label)
  level <- gsub(predictor, "", term, fixed = TRUE)
  ref   <- ref_levels[predictor]
  if (!is.na(ref)) paste0(label, ": ", level, " (ref: ", ref, ")")
  else paste0(label, ": ", level)
}

prep_uv_dat <- function(outcome, v, data) {
  dat <- data[, c(outcome, v)]
  for (drop_lvl in c("Variable", "Other")) {
    if (v %in% c("W_Pressure", "W_Frequency", "W_Continuity") &&
        drop_lvl %in% levels(dat[[v]])) {
      dat <- dat[is.na(dat[[v]]) | dat[[v]] != drop_lvl, ]
      dat[[v]] <- droplevels(dat[[v]])
    }
  }
  dat[complete.cases(dat), ]
}

run_lm_uv <- function(outcome, predictors, data, labels) {
  map_dfr(predictors, function(v) {
    dat <- prep_uv_dat(outcome, v, data)
    fit <- tryCatch(lm(as.formula(paste(outcome, "~ .")), data = dat),
                   error = function(e) NULL)
    if (is.null(fit)) return(NULL)
    tidy(fit, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(predictor = v, label = labels[v], AIC = AIC(fit), n = nrow(dat))
  })
}

pub_theme <- theme_classic(base_size = 11) +
  theme(
    plot.title         = element_text(face = "bold", size = 12),
    axis.text.y        = element_text(size = 9),
    axis.text.x        = element_text(size = 9),
    legend.position    = "none",
    strip.text         = element_text(face = "bold", size = 9),
    strip.background   = element_rect(fill = "grey92", colour = NA),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    plot.caption       = element_text(size = 9, colour = "grey40", hjust = 0)
  )
```

## Load & prepare data

```{r load-data}
df <- read.csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
               stringsAsFactors = FALSE, na.strings = c(""))

edu_levels <- c(
  "No formal education", "Incomplete primary (1–5 yrs)", "Complete primary",
  "Incomplete secondary (7–8 yrs)", "Complete secondary",
  "Incomplete high school (10–11 yrs)", "Complete high school",
  "Incomplete university (13–15 yrs)", "Complete university", "Postgraduate"
)

df_table <- df %>%
  mutate(
    Age         = clean_for_table(D_AGE),
    Yrs_in_nbhd = clean_for_table(D_LOC_TIME_CLEAN),
    HH_size     = clean_for_table(D_HH_SIZE),
    Parity      = clean_for_table(D_CHLD_CLEAN),
    Season      = factor(SEASON, levels = c(0, 1), labels = c("Rainy", "Dry")),
    Education   = factor(
      dplyr::recode(as.character(SES_EDU_CAT_CLEAN),
        "Incomplete primary"     = "Incomplete primary (1–5 yrs)",
        "Incomplete secondary"   = "Incomplete secondary (7–8 yrs)",
        "Incomplete high school" = "Incomplete high school (10–11 yrs)",
        "Incomplete university"  = "Incomplete university (13–15 yrs)"
      ),
      levels = edu_levels, ordered = TRUE
    ),
    SES_Score       = clean_for_table(SES_SC_Total),
    Edu_Score       = as.numeric(Education),
    Employed        = factor(clean_for_table(D_EMP_CLEAN),
                             levels = c(0, 1), labels = c("No", "Yes")),
    Employment_type = factor(clean_for_table(D_EMP_FORM_CLEAN),
                             levels = c(0, 1), labels = c("Informal", "Formal")),
    Chronic_pain    = factor(clean_for_table(HLTH_CPAIN_CAT),
                             levels = c(0, 1), labels = c("No", "Yes")),
    Chronic_dis     = factor(clean_for_table(HLTH_CDIS_CAT),
                             levels = c(0, 1), labels = c("No", "Yes")),
    HWISE_Total  = clean_for_table(HW_TOTAL),
    Water_worry  = clean_for_table(as.numeric(HW_WORRY_SUB)),
    Water_5sub   = clean_for_table(HW_MEX_5SUB),
    HWISE_Hi     = factor(
      case_when(
        is.na(clean_for_table(HW_TOTAL)) ~ NA_character_,
        clean_for_table(HW_TOTAL) <= 11  ~ "Low (\u226411)",
        TRUE                             ~ "High (\u226512)"
      ),
      levels = c("Low (\u226411)", "High (\u226512)")
    ),
    HW_Worry   = clean_for_table(HW_WORRY),
    HW_Interr  = clean_for_table(HW_INTERR),
    HW_Clothes = clean_for_table(HW_CLOTHES),
    HW_Plans   = clean_for_table(HW_PLANS),
    HW_Food    = clean_for_table(HW_FOOD),
    HW_Hands   = clean_for_table(HW_HANDS),
    HW_Body    = clean_for_table(HW_BODY),
    HW_Drink   = clean_for_table(HW_DRINK),
    HW_Angry   = clean_for_table(HW_ANGRY),
    HW_Sleep   = clean_for_table(HW_SLEEP),
    HW_None    = clean_for_table(HW_NONE),
    HW_Shame   = clean_for_table(HW_SHAME),
    PSS_Total  = clean_for_table(PSS_TOTAL)
  )

df_water <- df %>%
  mutate(
    W_Continuity = case_when(
      is_sentinel_str(MX3_CONT_INT) ~ NA_character_,
      str_detect(tolower(MX3_CONT_INT), "^contin") ~ "Continuous",
      str_detect(tolower(MX3_CONT_INT), "^intermi|cortan|deja") ~ "Intermittent",
      str_detect(tolower(MX3_CONT_INT), "disminu") ~ "Reduced pressure/flow",
      MX3_CONT_INT %in% c("", "N/A") | is.na(MX3_CONT_INT) ~ NA_character_,
      TRUE ~ "Other"
    ),
    W_Continuity = factor(W_Continuity,
      levels = c("Continuous", "Intermittent", "Reduced pressure/flow", "Other")),
    W_Freq_raw  = coalesce(MX4_WFREQ, MX4_W_FREQ),
    W_Frequency = case_when(
      is_sentinel_str(W_Freq_raw) ~ NA_character_,
      str_detect(tolower(W_Freq_raw), "^diari|every day") ~ "Daily",
      str_detect(tolower(W_Freq_raw), "^intermi") ~ "Intermittent",
      str_detect(tolower(W_Freq_raw), "1 a 3|1-3") ~ "1-3 days/week",
      str_detect(tolower(W_Freq_raw), "4 a 6|4-6") ~ "4-6 days/week",
      str_detect(tolower(W_Freq_raw), "^contin") ~ "Continuous",
      W_Freq_raw %in% c("", "N/A") | is.na(W_Freq_raw) ~ NA_character_,
      TRUE ~ "Other"
    ),
    W_Frequency = factor(W_Frequency,
      levels = c("Continuous", "Daily", "4-6 days/week", "1-3 days/week",
                 "Intermittent", "Other")),
    W_Hrs_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_HRS_PER_WEEK), NA,
             as.numeric(str_extract(MX5_WFREQ_HRS_PER_WEEK, "^[0-9\\.]+")))
    ),
    W_Hrs_week  = ifelse(W_Hrs_week %in% MISSING_CODES | W_Hrs_week > 168, NA, W_Hrs_week),
    W_Days_week = suppressWarnings(
      ifelse(is_sentinel_str(MX5_WFREQ_NDAYS), NA,
             as.numeric(str_extract(MX5_WFREQ_NDAYS, "^[0-9]+")))
    ),
    W_Days_week  = ifelse(W_Days_week %in% MISSING_CODES | W_Days_week > 7, NA, W_Days_week),
    W_Pressure   = case_when(
      is_sentinel_str(MX6_WPRESS) ~ NA_character_,
      str_detect(tolower(MX6_WPRESS), "^s[ií]|^yes|^buena|^bien") ~ "Yes",
      str_detect(tolower(MX6_WPRESS), "^no|^baja|^poca") ~ "No",
      MX6_WPRESS %in% c("", "N/A") | is.na(MX6_WPRESS) ~ NA_character_,
      TRUE ~ "Variable"
    ),
    W_Pressure = factor(W_Pressure, levels = c("Yes", "No", "Variable")),
    W_WS_LOC_F = factor(W_WS_LOC, levels = c("WS", "WI"),
                        labels = c("Water Secure", "Water Insecure")),
    W_Trust = factor(
      case_when(is.na(MX8_TRUST_CAT) ~ NA_character_,
                MX8_TRUST_CAT == 0 ~ "No", MX8_TRUST_CAT == 1 ~ "Yes",
                TRUE ~ NA_character_),
      levels = c("Yes", "No")),
    W_Garrafon_brand = factor(
      case_when(is.na(MX9_GARRAFON_BRAND) ~ NA_character_,
                MX9_GARRAFON_BRAND == 1 ~ "Yes", TRUE ~ "No"),
      levels = c("No", "Yes")),
    W_Has_tinaco = factor(
      case_when(is.na(MX10_HAS_TINACO) ~ NA_character_,
                MX10_HAS_TINACO == 1 ~ "Yes", TRUE ~ "No"),
      levels = c("No", "Yes")),
    W_Has_cisterna = factor(
      case_when(is.na(MX10_HAS_CISTERNA) ~ NA_character_,
                MX10_HAS_CISTERNA == 1 ~ "Yes", TRUE ~ "No"),
      levels = c("No", "Yes")),
    W_Bill_waived = factor(
      case_when(is.na(MX11_BILL_WAIVED) ~ NA_character_,
                MX11_BILL_WAIVED == 1 ~ "Yes", TRUE ~ "No"),
      levels = c("No", "Yes")),
    W_Cost_2mnths = {
      x <- suppressWarnings(as.numeric(as.character(MX11_COST_2MNTHS)))
      ifelse(is.na(x), NA_real_, x)
    }
  )

df_reg <- df_table %>%
  mutate(
    EMO_NEG = {
      r <- suppressWarnings(as.numeric(as.character(df$MX26_EM_HHW_CAT)))
      r <- ifelse(r %in% MISSING_CODES, NA, r)
      case_when(r == 0 ~ 0L, r == 1 ~ 1L, TRUE ~ NA_integer_)
    },
    W_WS_LOC_F       = df_water$W_WS_LOC_F,
    W_Continuity     = df_water$W_Continuity,
    W_Frequency      = df_water$W_Frequency,
    W_Days_week      = df_water$W_Days_week,
    W_Hrs_week       = df_water$W_Hrs_week,
    W_Pressure       = df_water$W_Pressure,
    W_Trust          = df_water$W_Trust,
    W_Garrafon_brand = df_water$W_Garrafon_brand,
    W_Has_tinaco     = df_water$W_Has_tinaco,
    W_Has_cisterna   = df_water$W_Has_cisterna,
    W_Bill_waived    = df_water$W_Bill_waived,
    W_Cost_2mnths    = df_water$W_Cost_2mnths
  )

ref_levels <- sapply(names(df_reg), function(v) {
  x <- df_reg[[v]]
  if (is.factor(x)) levels(x)[1] else NA_character_
})
```

## Run unadjusted PSS models

```{r pss-models}
hwise_items <- c("HW_Worry", "HW_Interr", "HW_Clothes", "HW_Plans", "HW_Food",
                 "HW_Hands", "HW_Body", "HW_Drink", "HW_Angry", "HW_Sleep",
                 "HW_None", "HW_Shame")

pss_predictors <- c(
  "Age", "Parity", "HH_size", "Yrs_in_nbhd",
  "SES_Score", "Edu_Score", "Employed", "Employment_type",
  "Chronic_pain", "Chronic_dis",
  "HWISE_Total", "Water_5sub", "HWISE_Hi", "Water_worry",
  "Season", "W_WS_LOC_F", "W_Continuity", "W_Frequency",
  "W_Days_week", "W_Hrs_week", "W_Pressure",
  "W_Trust", "W_Garrafon_brand", "W_Has_tinaco", "W_Has_cisterna",
  "W_Bill_waived", "W_Cost_2mnths"
)
pss_predictors_fp <- append(pss_predictors, hwise_items,
                             after = which(pss_predictors == "Water_worry"))

pss_results <- run_lm_uv("PSS_Total", pss_predictors_fp, df_reg, pred_labels) %>%
  mutate(
    group         = group_map[predictor],
    group         = factor(group, levels = group_order),
    var_order     = predictor_order[predictor],
    display_label = mapply(make_display_label, predictor, term, label)
  ) %>%
  arrange(group, var_order)

plot_df_pss <- pss_results %>%
  mutate(
    label_full = factor(display_label, levels = rev(unique(display_label))),
    sig        = sig_sym(p.value),
    n_label    = paste0("n=", n)
  )
```

## Build sparkline tiles

Each row gets a horizontal stacked bar showing the distribution of its predictor
variable, placed immediately to the right of the existing n= labels.

- **Continuous predictors**: 7-bin histogram — segments coloured light→dark blue
- **Binary factors**: two segments — blue (reference) / orange (comparison)
- **Multi-level factors**: one segment per level, blue→orange gradient

```{r spark-tiles}
# ── Sparkline helper ──────────────────────────────────────────────────────────
make_spark_tiles <- function(predictor_name, label_full_val, group_val,
                              df_data, x_start, x_width, bins_cont = 7) {
  var <- df_data[[predictor_name]]

  if (is.factor(var) || is.character(var)) {
    var_f <- if (is.factor(var)) var else factor(var)
    tab   <- table(na.omit(droplevels(var_f)))
    if (length(tab) == 0) return(NULL)
    props <- as.numeric(tab) / sum(tab)
    n_lvl <- length(props)
    pal   <- if (n_lvl == 2) {
      c("#6baed6", "#fd8d3c")
    } else {
      colorRampPalette(c("#6baed6", "#fdae6b", "#fd8d3c"))(n_lvl)
    }
    cuml  <- c(0, cumsum(props))
  } else {
    vals <- na.omit(as.numeric(var))
    if (length(vals) < 5) return(NULL)
    h    <- hist(vals, breaks = bins_cont, plot = FALSE)
    props <- h$counts / sum(h$counts)
    n_lvl <- length(props)
    pal   <- colorRampPalette(c("#c6dbef", "#2171b5"))(n_lvl)
    cuml  <- c(0, cumsum(props))
  }

  tibble(
    label_full = label_full_val,
    group      = group_val,
    x_center   = x_start + (cuml[-length(cuml)] + diff(cuml) / 2) * x_width,
    seg_width  = diff(cuml) * x_width,
    fill_color = pal
  )
}

# ── Compute sparkline x positions (just right of existing n= labels) ──────────
x_max_ci      <- max(plot_df_pss$conf.high, na.rm = TRUE)
x_n_label_pos <- x_max_ci * 1.08          # mirrors existing geom_text x
x_spark_start <- x_max_ci * 1.20          # small gap after n= text
x_spark_width <- x_max_ci * 0.14          # total width of the spark bar
x_axis_end    <- x_spark_start + x_spark_width * 1.15  # axis limit

# ── Build tiles for every row ─────────────────────────────────────────────────
spark_df_pss <- pmap_dfr(
  list(
    predictor  = plot_df_pss$predictor,
    label_full = as.character(plot_df_pss$label_full),
    group      = as.character(plot_df_pss$group)
  ),
  function(predictor, label_full, group) {
    make_spark_tiles(predictor, label_full, group,
                     df_reg, x_spark_start, x_spark_width)
  }
) %>%
  mutate(
    label_full = factor(label_full, levels = levels(plot_df_pss$label_full)),
    group      = factor(group, levels = group_order)
  )
```

## Forest plot with sparklines

```{r forest-sparklines, fig.width=10.5, fig.height=11}
p_pss_spark <- ggplot(plot_df_pss,
                      aes(x = estimate, y = label_full,
                          colour = p.value < 0.05)) +
  # ── original forest layers ─────────────────────────────────────────────────
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.2) +
  geom_text(aes(label = sig), nudge_y = 0.25, size = 3.5,
            fontface = "bold", show.legend = FALSE) +
  # n= label (unchanged position)
  geom_text(aes(x = x_n_label_pos, label = n_label),
            size = 2.8, colour = "grey40", hjust = 0) +
  # ── sparkline tiles ────────────────────────────────────────────────────────
  geom_tile(
    data        = spark_df_pss,
    aes(y       = label_full,
        x       = x_center,
        fill    = fill_color,
        width   = seg_width),
    height      = 0.65,
    inherit.aes = FALSE
  ) +
  # ── "Dist." header — rendered in every facet strip via geom_text on a
  #    small helper dataset placed at a high y value (Inf clips nicely) ──────
  geom_text(
    data = tibble(
      group      = factor(group_order, levels = group_order),
      label_full = factor(NA_character_, levels = levels(plot_df_pss$label_full))
    ),
    aes(x = x_spark_start + x_spark_width / 2, y = Inf,
        label = "Dist."),
    inherit.aes = FALSE,
    vjust = 1.8, size = 2.8, colour = "grey35"
  ) +
  # ── scales & facets ────────────────────────────────────────────────────────
  facet_grid(group ~ ., scales = "free_y", space = "free_y") +
  scale_colour_manual(values = c("FALSE" = "grey55", "TRUE" = "#c0392b")) +
  scale_fill_identity() +
  scale_x_continuous(
    expand = expansion(mult = c(0.05, 0)),
    limits = c(NA, x_axis_end)
  ) +
  scale_y_discrete(expand = expansion(add = 0.4)) +
  labs(
    title   = "Unadjusted Linear Regression — Perceived Stress Score (β)",
    caption = paste0(
      "Red = p < 0.05   Symbols: \u2020p<0.10  *p<0.05  **p<0.01  ***p<0.001\n",
      "Dist.: continuous = light\u2192dark blue (low\u2192high); ",
      "binary = blue (ref) / orange (comparison); ",
      "multi-level = blue\u2192orange gradient."
    ),
    x = "\u03b2 coefficient (95% CI)",
    y = NULL
  ) +
  pub_theme +
  theme(panel.spacing = unit(0.8, "lines"))

p_pss_spark
```

```{r save, include=FALSE}
ggsave(
  "./output/figures/06z.Unadj_coef_PSS_sparklines.png",
  plot   = p_pss_spark,
  width  = 10.5, height = 11, dpi = 300, bg = "white"
)
cat("Saved: ./output/figures/06z.Unadj_coef_PSS_sparklines.png\n")
```
