---
title: "07.1 PSS Models – Aim 2: Components of Water Insecurity"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(gtsummary)
library(lmtest)
library(flextable)
library(ggplot2)
library(officer)
library(ggstance)
```

## Overview

This script fits the **Aim 2 linear regression models** predicting **Perceived
Stress Scale total score (`PSS_TOTAL`)**. Each predictor is tested in a separate
model adjusted for core covariates. Model names follow **Table 2** in the
methods section.

| Role | Variables |
|------|-----------|
| Outcome | `PSS_TOTAL` (continuous, 0–56) |
| Core covariates (every model) | Age (`D_AGE`), SES composite (`SES_SC_Total`), Household size (`D_HH_SIZE`), Chronic disease (`Chronic_dis`) |
| Augmenting (one per model) | HWISE individual items (m_a5) · Hours of water supply/week (m_a6) · Intermittent supply (m_a7) · Season (m_a8) · Neighborhood comparison (m_a9) · Excessive time handling water (m_a10) · Sufficiency rating (m_a11) |

> **Multiple testing**: Given the number of models estimated within this
> outcome, all F-test p-values are adjusted using the **Benjamini-Hochberg (BH)**
> procedure, applied across all Aim 2 models.

---

# Load and Prepare Data

```{r load-data}
MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

clean_num <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES, NA_real_, x_num)
}

df_raw <- read_csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
                   show_col_types = FALSE)

cat("Dataset:", nrow(df_raw), "rows ×", ncol(df_raw), "cols\n")

# ── Check new column availability ─────────────────────────────────────────────
new_cols <- c("HLTH_CDIS_CAT", "HW_TOTAL", "HW_WORRY_SUB", "HW_MEX_5SUB",
              "MX6_WPRESS", "MX4_WFREQ", "MX4_W_FREQ", "W_WS_LOC")
cat("\nNew column availability:\n")
for (col in new_cols) {
  cat(sprintf("  %-20s %s\n", col,
              ifelse(col %in% names(df_raw), "FOUND", "*** MISSING ***")))
}
```

```{r prepare-variables}
hwise_items <- c("HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
                 "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
                 "HW_ANGRY", "HW_SLEEP",  "HW_NONE",    "HW_SHAME")

df <- df_raw %>%
  mutate(
    # ── Outcome ────────────────────────────────────────────────────────────────
    PSS_TOTAL    = clean_num(PSS_TOTAL),

    # ── Core covariates ────────────────────────────────────────────────────────
    D_AGE        = clean_num(D_AGE),
    SES_SC_Total = clean_num(SES_SC_Total),
    D_HH_SIZE    = clean_num(D_HH_SIZE),

    # ── HWISE individual items (m_a5) ──────────────────────────────────────────
    across(all_of(hwise_items), clean_num),

    # ── Hours of water supply per week (m_a6) ──────────────────────────────────
    HRS_PER_WEEK = clean_num(MX5_WFREQ_HRS_PER_WEEK),

    # ── Continuous vs. intermittent supply (m_a7) ──────────────────────────────
    # 0 = Continuous (reference), 1 = Intermittent
    CONT_INT = {
      raw     <- as.character(MX3_CONT_INT)
      raw_num <- suppressWarnings(as.numeric(raw))
      factor(case_when(
        raw_num %in% MISSING_CODES                         ~ NA_character_,
        raw_num == 0                                       ~ "Continuous",
        raw_num == 1                                       ~ "Intermittent",
        str_detect(str_to_lower(raw), "contin")            ~ "Continuous",
        str_detect(str_to_lower(raw), "inter")             ~ "Intermittent",
        is.na(raw) | str_trim(raw) == "" | raw == "NA"    ~ NA_character_,
        TRUE                                               ~ raw
      ), levels = c("Continuous", "Intermittent"))
    },

    # ── Season (m_a8) ──────────────────────────────────────────────────────────
    # 0 = Season 1 (rainy, reference), 1 = Season 2 (dry)
    SEASON = factor(SEASON, levels = c(0, 1), labels = c("Rainy", "Dry")),

    # ── Neighborhood water comparison (m_a9) ───────────────────────────────────
    # 0 = Better (reference), 1 = Same, 2 = Worse
    NB_COMP = factor(
      case_when(
        clean_num(MX28_WQ_COMP_CAT) == 0 ~ "Better",
        clean_num(MX28_WQ_COMP_CAT) == 1 ~ "Same",
        clean_num(MX28_WQ_COMP_CAT) == 2 ~ "Worse",
        TRUE                             ~ NA_character_
      ),
      levels = c("Better", "Same", "Worse")
    ),

    # ── Excessive time handling water (m_a10) ──────────────────────────────────
    # 0 = No (reference), 1 = Yes
    TIME_WATER = factor(
      case_when(
        clean_num(MX19_TIME_CAT) == 0 ~ "No",
        clean_num(MX19_TIME_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── Sufficiency rating (m_a11) ─────────────────────────────────────────────
    # Participant's self-rating of their own water sufficiency (1–4 scale)
    # 1 = not sufficient, 2 = moderately sufficient, 3 = almost, 4 = completely
    # Entered as numeric (higher = more sufficient → expected negative β with PSS)
    SUFF_RATING = clean_num(MX16_SCEN_99),

    # ── Chronic disease (core covariate) ──────────────────────────────────────
    # 0 = No (reference), 1 = Yes
    Chronic_dis = factor(
      case_when(
        clean_num(HLTH_CDIS_CAT) == 0 ~ "No",
        clean_num(HLTH_CDIS_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── HWISE categorical ────────────────────────────────────────────────────
    HWISE_Cat = factor(
      case_when(
        is.na(clean_num(HW_TOTAL))    ~ NA_character_,
        clean_num(HW_TOTAL) <= 3      ~ "Marginal (0-3)",
        clean_num(HW_TOTAL) <= 11     ~ "Low (4-11)",
        TRUE                          ~ "Moderate/Severe (\u226512)"
      ),
      levels = c("Marginal (0-3)", "Low (4-11)", "Moderate/Severe (\u226512)")
    ),

    # ── Water worry subscale ─────────────────────────────────────────────────
    Water_worry = clean_num(HW_WORRY_SUB),

    # ── 5-item subscale ───────────────────────────────────────────────────────
    Water_5sub  = clean_num(HW_MEX_5SUB),
    HWI_5sub_hi = factor(
      case_when(
        is.na(clean_num(HW_MEX_5SUB)) ~ NA_character_,
        clean_num(HW_MEX_5SUB) >= 3   ~ "High",
        TRUE                          ~ "Low/none"
      ),
      levels = c("Low/none", "High")
    ),

    # ── Water pressure ───────────────────────────────────────────────────────
    W_Pressure = factor(
      case_when(
        str_detect(str_to_lower(as.character(MX6_WPRESS)),
                   "s[i\u00ed]|yes|buena|bien")  ~ "Yes",
        str_detect(str_to_lower(as.character(MX6_WPRESS)),
                   "^no|baja|poca")               ~ "No/Low",
        TRUE                                      ~ NA_character_
      ),
      levels = c("Yes", "No/Low")
    ),

    # ── Supply frequency ─────────────────────────────────────────────────────
    W_Frequency = {
      freq_raw <- if ("MX4_W_FREQ" %in% names(df_raw)) {
        dplyr::coalesce(as.character(MX4_WFREQ), as.character(MX4_W_FREQ))
      } else {
        as.character(MX4_WFREQ)
      }
      raw <- str_to_lower(as.character(freq_raw))
      factor(
        case_when(
          str_detect(raw, "^contin")           ~ "Continuous",
          str_detect(raw, "^diari|every day")  ~ "Daily",
          str_detect(raw, "4 a 6|4-6")         ~ "4\u20136 days/week",
          str_detect(raw, "1 a 3|1-3")         ~ "1\u20133 days/week",
          str_detect(raw, "^intermi")          ~ "Intermittent",
          TRUE                                 ~ NA_character_
        ),
        levels = c("Continuous", "Daily", "4\u20136 days/week",
                   "1\u20133 days/week", "Intermittent")
      )
    },

    # ── Neighborhood water security ───────────────────────────────────────────
    W_WS_LOC_F = if ("W_WS_LOC" %in% names(df_raw)) {
      factor(W_WS_LOC, levels = c("WS", "WI"),
             labels = c("Water Secure", "Water Insecure"))
    } else {
      factor(NA_character_, levels = c("Water Secure", "Water Insecure"))
    }
  )
```

```{r check-missingness}
analytic_vars <- c("PSS_TOTAL", "D_AGE", "SES_SC_Total", "D_HH_SIZE", "SEASON",
                   "Chronic_dis",
                   hwise_items, "HRS_PER_WEEK", "CONT_INT",
                   "NB_COMP", "TIME_WATER", "SUFF_RATING",
                   "HWISE_Cat", "Water_worry", "Water_5sub", "HWI_5sub_hi",
                   "W_Pressure", "W_Frequency", "W_WS_LOC_F")

df %>%
  summarise(across(all_of(analytic_vars), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_missing") %>%
  mutate(pct_missing = round(100 * n_missing / nrow(df), 1)) %>%
  arrange(desc(n_missing)) %>%
  print(n = Inf)
```

---

# Model Specifications {#model-specs}

All models share the same **core covariates**: `D_AGE`, `SES_SC_Total`,
`D_HH_SIZE`. One water predictor is added per model. Model names correspond to
**Table 2** in the methods.

```{r model-specs}
core_formula <- PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis

# ── m_a5: Individual HWISE item models (one per item) ─────────────────────────
hwise_item_specs <- setNames(
  lapply(hwise_items, function(item) {
    list(
      label   = paste0("m_a5 – ", item),
      group   = "m_a5: HWISE items",
      formula = as.formula(
        paste("PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis +", item)
      )
    )
  }),
  paste0("m_a5_", tolower(hwise_items))
)

# ── Main model registry ────────────────────────────────────────────────────────
model_specs <- c(
  hwise_item_specs,
  list(
    m_a6 = list(
      label   = "m_a6: Hours of water supply/week",
      group   = "m_a6: Hours supply",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HRS_PER_WEEK
    ),
    m_a7 = list(
      label   = "m_a7: Intermittent supply",
      group   = "m_a7: Intermittent",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + CONT_INT
    ),
    m_a8 = list(
      label   = "m_a8: Season (dry vs. rainy)",
      group   = "m_a8: Season",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + SEASON
    ),
    m_a9 = list(
      label   = "m_a9: Neighborhood comparison",
      group   = "m_a9: Neighborhood comparison",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + NB_COMP
    ),
    m_a10 = list(
      label   = "m_a10: Excessive time handling water",
      group   = "m_a10: Time burden",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + TIME_WATER
    ),
    m_a11 = list(
      label   = "m_a11: Sufficiency rating",
      group   = "m_a11: Sufficiency",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + SUFF_RATING
    ),
    m_a12 = list(
      label   = "m_a12: HWISE (categorical)",
      group   = "m_a12: HWISE categorical",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HWISE_Cat
    ),
    m_a13 = list(
      label   = "m_a13: Water worry subscale",
      group   = "m_a13: Water worry",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + Water_worry
    ),
    m_a14 = list(
      label   = "m_a14: 5-item subscale (categorical)",
      group   = "m_a14: 5-item categorical",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HWI_5sub_hi
    ),
    m_a15 = list(
      label   = "m_a15: 5-item subscale (continuous)",
      group   = "m_a15: 5-item continuous",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + Water_5sub
    ),
    m_a16 = list(
      label   = "m_a16: Water pressure",
      group   = "m_a16: Water pressure",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + W_Pressure
    ),
    m_a17 = list(
      label   = "m_a17: Supply frequency",
      group   = "m_a17: Supply frequency",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + W_Frequency
    ),
    m_a18 = list(
      label   = "m_a18: Neighborhood (WS/WI)",
      group   = "m_a18: Neighborhood",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + W_WS_LOC_F
    )
  )
)

cat("Total models registered:", length(model_specs), "\n")
cat("Model IDs:\n")
cat(paste(" -", names(model_specs)), sep = "\n")
```

---

# Fit Models

```{r fit-models}
fitted_models <- lapply(model_specs, function(spec) {
  lm(spec$formula, data = df, na.action = na.omit)
})

# Sample sizes per model
tibble(
  model_id = names(model_specs),
  label    = map_chr(model_specs, "label"),
  n        = map_int(fitted_models, nobs)
) %>% print(n = Inf)
```

---

# Summary Table with BH Correction {#summary-table}

One row per model. The **focal predictor** is the augmenting water variable.
P-values are adjusted with the **Benjamini-Hochberg procedure** applied across
all Aim 2 models.

```{r summary-table}
core_terms <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE", "Chronic_disYes")

# ── Extract focal predictor rows from each model ───────────────────────────────
summary_raw <- map_dfr(names(fitted_models), function(nm) {
  mod  <- fitted_models[[nm]]
  spec <- model_specs[[nm]]
  fit  <- glance(mod)

  # Refit M0 on same rows for F-test
  rows_used <- as.integer(rownames(model.frame(mod)))
  m0_sub    <- lm(core_formula, data = df[rows_used, ])
  lrt_tab   <- anova(m0_sub, mod)
  f_stat    <- lrt_tab$F[2]
  f_df1     <- lrt_tab$Df[2]
  f_df2     <- lrt_tab$Res.Df[2]
  f_p       <- lrt_tab$`Pr(>F)`[2]

  focal_coefs <- tidy(mod, conf.int = TRUE) %>%
    filter(!term %in% core_terms)

  if (nrow(focal_coefs) == 0) return(NULL)

  focal_coefs %>%
    mutate(
      model_id  = nm,
      label     = spec$label,
      group     = spec$group,
      n         = nobs(mod),
      adj_R2    = round(fit$adj.r.squared, 3),
      delta_adj_R2 = round(fit$adj.r.squared - glance(m0_sub)$adj.r.squared, 3),
      AIC       = round(fit$AIC, 1),
      F_stat    = round(f_stat, 2),
      F_df1     = as.integer(f_df1),
      F_df2     = as.integer(f_df2),
      F_p       = round(f_p, 4)
    )
})

# ── Apply BH correction across all model F-test p-values ──────────────────────
# One p-value per model (not per coefficient)
model_fps <- summary_raw %>%
  distinct(model_id, F_p) %>%
  mutate(F_p_adj = round(p.adjust(F_p, method = "BH"), 4))

summary_table <- summary_raw %>%
  left_join(model_fps, by = c("model_id", "F_p")) %>%
  mutate(
    beta_ci = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_coef  = case_when(
      is.na(p.value)  ~ "—",
      p.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", p.value)
    ),
    p_F_fmt = case_when(
      is.na(F_p)      ~ "—",
      F_p < 0.001     ~ "<0.001",
      TRUE            ~ sprintf("%.3f", F_p)
    ),
    p_adj_fmt = case_when(
      is.na(F_p_adj)   ~ "—",
      F_p_adj < 0.001  ~ "<0.001",
      TRUE             ~ sprintf("%.3f", F_p_adj)
    )
  )

# ── Display ────────────────────────────────────────────────────────────────────
summary_table %>%
  arrange(AIC) %>%
  select(label, term, n, adj_R2, delta_adj_R2, AIC, beta_ci,
         p_coef, F_stat, p_F_fmt, p_adj_fmt) %>%
  flextable() %>%
  set_header_labels(
    label        = "Model",
    term         = "Predictor term",
    n            = "N",
    adj_R2       = "Adj. R²",
    delta_adj_R2 = "Δ Adj. R²",
    AIC          = "AIC",
    beta_ci      = "β [95% CI]",
    p_coef       = "p (coef.)",
    F_stat       = "F",
    p_F_fmt      = "p (F-test)",
    p_adj_fmt    = "p_adj (BH)"
  ) %>%
  bold(~ !is.na(p_coef) & as.numeric(gsub("<", "", p_coef)) < 0.05,
       j = "p_coef") %>%
  color(~ !is.na(p_adj_fmt) & p_adj_fmt != "—" &
          as.numeric(gsub("<", "0", p_adj_fmt)) < 0.05,
        j = "p_adj_fmt", color = "darkgreen") %>%
  color(~ !is.na(p_adj_fmt) & p_adj_fmt != "—" &
          as.numeric(gsub("<", "0", p_adj_fmt)) >= 0.05,
        j = "p_adj_fmt", color = "grey50") %>%
  autofit()
```

---

# Top 5 Models: Publication Table {#top5-table}

```{r top5-pub-table, fig.width=10}
library(flextable)
library(officer)

# ── Readable predictor labels ─────────────────────────────────────────────────
readable_terms <- c(
  HRS_PER_WEEK                     = "Hours of supply / week",
  CONTINTIntermittent              = "Intermittent vs. continuous supply",
  SEASONDry                        = "Dry vs. rainy season",
  NB_COMPSame                      = "Neighborhood: same as CDMX",
  NB_COMPWorse                     = "Neighborhood: worse than CDMX",
  TIME_WATERYes                    = "Excessive time managing water",
  SUFF_RATING                      = "Sufficiency rating (1–4)",
  `HWISE_CatLow (4-11)`            = "HWISE: Low (4–11)",
  `HWISE_CatModerate/Severe (≥12)` = "HWISE: Moderate/Severe (≥12)",
  Water_worry                      = "Water worry subscale",
  HWI_5sub_hiHigh                  = "5-item subscale: High (≥3)",
  Water_5sub                       = "5-item subscale (continuous)",
  W_PressureNo.Low                 = "Water pressure: No/Low",
  `W_PressureNo/Low`               = "Water pressure: No/Low",
  W_FrequencyDaily                 = "Supply frequency: Daily",
  `W_Frequency4–6 days/week`       = "Supply frequency: 4–6 days/wk",
  `W_Frequency1–3 days/week`       = "Supply frequency: 1–3 days/wk",
  W_FrequencyIntermittent          = "Supply frequency: Intermittent",
  `W_WS_LOC_FWater Insecure`       = "Neighborhood: Water Insecure"
)

# ── Select top 5 models by AIC ────────────────────────────────────────────────
top5_ids <- summary_table %>%
  distinct(model_id, AIC) %>%
  slice_min(AIC, n = 5, with_ties = FALSE) %>%
  arrange(AIC) %>%
  mutate(rank = row_number())

top5_data <- summary_table %>%
  filter(model_id %in% top5_ids$model_id) %>%
  left_join(select(top5_ids, model_id, rank), by = "model_id") %>%
  arrange(rank, estimate) %>%
  transmute(
    rank        = rank,
    Model       = label,
    Predictor   = coalesce(readable_terms[term], term),
    N           = n,
    beta_ci     = beta_ci,
    adj_r2      = adj_R2,
    delta_r2    = delta_adj_R2,
    AIC         = AIC,
    p_coef      = p_coef,
    p_adj       = p_adj_fmt
  )

# ── Define colour palette ─────────────────────────────────────────────────────
col_header  <- "#2c3e50"   # dark navy (matches UV1 style)
col_sig     <- "#fff3cd"   # amber highlight  (BH-significant rows)
col_ns      <- "#f4f6f9"   # light grey       (non-significant rows)

# ── Build flextable ───────────────────────────────────────────────────────────
ft <- top5_data %>%
  select(rank, Model, Predictor, N, beta_ci, adj_r2, delta_r2,
         AIC, p_coef, p_adj) %>%
  flextable() %>%
  set_header_labels(
    rank      = "#",
    Model     = "Model",
    Predictor = "Focal predictor",
    N         = "N",
    beta_ci   = "β [95% CI]",
    adj_r2    = "Adj. R²",
    delta_r2  = "Δ Adj. R²",
    AIC       = "AIC",
    p_coef    = "p (coef.)",
    p_adj     = "p_adj (BH)"
  ) %>%
  # ── header styling ──────────────────────────────────────────────────────────
  bg(bg = col_header, part = "header") %>%
  color(color = "white", part = "header") %>%
  bold(part = "header") %>%
  fontsize(size = 10, part = "header") %>%
  # ── rank column ─────────────────────────────────────────────────────────────
  align(j = "rank", align = "center", part = "all") %>%
  # ── Amber background for BH-significant rows ─────────────────────────────
  bg(
    i = ~ !p_adj %in% c("—") &
          suppressWarnings(
            as.numeric(gsub("<0.001", "0.0005", p_adj))
          ) < 0.05,
    bg = col_sig, part = "body"
  ) %>%
  bg(
    i = ~ p_adj == "—" | suppressWarnings(
            as.numeric(gsub("<0.001", "0.0005", p_adj))
          ) >= 0.05,
    bg = col_ns, part = "body"
  ) %>%
  # ── p-value colouring ───────────────────────────────────────────────────────
  color(
    i = ~ !p_coef %in% c("—") &
          suppressWarnings(as.numeric(gsub("<", "0", p_coef))) < 0.05,
    j = "p_coef", color = "#1A5276"
  ) %>%
  bold(
    i = ~ !p_coef %in% c("—") &
          suppressWarnings(as.numeric(gsub("<", "0", p_coef))) < 0.05,
    j = "p_coef"
  ) %>%
  color(
    i = ~ !p_adj %in% c("—") &
          suppressWarnings(
            as.numeric(gsub("<0.001", "0.0005", p_adj))
          ) < 0.05,
    j = "p_adj", color = "#1E8449"
  ) %>%
  bold(
    i = ~ !p_adj %in% c("—") &
          suppressWarnings(
            as.numeric(gsub("<0.001", "0.0005", p_adj))
          ) < 0.05,
    j = "p_adj"
  ) %>%
  color(
    i = ~ p_adj == "—" | suppressWarnings(
            as.numeric(gsub("<0.001", "0.0005", p_adj))
          ) >= 0.05,
    j = "p_adj", color = "grey55"
  ) %>%
  # ── merge repeated Model / N / Adj.R² / ΔAIC / AIC cells ──────────────────
  merge_v(j = c("rank", "Model", "N", "adj_r2", "delta_r2", "AIC",
                "p_adj")) %>%
  valign(j = c("rank", "Model", "N", "adj_r2", "delta_r2", "AIC",
               "p_adj"),
         valign = "top", part = "body") %>%
  # ── borders ─────────────────────────────────────────────────────────────────
  border_remove() %>%
  hline_top(part = "header", border = fp_border(color = col_header, width = 1.5)) %>%
  hline_bottom(part = "header", border = fp_border(color = col_header, width = 1.5)) %>%
  hline_bottom(part = "body",   border = fp_border(color = col_header, width = 1.5)) %>%
  hline(border = fp_border(color = "#bdc3c7", width = 0.5), part = "body") %>%
  # ── alignment ───────────────────────────────────────────────────────────────
  align(j = c("N", "adj_r2", "delta_r2", "AIC", "p_coef", "p_adj"),
        align = "center", part = "all") %>%
  align(j = c("beta_ci"), align = "right", part = "body") %>%
  # ── font ────────────────────────────────────────────────────────────────────
  font(fontname = "Calibri", part = "all") %>%
  fontsize(size = 9.5, part = "body") %>%
  # ── caption ─────────────────────────────────────────────────────────────────
  add_footer_lines(paste0(
    "Top 5 models ranked by AIC (lower = better fit). ",
    "β = unstandardised regression coefficient; 95% CI = confidence interval. ",
    "Adj. R² = adjusted R-squared of the full model; ",
    "Δ Adj. R² = change over core-covariate-only model. ",
    "p_adj (BH) = F-test p-value adjusted using the Benjamini-Hochberg procedure ",
    "across all Aim 2 PSS models. ",
    "Amber rows: BH-adjusted p < 0.05. Bold blue: unadjusted p < 0.05."
  )) %>%
  color(color = "#555555", part = "footer") %>%
  fontsize(size = 9, part = "footer") %>%
  italic(part = "footer") %>%
  set_table_properties(layout = "fixed")

ft  # display in HTML

# ── Save to Word ──────────────────────────────────────────────────────────────
save_as_docx(
  ft,
  path = "./output/tables/07.Table_Top5_PSS_models.docx",
  pr_section = prop_section(
    page_size   = page_size(orient = "landscape"),
    page_margins = page_mar(top = 0.8, bottom = 0.8,
                            left = 0.8, right = 0.8)
  )
)
cat("Saved: output/Table_Top5_PSS_models.docx\n")
```

---

# HWISE Items Summary (m_a5) {#hwise-summary}

```{r hwise-summary}
hwise_readable <- c(
  HW_WORRY   = "Worry about water",
  HW_INTERR  = "Interrupted supply",
  HW_CLOTHES = "Couldn't wash clothes",
  HW_PLANS   = "Plans disrupted",
  HW_FOOD    = "Food preparation affected",
  HW_HANDS   = "Couldn't wash hands",
  HW_BODY    = "Couldn't bathe",
  HW_DRINK   = "Not enough to drink",
  HW_ANGRY   = "Felt angry",
  HW_SLEEP   = "Sleep disrupted",
  HW_NONE    = "No water at all",
  HW_SHAME   = "Felt shame"
)

summary_table %>%
  filter(group == "m_a5: HWISE items") %>%
  mutate(item_label = hwise_readable[term]) %>%
  select(item_label, n, adj_R2, delta_adj_R2, beta_ci,
         p_coef, p_F_fmt, p_adj_fmt) %>%
  arrange(as.numeric(gsub("<0.001", "0", p_F_fmt))) %>%
  flextable() %>%
  set_header_labels(
    item_label   = "HWISE item",
    n            = "N",
    adj_R2       = "Adj. R²",
    delta_adj_R2 = "Δ Adj. R²",
    beta_ci      = "β [95% CI]",
    p_coef       = "p (coef.)",
    p_F_fmt      = "p (F-test)",
    p_adj_fmt    = "p_adj (BH)"
  ) %>%
  color(~ !is.na(p_adj_fmt) & p_adj_fmt != "—" &
          as.numeric(gsub("<", "0", p_adj_fmt)) < 0.05,
        j = "p_adj_fmt", color = "darkgreen") %>%
  color(~ !is.na(p_adj_fmt) & p_adj_fmt != "—" &
          as.numeric(gsub("<", "0", p_adj_fmt)) >= 0.05,
        j = "p_adj_fmt", color = "grey50") %>%
  autofit()
```

---

# Forest Plot: Non-HWISE Predictors (m_a6–m_a11) {#forest-plot-main}

```{r forest-plot-main, fig.width = 9, fig.height = 7}
pub_theme <- theme_classic(base_size = 11) +
  theme(
    plot.title         = element_text(face = "bold", size = 12),
    axis.text.y        = element_text(size = 9),
    axis.text.x        = element_text(size = 9),
    legend.position    = "none",
    strip.text         = element_text(face = "bold", size = 9),
    strip.background   = element_rect(fill = "grey92", colour = NA),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    plot.caption       = element_text(size = 9, colour = "grey40", hjust = 0)
  )

sig_sym <- function(p) {
  dplyr::case_when(
    p < 0.001 ~ "***", p < 0.01 ~ "**", p < 0.05 ~ "*",
    p < 0.1   ~ "\u2020", TRUE ~ ""
  )
}

term_labels <- c(
  HRS_PER_WEEK                         = "Hours of supply/week",
  CONTINTIntermittent                  = "Intermittent vs. continuous",
  SEASONDry                            = "Dry vs. rainy season",
  NB_COMPSame                          = "Neighborhood: same as CDMX",
  NB_COMPWorse                         = "Neighborhood: worse than CDMX",
  TIME_WATERYes                        = "Excessive time managing water",
  SUFF_RATING                          = "Sufficiency rating (1\u20134)",
  `HWISE_CatLow (4-11)`                = "HWISE: Low (4\u201311)",
  `HWISE_CatModerate/Severe (≥12)`     = "HWISE: Moderate/Severe (\u226512)",
  Water_worry                          = "Water worry subscale",
  HWI_5sub_hiHigh                      = "5-item subscale: High (\u22653)",
  Water_5sub                           = "5-item subscale (continuous)",
  W_PressureNo.Low                     = "Water pressure: No/Low",
  `W_PressureNo/Low`                   = "Water pressure: No/Low",
  W_FrequencyDaily                     = "Supply frequency: Daily",
  `W_Frequency4–6 days/week`           = "Supply frequency: 4\u20136 days/wk",
  `W_Frequency1–3 days/week`           = "Supply frequency: 1\u20133 days/wk",
  W_FrequencyIntermittent              = "Supply frequency: Intermittent",
  `W_WS_LOC_FWater Insecure`           = "Neighborhood: Water Insecure"
)

group_labels <- c(
  "m_a6: Hours supply"       = "Water supply",
  "m_a7: Intermittent"       = "Water supply",
  "m_a8: Season"             = "Water supply",
  "m_a9: Neighborhood comparison" = "Neighborhood",
  "m_a10: Time burden"       = "Water supply",
  "m_a11: Sufficiency"       = "Water supply",
  "m_a12: HWISE categorical" = "HWISE composite",
  "m_a13: Water worry"       = "HWISE composite",
  "m_a14: 5-item categorical" = "HWISE composite",
  "m_a15: 5-item continuous" = "HWISE composite",
  "m_a16: Water pressure"    = "Water supply",
  "m_a17: Supply frequency"  = "Water supply",
  "m_a18: Neighborhood"      = "Neighborhood"
)

main_coefs <- summary_table %>%
  filter(!grepl("m_a5", group)) %>%
  mutate(
    term_label  = coalesce(term_labels[term], term),
    panel_group = coalesce(group_labels[group], group),
    panel_group = factor(panel_group,
                         levels = c("HWISE composite", "Water supply",
                                    "Neighborhood")),
    sig         = sig_sym(p.value)
  ) %>%
  arrange(panel_group, estimate)

main_coefs$term_label <- factor(main_coefs$term_label,
                                levels = unique(main_coefs$term_label))

ggplot(main_coefs,
       aes(x = estimate, y = term_label,
           xmin = conf.low, xmax = conf.high,
           colour = p.value < 0.05)) +
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.5) +
  geom_text(aes(label = sig), nudge_y = 0.35, size = 3.5,
            fontface = "bold", show.legend = FALSE) +
  facet_grid(panel_group ~ ., scales = "free_y", space = "free_y") +
  scale_colour_manual(values = c("FALSE" = "grey60", "TRUE" = "#c0392b")) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.12))) +
  labs(
    title    = "Aim 2: Water predictors of PSS",
    subtitle = "Adjusted for age, SES composite, household size, and chronic disease",
    caption  = "Red = p < 0.05  |  Symbols: \u2020p<0.10  *p<0.05  **p<0.01  ***p<0.001\nEach predictor from a separate model; 95% CI shown.",
    x        = "\u03b2 coefficient (95% CI)",
    y        = NULL
  ) +
  pub_theme +
  theme(panel.spacing = unit(0.4, "lines"))
```

---

# Forest Plot: HWISE Individual Items (m_a5) {#forest-plot-hwise}

```{r forest-plot-hwise, fig.width = 8, fig.height = 6}
hwise_coefs <- summary_table %>%
  filter(group == "m_a5: HWISE items") %>%
  mutate(
    item_label  = coalesce(hwise_readable[term], term),
    sig         = sig_sym(p.value),
    significant = !is.na(p.value) & p.value < 0.05
  ) %>%
  arrange(estimate)

hwise_coefs$item_label <- factor(hwise_coefs$item_label,
                                 levels = unique(hwise_coefs$item_label))

ggplot(hwise_coefs,
       aes(x = estimate, y = item_label,
           xmin = conf.low, xmax = conf.high,
           colour = significant)) +
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.5) +
  geom_text(aes(label = sig), nudge_y = 0.35, size = 3.5,
            fontface = "bold", show.legend = FALSE) +
  scale_colour_manual(values = c("FALSE" = "grey60", "TRUE" = "#c0392b")) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.12))) +
  labs(
    title    = "Aim 2: HWISE individual items as predictors of PSS (m_a5)",
    subtitle = "Adjusted for age, SES composite, household size, and chronic disease",
    caption  = "Red = p < 0.05  |  Symbols: \u2020p<0.10  *p<0.05  **p<0.01  ***p<0.001\n95% CI shown.  |  Note: Each item tested in a separate model.",
    x        = "\u03b2 coefficient (95% CI)",
    y        = NULL
  ) +
  pub_theme
```

---

# Per-Model Coefficient Tables {#per-model-tables}

Full regression tables (all terms) for each Aim 2 model.

```{r results-tables, results="asis"}
core_label_list <- list(
  D_AGE        ~ "Age (years)",
  SES_SC_Total ~ "SES composite score",
  D_HH_SIZE    ~ "Household size"
)

for (nm in names(fitted_models)) {
  spec <- model_specs[[nm]]
  mod  <- fitted_models[[nm]]
  fit  <- glance(mod)

  cat("\n\n## ", spec$label, " {.unnumbered}\n\n", sep = "")
  cat("**N =**", nobs(mod),
      "| **R² =**", round(fit$r.squared, 3),
      "| **Adj. R² =**", round(fit$adj.r.squared, 3),
      "| **AIC =**", round(fit$AIC, 1), "\n\n")

  print(
    tbl_regression(
      mod,
      label    = core_label_list,
      conf.int = TRUE
    ) %>%
      bold_p(t = 0.05) %>%
      italicize_levels() %>%
      modify_caption(paste("**", spec$label, "**"))
  )
}
```
