---
title: "07. Results Table – Main Regression Models"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(gtsummary)
library(flextable)
library(officer)
```

## Overview

This script recalculates the **HWISE reduced score** (excluding the *angry* and
*shame* items) and fits the eight pre-specified regression models described in
the analysis plan. Results are consolidated into a single publication-ready
table.

| Panel | Outcome | Model type | Models |
|-------|---------|------------|--------|
| A | Perceived Stress Score (PSS, 0–56) | Linear regression | m_a1 – m_a4 |
| B | Water-related emotion (positive / negative) | Logistic regression | m_b1 – m_b4 |

Covariates in all models: **Age** (`D_AGE`), **SES score** (`SES_SC_Total`),
**Household size** (`D_HH_SIZE`). Panel B additionally adjusts for **PSS**
(`PSS_TOTAL`).

---

# Load and Prepare Data

```{r load-data}
MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

clean_num <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES, NA_real_, x_num)
}

df_raw <- read_csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
                   show_col_types = FALSE)

cat("Dataset:", nrow(df_raw), "rows ×", ncol(df_raw), "cols\n")
```

---

# Recalculate HWISE Reduced Score

The **HWISE reduced score** sums the 10 items that remain after removing the
emotion-laden items *HW_ANGRY* and *HW_SHAME*. A participant's reduced score
is set to `NA` if any of the 10 contributing items is missing.

```{r hwise-reduced}
# Items included in the reduced score (original 12 minus angry and shame)
hwise_items_reduced <- c(
  "HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
  "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
  "HW_SLEEP", "HW_NONE"
)

df <- df_raw %>%
  mutate(
    # Clean all HWISE items
    across(all_of(hwise_items_reduced), clean_num),

    # Reduced HWISE score: row sum, NA if any item is missing
    HWISE_REDUCED = rowSums(
      across(all_of(hwise_items_reduced)),
      na.rm = FALSE   # propagate NA if any item missing
    ),

    # ── Outcomes ───────────────────────────────────────────
    PSS_TOTAL = clean_num(PSS_TOTAL),

    EMO_NEG = {
      raw <- clean_num(MX26_EM_HHW_CAT)
      case_when(
        raw == 0 ~ 0L,   # positive emotion
        raw == 1 ~ 1L,   # negative emotion
        TRUE     ~ NA_integer_
      )
    },

    # ── Core covariates ───────────────────────────────────────────
    D_AGE        = clean_num(D_AGE),
    SES_SC_Total = clean_num(SES_SC_Total),
    D_HH_SIZE    = clean_num(D_HH_SIZE),

    # ── Water-insecure neighbourhood (binary) ──────────────────────
    # W_WS_LOC: "WI" = water-insecure, "WS" = water-secure
    WI_NEIGHBORHOOD = factor(
      case_when(
        W_WS_LOC == "WI" ~ "Water-insecure",
        W_WS_LOC == "WS" ~ "Water-secure",
        TRUE             ~ NA_character_
      ),
      levels = c("Water-secure", "Water-insecure")  #water-secure = reference
    ),

    # ── Season ──────────────────────────────────────────────────────────────────
    # 0 = dry season (reference), 1 = rainy season
    SEASON = factor(
      SEASON,
      levels = c(0, 1),
      labels = c("Dry", "Rainy")
    )
  )

cat("HWISE_REDUCED — summary:\n")
print(summary(df$HWISE_REDUCED))
cat("\nN with valid HWISE_REDUCED:", sum(!is.na(df$HWISE_REDUCED)), "\n")
cat("\nWI_NEIGHBORHOOD distribution:\n")
print(table(df$WI_NEIGHBORHOOD, useNA = "always"))
cat("\nSEASON distribution:\n")
print(table(df$SEASON, useNA = "always"))
cat("\nEMO_NEG distribution:\n")
print(table(df$EMO_NEG, useNA = "always"))
```

---

# Model Specifications

```{r model-specs}
# ── Panel A: PSS as outcome (linear regression) ─────────────────────────────
# Covariates: Age + SES + Household size (in every model)

m_a1 <- lm(
  PSS_TOTAL ~ HWISE_REDUCED + D_AGE + SES_SC_Total + D_HH_SIZE,
  data = df, na.action = na.omit
)

m_a2 <- lm(
  PSS_TOTAL ~ HWISE_REDUCED * SEASON + D_AGE + SES_SC_Total + D_HH_SIZE,
  data = df, na.action = na.omit
)

m_a3 <- lm(
  PSS_TOTAL ~ WI_NEIGHBORHOOD + D_AGE + SES_SC_Total + D_HH_SIZE,
  data = df, na.action = na.omit
)

m_a4 <- lm(
  PSS_TOTAL ~ WI_NEIGHBORHOOD * SEASON + D_AGE + SES_SC_Total + D_HH_SIZE,
  data = df, na.action = na.omit
)

# ── Panel B: Emotion type as outcome (logistic regression) ───────────────────
# Covariates: PSS + Age + SES + Household size (in every model)

m_b1 <- glm(
  EMO_NEG ~ HWISE_REDUCED + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE,
  data = df, family = binomial, na.action = na.omit
)

m_b2 <- glm(
  EMO_NEG ~ HWISE_REDUCED * SEASON + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE,
  data = df, family = binomial, na.action = na.omit
)

m_b3 <- glm(
  EMO_NEG ~ WI_NEIGHBORHOOD + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE,
  data = df, family = binomial, na.action = na.omit
)

m_b4 <- glm(
  EMO_NEG ~ WI_NEIGHBORHOOD * SEASON + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE,
  data = df, family = binomial, na.action = na.omit
)

# ── Sample sizes ─────────────────────────────────────────────────────────────
tibble(
  model = c("m_a1", "m_a2", "m_a3", "m_a4",
            "m_b1", "m_b2", "m_b3", "m_b4"),
  n = c(nobs(m_a1), nobs(m_a2), nobs(m_a3), nobs(m_a4),
        nobs(m_b1), nobs(m_b2), nobs(m_b3), nobs(m_b4))
) %>% print()
```

---

# Results Table

Variable labels shared across all models:

```{r var-labels}
shared_labels <- list(
  HWISE_REDUCED   = "HWISE reduced score",
  SEASON          = "Season (Rainy vs. Dry)",
  WI_NEIGHBORHOOD = "Water-insecure neighbourhood",
  PSS_TOTAL       = "PSS total",
  D_AGE           = "Age (years)",
  SES_SC_Total    = "SES score",
  D_HH_SIZE       = "Household size"
)

# Return only labels whose names match variables in the given model
labels_for <- function(model) {
  shared_labels[names(shared_labels) %in% names(model$model)]
}
```

## Panel A — Perceived Stress Score (linear regression)

β coefficients with 95% confidence intervals.

```{r panel-a-table, results="asis"}
# Individual model tables
t_a1 <- tbl_regression(m_a1, label = labels_for(m_a1)) %>%
  modify_header(estimate ~ "**β (95% CI)**") %>%
  bold_p(t = 0.05) %>%
  modify_caption("**m_a1** HWISE reduced")

t_a2 <- tbl_regression(m_a2, label = labels_for(m_a2)) %>%
  modify_header(estimate ~ "**β (95% CI)**") %>%
  bold_p(t = 0.05) %>%
  modify_caption("**m_a2** HWISE reduced × Season")

t_a3 <- tbl_regression(m_a3, label = labels_for(m_a3)) %>%
  modify_header(estimate ~ "**β (95% CI)**") %>%
  bold_p(t = 0.05) %>%
  modify_caption("**m_a3** WI neighbourhood")

t_a4 <- tbl_regression(m_a4, label = labels_for(m_a4)) %>%
  modify_header(estimate ~ "**β (95% CI)**") %>%
  bold_p(t = 0.05) %>%
  modify_caption("**m_a4** WI neighbourhood × Season")

# Merged Panel A table
panel_a <- tbl_merge(
  tbls        = list(t_a1, t_a2, t_a3, t_a4),
  tab_spanner = c("**m_a1**", "**m_a2**", "**m_a3**", "**m_a4**")
) %>%
  modify_caption(
    "**Panel A. Linear regression models — Perceived Stress Score (PSS) as outcome**"
  )

panel_a
```

## Panel B — Water-related emotion (logistic regression)

Odds ratios (OR) with 95% confidence intervals.

```{r panel-b-table, results="asis"}
t_b1 <- tbl_regression(m_b1, exponentiate = TRUE, label = labels_for(m_b1)) %>%
  modify_header(estimate ~ "**OR (95% CI)**") %>%
  bold_p(t = 0.05) %>%
  modify_caption("**m_b1** HWISE reduced")

t_b2 <- tbl_regression(m_b2, exponentiate = TRUE, label = labels_for(m_b2)) %>%
  modify_header(estimate ~ "**OR (95% CI)**") %>%
  bold_p(t = 0.05) %>%
  modify_caption("**m_b2** HWISE reduced × Season")

t_b3 <- tbl_regression(m_b3, exponentiate = TRUE, label = labels_for(m_b3)) %>%
  modify_header(estimate ~ "**OR (95% CI)**") %>%
  bold_p(t = 0.05) %>%
  modify_caption("**m_b3** WI neighbourhood")

t_b4 <- tbl_regression(m_b4, exponentiate = TRUE, label = labels_for(m_b4)) %>%
  modify_header(estimate ~ "**OR (95% CI)**") %>%
  bold_p(t = 0.05) %>%
  modify_caption("**m_b4** WI neighbourhood × Season")

# Merged Panel B table
panel_b <- tbl_merge(
  tbls        = list(t_b1, t_b2, t_b3, t_b4),
  tab_spanner = c("**m_b1**", "**m_b2**", "**m_b3**", "**m_b4**")
) %>%
  modify_caption(
    "**Panel B. Logistic regression models — Water-related emotion (negative vs. positive) as outcome**"
  )

panel_b
```

---

# Model Fit Statistics

```{r fit-stats}
# ── Panel A: R², adjusted R², AIC ────────────────────────────────────────────
panel_a_fit <- map_dfr(
  list(m_a1 = m_a1, m_a2 = m_a2, m_a3 = m_a3, m_a4 = m_a4),
  ~ glance(.x),
  .id = "model"
) %>%
  select(model, n = nobs, r.squared, adj.r.squared, AIC, BIC) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))

cat("Panel A — model fit:\n")
print(panel_a_fit)

# ── Panel B: McFadden R², AIC ─────────────────────────────────────────────────
mcfadden <- function(mod) round(1 - mod$deviance / mod$null.deviance, 3)

panel_b_fit <- tibble(
  model       = c("m_b1", "m_b2", "m_b3", "m_b4"),
  n           = c(nobs(m_b1), nobs(m_b2), nobs(m_b3), nobs(m_b4)),
  mcfadden_r2 = c(mcfadden(m_b1), mcfadden(m_b2),
                  mcfadden(m_b3), mcfadden(m_b4)),
  AIC         = round(c(AIC(m_b1), AIC(m_b2), AIC(m_b3), AIC(m_b4)), 1),
  BIC         = round(c(BIC(m_b1), BIC(m_b2), BIC(m_b3), BIC(m_b4)), 1)
)

cat("\nPanel B — model fit:\n")
print(panel_b_fit)
```

---

# Season-Stratified Estimates

Interaction models (m_a2, m_a4, m_b2, m_b4) are complemented by season-
stratified re-runs to facilitate interpretation.

```{r season-strat}
seasons <- c("Dry", "Rainy")

# ── Panel A: stratified ───────────────────────────────────────────────────────
strat_a <- map_dfr(seasons, function(s) {
  df_s <- filter(df, SEASON == s)
  list(
    hwise  = lm(PSS_TOTAL ~ HWISE_REDUCED + D_AGE + SES_SC_Total + D_HH_SIZE,
                data = df_s, na.action = na.omit),
    wi_nbh = lm(PSS_TOTAL ~ WI_NEIGHBORHOOD + D_AGE + SES_SC_Total + D_HH_SIZE,
                data = df_s, na.action = na.omit)
  ) %>%
    map_dfr(~ tidy(.x, conf.int = TRUE), .id = "predictor") %>%
    filter(!term %in% c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE")) %>%
    mutate(season = s)
}) %>%
  mutate(
    ci_fmt = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_fmt  = case_when(p.value < 0.001 ~ "<0.001",
                       TRUE ~ sprintf("%.3f", p.value))
  ) %>%
  select(season, predictor, term, n = statistic, ci_fmt, p_fmt)

cat("Panel A — season-stratified focal predictors:\n")
print(strat_a, n = Inf)

# ── Panel B: stratified ───────────────────────────────────────────────────────
strat_b <- map_dfr(seasons, function(s) {
  df_s <- filter(df, SEASON == s)
  list(
    hwise  = glm(EMO_NEG ~ HWISE_REDUCED + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE,
                 data = df_s, family = binomial, na.action = na.omit),
    wi_nbh = glm(EMO_NEG ~ WI_NEIGHBORHOOD + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE,
                 data = df_s, family = binomial, na.action = na.omit)
  ) %>%
    map_dfr(~ tidy(.x, exponentiate = TRUE, conf.int = TRUE), .id = "predictor") %>%
    filter(!term %in% c("(Intercept)", "D_AGE", "SES_SC_Total",
                        "D_HH_SIZE", "PSS_TOTAL")) %>%
    mutate(season = s)
}) %>%
  mutate(
    or_fmt = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_fmt  = case_when(p.value < 0.001 ~ "<0.001",
                       TRUE ~ sprintf("%.3f", p.value))
  ) %>%
  select(season, predictor, term, or_fmt, p_fmt)

cat("\nPanel B — season-stratified focal predictors (OR):\n")
print(strat_b, n = Inf)
```

---

# Export to Word

The two gtsummary tables are exported as a single Word document that can be
dropped into the manuscript.

```{r export-word}
# Stack both panels into one Word document
doc <- read_docx()

doc <- doc %>%
  body_add_par("Panel A. Linear regression — Perceived Stress Score (PSS)",
               style = "heading 1") %>%
  body_add_flextable(
    as_flex_table(panel_a) %>% autofit()
  ) %>%
  body_add_break() %>%
  body_add_par("Panel B. Logistic regression — Water-related emotion (negative vs. positive)",
               style = "heading 1") %>%
  body_add_flextable(
    as_flex_table(panel_b) %>% autofit()
  )

out_path <- "./output/07.table_results.docx"
print(doc, target = out_path)
cat("Table exported to:", out_path, "\n")
```
