---
title: "08. Aim 2 Models: Components of Water Insecurity"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(gtsummary)
library(lmtest)
library(flextable)
library(ggplot2)
library(officer)
library(ggstance)
```

## Overview

This script fits the **Aim 2 regression models** testing water insecurity
components as predictors of two outcomes:

- **Part 1**: Perceived Stress Scale total score (`PSS_TOTAL`) — linear regression
- **Part 2**: Water-related emotion valence (`EMO_NEG`) — logistic regression

HWISE individual items and the water worry subscale are included alongside
other water supply predictors in a single merged forest plot per outcome.
Other HWISE composite/subscale scores are excluded.

| Role | Variables |
|------|-----------|
| PSS outcome | `PSS_TOTAL` (continuous, 0–56) |
| EMO outcome | `EMO_NEG` — 0 = positive, 1 = negative (Q26) |
| Core covariates (PSS) | Age (`D_AGE`), SES composite (`SES_SC_Total`), Household size (`D_HH_SIZE`), Chronic disease (`Chronic_dis`) |
| Core covariates (EMO) | As above + `PSS_TOTAL` |
| Augmenting (one per model) | HWISE individual items · Water worry subscale (`HW_WORRY_SUB`) · Hours of water supply/week · Intermittent supply · Season · Neighborhood comparison · Excessive time handling water · Sufficiency rating · Water pressure · Supply frequency · Neighborhood (WS/WI) · Years in neighbourhood (`D_LOC_TIME`) |

> **Multiple testing**: All p-values are adjusted using the
> **Benjamini-Hochberg (BH)** procedure, applied across all Aim 2 models
> within each outcome.

---

# Load and Prepare Data

```{r load-data}
MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

clean_num <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES, NA_real_, x_num)
}

df_raw <- read_csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
                   show_col_types = FALSE)

cat("Dataset:", nrow(df_raw), "rows ×", ncol(df_raw), "cols\n")

# ── Check new column availability ─────────────────────────────────────────────
new_cols <- c("HLTH_CDIS_CAT", "HW_TOTAL", "HW_WORRY_SUB", "HW_MEX_5SUB",
              "MX6_WPRESS", "MX4_WFREQ", "MX4_W_FREQ", "W_WS_LOC",
              "D_LOC_TIME_CLEAN")
cat("\nNew column availability:\n")
for (col in new_cols) {
  cat(sprintf("  %-20s %s\n", col,
              ifelse(col %in% names(df_raw), "FOUND", "*** MISSING ***")))
}
```

```{r prepare-variables}
hwise_items <- c("HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
                 "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
                 "HW_ANGRY", "HW_SLEEP",  "HW_NONE",    "HW_SHAME")

df <- df_raw %>%
  mutate(
    # ── Outcomes ───────────────────────────────────────────────────────────────
    PSS_TOTAL = clean_num(PSS_TOTAL),

    EMO_NEG = {
      raw <- clean_num(MX26_EM_HHW_CAT)
      case_when(
        raw == 0 ~ 0L,
        raw == 1 ~ 1L,
        TRUE     ~ NA_integer_
      )
    },

    # ── Core covariates ────────────────────────────────────────────────────────
    D_AGE        = clean_num(D_AGE),
    SES_SC_Total = clean_num(SES_SC_Total),
    D_HH_SIZE    = clean_num(D_HH_SIZE),
    D_LOC_TIME   = clean_num(D_LOC_TIME_CLEAN),

    # ── HWISE individual items ──────────────────────────────────────────────────
    across(all_of(hwise_items), clean_num),

    # ── Water worry subscale ───────────────────────────────────────────────────
    Water_worry = clean_num(HW_WORRY_SUB),

    # ── Hours of water supply per week ─────────────────────────────────────────
    HRS_PER_WEEK = clean_num(MX5_WFREQ_HRS_PER_WEEK),

    # ── Continuous vs. intermittent supply ─────────────────────────────────────
    # 0 = Continuous (reference), 1 = Intermittent
    CONT_INT = {
      raw     <- as.character(MX3_CONT_INT)
      raw_num <- suppressWarnings(as.numeric(raw))
      factor(case_when(
        raw_num %in% MISSING_CODES                         ~ NA_character_,
        raw_num == 0                                       ~ "Continuous",
        raw_num == 1                                       ~ "Intermittent",
        str_detect(str_to_lower(raw), "contin")            ~ "Continuous",
        str_detect(str_to_lower(raw), "inter")             ~ "Intermittent",
        is.na(raw) | str_trim(raw) == "" | raw == "NA"    ~ NA_character_,
        TRUE                                               ~ raw
      ), levels = c("Continuous", "Intermittent"))
    },

    # ── Season ─────────────────────────────────────────────────────────────────
    # 0 = Season 1 (rainy, reference), 1 = Season 2 (dry)
    SEASON = factor(SEASON, levels = c(0, 1), labels = c("Rainy", "Dry")),

    # ── Neighborhood water comparison ──────────────────────────────────────────
    # 0 = Better (reference), 1 = Same, 2 = Worse
    NB_COMP = factor(
      case_when(
        clean_num(MX28_WQ_COMP_CAT) == 0 ~ "Better",
        clean_num(MX28_WQ_COMP_CAT) == 1 ~ "Same",
        clean_num(MX28_WQ_COMP_CAT) == 2 ~ "Worse",
        TRUE                             ~ NA_character_
      ),
      levels = c("Better", "Same", "Worse")
    ),

    # ── Excessive time handling water ──────────────────────────────────────────
    # 0 = No (reference), 1 = Yes
    TIME_WATER = factor(
      case_when(
        clean_num(MX19_TIME_CAT) == 0 ~ "No",
        clean_num(MX19_TIME_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── Sufficiency rating ─────────────────────────────────────────────────────
    # 1 = not sufficient, 4 = completely sufficient
    SUFF_RATING = clean_num(MX16_SCEN_99),

    # ── Chronic disease ────────────────────────────────────────────────────────
    # 0 = No (reference), 1 = Yes
    Chronic_dis = factor(
      case_when(
        clean_num(HLTH_CDIS_CAT) == 0 ~ "No",
        clean_num(HLTH_CDIS_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── Water pressure ─────────────────────────────────────────────────────────
    W_Pressure = factor(
      case_when(
        str_detect(str_to_lower(as.character(MX6_WPRESS)),
                   "s[i\u00ed]|yes|buena|bien")  ~ "Yes",
        str_detect(str_to_lower(as.character(MX6_WPRESS)),
                   "^no|baja|poca")               ~ "No/Low",
        TRUE                                      ~ NA_character_
      ),
      levels = c("Yes", "No/Low")
    ),

    # ── Supply frequency ───────────────────────────────────────────────────────
    W_Frequency = {
      freq_raw <- if ("MX4_W_FREQ" %in% names(df_raw)) {
        dplyr::coalesce(as.character(MX4_WFREQ), as.character(MX4_W_FREQ))
      } else {
        as.character(MX4_WFREQ)
      }
      raw <- str_to_lower(as.character(freq_raw))
      factor(
        case_when(
          str_detect(raw, "^contin")           ~ "Continuous",
          str_detect(raw, "^diari|every day")  ~ "Daily",
          str_detect(raw, "4 a 6|4-6")         ~ "4\u20136 days/week",
          str_detect(raw, "1 a 3|1-3")         ~ "1\u20133 days/week",
          str_detect(raw, "^intermi")          ~ "Intermittent",
          TRUE                                 ~ NA_character_
        ),
        levels = c("Continuous", "Daily", "4\u20136 days/week",
                   "1\u20133 days/week", "Intermittent")
      )
    },

    # ── Neighborhood water security ────────────────────────────────────────────
    W_WS_LOC_F = if ("W_WS_LOC" %in% names(df_raw)) {
      factor(W_WS_LOC, levels = c("WS", "WI"),
             labels = c("Water Secure", "Water Insecure"))
    } else {
      factor(NA_character_, levels = c("Water Secure", "Water Insecure"))
    }
  )
```

```{r check-missingness}
analytic_vars <- c("PSS_TOTAL", "EMO_NEG", "D_AGE", "SES_SC_Total", "D_HH_SIZE",
                   "SEASON", "Chronic_dis",
                   hwise_items, "Water_worry", "HRS_PER_WEEK", "CONT_INT",
                   "NB_COMP", "TIME_WATER", "SUFF_RATING",
                   "W_Pressure", "W_Frequency", "W_WS_LOC_F", "D_LOC_TIME")

df %>%
  summarise(across(all_of(analytic_vars), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_missing") %>%
  mutate(pct_missing = round(100 * n_missing / nrow(df), 1)) %>%
  arrange(desc(n_missing)) %>%
  print(n = Inf)
```

```{r check-outcomes}
cat("PSS_TOTAL summary:\n")
print(summary(df$PSS_TOTAL))

cat("\nEMO_NEG distribution:\n")
print(table(df$EMO_NEG, useNA = "always"))
cat("\nProportion negative (valid responses):",
    round(mean(df$EMO_NEG == 1, na.rm = TRUE), 3), "\n")
```

---

# Shared Helpers

```{r shared-helpers}
# ── Readable HWISE item labels ─────────────────────────────────────────────────
hwise_readable <- c(
  HW_WORRY   = "Worry about water",
  HW_INTERR  = "Interrupted supply",
  HW_CLOTHES = "Couldn't wash clothes",
  HW_PLANS   = "Plans disrupted",
  HW_FOOD    = "Food preparation affected",
  HW_HANDS   = "Couldn't wash hands",
  HW_BODY    = "Couldn't bathe",
  HW_DRINK   = "Not enough to drink",
  HW_ANGRY   = "Felt angry",
  HW_SLEEP   = "Went to sleep thirsty",
  HW_NONE    = "No water at all",
  HW_SHAME   = "Felt shame"
)

# ── Non-HWISE term labels ──────────────────────────────────────────────────────
term_labels <- c(
  Water_worry                    = "Water worry subscale score",
  HRS_PER_WEEK                   = "Hours of supply per week",
  CONT_INTIntermittent           = "Supply continuity: Intermittent (ref: Continuous)",
  SEASONDry                      = "Season: Dry (ref: Rainy)",
  NB_COMPSame                    = "Neighborhood comparison: Same as CDMX",
  NB_COMPWorse                   = "Neighborhood comparison: Worse than CDMX",
  TIME_WATERYes                  = "Excessive time managing water: Yes (ref: No)",
  SUFF_RATING                    = "Sufficiency rating (1\u20134)",
  `W_PressureNo/Low`             = "Adequate water pressure: No/Low (ref: Yes)",
  W_FrequencyDaily               = "Supply frequency: Daily (ref: Continuous)",
  `W_Frequency4–6 days/week`     = "Supply frequency: 4\u20136 days/wk (ref: Continuous)",
  `W_Frequency1–3 days/week`     = "Supply frequency: 1\u20133 days/wk (ref: Continuous)",
  W_FrequencyIntermittent        = "Supply frequency: Intermittent (ref: Continuous)",
  `W_WS_LOC_FWater Insecure`     = "Neighborhood: Water Insecure (ref: Water Secure)",
  D_LOC_TIME                     = "Years in neighbourhood"
)

# ── Publication plot theme ─────────────────────────────────────────────────────
pub_theme <- theme_classic(base_size = 11) +
  theme(
    plot.title         = element_text(face = "bold", size = 12),
    axis.text.y        = element_text(size = 9),
    axis.text.x        = element_text(size = 9),
    legend.position    = "none",
    strip.text         = element_text(face = "bold", size = 9),
    strip.background   = element_rect(fill = "grey92", colour = NA),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    plot.caption       = element_text(size = 9, colour = "grey40", hjust = 0)
  )

sig_sym <- function(p) {
  dplyr::case_when(
    p < 0.001 ~ "***", p < 0.01 ~ "**", p < 0.05 ~ "*",
    p < 0.1   ~ "\u2020", TRUE ~ ""
  )
}

# Append significance symbol to a formatted p-value string
sig_fmt <- function(p_num, p_str) {
  dplyr::case_when(
    is.na(p_num)  ~ p_str,
    p_num < 0.001 ~ paste0(p_str, " ***"),
    p_num < 0.01  ~ paste0(p_str, " **"),
    p_num < 0.05  ~ paste0(p_str, " *"),
    p_num < 0.1   ~ paste0(p_str, " \u2020"),
    TRUE          ~ p_str
  )
}

# ── Pretty flextable style ─────────────────────────────────────────────────────
pretty_ft_aim2 <- function(ft) {
  ft %>%
    bold(part = "header") %>%
    bold(j = "Model") %>%
    bg(part = "header", bg = "#2c3e50") %>%
    color(part = "header", color = "white") %>%
    theme_booktabs() %>%
    set_table_properties(layout = "autofit", width = 1)
}
```

---

# Part 1: PSS Models (Linear Regression) {#pss-models}

## Model Specifications

```{r pss-model-specs}
core_pss <- PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis

# ── m_a5: Individual HWISE item models (one per item) ─────────────────────────
hwise_specs_pss <- setNames(
  lapply(hwise_items, function(item) {
    list(
      label   = paste0("m_a5 \u2013 ", item),
      group   = "m_a5: HWISE items",
      formula = as.formula(
        paste("PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis +", item)
      )
    )
  }),
  paste0("m_a5_", tolower(hwise_items))
)

model_specs_pss <- c(
  hwise_specs_pss,
  list(
    m_a12 = list(
      label   = "m_a12: Water worry subscale",
      group   = "m_a12: Water worry",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + Water_worry
    ),
    m_a6 = list(
      label   = "m_a6: Hours of water supply/week",
      group   = "m_a6: Hours supply",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HRS_PER_WEEK
    ),
    m_a7 = list(
      label   = "m_a7: Intermittent supply",
      group   = "m_a7: Intermittent",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + CONT_INT
    ),
    m_a8 = list(
      label   = "m_a8: Season (dry vs. rainy)",
      group   = "m_a8: Season",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + SEASON
    ),
    m_a9 = list(
      label   = "m_a9: Neighborhood comparison",
      group   = "m_a9: Neighborhood comparison",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + NB_COMP
    ),
    m_a10 = list(
      label   = "m_a10: Excessive time handling water",
      group   = "m_a10: Time burden",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + TIME_WATER
    ),
    m_a11 = list(
      label   = "m_a11: Sufficiency rating",
      group   = "m_a11: Sufficiency",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + SUFF_RATING
    ),
    m_a16 = list(
      label   = "m_a16: Water pressure",
      group   = "m_a16: Water pressure",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + W_Pressure
    ),
    m_a17 = list(
      label   = "m_a17: Supply frequency",
      group   = "m_a17: Supply frequency",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + W_Frequency
    ),
    m_a18 = list(
      label   = "m_a18: Neighborhood (WS/WI)",
      group   = "m_a18: Neighborhood",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + W_WS_LOC_F
    ),
    m_a_loc = list(
      label   = "m_a_loc: Years in neighbourhood",
      group   = "m_a_loc: Years in neighbourhood",
      formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + D_LOC_TIME
    )
  )
)

cat("Total PSS models registered:", length(model_specs_pss), "\n")
cat("Model IDs:\n")
cat(paste(" -", names(model_specs_pss)), sep = "\n")
```

## Fit PSS Models

```{r pss-fit-models}
fitted_pss <- lapply(model_specs_pss, function(spec) {
  lm(spec$formula, data = df, na.action = na.omit)
})

tibble(
  model_id = names(model_specs_pss),
  label    = map_chr(model_specs_pss, "label"),
  n        = map_int(fitted_pss, nobs)
) %>% print(n = Inf)
```

## Summary Table

```{r pss-summary-table}
core_terms_pss <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE", "Chronic_disYes")

summary_raw_pss <- map_dfr(names(fitted_pss), function(nm) {
  mod  <- fitted_pss[[nm]]
  spec <- model_specs_pss[[nm]]
  fit  <- glance(mod)

  rows_used <- as.integer(rownames(model.frame(mod)))
  m0_sub    <- lm(core_pss, data = df[rows_used, ])
  lrt_tab   <- anova(m0_sub, mod)
  f_stat    <- lrt_tab$F[2]
  f_df1     <- lrt_tab$Df[2]
  f_df2     <- lrt_tab$Res.Df[2]
  f_p       <- lrt_tab$`Pr(>F)`[2]

  focal_coefs <- tidy(mod, conf.int = TRUE) %>%
    filter(!term %in% core_terms_pss)

  if (nrow(focal_coefs) == 0) return(NULL)

  focal_coefs %>%
    mutate(
      model_id     = nm,
      label        = spec$label,
      group        = spec$group,
      n            = nobs(mod),
      adj_R2       = round(fit$adj.r.squared, 3),
      delta_adj_R2 = round(fit$adj.r.squared - glance(m0_sub)$adj.r.squared, 3),
      AIC          = round(fit$AIC, 1),
      F_stat       = round(f_stat, 2),
      F_df1        = as.integer(f_df1),
      F_df2        = as.integer(f_df2),
      F_p          = round(f_p, 4)
    )
})

summary_pss <- summary_raw_pss %>%
  mutate(
    beta_ci   = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_coef    = case_when(
      is.na(p.value)  ~ "\u2014",
      p.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", p.value)
    ),
    p_F_fmt   = case_when(
      is.na(F_p)  ~ "\u2014",
      F_p < 0.001 ~ "<0.001",
      TRUE        ~ sprintf("%.3f", F_p)
    ),
    p_adj_fmt = p_F_fmt
  )

tbl_sum_pss <- summary_pss %>%
  arrange(AIC) %>%
  mutate(
    p_coef_s  = sig_fmt(p.value,  p_coef),
    p_F_s     = sig_fmt(F_p,      p_F_fmt),
    p_adj_s   = sig_fmt(F_p,      p_adj_fmt)
  )

b_pc_s  <- which(!is.na(tbl_sum_pss$p.value) & tbl_sum_pss$p.value < 0.05)
b_pF_s  <- which(!is.na(tbl_sum_pss$F_p)     & tbl_sum_pss$F_p     < 0.05)
b_pa_s  <- which(!is.na(tbl_sum_pss$F_p) & tbl_sum_pss$F_p < 0.05)

tbl_sum_pss %>%
  select(label, term, n, adj_R2, delta_adj_R2, AIC, beta_ci,
         p_coef_s, F_stat, p_F_s, p_adj_s) %>%
  flextable() %>%
  set_header_labels(
    label        = "Model",
    term         = "Predictor term",
    n            = "N",
    adj_R2       = "Adj. R\u00b2",
    delta_adj_R2 = "\u0394 Adj. R\u00b2",
    AIC          = "AIC",
    beta_ci      = "\u03b2 [95% CI]",
    p_coef_s     = "p (coef.)",
    F_stat       = "F",
    p_F_s        = "p (F-test)",
    p_adj_s      = "p (F-test) "
  ) %>%
  bold(i = b_pc_s, j = "p_coef_s") %>%
  bold(i = b_pF_s, j = "p_F_s") %>%
  bold(i = b_pa_s, j = "p_adj_s") %>%
  set_table_properties(layout = "autofit", width = 1)
```

## All PSS Models: Publication Table {#pss-all-models}

```{r pss-all-models, fig.width=10}
readable_terms_pss <- c(
  Water_worry                = "Water worry subscale score",
  HRS_PER_WEEK               = "Hours of supply / week",
  CONT_INTIntermittent       = "Intermittent vs. continuous supply",
  SEASONDry                  = "Dry vs. rainy season",
  NB_COMPSame                = "Neighborhood: same as CDMX",
  NB_COMPWorse               = "Neighborhood: worse than CDMX",
  TIME_WATERYes              = "Excessive time managing water",
  SUFF_RATING                = "Sufficiency rating (1\u20134)",
  `W_PressureNo/Low`         = "Water pressure: No/Low",
  W_FrequencyDaily           = "Supply frequency: Daily",
  `W_Frequency4–6 days/week` = "Supply frequency: 4\u20136 days/wk",
  `W_Frequency1–3 days/week` = "Supply frequency: 1\u20133 days/wk",
  W_FrequencyIntermittent    = "Supply frequency: Intermittent",
  `W_WS_LOC_FWater Insecure` = "Neighborhood: Water Insecure",
  D_LOC_TIME                 = "Years in neighbourhood"
)
# Add HWISE item labels
readable_terms_pss <- c(readable_terms_pss, hwise_readable)

all_ids_pss <- summary_pss %>%
  distinct(model_id, AIC) %>%
  arrange(AIC) %>%
  mutate(rank = row_number())

top5_pss <- summary_pss %>%
  left_join(select(all_ids_pss, model_id, rank), by = "model_id") %>%
  arrange(rank, estimate) %>%
  transmute(
    rank        = rank,
    Model       = label,
    Predictor   = coalesce(readable_terms_pss[term], term),
    N           = n,
    beta_ci     = beta_ci,
    adj_r2      = adj_R2,
    delta_r2    = delta_adj_R2,
    AIC         = AIC,
    p_coef      = sig_fmt(p.value,  p_coef),
    p_adj       = sig_fmt(F_p,      p_adj_fmt),
    p_coef_num  = p.value,
    p_adj_num   = F_p
  )

b_pc_p  <- which(!is.na(top5_pss$p_coef_num) & top5_pss$p_coef_num < 0.05)
b_pa_p  <- which(!is.na(top5_pss$p_adj_num)  & top5_pss$p_adj_num  < 0.05)

ft_pss <- top5_pss %>%
  select(-p_coef_num, -p_adj_num) %>%
  flextable() %>%
  set_header_labels(
    rank      = "#",
    Model     = "Model",
    Predictor = "Focal predictor",
    N         = "N",
    beta_ci   = "\u03b2 [95% CI]",
    adj_r2    = "Adj. R\u00b2",
    delta_r2  = "\u0394 Adj. R\u00b2",
    AIC       = "AIC",
    p_coef    = "p (coef.)",
    p_adj     = "p (F-test)"
  ) %>%
  bold(i = b_pc_p, j = "p_coef") %>%
  bold(i = b_pa_p, j = "p_adj") %>%
  merge_v(j = c("rank", "Model", "N", "adj_r2", "delta_r2", "AIC", "p_adj")) %>%
  valign(j = c("rank", "Model", "N", "adj_r2", "delta_r2", "AIC", "p_adj"),
         valign = "top", part = "body") %>%
  set_caption(paste0(
    "All Aim 2 PSS models ranked by AIC. ",
    "\u03b2 = unstandardised regression coefficient; 95% CI = confidence interval. ",
    "Adj. R\u00b2 = adjusted R-squared of the full model; ",
    "\u0394 Adj. R\u00b2 = change over core-covariate-only model. ",
    "All models adjusted for age, SES, household size, and chronic disease. ",
    "p (F-test) = F-test p-value for the focal predictor."
  )) %>%
  pretty_ft_aim2()

ft_pss
```

## Forest Plot: All Predictors (HWISE items + Water supply) {#pss-forest}

```{r pss-forest, fig.width = 10, fig.height = 9}
group_labels_pss <- c(
  "m_a5: HWISE items"                    = "HWISE items",
  "m_a12: Water worry"                   = "HWISE items",
  "m_a6: Hours supply"                   = "Water supply",
  "m_a7: Intermittent"                   = "Water supply",
  "m_a8: Season"                         = "Water supply",
  "m_a9: Neighborhood comparison"        = "Water supply",
  "m_a10: Time burden"                   = "Water supply",
  "m_a11: Sufficiency"                   = "Water supply",
  "m_a16: Water pressure"                = "Water supply",
  "m_a17: Supply frequency"              = "Water supply",
  "m_a18: Neighborhood"                  = "Water supply",
  "m_a_loc: Years in neighbourhood"      = "Water supply"
)

fp_order_pss <- c(
  "Water worry subscale score"                           =  1,
  "Hours of supply per week"                             =  2,
  "Supply continuity: Intermittent (ref: Continuous)"    =  3,
  "Supply frequency: Daily (ref: Continuous)"            =  4,
  "Supply frequency: 4\u20136 days/wk (ref: Continuous)" =  5,
  "Supply frequency: 1\u20133 days/wk (ref: Continuous)" =  6,
  "Supply frequency: Intermittent (ref: Continuous)"     =  7,
  "Season: Dry (ref: Rainy)"                             =  8,
  "Neighborhood comparison: Same as CDMX"                =  9,
  "Neighborhood comparison: Worse than CDMX"             = 10,
  "Excessive time managing water: Yes (ref: No)"         = 11,
  "Sufficiency rating (1\u20134)"                        = 12,
  "Adequate water pressure: No/Low (ref: Yes)"           = 13,
  "Neighborhood: Water Insecure (ref: Water Secure)"     = 14,
  "Years in neighbourhood"                               = 15
)

coefs_pss <- summary_pss %>%
  mutate(
    term_label  = coalesce(term_labels[term], hwise_readable[term], term),
    panel_group = coalesce(group_labels_pss[group], group),
    panel_group = factor(panel_group,
                         levels = c("HWISE items", "Water supply")),
    sig         = sig_sym(p.value),
    fp_order    = coalesce(fp_order_pss[term_label], estimate + 100L)
  ) %>%
  arrange(panel_group, fp_order)

coefs_pss$term_label <- factor(coefs_pss$term_label,
                               levels = unique(coefs_pss$term_label))

x_all_pss <- c(coefs_pss$conf.low, coefs_pss$conf.high)
x_lim_pss <- c(min(x_all_pss, na.rm = TRUE), max(x_all_pss, na.rm = TRUE))

p_pss <- ggplot(coefs_pss,
       aes(x = estimate, y = term_label,
           xmin = conf.low, xmax = conf.high,
           colour = p.value < 0.05)) +
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.5) +
  geom_text(aes(label = sig), nudge_y = 0.35, size = 3.5,
            fontface = "bold", show.legend = FALSE) +
  geom_text(aes(x = Inf, label = paste0("n=", n)),
            hjust = 1.05, size = 2.8, colour = "grey40",
            show.legend = FALSE) +
  facet_grid(panel_group ~ ., scales = "free_y", space = "free_y") +
  scale_colour_manual(values = c("FALSE" = "grey60", "TRUE" = "#c0392b")) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.12))) +
  coord_cartesian(xlim = x_lim_pss) +
  labs(
    title    = "Adjusted linear regression \u2014 Perceived Stress Score (\u03b2)",
    subtitle = "Adjusted for age, SES, household size, and chronic disease",
    caption  = "Red = p < 0.05  |  Symbols: \u2020p<0.10  *p<0.05  **p<0.01  ***p<0.001\nEach predictor from a separate model; 95% CI shown.",
    x        = "\u03b2 coefficient (95% CI)",
    y        = NULL
  ) +
  pub_theme +
  theme(panel.spacing = unit(0.4, "lines"))

p_pss
ggsave("./output/figures/08.aim2_pss_forest.png", p_pss,
       width = 10, height = 9, dpi = 300, bg = "white")
cat("Saved: ./output/figures/08.aim2_pss_forest.png\n")
```

## Per-Model Coefficient Tables (PSS) {#pss-coef-tables}

```{r pss-coef-tables, results="asis"}
core_label_list_pss <- list(
  D_AGE        ~ "Age (years)",
  SES_SC_Total ~ "SES composite score",
  D_HH_SIZE    ~ "Household size",
  Chronic_dis  ~ "Chronic disease"
)

for (nm in names(fitted_pss)) {
  spec <- model_specs_pss[[nm]]
  mod  <- fitted_pss[[nm]]
  fit  <- glance(mod)

  cat("\n\n## ", spec$label, " {.unnumbered}\n\n", sep = "")
  cat("**N =**", nobs(mod),
      "| **R\u00b2 =**", round(fit$r.squared, 3),
      "| **Adj. R\u00b2 =**", round(fit$adj.r.squared, 3),
      "| **AIC =**", round(fit$AIC, 1), "\n\n")

  print(
    tbl_regression(
      mod,
      label    = core_label_list_pss,
      conf.int = TRUE
    ) %>%
      bold_p(t = 0.05) %>%
      italicize_levels() %>%
      modify_caption(paste("**", spec$label, "**"))
  )
}
```

---

# Part 2: Emotion Models (Logistic Regression) {#emo-models}

## Model Specifications

```{r emo-model-specs}
core_emo <- EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis

# ── m_b5: Individual HWISE item models (one per item) ─────────────────────────
hwise_specs_emo <- setNames(
  lapply(hwise_items, function(item) {
    list(
      label   = paste0("m_b5 \u2013 ", item),
      group   = "m_b5: HWISE items",
      formula = as.formula(
        paste("EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis +", item)
      )
    )
  }),
  paste0("m_b5_", tolower(hwise_items))
)

model_specs_emo <- c(
  hwise_specs_emo,
  list(
    m_b12 = list(
      label   = "m_b12: Water worry subscale",
      group   = "m_b12: Water worry",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + Water_worry
    ),
    m_b6 = list(
      label   = "m_b6: Hours of water supply/week",
      group   = "m_b6: Hours supply",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HRS_PER_WEEK
    ),
    m_b7 = list(
      label   = "m_b7: Intermittent supply",
      group   = "m_b7: Intermittent",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + CONT_INT
    ),
    m_b8 = list(
      label   = "m_b8: Season (dry vs. rainy)",
      group   = "m_b8: Season",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + SEASON
    ),
    m_b9 = list(
      label   = "m_b9: Neighborhood comparison",
      group   = "m_b9: Neighborhood comparison",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + NB_COMP
    ),
    m_b10 = list(
      label   = "m_b10: Excessive time handling water",
      group   = "m_b10: Time burden",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + TIME_WATER
    ),
    m_b11 = list(
      label   = "m_b11: Sufficiency rating",
      group   = "m_b11: Sufficiency",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + SUFF_RATING
    ),
    m_b16 = list(
      label   = "m_b16: Water pressure",
      group   = "m_b16: Water pressure",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + W_Pressure
    ),
    m_b17 = list(
      label   = "m_b17: Supply frequency",
      group   = "m_b17: Supply frequency",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + W_Frequency
    ),
    m_b18 = list(
      label   = "m_b18: Neighborhood (WS/WI)",
      group   = "m_b18: Neighborhood",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + W_WS_LOC_F
    ),
    m_b_loc = list(
      label   = "m_b_loc: Years in neighbourhood",
      group   = "m_b_loc: Years in neighbourhood",
      formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + D_LOC_TIME
    )
  )
)

cat("Total EMO models registered:", length(model_specs_emo), "\n")
cat("Model IDs:\n")
cat(paste(" -", names(model_specs_emo)), sep = "\n")
```

## Fit EMO Models

```{r emo-fit-models}
fitted_emo <- lapply(model_specs_emo, function(spec) {
  glm(spec$formula, data = df, family = binomial, na.action = na.omit)
})

tibble(
  model_id  = names(model_specs_emo),
  label     = map_chr(model_specs_emo, "label"),
  n         = map_int(fitted_emo, nobs),
  converged = map_lgl(fitted_emo, ~ .x$converged)
) %>% print(n = Inf)
```

## Summary Table

```{r emo-summary-table}
mcfadden_r2 <- function(mod) round(1 - mod$deviance / mod$null.deviance, 3)

core_terms_emo <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE",
                    "PSS_TOTAL", "Chronic_disYes")

summary_raw_emo <- map_dfr(names(fitted_emo), function(nm) {
  mod  <- fitted_emo[[nm]]
  spec <- model_specs_emo[[nm]]

  rows_used  <- as.integer(rownames(model.frame(mod)))
  m0_sub     <- glm(core_emo, data = df[rows_used, ],
                    family = binomial, na.action = na.omit)
  lrt_result <- lrtest(m0_sub, mod)
  fit        <- glance(mod)

  focal_coefs <- tidy(mod, exponentiate = TRUE, conf.int = TRUE) %>%
    filter(!term %in% core_terms_emo)

  if (nrow(focal_coefs) == 0) return(NULL)

  focal_coefs %>%
    mutate(
      model_id = nm,
      label    = spec$label,
      group    = spec$group,
      n        = nobs(mod),
      mcfadden = mcfadden_r2(mod),
      delta_r2 = round(mcfadden_r2(mod) - mcfadden_r2(m0_sub), 3),
      AIC      = round(fit$AIC, 1),
      LRT_chi2 = round(lrt_result$Chisq[2], 2),
      LRT_df   = as.integer(lrt_result$Df[2]),
      LRT_p    = round(lrt_result$`Pr(>Chisq)`[2], 4)
    )
})

summary_emo <- summary_raw_emo %>%
  mutate(
    or_ci     = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_coef    = case_when(
      is.na(p.value)  ~ "\u2014",
      p.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", p.value)
    ),
    p_lrt_fmt = case_when(
      is.na(LRT_p)  ~ "\u2014",
      LRT_p < 0.001 ~ "<0.001",
      TRUE          ~ sprintf("%.3f", LRT_p)
    ),
    p_adj_fmt = p_lrt_fmt
  )

tbl_sum_emo <- summary_emo %>%
  arrange(AIC) %>%
  mutate(
    p_coef_s  = sig_fmt(p.value,   p_coef),
    p_lrt_s   = sig_fmt(LRT_p,     p_lrt_fmt),
    p_adj_s   = sig_fmt(LRT_p,     p_adj_fmt)
  )

b_pc_e  <- which(!is.na(tbl_sum_emo$p.value)   & tbl_sum_emo$p.value   < 0.05)
b_pl_e  <- which(!is.na(tbl_sum_emo$LRT_p)     & tbl_sum_emo$LRT_p     < 0.05)
b_pa_e  <- which(!is.na(tbl_sum_emo$LRT_p) & tbl_sum_emo$LRT_p < 0.05)

tbl_sum_emo %>%
  select(label, term, n, mcfadden, delta_r2, AIC, or_ci,
         p_coef_s, LRT_chi2, p_lrt_s, p_adj_s) %>%
  flextable() %>%
  set_header_labels(
    label     = "Model",
    term      = "Predictor term",
    n         = "N",
    mcfadden  = "McFadden R\u00b2",
    delta_r2  = "\u0394 McFadden R\u00b2",
    AIC       = "AIC",
    or_ci     = "OR [95% CI]",
    p_coef_s  = "p (coef.)",
    LRT_chi2  = "LRT \u03c7\u00b2",
    p_lrt_s   = "p (LRT)",
    p_adj_s   = "p (LRT) "
  ) %>%
  bold(i = b_pc_e, j = "p_coef_s") %>%
  bold(i = b_pl_e, j = "p_lrt_s") %>%
  bold(i = b_pa_e, j = "p_adj_s") %>%
  set_table_properties(layout = "autofit", width = 1)
```

## All EMO Models: Publication Table {#emo-all-models}

```{r emo-all-models}
readable_terms_emo <- c(
  Water_worry                = "Water worry subscale score",
  HRS_PER_WEEK               = "Hours of supply / week",
  CONT_INTIntermittent       = "Intermittent vs. continuous supply",
  SEASONDry                  = "Dry vs. rainy season",
  NB_COMPSame                = "Neighborhood: same as CDMX",
  NB_COMPWorse               = "Neighborhood: worse than CDMX",
  TIME_WATERYes              = "Excessive time managing water",
  SUFF_RATING                = "Sufficiency rating (1\u20134)",
  `W_PressureNo/Low`         = "Water pressure: No/Low",
  W_FrequencyDaily           = "Supply frequency: Daily",
  `W_Frequency4–6 days/week` = "Supply frequency: 4\u20136 days/wk",
  `W_Frequency1–3 days/week` = "Supply frequency: 1\u20133 days/wk",
  W_FrequencyIntermittent    = "Supply frequency: Intermittent",
  `W_WS_LOC_FWater Insecure` = "Neighborhood: Water Insecure",
  D_LOC_TIME                 = "Years in neighbourhood"
)
readable_terms_emo <- c(readable_terms_emo, hwise_readable)

all_ids_emo <- summary_emo %>%
  distinct(model_id, AIC) %>%
  arrange(AIC) %>%
  mutate(rank = row_number())

top5_emo <- summary_emo %>%
  left_join(select(all_ids_emo, model_id, rank), by = "model_id") %>%
  arrange(rank, estimate) %>%
  transmute(
    rank        = rank,
    Model       = label,
    Predictor   = coalesce(readable_terms_emo[term], term),
    N           = n,
    or_ci       = or_ci,
    mcfadden    = mcfadden,
    delta_r2    = delta_r2,
    AIC         = AIC,
    p_coef      = sig_fmt(p.value,   p_coef),
    p_adj       = sig_fmt(LRT_p,     p_adj_fmt),
    p_coef_num  = p.value,
    p_adj_num   = LRT_p
  )

b_pc_ep <- which(!is.na(top5_emo$p_coef_num) & top5_emo$p_coef_num < 0.05)
b_pa_ep <- which(!is.na(top5_emo$p_adj_num)  & top5_emo$p_adj_num  < 0.05)

ft_emo <- top5_emo %>%
  select(-p_coef_num, -p_adj_num) %>%
  flextable() %>%
  set_header_labels(
    rank      = "#",
    Model     = "Model",
    Predictor = "Focal predictor",
    N         = "N",
    or_ci     = "OR [95% CI]",
    mcfadden  = "McFadden R\u00b2",
    delta_r2  = "\u0394 McFadden R\u00b2",
    AIC       = "AIC",
    p_coef    = "p (coef.)",
    p_adj     = "p (LRT)"
  ) %>%
  bold(i = b_pc_ep, j = "p_coef") %>%
  bold(i = b_pa_ep, j = "p_adj") %>%
  merge_v(j = c("rank", "Model", "N", "mcfadden", "delta_r2", "AIC", "p_adj")) %>%
  valign(j = c("rank", "Model", "N", "mcfadden", "delta_r2", "AIC", "p_adj"),
         valign = "top", part = "body") %>%
  set_caption(paste0(
    "All Aim 2 emotion models ranked by AIC. ",
    "OR = odds ratio; 95% CI = confidence interval. ",
    "McFadden R\u00b2 = pseudo-R\u00b2 of the full model; ",
    "\u0394 McFadden R\u00b2 = change over core-covariate-only model. ",
    "All models adjusted for age, SES, household size, ",
    "perceived stress (PSS), and chronic disease. ",
    "p (LRT) = likelihood ratio test p-value for the focal predictor."
  )) %>%
  pretty_ft_aim2()

ft_emo

out_combined <- "./output/tables/08.Tables_Aim2_Models.docx"
doc_combined <- read_docx() %>%
  body_add_par("Aim 2 \u2014 All PSS Models (Linear Regression)", style = "heading 1") %>%
  body_add_flextable(ft_pss) %>%
  body_add_break() %>%
  body_add_par("Aim 2 \u2014 All Emotion Models (Logistic Regression)", style = "heading 1") %>%
  body_add_flextable(ft_emo)
print(doc_combined, target = out_combined)
cat("Saved:", out_combined, "\n")
```

## Forest Plot: All Predictors (HWISE items + Water supply) {#emo-forest}

```{r emo-forest, fig.width = 10, fig.height = 9}
group_labels_emo <- c(
  "m_b5: HWISE items"                    = "HWISE items",
  "m_b12: Water worry"                   = "HWISE items",
  "m_b6: Hours supply"                   = "Water supply",
  "m_b7: Intermittent"                   = "Water supply",
  "m_b8: Season"                         = "Water supply",
  "m_b9: Neighborhood comparison"        = "Water supply",
  "m_b10: Time burden"                   = "Water supply",
  "m_b11: Sufficiency"                   = "Water supply",
  "m_b16: Water pressure"                = "Water supply",
  "m_b17: Supply frequency"              = "Water supply",
  "m_b18: Neighborhood"                  = "Water supply",
  "m_b_loc: Years in neighbourhood"      = "Water supply"
)

fp_order_emo <- c(
  "Hours of supply per week"                              =  1,
  "Supply continuity: Intermittent (ref: Continuous)"    =  2,
  "Supply frequency: Daily (ref: Continuous)"            =  3,
  "Supply frequency: 4\u20136 days/wk (ref: Continuous)" =  4,
  "Supply frequency: 1\u20133 days/wk (ref: Continuous)" =  5,
  "Supply frequency: Intermittent (ref: Continuous)"     =  6,
  "Season: Dry (ref: Rainy)"                             =  7,
  "Neighborhood comparison: Same as CDMX"                =  8,
  "Neighborhood comparison: Worse than CDMX"             =  9,
  "Excessive time managing water: Yes (ref: No)"         = 10,
  "Sufficiency rating (1\u20134)"                        = 11,
  "Adequate water pressure: No/Low (ref: Yes)"           = 12,
  "Neighborhood: Water Insecure (ref: Water Secure)"     = 13,
  "Years in neighbourhood"                               = 14
)

coefs_emo <- summary_emo %>%
  mutate(
    term_label  = coalesce(term_labels[term], hwise_readable[term], term),
    panel_group = coalesce(group_labels_emo[group], group),
    panel_group = factor(panel_group,
                         levels = c("HWISE items", "Water supply")),
    sig         = sig_sym(p.value),
    fp_order    = coalesce(fp_order_emo[term_label], estimate + 100)
  ) %>%
  arrange(panel_group, fp_order)

coefs_emo$term_label <- factor(coefs_emo$term_label,
                               levels = unique(coefs_emo$term_label))

x_all_emo <- c(coefs_emo$conf.low, coefs_emo$conf.high)
x_lim_emo <- c(min(x_all_emo, na.rm = TRUE), max(x_all_emo, na.rm = TRUE))

p_emo <- ggplot(coefs_emo,
       aes(x = estimate, y = term_label,
           xmin = conf.low, xmax = conf.high,
           colour = p.value < 0.05)) +
  geom_vline(xintercept = 1, linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.5) +
  geom_text(aes(label = sig), nudge_y = 0.35, size = 3.5,
            fontface = "bold", show.legend = FALSE) +
  geom_text(aes(x = Inf, label = paste0("n=", n)),
            hjust = 1.05, size = 2.8, colour = "grey40",
            show.legend = FALSE) +
  facet_grid(panel_group ~ ., scales = "free_y", space = "free_y") +
  scale_colour_manual(values = c("FALSE" = "grey60", "TRUE" = "#2471a3")) +
  scale_x_log10(expand = expansion(mult = c(0.05, 0.12))) +
  coord_cartesian(xlim = x_lim_emo) +
  labs(
    title    = "Adjusted logistic regression \u2014 Negative emotional response (OR)",
    subtitle = paste0("Adjusted for age, SES, household size,",
                      " perceived stress (PSS), and chronic disease"),
    caption  = "Blue = p < 0.05  |  Symbols: \u2020p<0.10  *p<0.05  **p<0.01  ***p<0.001\nLog scale  |  Each predictor from a separate model; 95% CI shown.",
    x        = "Odds Ratio (95% CI, log scale)",
    y        = NULL
  ) +
  pub_theme +
  theme(panel.spacing = unit(0.4, "lines"))

p_emo
ggsave("./output/figures/08.aim2_emo_forest.png", p_emo,
       width = 10, height = 9, dpi = 300, bg = "white")
cat("Saved: ./output/figures/08.aim2_emo_forest.png\n")
```

## Per-Model Coefficient Tables (EMO) {#emo-coef-tables}

```{r emo-coef-tables, results="asis"}
core_label_list_emo <- list(
  D_AGE        ~ "Age (years)",
  SES_SC_Total ~ "SES composite score",
  D_HH_SIZE    ~ "Household size",
  PSS_TOTAL    ~ "PSS Total (perceived stress)",
  Chronic_dis  ~ "Chronic disease"
)

for (nm in names(fitted_emo)) {
  spec <- model_specs_emo[[nm]]
  mod  <- fitted_emo[[nm]]
  fit  <- glance(mod)

  cat("\n\n## ", spec$label, " {.unnumbered}\n\n", sep = "")
  cat("**N =**", nobs(mod),
      "| **McFadden R\u00b2 =**", mcfadden_r2(mod),
      "| **AIC =**", round(fit$AIC, 1), "\n\n")

  print(
    tbl_regression(
      mod,
      exponentiate = TRUE,
      label        = core_label_list_emo,
      conf.int     = TRUE
    ) %>%
      bold_p(t = 0.05) %>%
      italicize_levels() %>%
      modify_caption(paste("**", spec$label, "**  (OR, 95% CI)"))
  )
}
```
