---
title: "08. Multivariate Regression Models"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(flextable)
library(officer)
library(ggplot2)
library(forcats)
library(ggrepel)
library(patchwork)
library(psych)
library(FactoMineR)
library(factoextra)

# Create output directories if needed
dir.create("./output/figures", recursive = TRUE, showWarnings = FALSE)
dir.create("./output/tables",  recursive = TRUE, showWarnings = FALSE)

# Covariates to exclude from coefficient plots and compact tables
COVARIATES <- c("D_AGE", "SES_TOTAL", "D_HH_SIZE", "CHRONIC_DISYes", "PSS_TOTAL")
# Note: SES_TOTAL = SES_SC_Total (composite SES score); EDU_SCORE used in sensitivity only

# Significance symbol helper
sig_symbol <- function(p) {
  case_when(
    p < 0.001 ~ "***",
    p < 0.01  ~ "**",
    p < 0.05  ~ "*",
    p < 0.1   ~ "†",
    TRUE       ~ ""
  )
}
```

## Overview

This script recalculates the **HWISE reduced score** (excluding the *angry* and
*shame* items) and fits the eight pre-specified regression models described in
the analysis plan. Results are consolidated into a single publication-ready
table.

| Panel | Outcome | Model type | Models |
|-------|---------|------------|--------|
| A | Perceived Stress Score (PSS, 0–56) | Linear regression | m_a1 – m_a4 |
| B | Water-related emotion (positive / negative) | Logistic regression | m_b1 – m_b4 |

Covariates in all models: **Age** (`D_AGE`), **SES score** (`SES_TOTAL`),
**Household size** (`D_HH_SIZE`), **Chronic disease** (`CHRONIC_DIS`). Panel B additionally adjusts for **PSS**
(`PSS_TOTAL`).

HWISE_cat was defined using the cutoffs proposed by Young et al. (2021) 
for the full 12-item HWISE scale, which are based on the original validation 
sample from 8 countries. However, since we are using a reduced 10-item version
of HWISE, these cutoffs may not be directly applicable. For the purposes of 
this analysis, we will use the same cutoffs as a starting point,
but we will also explore the distribution of the reduced HWISE scores in 
our sample to see if they align with these categories or if 
adjustments are needed.

---

# Load and Prepare Data

```{r load-data}
MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

clean_num <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES, NA_real_, x_num)
}

df_raw <- read_csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
                   show_col_types = FALSE)

cat("Dataset:", nrow(df_raw), "rows ×", ncol(df_raw), "cols\n")
```

---

# Recalculate HWISE Reduced Score

The **HWISE reduced score** sums the 10 items that remain after removing the
emotion-laden items *HW_ANGRY* and *HW_SHAME*. A participant's reduced score
is set to `NA` if any of the 10 contributing items is missing.

```{r hwise-reduced}
# Items included in the reduced score (original 12 minus angry and shame)
hwise_items_reduced <- c(
  "HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
  "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
  "HW_SLEEP", "HW_NONE"
)

hwise_items_full <- c(hwise_items_reduced, "HW_ANGRY", "HW_SHAME")

df <- df_raw %>%
  mutate(
    # Clean all HWISE items (reduced + emotion items)
    across(all_of(hwise_items_full), clean_num),

    # Reduced HWISE score: row sum, NA if any item is missing
    HWISE_REDUCED = rowSums(
      across(all_of(hwise_items_reduced)),
      na.rm = FALSE   # propagate NA if any item missing
    ),

    # Full HWISE score (all 12 items)
    HWISE_FULL = rowSums(
      across(all_of(hwise_items_full)),
      na.rm = FALSE
    ),

    # ── Outcomes ───────────────────────────────────────────
    PSS_TOTAL = clean_num(PSS_TOTAL),

    EMO_NEG = {
      raw <- clean_num(MX26_EM_HHW_CAT)
      case_when(
        raw == 0 ~ 0L,   # positive emotion
        raw == 1 ~ 1L,   # negative emotion
        TRUE     ~ NA_integer_
      )
    },

    # ── Core covariates ───────────────────────────────────────────
    D_AGE       = clean_num(D_AGE),
    SES_TOTAL   = clean_num(SES_SC_Total),        # composite SES score
    D_HH_SIZE   = clean_num(D_HH_SIZE),
    CHRONIC_DIS = factor(clean_num(HLTH_CDIS_CAT),
                         levels = c(0, 1), labels = c("No", "Yes")),

    # ── Water-insecure neighbourhood (binary) ──────────────────────
    # W_WS_LOC: "WI" = water-insecure, "WS" = water-secure
    WI_NEIGHBORHOOD = factor(
      case_when(
        W_WS_LOC == "WI" ~ "Water-insecure",
        W_WS_LOC == "WS" ~ "Water-secure",
        TRUE             ~ NA_character_
      ),
      levels = c("Water-secure", "Water-insecure")  #water-secure = reference
    ),

    # ── Season ──────────────────────────────────────────────────────────────────
    # 0 = rainy season (reference), 1 = dry season
    SEASON = factor(
      SEASON,
      levels = c(0, 1),
      labels = c("Rainy", "Dry")
    )
  )

cat("HWISE_REDUCED — summary:\n")
print(summary(df$HWISE_REDUCED))
cat("\nN with valid HWISE_REDUCED:", sum(!is.na(df$HWISE_REDUCED)), "\n")
cat("N with valid HWISE_FULL:   ", sum(!is.na(df$HWISE_FULL)), "\n")
cat("\nWI_NEIGHBORHOOD distribution:\n")
print(table(df$WI_NEIGHBORHOOD, useNA = "always"))
cat("\nSEASON distribution:\n")
print(table(df$SEASON, useNA = "always"))
cat("\nEMO_NEG distribution:\n")
print(table(df$EMO_NEG, useNA = "always"))
```

---

# Model Specifications

```{r model-specs}
# ── Panel A: PSS as outcome (linear regression) ─────────────────────────────
# Covariates: Age + SES composite score + Household size (in every model)

m_a1 <- lm(
  PSS_TOTAL ~ HWISE_REDUCED + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

m_a2 <- lm(
  PSS_TOTAL ~ HWISE_REDUCED * SEASON + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

m_a3 <- lm(
  PSS_TOTAL ~ WI_NEIGHBORHOOD + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

m_a4 <- lm(
  PSS_TOTAL ~ WI_NEIGHBORHOOD * SEASON + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

m_a5 <- lm(
  PSS_TOTAL ~ HWISE_REDUCED + SEASON + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

m_a6 <- lm(
  PSS_TOTAL ~ WI_NEIGHBORHOOD + SEASON + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

m_a7 <- lm(
  PSS_TOTAL ~ SEASON + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

m_a8 <- lm(
  PSS_TOTAL ~ HWISE_REDUCED * WI_NEIGHBORHOOD + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

m_a9 <- lm(
  PSS_TOTAL ~ HWISE_REDUCED + WI_NEIGHBORHOOD + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

m_a10 <- lm(
  PSS_TOTAL ~ HWISE_FULL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, na.action = na.omit
)

# ── Panel B: Emotion type as outcome (logistic regression) ───────────────────
# Covariates: PSS + Age + SES + Household size + Employment type (in every model)

m_b1 <- glm(
  EMO_NEG ~ HWISE_REDUCED + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, family = binomial, na.action = na.omit
)

m_b2 <- glm(
  EMO_NEG ~ HWISE_REDUCED * SEASON + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, family = binomial, na.action = na.omit
)

m_b3 <- glm(
  EMO_NEG ~ WI_NEIGHBORHOOD + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, family = binomial, na.action = na.omit
)

m_b4 <- glm(
  EMO_NEG ~ WI_NEIGHBORHOOD * SEASON + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, family = binomial, na.action = na.omit
)

m_b5 <- glm(
  EMO_NEG ~ HWISE_REDUCED + SEASON + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, family = binomial, na.action = na.omit
)

m_b6 <- glm(
  EMO_NEG ~ WI_NEIGHBORHOOD + SEASON + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, family = binomial, na.action = na.omit
)

m_b7 <- glm(
  EMO_NEG ~ SEASON + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, family = binomial, na.action = na.omit
)

m_b8 <- glm(
  EMO_NEG ~ HWISE_REDUCED * WI_NEIGHBORHOOD + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, family = binomial, na.action = na.omit
)

m_b9 <- glm(
  EMO_NEG ~ HWISE_REDUCED + WI_NEIGHBORHOOD + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
  data = df, family = binomial, na.action = na.omit
)

# ── Sample sizes ─────────────────────────────────────────────────────────────
tibble(
  model = c("m_a1", "m_a2", "m_a3", "m_a4", "m_a5", "m_a6", "m_a7", "m_a8", "m_a9", "m_a10",
            "m_b1", "m_b2", "m_b3", "m_b4", "m_b5", "m_b6", "m_b7", "m_b8", "m_b9"),
  n = c(nobs(m_a1), nobs(m_a2), nobs(m_a3), nobs(m_a4), nobs(m_a5), nobs(m_a6), nobs(m_a7), nobs(m_a8), nobs(m_a9), nobs(m_a10),
        nobs(m_b1), nobs(m_b2), nobs(m_b3), nobs(m_b4), nobs(m_b5), nobs(m_b6), nobs(m_b7), nobs(m_b8), nobs(m_b9))
) %>% print()
```

---

# Results

## Panel A — Coefficient Plot

```{r coef-plot-a, fig.width=8, fig.height=5}
term_labels_a <- c(
  "HWISE_REDUCED"                                       = "HWISE reduced score",
  "HWISE_FULL"                                          = "HWISE full score (12 items)",
  "SEASONDry"                                           = "Season (Dry)",
  "WI_NEIGHBORHOODWater-insecure"                       = "Water-insecure neighbourhood",
  "HWISE_REDUCED:SEASONDry"                             = "HWISE × Season (Dry)",
  "WI_NEIGHBORHOODWater-insecure:SEASONDry"             = "WI neighbourhood × Season (Dry)",
  "HWISE_REDUCED:WI_NEIGHBORHOODWater-insecure"         = "HWISE × WI neighbourhood"
)

coef_a <- bind_rows(
  tidy(m_a1, conf.int = TRUE) %>% mutate(model = "a1: HWISE reduced"),
  tidy(m_a2, conf.int = TRUE) %>% mutate(model = "a2: HWISE × Season"),
  tidy(m_a3, conf.int = TRUE) %>% mutate(model = "a3: WI neighbourhood"),
  tidy(m_a4, conf.int = TRUE) %>% mutate(model = "a4: WI neighbourhood × Season")
) %>%
  filter(!term %in% c("(Intercept)"), !term %in% COVARIATES) %>%
  mutate(
    term_label = recode(term, !!!term_labels_a, .default = term),
    sig        = sig_symbol(p.value),
    model      = factor(model, levels = c(
      "a1: HWISE reduced", "a2: HWISE × Season",
      "a3: WI neighbourhood", "a4: WI neighbourhood × Season"
    ))
  )

p_a <- ggplot(coef_a, aes(x = estimate, y = fct_rev(term_label), colour = model)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high),
                  position = position_dodge(width = 0.6), size = 0.5) +
  geom_text(aes(label = sig, x = conf.high),
            position = position_dodge(width = 0.6),
            hjust = -0.3, size = 3.5, show.legend = FALSE) +
  scale_colour_manual(
    values = c("#E41A1C","#377EB8","#4DAF4A","#984EA3"),
    name   = "Model",
    labels = scales::label_wrap(15)
  ) +
  scale_y_discrete(labels = scales::label_wrap(15)) +
  labs(
    title   = "Panel A — Linear regression: Perceived Stress Score",
    x       = "β coefficient (95% CI)",
    y       = NULL,
    caption = "† p<0.1  * p<0.05  ** p<0.01  *** p<0.001\nAge, SES score, and household size excluded."
  ) +
  theme_bw(base_size = 12) +
  theme(
    legend.position      = c(0.98, 0.98),
    legend.justification = c("right", "top"),
    legend.background    = element_rect(fill = scales::alpha("white", 0.85), colour = "grey80"),
    legend.key.size      = unit(0.4, "cm"),
    legend.text          = element_text(size = 9),
    panel.grid.minor     = element_blank(),
    plot.caption         = element_text(size = 9, colour = "grey40")
  )

p_a
ggsave("./output/figures/07.panel_a_coef_plot.png", p_a,
       width = 8, height = 5, dpi = 300, bg = "white")
cat("Saved: ./output/figures/07.panel_a_coef_plot.png\n")
```

## Panel A — Compact Linear Regression Table

One row per predictor per model; covariates (Age, SES, Household size, Chronic disease) excluded.

```{r panel-a-compact}
compact_a <- bind_rows(
  tidy(m_a1, conf.int = TRUE) %>% mutate(Model = "a1: HWISE reduced"),
  tidy(m_a2, conf.int = TRUE) %>% mutate(Model = "a2: HWISE × Season"),
  tidy(m_a3, conf.int = TRUE) %>% mutate(Model = "a3: WI neighbourhood"),
  tidy(m_a4, conf.int = TRUE) %>% mutate(Model = "a4: WI neighbourhood × Season"),
  tidy(m_a5, conf.int = TRUE) %>% mutate(Model = "a5: HWISE + Season"),
  tidy(m_a6, conf.int = TRUE) %>% mutate(Model = "a6: WI neighbourhood + Season"),
  tidy(m_a7, conf.int = TRUE) %>% mutate(Model = "a7: Season only"),
  tidy(m_a8, conf.int = TRUE) %>% mutate(Model = "a8: HWISE × WI neighbourhood"),
  tidy(m_a9,  conf.int = TRUE) %>% mutate(Model = "a9: HWISE + WI neighbourhood"),
  tidy(m_a10, conf.int = TRUE) %>% mutate(Model = "a10: HWISE full score")
) %>%
  filter(!term %in% c("(Intercept)"), !term %in% COVARIATES) %>%
  mutate(
    Variable  = recode(term, !!!term_labels_a, .default = term),
    Beta      = sprintf("%.2f", estimate),
    `95% CI`  = sprintf("[%.2f, %.2f]", conf.low, conf.high),
    `p-value` = case_when(
      p.value < 0.001 ~ paste0(sprintf("%.3f", p.value), " ***"),
      p.value < 0.01  ~ paste0(sprintf("%.3f", p.value), " **"),
      p.value < 0.05  ~ paste0(sprintf("%.3f", p.value), " *"),
      p.value < 0.1   ~ paste0(sprintf("%.3f", p.value), " †"),
      TRUE            ~ sprintf("%.3f", p.value)
    ),
    Model = factor(Model, levels = c(
      "a1: HWISE reduced", "a2: HWISE × Season",
      "a3: WI neighbourhood", "a4: WI neighbourhood × Season",
      "a5: HWISE + Season", "a6: WI neighbourhood + Season",
      "a7: Season only", "a8: HWISE × WI neighbourhood",
      "a9: HWISE + WI neighbourhood", "a10: HWISE full score"
    ))
  ) %>%
  select(Model, Variable, Beta, `95% CI`, `p-value`) %>%
  arrange(Model)

compact_a %>%
  flextable() %>%
  merge_v(j = "Model") %>%
  valign(j = "Model", valign = "top") %>%
  bold(j = "Model") %>%
  set_caption("Panel A. Linear regression — Perceived Stress Score (β). Covariates excluded.") %>%
  autofit()
```

## Panel B — Coefficient Plot (OR scale)

```{r coef-plot-b, fig.width=8, fig.height=5}
term_labels_b <- c(
  "HWISE_REDUCED"                                       = "HWISE reduced score",
  "SEASONDry"                                           = "Season (Dry)",
  "WI_NEIGHBORHOODWater-insecure"                       = "Water-insecure neighbourhood",
  "HWISE_REDUCED:SEASONDry"                             = "HWISE × Season (Dry)",
  "WI_NEIGHBORHOODWater-insecure:SEASONDry"             = "WI neighbourhood × Season (Dry)",
  "HWISE_REDUCED:WI_NEIGHBORHOODWater-insecure"         = "HWISE × WI neighbourhood"
)

coef_b <- bind_rows(
  tidy(m_b1, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "b1: HWISE reduced"),
  tidy(m_b2, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "b2: HWISE × Season"),
  tidy(m_b3, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "b3: WI neighbourhood"),
  tidy(m_b4, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "b4: WI neighbourhood × Season")
) %>%
  filter(!term %in% c("(Intercept)"), !term %in% COVARIATES) %>%
  mutate(
    term_label = recode(term, !!!term_labels_b, .default = term),
    sig        = sig_symbol(p.value),
    model      = factor(model, levels = c(
      "b1: HWISE reduced", "b2: HWISE × Season",
      "b3: WI neighbourhood", "b4: WI neighbourhood × Season"
    ))
  )

p_b <- ggplot(coef_b, aes(x = estimate, y = fct_rev(term_label), colour = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(xmin = conf.low, xmax = conf.high),
                  position = position_dodge(width = 0.6), size = 0.5) +
  geom_text(aes(label = sig, x = conf.high),
            position = position_dodge(width = 0.6),
            hjust = -0.3, size = 3.5, show.legend = FALSE) +
  scale_x_log10() +
  scale_colour_manual(
    values = c("#E41A1C","#377EB8","#4DAF4A","#984EA3"),
    name   = "Model",
    labels = scales::label_wrap(22)
  ) +
  scale_y_discrete(labels = scales::label_wrap(25)) +
  labs(
    title   = "Panel B — Logistic regression: Water-related emotion (OR)",
    x       = "Odds Ratio (95% CI, log scale)",
    y       = NULL,
    caption = "† p<0.1  * p<0.05  ** p<0.01  *** p<0.001\nAge, SES score, household size, and PSS excluded."
  ) +
  theme_bw(base_size = 12) +
  theme(
    legend.position      = c(0.98, 0.98),
    legend.justification = c("right", "top"),
    legend.background    = element_rect(fill = scales::alpha("white", 0.85), colour = "grey80"),
    legend.key.size      = unit(0.4, "cm"),
    legend.text          = element_text(size = 9),
    panel.grid.minor     = element_blank(),
    plot.caption         = element_text(size = 9, colour = "grey40")
  )

p_b
ggsave("./output/figures/07.panel_b_coef_plot.png", p_b,
       width = 8, height = 5, dpi = 300, bg = "white")
cat("Saved: ./output/figures/07.panel_b_coef_plot.png\n")
```

## Panel B — Compact Logistic Regression Table

One row per predictor per model; covariates (Age, SES, Household size, Chronic disease, PSS) excluded.

```{r panel-b-compact}
term_labels_b_named <- c(
  term_labels_b,
  "PSS_TOTAL" = "PSS total"
)

compact_b <- bind_rows(
  tidy(m_b1, exponentiate = TRUE, conf.int = TRUE) %>% mutate(Model = "b1: HWISE reduced"),
  tidy(m_b2, exponentiate = TRUE, conf.int = TRUE) %>% mutate(Model = "b2: HWISE × Season"),
  tidy(m_b3, exponentiate = TRUE, conf.int = TRUE) %>% mutate(Model = "b3: WI neighbourhood"),
  tidy(m_b4, exponentiate = TRUE, conf.int = TRUE) %>% mutate(Model = "b4: WI neighbourhood × Season"),
  tidy(m_b5, exponentiate = TRUE, conf.int = TRUE) %>% mutate(Model = "b5: HWISE + Season"),
  tidy(m_b6, exponentiate = TRUE, conf.int = TRUE) %>% mutate(Model = "b6: WI neighbourhood + Season"),
  tidy(m_b7, exponentiate = TRUE, conf.int = TRUE) %>% mutate(Model = "b7: Season only"),
  tidy(m_b8, exponentiate = TRUE, conf.int = TRUE) %>% mutate(Model = "b8: HWISE × WI neighbourhood"),
  tidy(m_b9, exponentiate = TRUE, conf.int = TRUE) %>% mutate(Model = "b9: HWISE + WI neighbourhood")
) %>%
  filter(!term %in% c("(Intercept)"), !term %in% COVARIATES) %>%
  mutate(
    Variable = recode(term, !!!term_labels_b_named, .default = term),
    OR       = sprintf("%.2f", estimate),
    `95% CI` = sprintf("[%.2f, %.2f]", conf.low, conf.high),
    `p-value` = case_when(
      p.value < 0.001 ~ paste0(sprintf("%.3f", p.value), " ***"),
      p.value < 0.01  ~ paste0(sprintf("%.3f", p.value), " **"),
      p.value < 0.05  ~ paste0(sprintf("%.3f", p.value), " *"),
      p.value < 0.1   ~ paste0(sprintf("%.3f", p.value), " †"),
      TRUE            ~ sprintf("%.3f", p.value)
    ),
    Model = factor(Model, levels = c(
      "b1: HWISE reduced", "b2: HWISE × Season",
      "b3: WI neighbourhood", "b4: WI neighbourhood × Season",
      "b5: HWISE + Season", "b6: WI neighbourhood + Season",
      "b7: Season only", "b8: HWISE × WI neighbourhood",
      "b9: HWISE + WI neighbourhood"
    ))
  ) %>%
  select(Model, Variable, OR, `95% CI`, `p-value`) %>%
  arrange(Model)

compact_b %>%
  flextable() %>%
  merge_v(j = "Model") %>%
  valign(j = "Model", valign = "top") %>%
  bold(j = "Model") %>%
  set_caption("Panel B. Logistic regression — Water-related emotion (OR). Covariates excluded.") %>%
  autofit()
```

---

# Model Fit Statistics

```{r fit-stats}
mcfadden <- function(mod) round(1 - mod$deviance / mod$null.deviance, 3)

model_labels <- c(
  m_a1 = "HWISE reduced",
  m_a2 = "HWISE × Season",
  m_a3 = "WI neighbourhood",
  m_a4 = "WI neighbourhood × Season",
  m_a5 = "HWISE + Season",
  m_a6 = "WI neighbourhood + Season",
  m_a7 = "Season only",
  m_a8  = "HWISE × WI neighbourhood",
  m_a9  = "HWISE + WI neighbourhood",
  m_a10 = "HWISE full score (12 items)",
  m_b1  = "HWISE reduced",
  m_b2 = "HWISE × Season",
  m_b3 = "WI neighbourhood",
  m_b4 = "WI neighbourhood × Season",
  m_b5 = "HWISE + Season",
  m_b6 = "WI neighbourhood + Season",
  m_b7 = "Season only",
  m_b8 = "HWISE × WI neighbourhood",
  m_b9 = "HWISE + WI neighbourhood"
)

# ── Panel A ───────────────────────────────────────────────────────────────────
panel_a_fit <- map_dfr(
  list(m_a1 = m_a1, m_a2 = m_a2, m_a3 = m_a3, m_a4 = m_a4,
       m_a5 = m_a5, m_a6 = m_a6, m_a7 = m_a7, m_a8 = m_a8, m_a9 = m_a9, m_a10 = m_a10),
  ~ glance(.x), .id = "model"
) %>%
  mutate(
    Panel      = "A — Perceived Stress Score (linear)",
    Predictors = model_labels[model],
    n          = nobs,
    R2         = round(r.squared, 3),
    Adj_R2     = round(adj.r.squared, 3),
    AIC        = round(AIC, 1),
    BIC        = round(BIC, 1),
    ΔAIC       = round(AIC - min(AIC), 1)
  ) %>%
  arrange(AIC) %>%
  select(Panel, Model = model, Predictors, n, R2, Adj_R2, AIC, BIC, ΔAIC)

# ── Panel B ───────────────────────────────────────────────────────────────────
panel_b_fit <- tibble(
  model = c("m_b1","m_b2","m_b3","m_b4","m_b5","m_b6","m_b7","m_b8","m_b9"),
  n     = c(nobs(m_b1), nobs(m_b2), nobs(m_b3), nobs(m_b4),
            nobs(m_b5), nobs(m_b6), nobs(m_b7), nobs(m_b8), nobs(m_b9)),
  McFadden_R2 = c(mcfadden(m_b1), mcfadden(m_b2), mcfadden(m_b3), mcfadden(m_b4),
                  mcfadden(m_b5), mcfadden(m_b6), mcfadden(m_b7), mcfadden(m_b8), mcfadden(m_b9)),
  AIC = round(c(AIC(m_b1), AIC(m_b2), AIC(m_b3), AIC(m_b4),
                AIC(m_b5), AIC(m_b6), AIC(m_b7), AIC(m_b8), AIC(m_b9)), 1),
  BIC = round(c(BIC(m_b1), BIC(m_b2), BIC(m_b3), BIC(m_b4),
                BIC(m_b5), BIC(m_b6), BIC(m_b7), BIC(m_b8), BIC(m_b9)), 1)
) %>%
  mutate(
    Panel      = "B — Water-related emotion (logistic)",
    Predictors = model_labels[model],
    ΔAIC       = round(AIC - min(AIC), 1)
  ) %>%
  arrange(AIC) %>%
  select(Panel, Model = model, Predictors, n, McFadden_R2, AIC, BIC, ΔAIC)

# ── Pretty flextable — Panel A ────────────────────────────────────────────────
panel_a_fit %>%
  flextable() %>%
  set_header_labels(
    Panel = "Panel", Model = "Model", Predictors = "Predictors",
    n = "N", R2 = "R²", Adj_R2 = "Adj. R²",
    AIC = "AIC", BIC = "BIC", ΔAIC = "ΔAIC"
  ) %>%
  merge_v(j = "Panel") %>%
  valign(j = "Panel", valign = "top") %>%
  bold(j = c("AIC", "ΔAIC")) %>%
  bg(i = ~ ΔAIC == 0, bg = "#d4edda") %>%   # highlight best model
  color(i = ~ ΔAIC > 10, j = "ΔAIC", color = "#999999") %>%
  set_caption("Panel A — Model fit statistics sorted by AIC (best to worst). Green = best model; grey ΔAIC > 10 = poor support.") %>%
  colformat_num(j = c("R2", "Adj_R2"), digits = 3) %>%
  colformat_num(j = c("AIC", "BIC", "ΔAIC"), digits = 1) %>%
  autofit()
```

```{r fit-stats-b}
# ── Pretty flextable — Panel B ────────────────────────────────────────────────
panel_b_fit %>%
  flextable() %>%
  set_header_labels(
    Panel = "Panel", Model = "Model", Predictors = "Predictors",
    n = "N", McFadden_R2 = "McFadden R²",
    AIC = "AIC", BIC = "BIC", ΔAIC = "ΔAIC"
  ) %>%
  merge_v(j = "Panel") %>%
  valign(j = "Panel", valign = "top") %>%
  bold(j = c("AIC", "ΔAIC")) %>%
  bg(i = ~ ΔAIC == 0, bg = "#d4edda") %>%
  color(i = ~ ΔAIC > 10, j = "ΔAIC", color = "#999999") %>%
  set_caption("Panel B — Model fit statistics sorted by AIC (best to worst). Green = best model; grey ΔAIC > 10 = poor support.") %>%
  colformat_num(j = "McFadden_R2", digits = 3) %>%
  colformat_num(j = c("AIC", "BIC", "ΔAIC"), digits = 1) %>%
  autofit()
```

---

# Season-Stratified Estimates

Interaction models (m_a2, m_a4, m_b2, m_b4) are complemented by season-
stratified re-runs to facilitate interpretation.

```{r season-strat}
seasons <- c("Dry", "Rainy")

# ── Panel A: stratified ───────────────────────────────────────────────────────
strat_a <- map_dfr(seasons, function(s) {
  df_s <- filter(df, SEASON == s)
  list(
    hwise  = lm(PSS_TOTAL ~ HWISE_REDUCED + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
                data = df_s, na.action = na.omit),
    wi_nbh = lm(PSS_TOTAL ~ WI_NEIGHBORHOOD + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
                data = df_s, na.action = na.omit)
  ) %>%
    map_dfr(~ tidy(.x, conf.int = TRUE), .id = "predictor") %>%
    filter(!term %in% c("(Intercept)", "D_AGE", "SES_TOTAL", "D_HH_SIZE", "CHRONIC_DISYes")) %>%
    mutate(season = s)
}) %>%
  mutate(
    ci_fmt = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_fmt  = case_when(p.value < 0.001 ~ "<0.001",
                       TRUE ~ sprintf("%.3f", p.value))
  ) %>%
  select(season, predictor, term, n = statistic, ci_fmt, p_fmt)

cat("Panel A — season-stratified focal predictors:\n")
print(strat_a, n = Inf)

# ── Panel B: stratified ───────────────────────────────────────────────────────
strat_b <- map_dfr(seasons, function(s) {
  df_s <- filter(df, SEASON == s)
  list(
    hwise  = glm(EMO_NEG ~ HWISE_REDUCED + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
                 data = df_s, family = binomial, na.action = na.omit),
    wi_nbh = glm(EMO_NEG ~ WI_NEIGHBORHOOD + PSS_TOTAL + D_AGE + SES_TOTAL + D_HH_SIZE + CHRONIC_DIS,
                 data = df_s, family = binomial, na.action = na.omit)
  ) %>%
    map_dfr(~ tidy(.x, exponentiate = TRUE, conf.int = TRUE), .id = "predictor") %>%
    filter(!term %in% c("(Intercept)", "D_AGE", "SES_TOTAL",
                        "D_HH_SIZE", "PSS_TOTAL", "CHRONIC_DISYes")) %>%
    mutate(season = s)
}) %>%
  mutate(
    or_fmt = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_fmt  = case_when(p.value < 0.001 ~ "<0.001",
                       TRUE ~ sprintf("%.3f", p.value))
  ) %>%
  select(season, predictor, term, or_fmt, p_fmt)

cat("\nPanel B — season-stratified focal predictors (OR):\n")
print(strat_b, n = Inf)
```

---

# Export to Word

The compact regression tables are exported as Word documents with publication-ready formatting.

```{r export-word}
# ── Helper: build a pretty flextable ─────────────────────────────────────────
pretty_ft <- function(df, caption_txt) {
  df %>%
    flextable() %>%
    merge_v(j = "Model") %>%
    valign(j = "Model", valign = "top") %>%
    bold(part = "header") %>%
    bold(j = "Model") %>%
    bg(part = "header", bg = "#2c3e50") %>%
    color(part = "header", color = "white") %>%
    set_caption(caption_txt) %>%
    theme_booktabs() %>%
    autofit()
}

# ── All-models Word document ──────────────────────────────────────────────────
doc_all <- read_docx()

doc_all <- doc_all %>%
  body_add_par("Panel A. Linear regression — Perceived Stress Score (PSS)",
               style = "heading 1") %>%
  body_add_flextable(
    pretty_ft(compact_a,
              "Panel A. Linear regression — Perceived Stress Score (β). Covariates excluded.")
  ) %>%
  body_add_break() %>%
  body_add_par("Panel B. Logistic regression — Water-related emotion (negative vs. positive)",
               style = "heading 1") %>%
  body_add_flextable(
    pretty_ft(compact_b,
              "Panel B. Logistic regression — Water-related emotion (OR). Covariates excluded.")
  )

out_all <- "./output/tables/08.multivariate_models.docx"
print(doc_all, target = out_all)
cat("All-models table exported to:", out_all, "\n")

# ── Filtered Word document: m_a1–m_a4, m_b1–m_b4 only ───────────────────────
core_models_a <- c(
  "a1: HWISE reduced", "a2: HWISE × Season",
  "a3: WI neighbourhood", "a4: WI neighbourhood × Season"
)
core_models_b <- c(
  "b1: HWISE reduced", "b2: HWISE × Season",
  "b3: WI neighbourhood", "b4: WI neighbourhood × Season"
)

compact_a_core <- compact_a %>%
  filter(Model %in% core_models_a) %>%
  mutate(Model = droplevels(Model))

compact_b_core <- compact_b %>%
  filter(Model %in% core_models_b) %>%
  mutate(Model = droplevels(Model))

doc_core <- read_docx()

doc_core <- doc_core %>%
  body_add_par("Panel A. Linear regression — Perceived Stress Score (PSS)",
               style = "heading 1") %>%
  body_add_flextable(
    pretty_ft(compact_a_core,
              "Panel A. Linear regression — Perceived Stress Score (β). Models m_a1–m_a4. Covariates excluded.")
  ) %>%
  body_add_break() %>%
  body_add_par("Panel B. Logistic regression — Water-related emotion (negative vs. positive)",
               style = "heading 1") %>%
  body_add_flextable(
    pretty_ft(compact_b_core,
              "Panel B. Logistic regression — Water-related emotion (OR). Models m_b1–m_b4. Covariates excluded.")
  )

out_core <- "./output/tables/08.multivariate_models_core.docx"
print(doc_core, target = out_core)
cat("Core-models table exported to:", out_core, "\n")
```

---

# Principal Component Analysis

PCA is run on the 10 individual HWISE-reduced items plus the continuous
covariates (PSS, Age, SES, Household size). All variables are standardised
(mean = 0, SD = 1) before decomposition. Cases with any missing value across
these variables are listwise-deleted.

```{r pca-prep}
pca_vars <- c(
  "HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
  "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
  "HW_SLEEP", "HW_NONE",
  "PSS_TOTAL", "D_AGE", "SES_TOTAL", "D_HH_SIZE"
)

# Pretty labels for PCA axes
pca_labels <- c(
  HW_WORRY    = "Worry (HWISE)",
  HW_INTERR   = "Interruption (HWISE)",
  HW_CLOTHES  = "Clothes (HWISE)",
  HW_PLANS    = "Plans (HWISE)",
  HW_FOOD     = "Food (HWISE)",
  HW_HANDS    = "Hands (HWISE)",
  HW_BODY     = "Body (HWISE)",
  HW_DRINK    = "Drink (HWISE)",
  HW_SLEEP    = "Sleep (HWISE)",
  HW_NONE     = "None (HWISE)",
  PSS_TOTAL   = "PSS total",
  D_AGE       = "Age",
  SES_TOTAL = "SES score",
  D_HH_SIZE   = "Household size"
)

df_pca <- df %>%
  select(all_of(pca_vars), WI_NEIGHBORHOOD, SEASON) %>%
  drop_na(all_of(pca_vars))

pca_fit <- prcomp(df_pca[, pca_vars], scale. = TRUE, center = TRUE)

# Variance explained
var_exp <- summary(pca_fit)$importance
cat(sprintf("PC1: %.1f%%   PC2: %.1f%%   PC3: %.1f%%\n",
            var_exp[2, 1] * 100,
            var_exp[2, 2] * 100,
            var_exp[2, 3] * 100))
```

## Scree Plot

```{r pca-scree, fig.width=6, fig.height=4}
scree_df <- tibble(
  PC       = paste0("PC", seq_along(pca_fit$sdev)),
  variance = pca_fit$sdev^2,
  pct      = variance / sum(variance) * 100,
  cum_pct  = cumsum(pct)
) %>%
  mutate(PC = factor(PC, levels = PC))

p_scree <- ggplot(scree_df, aes(x = PC, y = pct)) +
  geom_col(fill = "#377EB8", alpha = 0.8, width = 0.6) +
  geom_line(aes(group = 1), colour = "#E41A1C", linewidth = 0.7) +
  geom_point(colour = "#E41A1C", size = 2) +
  geom_hline(yintercept = 100 / length(pca_vars),
             linetype = "dashed", colour = "grey50") +
  annotate("text", x = length(pca_vars) - 1,
           y = 100 / length(pca_vars) + 0.8,
           label = "Average eigenvalue", size = 3, colour = "grey40") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(
    title   = "Scree plot — variance explained per component",
    x       = NULL,
    y       = "Variance explained (%)",
    caption = sprintf("Dashed line = average eigenvalue (%.1f%%)",
                      100 / length(pca_vars))
  ) +
  theme_bw(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        plot.caption = element_text(size = 9, colour = "grey40"))

p_scree
ggsave("./output/figures/07.pca_scree.png", p_scree,
       width = 6, height = 4, dpi = 300, bg = "white")
```

## Variable Loadings (PC1 & PC2)

```{r pca-loadings, fig.width=7, fig.height=6}
loadings_df <- as_tibble(pca_fit$rotation[, 1:2], rownames = "variable") %>%
  mutate(
    label    = pca_labels[variable],
    hwise    = str_detect(variable, "^HW_"),
    category = if_else(hwise, "HWISE item", "Covariate")
  )

p_load <- ggplot(loadings_df, aes(x = PC1, y = PC2,
                                   colour = category, label = label)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey70") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  geom_segment(aes(xend = 0, yend = 0),
               arrow = arrow(length = unit(0.2, "cm"), ends = "first"),
               alpha = 0.6) +
  geom_point(size = 3) +
  ggrepel::geom_text_repel(size = 3.2, max.overlaps = 20,
                            show.legend = FALSE) +
  scale_colour_manual(values = c("HWISE item" = "#377EB8",
                                  "Covariate"  = "#E41A1C"),
                      name = NULL) +
  labs(
    title   = "PCA loadings — PC1 vs PC2",
    x       = sprintf("PC1 (%.1f%% variance)", var_exp[2, 1] * 100),
    y       = sprintf("PC2 (%.1f%% variance)", var_exp[2, 2] * 100),
    caption = "Arrows point toward the origin; length = loading magnitude."
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position   = "bottom",
        panel.grid.minor  = element_blank(),
        plot.caption      = element_text(size = 9, colour = "grey40"))

p_load
ggsave("./output/figures/07.pca_loadings.png", p_load,
       width = 7, height = 6, dpi = 300, bg = "white")
```

## Individual Scores — Coloured by Group

```{r pca-scores, fig.width=10, fig.height=4}
scores_df <- as_tibble(pca_fit$x[, 1:2]) %>%
  bind_cols(df_pca %>% select(WI_NEIGHBORHOOD, SEASON))

# Helper to build one scores plot
scores_plot <- function(data, colour_var, title_str, palette) {
  ggplot(data, aes(x = PC1, y = PC2, colour = .data[[colour_var]])) +
    geom_point(alpha = 0.45, size = 1.2) +
    stat_ellipse(aes(fill = .data[[colour_var]]),
                 geom = "polygon", alpha = 0.08, linewidth = 0.7) +
    scale_colour_manual(values = palette, name = NULL, na.translate = FALSE) +
    scale_fill_manual(values   = palette, name = NULL, na.translate = FALSE) +
    labs(
      title = title_str,
      x     = sprintf("PC1 (%.1f%%)", var_exp[2, 1] * 100),
      y     = sprintf("PC2 (%.1f%%)", var_exp[2, 2] * 100)
    ) +
    theme_bw(base_size = 11) +
    theme(legend.position  = "bottom",
          panel.grid.minor = element_blank())
}

p_wi <- scores_plot(
  scores_df, "WI_NEIGHBORHOOD",
  "PC scores by neighbourhood type",
  c("Water-secure" = "#377EB8", "Water-insecure" = "#E41A1C")
)

p_sea <- scores_plot(
  scores_df, "SEASON",
  "PC scores by season",
  c("Rainy" = "#4DAF4A", "Dry" = "#FF7F00")
)

p_scores <- p_wi + p_sea +
  patchwork::plot_annotation(
    title   = "PCA individual scores",
    caption = "Shaded ellipses = 95% normal confidence regions."
  )

p_scores
ggsave("./output/figures/07.pca_scores.png", p_scores,
       width = 10, height = 4, dpi = 300, bg = "white")
```

## Loadings Heatmap

```{r pca-heatmap, fig.width=7, fig.height=5}
heat_df <- as_tibble(pca_fit$rotation[, 1:4], rownames = "variable") %>%
  mutate(label = pca_labels[variable]) %>%
  pivot_longer(PC1:PC4, names_to = "PC", values_to = "loading")

p_heat <- ggplot(heat_df, aes(x = PC, y = fct_rev(label), fill = loading)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%.2f", loading)), size = 2.8) +
  scale_fill_gradient2(low = "#E41A1C", mid = "white", high = "#377EB8",
                       midpoint = 0, name = "Loading") +
  labs(
    title   = "PCA loadings heatmap — first four components",
    x       = NULL, y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(panel.grid = element_blank(),
        axis.text.y = element_text(size = 9))

p_heat
ggsave("./output/figures/07.pca_heatmap.png", p_heat,
       width = 7, height = 5, dpi = 300, bg = "white")
```

---

# Polychoric PCA

Standard PCA assumes continuous, normally distributed variables. The HWISE
items are ordinal (0–3 Likert scale), so PCA on the **polychoric correlation
matrix** is more appropriate. Polychoric correlations estimate the latent
Pearson correlation between pairs of assumed-continuous underlying variables
that are observed as ordered categories.

PCA here is restricted to the **10 HWISE-reduced items** — the natural domain
for this approach — using `psych::polychoric()` followed by an eigendecomposition
of the resulting correlation matrix.

```{r poly-pca-prep}
library(psych)

hwise_items <- c(
  "HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
  "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
  "HW_SLEEP", "HW_NONE"
)

hwise_labels <- c(
  HW_WORRY   = "Worry",   HW_INTERR  = "Interruption",
  HW_CLOTHES = "Clothes", HW_PLANS   = "Plans",
  HW_FOOD    = "Food",    HW_HANDS   = "Hands",
  HW_BODY    = "Body",    HW_DRINK   = "Drink",
  HW_SLEEP   = "Sleep",   HW_NONE    = "None"
)

df_hwise <- df %>%
  select(all_of(hwise_items), WI_NEIGHBORHOOD, SEASON) %>%
  drop_na(all_of(hwise_items))

cat("N complete cases for polychoric PCA:", nrow(df_hwise), "\n")

# Polychoric correlation matrix
poly_cor <- polychoric(df_hwise[, hwise_items])
cat("\nPolychoric correlation matrix computed.\n")
```

## Polychoric Correlation Heatmap

```{r poly-cor-heatmap, fig.width=6, fig.height=5}
poly_mat <- poly_cor$rho
rownames(poly_mat) <- hwise_labels[rownames(poly_mat)]
colnames(poly_mat) <- hwise_labels[colnames(poly_mat)]

poly_cor_long <- as_tibble(poly_mat, rownames = "item1") %>%
  pivot_longer(-item1, names_to = "item2", values_to = "r") %>%
  mutate(
    item1 = factor(item1, levels = rev(hwise_labels)),
    item2 = factor(item2, levels = hwise_labels)
  )

p_poly_cor <- ggplot(poly_cor_long, aes(x = item2, y = item1, fill = r)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%.2f", r)), size = 2.6) +
  scale_fill_gradient2(low = "#E41A1C", mid = "white", high = "#377EB8",
                       midpoint = 0, limits = c(-1, 1), name = "ρ") +
  labs(title = "Polychoric correlation matrix — HWISE reduced items",
       x = NULL, y = NULL) +
  theme_bw(base_size = 11) +
  theme(axis.text.x  = element_text(angle = 45, hjust = 1),
        panel.grid   = element_blank())

p_poly_cor
ggsave("./output/figures/07.poly_cor_heatmap.png", p_poly_cor,
       width = 6, height = 5, dpi = 300, bg = "white")
```

## Scree Plot

```{r poly-pca-scree, fig.width=6, fig.height=4}
eig      <- eigen(poly_cor$rho)
eig_vals <- eig$values
pct_poly <- eig_vals / sum(eig_vals) * 100

scree_poly <- tibble(
  PC      = factor(paste0("PC", seq_along(eig_vals)),
                   levels = paste0("PC", seq_along(eig_vals))),
  eigenvalue = eig_vals,
  pct        = pct_poly,
  cum_pct    = cumsum(pct_poly)
)

p_poly_scree <- ggplot(scree_poly, aes(x = PC, y = eigenvalue)) +
  geom_col(fill = "#4DAF4A", alpha = 0.8, width = 0.6) +
  geom_line(aes(group = 1), colour = "#984EA3", linewidth = 0.7) +
  geom_point(colour = "#984EA3", size = 2.5) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey50") +
  annotate("text", x = length(eig_vals) - 1, y = 1.12,
           label = "Kaiser criterion (λ = 1)", size = 3, colour = "grey40") +
  geom_text(aes(label = sprintf("%.1f%%", pct), y = eigenvalue + 0.05),
            size = 2.8, vjust = 0) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(
    title   = "Polychoric PCA — scree plot",
    x       = NULL,
    y       = "Eigenvalue",
    caption = "Dashed line = Kaiser criterion (eigenvalue > 1)."
  ) +
  theme_bw(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        plot.caption = element_text(size = 9, colour = "grey40"))

p_poly_scree
ggsave("./output/figures/07.poly_pca_scree.png", p_poly_scree,
       width = 6, height = 4, dpi = 300, bg = "white")
```

## Variable Loadings (PC1 & PC2)

```{r poly-pca-loadings, fig.width=6, fig.height=5}
# Scale eigenvectors to loadings: loading = eigenvector * sqrt(eigenvalue)
poly_loadings <- sweep(eig$vectors, 2, sqrt(eig_vals), `*`)
rownames(poly_loadings) <- hwise_labels
colnames(poly_loadings) <- paste0("PC", seq_len(ncol(poly_loadings)))

load_df <- as_tibble(poly_loadings[, 1:2], rownames = "item") %>%
  mutate(item = factor(item, levels = hwise_labels))

p_poly_load <- ggplot(load_df, aes(x = PC1, y = PC2, label = item)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey70") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  geom_segment(aes(xend = 0, yend = 0),
               arrow = arrow(length = unit(0.2, "cm"), ends = "first"),
               colour = "#4DAF4A", alpha = 0.7) +
  geom_point(colour = "#4DAF4A", size = 3) +
  ggrepel::geom_text_repel(size = 3.3, max.overlaps = 20) +
  labs(
    title   = "Polychoric PCA loadings — PC1 vs PC2",
    x       = sprintf("PC1 (%.1f%% variance)", pct_poly[1]),
    y       = sprintf("PC2 (%.1f%% variance)", pct_poly[2]),
    caption = "Loadings = eigenvector × √eigenvalue."
  ) +
  theme_bw(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        plot.caption = element_text(size = 9, colour = "grey40"))

p_poly_load
ggsave("./output/figures/07.poly_pca_loadings.png", p_poly_load,
       width = 6, height = 5, dpi = 300, bg = "white")
```

## Individual Scores — Coloured by Group

```{r poly-pca-scores, fig.width=10, fig.height=4}
# Project standardised item scores onto eigenvectors
hwise_scaled  <- scale(df_hwise[, hwise_items])
poly_scores   <- hwise_scaled %*% eig$vectors
colnames(poly_scores) <- paste0("PC", seq_len(ncol(poly_scores)))

scores_poly <- as_tibble(poly_scores[, 1:2]) %>%
  bind_cols(df_hwise %>% select(WI_NEIGHBORHOOD, SEASON))

p_pwi <- scores_plot(
  scores_poly, "WI_NEIGHBORHOOD",
  "PC scores by neighbourhood type",
  c("Water-secure" = "#377EB8", "Water-insecure" = "#E41A1C")
)

p_psea <- scores_plot(
  scores_poly, "SEASON",
  "PC scores by season",
  c("Rainy" = "#4DAF4A", "Dry" = "#FF7F00")
)

p_poly_scores <- p_pwi + p_psea +
  patchwork::plot_annotation(
    title   = "Polychoric PCA — individual scores",
    caption = "Shaded ellipses = 95% normal confidence regions."
  )

p_poly_scores
ggsave("./output/figures/07.poly_pca_scores.png", p_poly_scores,
       width = 10, height = 4, dpi = 300, bg = "white")
```

## Loadings Heatmap (first 4 PCs)

```{r poly-pca-heatmap, fig.width=7, fig.height=5}
heat_poly <- as_tibble(poly_loadings[, 1:4], rownames = "item") %>%
  mutate(item = factor(item, levels = hwise_labels)) %>%
  pivot_longer(PC1:PC4, names_to = "PC", values_to = "loading")

p_poly_heat <- ggplot(heat_poly, aes(x = PC, y = fct_rev(item), fill = loading)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%.2f", loading)), size = 2.8) +
  scale_fill_gradient2(low = "#E41A1C", mid = "white", high = "#377EB8",
                       midpoint = 0, name = "Loading") +
  labs(
    title = "Polychoric PCA loadings heatmap — first four components",
    x = NULL, y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(panel.grid = element_blank(),
        axis.text.y = element_text(size = 10))

p_poly_heat
ggsave("./output/figures/07.poly_pca_heatmap.png", p_poly_heat,
       width = 7, height = 5, dpi = 300, bg = "white")
```

---

# Mixed-Data PCA (FAMD)

**Factor Analysis of Mixed Data (FAMD)** handles continuous and categorical
variables simultaneously. Continuous variables are standardised; categorical
variables are dummy-coded and weighted so both types contribute equally to the
decomposition (`FactoMineR::FAMD`).

Variables included:

| Type | Variables |
|------|-----------|
| Continuous | 10 HWISE-reduced items, PSS total, Age, SES score, Household size |
| Categorical | WI neighbourhood, Season, Water-related emotion (EMO_NEG) |

```{r famd-prep}
library(FactoMineR)
library(factoextra)

famd_vars_cont <- c(
  "HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
  "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
  "HW_SLEEP", "HW_NONE",
  "PSS_TOTAL", "D_AGE", "SES_TOTAL", "D_HH_SIZE"
)
famd_vars_cat <- c("WI_NEIGHBORHOOD", "SEASON", "EMO_NEG")

# Pretty labels reused across plots
famd_labels <- c(
  HW_WORRY    = "Worry",        HW_INTERR  = "Interruption",
  HW_CLOTHES  = "Clothes",      HW_PLANS   = "Plans",
  HW_FOOD     = "Food",         HW_HANDS   = "Hands",
  HW_BODY     = "Body",         HW_DRINK   = "Drink",
  HW_SLEEP    = "Sleep",        HW_NONE    = "None",
  PSS_TOTAL   = "PSS total",    D_AGE      = "Age",
  SES_TOTAL = "SES score",   D_HH_SIZE  = "Household size",
  WI_NEIGHBORHOOD = "WI neighbourhood",
  SEASON      = "Season",
  EMO_NEG     = "Neg. emotion"
)

df_famd <- df %>%
  select(all_of(c(famd_vars_cont, famd_vars_cat))) %>%
  mutate(
    EMO_NEG = factor(EMO_NEG, levels = c(0, 1),
                     labels = c("Positive", "Negative"))
  ) %>%
  drop_na()

cat("N complete cases for FAMD:", nrow(df_famd), "\n")
cat("Variables:", ncol(df_famd), "\n")

famd_fit <- FAMD(
  df_famd,
  ncp   = 10,
  graph = FALSE
)
```

## Scree Plot

```{r famd-scree, fig.width=6, fig.height=4}
eig_famd <- as.data.frame(famd_fit$eig)
colnames(eig_famd) <- c("eigenvalue", "pct", "cum_pct")

scree_famd <- tibble(
  PC      = factor(paste0("Dim", seq_len(nrow(eig_famd))),
                   levels = paste0("Dim", seq_len(nrow(eig_famd)))),
  eigenvalue  = eig_famd$eigenvalue,
  pct         = eig_famd$pct,
  cum_pct     = eig_famd$cum_pct
)

p_famd_scree <- ggplot(scree_famd, aes(x = PC, y = eigenvalue)) +
  geom_col(fill = "#984EA3", alpha = 0.8, width = 0.6) +
  geom_line(aes(group = 1), colour = "#FF7F00", linewidth = 0.7) +
  geom_point(colour = "#FF7F00", size = 2.5) +
  geom_hline(yintercept = 1, linetype = "dashed", colour = "grey50") +
  annotate("text", x = nrow(eig_famd) - 1.5, y = 1.12,
           label = "Kaiser criterion (λ = 1)", size = 3, colour = "grey40") +
  geom_text(aes(label = sprintf("%.1f%%", pct), y = eigenvalue + 0.05),
            size = 2.6, vjust = 0) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(
    title   = "FAMD — scree plot",
    x       = NULL, y = "Eigenvalue",
    caption = sprintf("Cumulative variance Dim1–Dim2: %.1f%%",
                      scree_famd$cum_pct[2])
  ) +
  theme_bw(base_size = 12) +
  theme(panel.grid.minor = element_blank(),
        plot.caption = element_text(size = 9, colour = "grey40"))

p_famd_scree
ggsave("./output/figures/07.famd_scree.png", p_famd_scree,
       width = 6, height = 4, dpi = 300, bg = "white")
```

## Variable Contributions to Dim 1 & 2

```{r famd-var-contrib, fig.width=8, fig.height=5}
var_famd <- famd_fit$var

contrib_df <- bind_rows(
  as_tibble(var_famd$contrib[, 1:2], rownames = "variable") %>%
    mutate(type = "Contribution (%)"),
  as_tibble(var_famd$cos2[, 1:2], rownames = "variable") %>%
    mutate(type = "Cos² (quality)")
) %>%
  pivot_longer(c(Dim.1, Dim.2), names_to = "dim", values_to = "value") %>%
  mutate(
    label    = coalesce(famd_labels[variable], variable),
    dim      = recode(dim, Dim.1 = "Dimension 1", Dim.2 = "Dimension 2"),
    var_type = case_when(
      variable %in% famd_vars_cont ~ "Continuous",
      TRUE                         ~ "Categorical"
    )
  )

p_famd_contrib <- contrib_df %>%
  filter(type == "Contribution (%)") %>%
  mutate(label = fct_reorder(label, value, .fun = max)) %>%
  ggplot(aes(x = value, y = label, fill = var_type)) +
  geom_col(alpha = 0.85) +
  geom_vline(
    data = ~ distinct(., dim) %>%
      mutate(threshold = 100 / length(famd_vars_cont) +
               length(famd_vars_cat)),
    aes(xintercept = 100 / (length(famd_vars_cont) + length(famd_vars_cat))),
    linetype = "dashed", colour = "grey40"
  ) +
  facet_wrap(~ dim) +
  scale_fill_manual(values = c(Continuous = "#377EB8", Categorical = "#E41A1C"),
                    name = "Variable type") +
  labs(
    title   = "FAMD — variable contributions to Dim 1 & 2",
    x       = "Contribution (%)", y = NULL,
    caption = "Dashed line = average contribution if all variables contributed equally."
  ) +
  theme_bw(base_size = 11) +
  theme(legend.position  = "bottom",
        panel.grid.minor = element_blank(),
        plot.caption     = element_text(size = 9, colour = "grey40"))

p_famd_contrib
ggsave("./output/figures/07.famd_var_contrib.png", p_famd_contrib,
       width = 8, height = 5, dpi = 300, bg = "white")
```

## Variable Map (Dim 1 vs Dim 2)

```{r famd-var-map, fig.width=7, fig.height=6}
coords_df <- as_tibble(var_famd$coord[, 1:2], rownames = "variable") %>%
  mutate(
    label    = coalesce(famd_labels[variable], variable),
    var_type = if_else(variable %in% famd_vars_cont, "Continuous", "Categorical"),
    cos2_sum = var_famd$cos2[, 1] + var_famd$cos2[, 2]
  )

p_famd_var <- ggplot(coords_df,
                     aes(x = Dim.1, y = Dim.2,
                         colour = var_type, size = cos2_sum,
                         label = label)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey70") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  annotate("path",
           x = cos(seq(0, 2 * pi, length.out = 100)),
           y = sin(seq(0, 2 * pi, length.out = 100)),
           colour = "grey70", linetype = "dotted") +
  geom_segment(aes(xend = 0, yend = 0),
               arrow = arrow(length = unit(0.18, "cm"), ends = "first"),
               alpha = 0.5, show.legend = FALSE) +
  geom_point() +
  ggrepel::geom_text_repel(size = 3.1, max.overlaps = 25,
                            show.legend = FALSE) +
  scale_colour_manual(values = c(Continuous = "#377EB8", Categorical = "#E41A1C"),
                      name = "Variable type") +
  scale_size_continuous(range = c(1.5, 5), name = "Cos² Dim1+2") +
  labs(
    title   = "FAMD — variable map (Dim 1 vs Dim 2)",
    x       = sprintf("Dim 1 (%.1f%%)", eig_famd$pct[1]),
    y       = sprintf("Dim 2 (%.1f%%)", eig_famd$pct[2]),
    caption = "Point size = representation quality (cos²). Unit circle shown."
  ) +
  theme_bw(base_size = 12) +
  theme(legend.position  = "bottom",
        panel.grid.minor = element_blank(),
        plot.caption     = element_text(size = 9, colour = "grey40"))

p_famd_var
ggsave("./output/figures/07.famd_var_map.png", p_famd_var,
       width = 7, height = 6, dpi = 300, bg = "white")
```

## Individual Scores — Coloured by Group

```{r famd-scores, fig.width=10, fig.height=4}
scores_famd <- as_tibble(famd_fit$ind$coord[, 1:2],
                          .name_repair = ~ c("Dim1", "Dim2")) %>%
  bind_cols(df_famd %>% select(WI_NEIGHBORHOOD, SEASON))

p_fwi <- ggplot(scores_famd,
                aes(x = Dim1, y = Dim2, colour = WI_NEIGHBORHOOD)) +
  geom_point(alpha = 0.4, size = 1.2) +
  stat_ellipse(aes(fill = WI_NEIGHBORHOOD),
               geom = "polygon", alpha = 0.08, linewidth = 0.7) +
  scale_colour_manual(values = c("Water-secure"   = "#377EB8",
                                  "Water-insecure" = "#E41A1C"),
                      name = NULL, na.translate = FALSE) +
  scale_fill_manual(values   = c("Water-secure"   = "#377EB8",
                                  "Water-insecure" = "#E41A1C"),
                    name = NULL, na.translate = FALSE) +
  labs(title = "By neighbourhood type",
       x = sprintf("Dim 1 (%.1f%%)", eig_famd$pct[1]),
       y = sprintf("Dim 2 (%.1f%%)", eig_famd$pct[2])) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())

p_fsea <- ggplot(scores_famd,
                 aes(x = Dim1, y = Dim2, colour = SEASON)) +
  geom_point(alpha = 0.4, size = 1.2) +
  stat_ellipse(aes(fill = SEASON),
               geom = "polygon", alpha = 0.08, linewidth = 0.7) +
  scale_colour_manual(values = c("Rainy" = "#4DAF4A", "Dry" = "#FF7F00"),
                      name = NULL, na.translate = FALSE) +
  scale_fill_manual(values   = c("Rainy" = "#4DAF4A", "Dry" = "#FF7F00"),
                    name = NULL, na.translate = FALSE) +
  labs(title = "By season",
       x = sprintf("Dim 1 (%.1f%%)", eig_famd$pct[1]),
       y = sprintf("Dim 2 (%.1f%%)", eig_famd$pct[2])) +
  theme_bw(base_size = 11) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())

p_famd_scores <- p_fwi + p_fsea +
  patchwork::plot_annotation(
    title   = "FAMD — individual scores",
    caption = "Shaded ellipses = 95% normal confidence regions."
  )

p_famd_scores
ggsave("./output/figures/07.famd_scores.png", p_famd_scores,
       width = 10, height = 4, dpi = 300, bg = "white")
```

## Cos² Heatmap — Variable Quality Across Dimensions

```{r famd-cos2-heatmap, fig.width=7, fig.height=5}
n_dims  <- min(5, ncol(var_famd$cos2))
cos2_df <- as_tibble(var_famd$cos2[, 1:n_dims], rownames = "variable") %>%
  mutate(label = coalesce(famd_labels[variable], variable),
         var_type = if_else(variable %in% famd_vars_cont,
                            "Continuous", "Categorical")) %>%
  arrange(var_type, desc(Dim.1)) %>%
  pivot_longer(-c(variable, label, var_type), names_to = "dim", values_to = "cos2") %>%
  mutate(
    label = factor(label, levels = unique(label)),
    dim   = str_replace(dim, "\\.", " ")
  )

p_famd_cos2 <- ggplot(cos2_df,
                       aes(x = dim, y = fct_rev(label), fill = cos2)) +
  geom_tile(colour = "white", linewidth = 0.4) +
  geom_text(aes(label = sprintf("%.2f", cos2)), size = 2.6) +
  facet_grid(var_type ~ ., scales = "free_y", space = "free_y") +
  scale_fill_gradient(low = "white", high = "#984EA3", limits = c(0, 1),
                      name = "Cos²") +
  labs(
    title   = "FAMD — variable quality (cos²) across first five dimensions",
    x       = NULL, y = NULL,
    caption = "Higher cos² = variable better represented on that dimension."
  ) +
  theme_bw(base_size = 11) +
  theme(panel.grid    = element_blank(),
        strip.text    = element_text(face = "bold"),
        plot.caption  = element_text(size = 9, colour = "grey40"))

p_famd_cos2
ggsave("./output/figures/07.famd_cos2_heatmap.png", p_famd_cos2,
       width = 7, height = 5, dpi = 300, bg = "white")
```

---

# Sensitivity Analysis: SES composite vs Education score

Comparison of primary-predictor estimates across all models when using
**composite SES score** (`SES_SC_Total`, main analysis) vs **ordinal education score**
(`SES_EDU_SC_CLEAN`, sensitivity) as the SES covariate.

```{r ses-sensitivity}
# ── Prepare data with EDU_SCORE as alternative covariate ────────────────────
df_edu <- df %>%
  mutate(EDU_SCORE = clean_num(SES_EDU_SC_CLEAN))

# ── Helper: refit model swapping SES_TOTAL → EDU_SCORE ──────────────────────
swap_to_edu <- function(mod) {
  f_old  <- formula(mod)
  f_new  <- as.formula(gsub("SES_TOTAL", "EDU_SCORE", deparse(f_old)))
  if (inherits(mod, "glm")) {
    glm(f_new, data = df_edu, family = binomial)
  } else {
    lm(f_new, data = df_edu)
  }
}

# ── Refit all 19 models ──────────────────────────────────────────────────────
panel_a_mods <- list(
  m_a1, m_a2, m_a3, m_a4, m_a5, m_a6, m_a7, m_a8, m_a9, m_a10
)
panel_b_mods <- list(
  m_b1, m_b2, m_b3, m_b4, m_b5, m_b6, m_b7, m_b8, m_b9
)

panel_a_edu <- lapply(panel_a_mods, swap_to_edu)
panel_b_edu <- lapply(panel_b_mods, swap_to_edu)

# ── Terms to keep (primary predictors only) ──────────────────────────────────
PRIMARY_TERMS <- c(
  "HWISE_REDUCED", "HWISE_FULL",
  "WI_NEIGHBORHOODWater-insecure",
  "SEASONDry",
  "HWISE_REDUCED:SEASONDry",
  "WI_NEIGHBORHOODWater-insecure:SEASONDry",
  "HWISE_REDUCED:WI_NEIGHBORHOODWater-insecure"
)

TERM_LABELS <- c(
  "HWISE_REDUCED"                               = "HWISE reduced",
  "HWISE_FULL"                                  = "HWISE full",
  "WI_NEIGHBORHOODWater-insecure"               = "WI neighbourhood",
  "SEASONDry"                                   = "Season (Dry)",
  "HWISE_REDUCED:SEASONDry"                     = "HWISE × Season",
  "WI_NEIGHBORHOODWater-insecure:SEASONDry"     = "WI nbhd × Season",
  "HWISE_REDUCED:WI_NEIGHBORHOODWater-insecure" = "HWISE × WI nbhd"
)

# ── Extract coefficients for one panel ──────────────────────────────────────
extract_comparison <- function(mods_edu, mods_ses, model_names, is_logistic = FALSE) {
  bind_rows(
    lapply(seq_along(mods_edu), function(i) {
      get_coefs <- function(mod, covariate_label) {
        tidy(mod, conf.int = TRUE, exponentiate = is_logistic) %>%
          filter(term %in% PRIMARY_TERMS) %>%
          mutate(
            model      = model_names[i],
            covariate  = covariate_label,
            term_label = recode(term, !!!TERM_LABELS, .default = term)
          ) %>%
          select(model, covariate, term_label, estimate, conf.low, conf.high, p.value)
      }
      bind_rows(
        get_coefs(mods_edu[[i]], "SES composite"),
        get_coefs(mods_ses[[i]], "EDU score")
      )
    })
  )
}

model_names_a <- c(
  "a1: HWISE reduced", "a2: HWISE × Season",
  "a3: WI neighbourhood", "a4: WI nbhd × Season",
  "a5: HWISE + Season", "a6: WI nbhd + Season",
  "a7: Season only", "a8: HWISE × WI nbhd",
  "a9: HWISE + WI nbhd", "a10: HWISE full"
)

model_names_b <- c(
  "b1: HWISE reduced", "b2: HWISE × Season",
  "b3: WI neighbourhood", "b4: WI nbhd × Season",
  "b5: HWISE + Season", "b6: WI nbhd + Season",
  "b7: Season only", "b8: HWISE × WI nbhd",
  "b9: HWISE + WI nbhd"
)

cmp_a <- extract_comparison(panel_a_mods, panel_a_edu, model_names_a, is_logistic = FALSE)
cmp_b <- extract_comparison(panel_b_mods, panel_b_edu, model_names_b, is_logistic = TRUE)

# ── Format helper ─────────────────────────────────────────────────────────────
fmt_est <- function(est, lo, hi, p, digits = 3) {
  sprintf("%.*f [%.*f, %.*f]%s",
          digits, est, digits, lo, digits, hi,
          ifelse(p < 0.001, "***", ifelse(p < 0.01, "**",
                 ifelse(p < 0.05, "*", ifelse(p < 0.1, "†", "")))))
}

make_wide_table <- function(cmp) {
  cmp %>%
    mutate(cell = mapply(fmt_est, estimate, conf.low, conf.high, p.value)) %>%
    select(model, term_label, covariate, cell) %>%
    pivot_wider(names_from = covariate, values_from = cell) %>%
    arrange(model, term_label) %>%
    rename(
      Model           = model,
      Predictor       = term_label
    )
}

tbl_cmp_a <- make_wide_table(cmp_a)
tbl_cmp_b <- make_wide_table(cmp_b)
```

## Panel A — PSS (linear): SES composite vs EDU score

> Values shown as **β [95% CI]**. Significance: † p<0.1, \* p<0.05, \*\* p<0.01, \*\*\* p<0.001.

```{r ses-cmp-table-a}
flextable(tbl_cmp_a) %>%
  set_header_labels(
    Model           = "Model",
    Predictor       = "Predictor",
    `SES composite` = "SES composite (β [95% CI])",
    `EDU score`     = "EDU score (β [95% CI])"
  ) %>%
  bold(part = "header") %>%
  bg(part = "header", bg = "#2c3e50") %>%
  color(part = "header", color = "white") %>%
  merge_v(j = "Model") %>%
  valign(j = "Model", valign = "top") %>%
  autofit() %>%
  theme_booktabs()
```

## Panel B — EMO_NEG (logistic): SES composite vs EDU score

> Values shown as **OR [95% CI]**. Significance: † p<0.1, \* p<0.05, \*\* p<0.01, \*\*\* p<0.001.

```{r ses-cmp-table-b}
flextable(tbl_cmp_b) %>%
  set_header_labels(
    Model           = "Model",
    Predictor       = "Predictor",
    `SES composite` = "SES composite (OR [95% CI])",
    `EDU score`     = "EDU score (OR [95% CI])"
  ) %>%
  bold(part = "header") %>%
  bg(part = "header", bg = "#2c3e50") %>%
  color(part = "header", color = "white") %>%
  merge_v(j = "Model") %>%
  valign(j = "Model", valign = "top") %>%
  autofit() %>%
  theme_booktabs()
```

## Coefficient plot comparison — Panel A

```{r ses-cmp-plot-a, fig.width=10, fig.height=6}
p_cmp_a <- ggplot(cmp_a, aes(
    x = estimate, xmin = conf.low, xmax = conf.high,
    y = fct_rev(term_label), colour = covariate
  )) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(position = position_dodge(width = 0.5), size = 0.5) +
  facet_wrap(~ model, scales = "free_x", ncol = 3) +
  scale_colour_manual(
    values = c("SES composite" = "#2471a3", "EDU score" = "#c0392b"),
    name   = "SES covariate"
  ) +
  labs(
    title   = "Panel A — β coefficients: SES composite vs EDU score",
    caption = "† p<0.1  * p<0.05  ** p<0.01  *** p<0.001",
    x = "β (95% CI)", y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(
    legend.position  = "bottom",
    panel.grid.minor = element_blank(),
    strip.text       = element_text(face = "bold", size = 8),
    plot.caption     = element_text(size = 9, colour = "grey40")
  )

p_cmp_a
ggsave("./output/figures/07.ses_comparison_panel_a.png", p_cmp_a,
       width = 10, height = 6, dpi = 300, bg = "white")
```

## Coefficient plot comparison — Panel B

```{r ses-cmp-plot-b, fig.width=10, fig.height=6}
p_cmp_b <- ggplot(cmp_b, aes(
    x = estimate, xmin = conf.low, xmax = conf.high,
    y = fct_rev(term_label), colour = covariate
  )) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
  geom_pointrange(position = position_dodge(width = 0.5), size = 0.5) +
  facet_wrap(~ model, scales = "free_x", ncol = 3) +
  scale_colour_manual(
    values = c("SES composite" = "#2471a3", "EDU score" = "#c0392b"),
    name   = "SES covariate"
  ) +
  scale_x_log10() +
  labs(
    title   = "Panel B — OR: SES composite vs EDU score",
    caption = "† p<0.1  * p<0.05  ** p<0.01  *** p<0.001  |  Log scale",
    x = "OR (95% CI, log scale)", y = NULL
  ) +
  theme_bw(base_size = 11) +
  theme(
    legend.position  = "bottom",
    panel.grid.minor = element_blank(),
    strip.text       = element_text(face = "bold", size = 8),
    plot.caption     = element_text(size = 9, colour = "grey40")
  )

p_cmp_b
ggsave("./output/figures/07.ses_comparison_panel_b.png", p_cmp_b,
       width = 10, height = 6, dpi = 300, bg = "white")
```
