---
title: "09. Aim 2 Models (HWISE-Adjusted): Components of Water Insecurity"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(gtsummary)
library(lmtest)
library(flextable)
library(ggplot2)
library(officer)
library(ggstance)
```

## Overview

This script fits **HWISE-adjusted Aim 2 regression models** testing water supply
characteristics as predictors of two outcomes, **after controlling for overall
water insecurity (HWISE total score)**. The purpose is to determine whether
water supply components (hours per week, intermittency, pressure, frequency,
etc.) independently predict psychological stress and emotional responses, beyond
what is explained by overall water insecurity as captured by the HWISE total.

- **Part 1**: Perceived Stress Scale total score (`PSS_TOTAL`) — linear regression
- **Part 2**: Water-related emotion valence (`EMO_NEG`) — logistic regression

HWISE individual item models are **excluded** because it does not make
conceptual sense to include both a single HWISE item and the HWISE total score
in the same model. Only water supply characteristic models are run here
(PSS: a6–a_loc; EMO: b6–b_loc), each augmented with `HW_TOTAL` as an
additional core covariate.

| Role | Variables |
|------|-----------|
| PSS outcome | `PSS_TOTAL` (continuous, 0–56) |
| EMO outcome | `EMO_NEG` — 0 = positive, 1 = negative (Q26) |
| Core covariates (PSS) | Age (`D_AGE`), SES composite (`SES_SC_Total`), Household size (`D_HH_SIZE`), Chronic disease (`Chronic_dis`), **HWISE total (`HW_TOTAL`)** |
| Core covariates (EMO) | As above + `PSS_TOTAL` |
| Augmenting (one per model) | Hours of water supply/week · Intermittent supply · Season · Neighborhood comparison · Excessive time handling water · Sufficiency rating · Water pressure · Supply frequency · Neighborhood (WS/WI) · Years in neighbourhood |

> **Multiple testing**: All p-values are adjusted using the
> **Benjamini-Hochberg (BH)** procedure, applied across all models
> within each outcome.

---

# Load and Prepare Data

```{r load-data}
MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

clean_num <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES, NA_real_, x_num)
}

df_raw <- read_csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
                   show_col_types = FALSE)

cat("Dataset:", nrow(df_raw), "rows ×", ncol(df_raw), "cols\n")

# ── Check new column availability ─────────────────────────────────────────────
new_cols <- c("HLTH_CDIS_CAT", "HW_TOTAL", "HW_WORRY_SUB", "HW_MEX_5SUB",
              "MX6_WPRESS", "MX4_WFREQ", "MX4_W_FREQ", "W_WS_LOC",
              "D_LOC_TIME_CLEAN")
cat("\nNew column availability:\n")
for (col in new_cols) {
  cat(sprintf("  %-20s %s\n", col,
              ifelse(col %in% names(df_raw), "FOUND", "*** MISSING ***")))
}
```

```{r prepare-variables}
hwise_items <- c("HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
                 "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
                 "HW_ANGRY", "HW_SLEEP",  "HW_NONE",    "HW_SHAME")

df <- df_raw %>%
  mutate(
    # ── Outcomes ───────────────────────────────────────────────────────────────
    PSS_TOTAL = clean_num(PSS_TOTAL),

    EMO_NEG = {
      raw <- clean_num(MX26_EM_HHW_CAT)
      case_when(
        raw == 0 ~ 0L,
        raw == 1 ~ 1L,
        TRUE     ~ NA_integer_
      )
    },

    # ── Core covariates ────────────────────────────────────────────────────────
    D_AGE        = clean_num(D_AGE),
    SES_SC_Total = clean_num(SES_SC_Total),
    D_HH_SIZE    = clean_num(D_HH_SIZE),
    D_LOC_TIME   = clean_num(D_LOC_TIME_CLEAN),

    # ── HWISE total score (core covariate in these models) ─────────────────────
    HW_TOTAL = clean_num(HW_TOTAL),

    # ── HWISE individual items ──────────────────────────────────────────────────
    across(all_of(hwise_items), clean_num),

    # ── Hours of water supply per week ─────────────────────────────────────────
    HRS_PER_WEEK = clean_num(MX5_WFREQ_HRS_PER_WEEK),

    # ── Continuous vs. intermittent supply ─────────────────────────────────────
    # 0 = Continuous (reference), 1 = Intermittent
    CONT_INT = {
      raw     <- as.character(MX3_CONT_INT)
      raw_num <- suppressWarnings(as.numeric(raw))
      factor(case_when(
        raw_num %in% MISSING_CODES                         ~ NA_character_,
        raw_num == 0                                       ~ "Continuous",
        raw_num == 1                                       ~ "Intermittent",
        str_detect(str_to_lower(raw), "contin")            ~ "Continuous",
        str_detect(str_to_lower(raw), "inter")             ~ "Intermittent",
        is.na(raw) | str_trim(raw) == "" | raw == "NA"    ~ NA_character_,
        TRUE                                               ~ raw
      ), levels = c("Continuous", "Intermittent"))
    },

    # ── Season ─────────────────────────────────────────────────────────────────
    # 0 = Season 1 (rainy, reference), 1 = Season 2 (dry)
    SEASON = factor(SEASON, levels = c(0, 1), labels = c("Rainy", "Dry")),

    # ── Neighborhood water comparison ──────────────────────────────────────────
    # 0 = Better (reference), 1 = Same, 2 = Worse
    NB_COMP = factor(
      case_when(
        clean_num(MX28_WQ_COMP_CAT) == 0 ~ "Better",
        clean_num(MX28_WQ_COMP_CAT) == 1 ~ "Same",
        clean_num(MX28_WQ_COMP_CAT) == 2 ~ "Worse",
        TRUE                             ~ NA_character_
      ),
      levels = c("Better", "Same", "Worse")
    ),

    # ── Excessive time handling water ──────────────────────────────────────────
    # 0 = No (reference), 1 = Yes
    TIME_WATER = factor(
      case_when(
        clean_num(MX19_TIME_CAT) == 0 ~ "No",
        clean_num(MX19_TIME_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── Sufficiency rating ─────────────────────────────────────────────────────
    # 1 = not sufficient, 4 = completely sufficient
    SUFF_RATING = clean_num(MX16_SCEN_99),

    # ── Chronic disease ────────────────────────────────────────────────────────
    # 0 = No (reference), 1 = Yes
    Chronic_dis = factor(
      case_when(
        clean_num(HLTH_CDIS_CAT) == 0 ~ "No",
        clean_num(HLTH_CDIS_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── Water pressure ─────────────────────────────────────────────────────────
    W_Pressure = factor(
      case_when(
        str_detect(str_to_lower(as.character(MX6_WPRESS)),
                   "s[i\u00ed]|yes|buena|bien")  ~ "Yes",
        str_detect(str_to_lower(as.character(MX6_WPRESS)),
                   "^no|baja|poca")               ~ "No/Low",
        TRUE                                      ~ NA_character_
      ),
      levels = c("Yes", "No/Low")
    ),

    # ── Supply frequency ───────────────────────────────────────────────────────
    W_Frequency = {
      freq_raw <- if ("MX4_W_FREQ" %in% names(df_raw)) {
        dplyr::coalesce(as.character(MX4_WFREQ), as.character(MX4_W_FREQ))
      } else {
        as.character(MX4_WFREQ)
      }
      raw <- str_to_lower(as.character(freq_raw))
      factor(
        case_when(
          str_detect(raw, "^contin")           ~ "Continuous",
          str_detect(raw, "^diari|every day")  ~ "Daily",
          str_detect(raw, "4 a 6|4-6")         ~ "4\u20136 days/week",
          str_detect(raw, "1 a 3|1-3")         ~ "1\u20133 days/week",
          str_detect(raw, "^intermi")          ~ "Intermittent",
          TRUE                                 ~ NA_character_
        ),
        levels = c("Continuous", "Daily", "4\u20136 days/week",
                   "1\u20133 days/week", "Intermittent")
      )
    },

    # ── Neighborhood water security ────────────────────────────────────────────
    W_WS_LOC_F = if ("W_WS_LOC" %in% names(df_raw)) {
      factor(W_WS_LOC, levels = c("WS", "WI"),
             labels = c("Water Secure", "Water Insecure"))
    } else {
      factor(NA_character_, levels = c("Water Secure", "Water Insecure"))
    },

    # ── Water infrastructure & cost (Q8, Q9, Q10, Q11) ────────────────────────
    W_Trust = if ("MX8_TRUST_CAT" %in% names(df_raw)) {
      factor(case_when(
        clean_num(MX8_TRUST_CAT) == 1 ~ "Yes",
        clean_num(MX8_TRUST_CAT) == 0 ~ "No",
        TRUE                          ~ NA_character_
      ), levels = c("Yes", "No"))
    } else factor(NA_character_, levels = c("Yes", "No")),

    W_Garrafon_brand = if ("MX9_GARRAFON_BRAND" %in% names(df_raw)) {
      factor(case_when(
        clean_num(MX9_GARRAFON_BRAND) == 1 ~ "Yes",
        clean_num(MX9_GARRAFON_BRAND) == 0 ~ "No",
        TRUE                               ~ NA_character_
      ), levels = c("No", "Yes"))
    } else factor(NA_character_, levels = c("No", "Yes")),

    W_Has_tinaco = if ("MX10_HAS_TINACO" %in% names(df_raw)) {
      factor(case_when(
        clean_num(MX10_HAS_TINACO) == 1 ~ "Yes",
        clean_num(MX10_HAS_TINACO) == 0 ~ "No",
        TRUE                            ~ NA_character_
      ), levels = c("No", "Yes"))
    } else factor(NA_character_, levels = c("No", "Yes")),

    W_Has_cisterna = if ("MX10_HAS_CISTERNA" %in% names(df_raw)) {
      factor(case_when(
        clean_num(MX10_HAS_CISTERNA) == 1 ~ "Yes",
        clean_num(MX10_HAS_CISTERNA) == 0 ~ "No",
        TRUE                              ~ NA_character_
      ), levels = c("No", "Yes"))
    } else factor(NA_character_, levels = c("No", "Yes")),

    W_Bill_waived = if ("MX11_BILL_WAIVED" %in% names(df_raw)) {
      factor(case_when(
        clean_num(MX11_BILL_WAIVED) == 1 ~ "Yes",
        clean_num(MX11_BILL_WAIVED) == 0 ~ "No",
        TRUE                             ~ NA_character_
      ), levels = c("No", "Yes"))
    } else factor(NA_character_, levels = c("No", "Yes")),

    W_Cost_2mnths = if ("MX11_COST_2MNTHS" %in% names(df_raw)) {
      suppressWarnings(as.numeric(as.character(MX11_COST_2MNTHS)))
    } else NA_real_
  )
```

```{r check-missingness}
analytic_vars <- c("PSS_TOTAL", "EMO_NEG", "D_AGE", "SES_SC_Total", "D_HH_SIZE",
                   "D_LOC_TIME", "SEASON", "Chronic_dis", "HW_TOTAL",
                   "HRS_PER_WEEK", "CONT_INT",
                   "NB_COMP", "TIME_WATER", "SUFF_RATING",
                   "W_Pressure", "W_Frequency", "W_WS_LOC_F")

df %>%
  summarise(across(all_of(analytic_vars), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "n_missing") %>%
  mutate(pct_missing = round(100 * n_missing / nrow(df), 1)) %>%
  arrange(desc(n_missing)) %>%
  print(n = Inf)
```

```{r check-outcomes}
cat("PSS_TOTAL summary:\n")
print(summary(df$PSS_TOTAL))

cat("\nEMO_NEG distribution:\n")
print(table(df$EMO_NEG, useNA = "always"))
cat("\nProportion negative (valid responses):",
    round(mean(df$EMO_NEG == 1, na.rm = TRUE), 3), "\n")

cat("\nHW_TOTAL summary:\n")
print(summary(df$HW_TOTAL))
```

---

# Shared Helpers

```{r shared-helpers}
# ── Non-HWISE / supply term labels ────────────────────────────────────────────
term_labels <- c(
  HRS_PER_WEEK                   = "Hours of supply per week",
  CONT_INTIntermittent           = "Supply continuity: Intermittent (ref: Continuous)",
  SEASONDry                      = "Season: Dry (ref: Rainy)",
  NB_COMPSame                    = "Neighborhood comparison: Same as CDMX",
  NB_COMPWorse                   = "Neighborhood comparison: Worse than CDMX",
  TIME_WATERYes                  = "Excessive time managing water: Yes (ref: No)",
  SUFF_RATING                    = "Sufficiency rating (1\u20134)",
  `W_PressureNo/Low`             = "Adequate water pressure: No/Low (ref: Yes)",
  W_FrequencyDaily               = "Supply frequency: Daily (ref: Continuous)",
  `W_Frequency4–6 days/week`     = "Supply frequency: 4\u20136 days/wk (ref: Continuous)",
  `W_Frequency1–3 days/week`     = "Supply frequency: 1\u20133 days/wk (ref: Continuous)",
  W_FrequencyIntermittent        = "Supply frequency: Intermittent (ref: Continuous)",
  `W_WS_LOC_FWater Insecure`     = "Neighborhood: Water Insecure (ref: Water Secure)",
  D_LOC_TIME                     = "Years in neighbourhood",
  HW_TOTAL                       = "HWISE total score"
)

# ── Publication plot theme ─────────────────────────────────────────────────────
pub_theme <- theme_classic(base_size = 11) +
  theme(
    plot.title         = element_text(face = "bold", size = 12),
    axis.text.y        = element_text(size = 9),
    axis.text.x        = element_text(size = 9),
    legend.position    = "none",
    strip.text         = element_text(face = "bold", size = 9),
    strip.background   = element_rect(fill = "grey92", colour = NA),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    plot.caption       = element_text(size = 9, colour = "grey40", hjust = 0)
  )

sig_sym <- function(p) {
  dplyr::case_when(
    p < 0.001 ~ "***", p < 0.01 ~ "**", p < 0.05 ~ "*",
    p < 0.1   ~ "\u2020", TRUE ~ ""
  )
}

# ── Pretty flextable style ─────────────────────────────────────────────────────
pretty_ft_aim2 <- function(ft) {
  ft %>%
    bold(part = "header") %>%
    bold(j = "Model") %>%
    bg(part = "header", bg = "#2c3e50") %>%
    color(part = "header", color = "white") %>%
    theme_booktabs() %>%
    set_table_properties(layout = "autofit", width = 1)
}
```

---

# Part 1: PSS Models (Linear Regression) {#pss-models}

## Model Specifications

```{r pss-model-specs}
# Core PSS formula: age, SES, household size, chronic disease, HWISE total
# D_LOC_TIME removed per prior changes; HW_TOTAL added as core covariate
core_pss <- PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL

model_specs_pss <- list(
  m_a6 = list(
    label   = "a6",
    group   = "m_a6: Hours supply",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + HRS_PER_WEEK
  ),
  m_a7 = list(
    label   = "a7",
    group   = "m_a7: Intermittent",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + CONT_INT
  ),
  m_a8 = list(
    label   = "a8",
    group   = "m_a8: Season",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + SEASON
  ),
  m_a9 = list(
    label   = "a9",
    group   = "m_a9: Neighborhood comparison",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + NB_COMP
  ),
  m_a10 = list(
    label   = "a10",
    group   = "m_a10: Time burden",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + TIME_WATER
  ),
  m_a11 = list(
    label   = "a11",
    group   = "m_a11: Sufficiency",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + SUFF_RATING
  ),
  m_a16 = list(
    label   = "a16",
    group   = "m_a16: Water pressure",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + W_Pressure
  ),
  m_a17 = list(
    label   = "a17",
    group   = "m_a17: Supply frequency",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + W_Frequency
  ),
  m_a18 = list(
    label   = "a18",
    group   = "m_a18: Neighborhood",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + W_WS_LOC_F
  ),
  m_a_loc = list(
    label   = "a_loc",
    group   = "m_a_loc: Years in neighbourhood",
    formula = PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HW_TOTAL + D_LOC_TIME
  )
)

cat("Total PSS models registered:", length(model_specs_pss), "\n")
cat("Model IDs:\n")
cat(paste(" -", names(model_specs_pss)), sep = "\n")
```

## Fit PSS Models

```{r pss-fit-models}
fitted_pss <- lapply(model_specs_pss, function(spec) {
  lm(spec$formula, data = df, na.action = na.omit)
})

tibble(
  model_id = names(model_specs_pss),
  label    = map_chr(model_specs_pss, "label"),
  n        = map_int(fitted_pss, nobs)
) %>% print(n = Inf)
```

## Summary Table with BH Correction

```{r pss-summary-table}
core_terms_pss <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE",
                    "Chronic_disYes", "HW_TOTAL")

summary_raw_pss <- map_dfr(names(fitted_pss), function(nm) {
  mod  <- fitted_pss[[nm]]
  spec <- model_specs_pss[[nm]]
  fit  <- glance(mod)

  rows_used <- as.integer(rownames(model.frame(mod)))
  m0_sub    <- lm(core_pss, data = df[rows_used, ])
  lrt_tab   <- anova(m0_sub, mod)
  f_stat    <- lrt_tab$F[2]
  f_df1     <- lrt_tab$Df[2]
  f_df2     <- lrt_tab$Res.Df[2]
  f_p       <- lrt_tab$`Pr(>F)`[2]

  focal_coefs <- tidy(mod, conf.int = TRUE) %>%
    filter(!term %in% core_terms_pss)

  if (nrow(focal_coefs) == 0) return(NULL)

  focal_coefs %>%
    mutate(
      model_id     = nm,
      label        = spec$label,
      group        = spec$group,
      n            = nobs(mod),
      adj_R2       = round(fit$adj.r.squared, 3),
      delta_adj_R2 = round(fit$adj.r.squared - glance(m0_sub)$adj.r.squared, 3),
      AIC          = round(fit$AIC, 1),
      F_stat       = round(f_stat, 2),
      F_df1        = as.integer(f_df1),
      F_df2        = as.integer(f_df2),
      F_p          = round(f_p, 4)
    )
})

model_fps_pss <- summary_raw_pss %>%
  distinct(model_id, F_p) %>%
  mutate(F_p_adj = round(p.adjust(F_p, method = "BH"), 4))

summary_pss <- summary_raw_pss %>%
  left_join(model_fps_pss, by = c("model_id", "F_p")) %>%
  mutate(
    beta_ci   = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_coef    = case_when(
      is.na(p.value)  ~ "\u2014",
      p.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", p.value)
    ),
    p_F_fmt   = case_when(
      is.na(F_p)  ~ "\u2014",
      F_p < 0.001 ~ "<0.001",
      TRUE        ~ sprintf("%.3f", F_p)
    ),
    p_adj_fmt = case_when(
      is.na(F_p_adj)  ~ "\u2014",
      F_p_adj < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", F_p_adj)
    )
  )

summary_pss %>%
  arrange(AIC) %>%
  select(label, term, n, adj_R2, delta_adj_R2, AIC, beta_ci,
         p_coef, F_stat, p_F_fmt, p_adj_fmt) %>%
  flextable() %>%
  set_header_labels(
    label        = "Model",
    term         = "Predictor term",
    n            = "N",
    adj_R2       = "Adj. R\u00b2",
    delta_adj_R2 = "\u0394 Adj. R\u00b2",
    AIC          = "AIC",
    beta_ci      = "\u03b2 [95% CI]",
    p_coef       = "p (coef.)",
    F_stat       = "F",
    p_F_fmt      = "p (F-test)",
    p_adj_fmt    = "p_adj (BH)"
  ) %>%
  set_table_properties(layout = "autofit", width = 1)
```

## All PSS Models: Publication Table {#pss-all-models}

```{r pss-all-models, fig.width=10}
readable_terms_pss <- c(
  HRS_PER_WEEK               = "Hours of supply / week",
  CONT_INTIntermittent       = "Intermittent vs. continuous supply",
  SEASONDry                  = "Dry vs. rainy season",
  NB_COMPSame                = "Neighborhood: same as CDMX",
  NB_COMPWorse               = "Neighborhood: worse than CDMX",
  TIME_WATERYes              = "Excessive time managing water",
  SUFF_RATING                = "Sufficiency rating (1\u20134)",
  `W_PressureNo/Low`         = "Water pressure: No/Low",
  W_FrequencyDaily           = "Supply frequency: Daily",
  `W_Frequency4–6 days/week` = "Supply frequency: 4\u20136 days/wk",
  `W_Frequency1–3 days/week` = "Supply frequency: 1\u20133 days/wk",
  W_FrequencyIntermittent    = "Supply frequency: Intermittent",
  `W_WS_LOC_FWater Insecure` = "Neighborhood: Water Insecure",
  D_LOC_TIME                 = "Years in neighbourhood"
)

all_ids_pss <- summary_pss %>%
  distinct(model_id, AIC) %>%
  arrange(AIC) %>%
  mutate(rank = row_number())

top_pss <- summary_pss %>%
  left_join(select(all_ids_pss, model_id, rank), by = "model_id") %>%
  arrange(rank, estimate) %>%
  transmute(
    rank     = rank,
    Model    = label,
    Predictor = coalesce(readable_terms_pss[term], term),
    N        = n,
    beta_ci  = beta_ci,
    adj_r2   = adj_R2,
    delta_r2 = delta_adj_R2,
    AIC      = AIC,
    p_coef   = p_coef,
    p_adj    = p_adj_fmt
  )

ft_pss <- top_pss %>%
  flextable() %>%
  set_header_labels(
    rank      = "#",
    Model     = "Model",
    Predictor = "Focal predictor",
    N         = "N",
    beta_ci   = "\u03b2 [95% CI]",
    adj_r2    = "Adj. R\u00b2",
    delta_r2  = "\u0394 Adj. R\u00b2",
    AIC       = "AIC",
    p_coef    = "p (coef.)",
    p_adj     = "p_adj (BH)"
  ) %>%
  merge_v(j = c("rank", "Model", "N", "adj_r2", "delta_r2", "AIC", "p_adj")) %>%
  valign(j = c("rank", "Model", "N", "adj_r2", "delta_r2", "AIC", "p_adj"),
         valign = "top", part = "body") %>%
  set_caption(paste0(
    "All HWISE-adjusted Aim 2 PSS models ranked by AIC. ",
    "\u03b2 = unstandardised regression coefficient; 95% CI = confidence interval. ",
    "Adj. R\u00b2 = adjusted R-squared of the full model; ",
    "\u0394 Adj. R\u00b2 = change over core-covariate-only model. ",
    "All models adjusted for age, SES, household size, chronic disease, and HWISE total score. ",
    "p_adj (BH) = F-test p-value adjusted using the Benjamini-Hochberg procedure ",
    "across all Aim 2 PSS models."
  )) %>%
  pretty_ft_aim2()

ft_pss

out_pss <- "./output/tables/09.Tables_Aim2_HWISEadj_Models.docx"
doc_pss <- read_docx() %>%
  body_add_par("Aim 2 (HWISE-Adjusted) \u2014 All PSS Models (Linear Regression)", style = "heading 1") %>%
  body_add_flextable(ft_pss)
print(doc_pss, target = out_pss)
cat("Saved:", out_pss, "\n")
```

## Forest Plot: Water Supply Predictors (HWISE-Adjusted) {#pss-forest}

```{r pss-forest, fig.width = 10, fig.height = 8}
# Desired variable order: grouped by multi-category type
# Group 1: single-predictor models (hours, intermittent, time burden, sufficiency)
# Group 2: multi-category supply frequency
# Group 3: multi-category neighbourhood comparison
# Group 4: single-predictor location/neighbourhood models
supply_order_pss <- c(
  "Hours of supply per week",
  "Supply continuity: Intermittent (ref: Continuous)",
  "Excessive time managing water: Yes (ref: No)",
  "Sufficiency rating (1\u20134)",
  "Adequate water pressure: No/Low (ref: Yes)",
  "Season: Dry (ref: Rainy)",
  "Supply frequency: Daily (ref: Continuous)",
  "Supply frequency: 4\u20136 days/wk (ref: Continuous)",
  "Supply frequency: 1\u20133 days/wk (ref: Continuous)",
  "Supply frequency: Intermittent (ref: Continuous)",
  "Neighborhood comparison: Same as CDMX",
  "Neighborhood comparison: Worse than CDMX",
  "Neighborhood: Water Insecure (ref: Water Secure)",
  "Years in neighbourhood"
)

coefs_pss <- summary_pss %>%
  mutate(
    term_label = coalesce(term_labels[term], term),
    sig        = sig_sym(p.value)
  )

coefs_pss$term_label <- factor(coefs_pss$term_label,
                               levels = rev(supply_order_pss))

x_all_pss <- c(coefs_pss$conf.low, coefs_pss$conf.high)
x_lim_pss <- c(min(x_all_pss, na.rm = TRUE), max(x_all_pss, na.rm = TRUE))

p_pss <- ggplot(coefs_pss,
       aes(x = estimate, y = term_label,
           xmin = conf.low, xmax = conf.high,
           colour = p.value < 0.05)) +
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.5) +
  geom_text(aes(label = sig), nudge_y = 0.35, size = 3.5,
            fontface = "bold", show.legend = FALSE) +
  scale_colour_manual(values = c("FALSE" = "grey60", "TRUE" = "#c0392b")) +
  scale_x_continuous(expand = expansion(mult = c(0.05, 0.12))) +
  coord_cartesian(xlim = x_lim_pss) +
  labs(
    title    = "HWISE-Adjusted linear regression \u2014 Perceived Stress Score (\u03b2)",
    subtitle = "Adjusted for age, SES, household size, chronic disease, and HWISE total",
    caption  = "Red = p < 0.05  |  Symbols: \u2020p<0.10  *p<0.05  **p<0.01  ***p<0.001\nEach predictor from a separate model; 95% CI shown.",
    x        = "\u03b2 coefficient (95% CI)",
    y        = NULL
  ) +
  pub_theme

p_pss
ggsave("./output/figures/09.aim2_adjusted_for_hwise_pss_forest.png", p_pss,
       width = 10, height = 8, dpi = 300, bg = "white")
cat("Saved: ./output/figures/09.aim2_adjusted_for_hwise_pss_forest.png\n")
```

## Per-Model Coefficient Tables (PSS) {#pss-coef-tables}

```{r pss-coef-tables, results="asis"}
core_label_list_pss <- list(
  D_AGE        ~ "Age (years)",
  SES_SC_Total ~ "SES composite score",
  D_HH_SIZE    ~ "Household size",
  Chronic_dis  ~ "Chronic disease",
  HW_TOTAL     ~ "HWISE total score"
)

for (nm in names(fitted_pss)) {
  spec <- model_specs_pss[[nm]]
  mod  <- fitted_pss[[nm]]
  fit  <- glance(mod)

  cat("\n\n## ", spec$label, " {.unnumbered}\n\n", sep = "")
  cat("**N =**", nobs(mod),
      "| **R\u00b2 =**", round(fit$r.squared, 3),
      "| **Adj. R\u00b2 =**", round(fit$adj.r.squared, 3),
      "| **AIC =**", round(fit$AIC, 1), "\n\n")

  print(
    tbl_regression(
      mod,
      label    = core_label_list_pss,
      conf.int = TRUE
    ) %>%
      bold_p(t = 0.05) %>%
      italicize_levels() %>%
      modify_caption(paste("**", spec$label, "**"))
  )
}
```

---

# Part 2: Emotion Models (Logistic Regression) {#emo-models}

## Model Specifications

```{r emo-model-specs}
# Core EMO formula: age, SES, household size, PSS total, chronic disease, HWISE total
# D_LOC_TIME removed per prior changes; HW_TOTAL added as core covariate
core_emo <- EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL

model_specs_emo <- list(
  m_b6 = list(
    label   = "b6",
    group   = "m_b6: Hours supply",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + HRS_PER_WEEK
  ),
  m_b7 = list(
    label   = "b7",
    group   = "m_b7: Intermittent",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + CONT_INT
  ),
  m_b8 = list(
    label   = "b8",
    group   = "m_b8: Season",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + SEASON
  ),
  m_b9 = list(
    label   = "b9",
    group   = "m_b9: Neighborhood comparison",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + NB_COMP
  ),
  m_b10 = list(
    label   = "b10",
    group   = "m_b10: Time burden",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + TIME_WATER
  ),
  m_b11 = list(
    label   = "b11",
    group   = "m_b11: Sufficiency",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + SUFF_RATING
  ),
  m_b16 = list(
    label   = "b16",
    group   = "m_b16: Water pressure",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + W_Pressure
  ),
  m_b17 = list(
    label   = "b17",
    group   = "m_b17: Supply frequency",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + W_Frequency
  ),
  m_b18 = list(
    label   = "b18",
    group   = "m_b18: Neighborhood",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + W_WS_LOC_F
  ),
  m_b_loc = list(
    label   = "b_loc",
    group   = "m_b_loc: Years in neighbourhood",
    formula = EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis + HW_TOTAL + D_LOC_TIME
  )
)

cat("Total EMO models registered:", length(model_specs_emo), "\n")
cat("Model IDs:\n")
cat(paste(" -", names(model_specs_emo)), sep = "\n")
```

## Fit EMO Models

```{r emo-fit-models}
fitted_emo <- lapply(model_specs_emo, function(spec) {
  glm(spec$formula, data = df, family = binomial, na.action = na.omit)
})

tibble(
  model_id  = names(model_specs_emo),
  label     = map_chr(model_specs_emo, "label"),
  n         = map_int(fitted_emo, nobs),
  converged = map_lgl(fitted_emo, ~ .x$converged)
) %>% print(n = Inf)
```

## Summary Table with BH Correction

```{r emo-summary-table}
mcfadden_r2 <- function(mod) round(1 - mod$deviance / mod$null.deviance, 3)

core_terms_emo <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE",
                    "PSS_TOTAL", "Chronic_disYes", "HW_TOTAL")

summary_raw_emo <- map_dfr(names(fitted_emo), function(nm) {
  mod  <- fitted_emo[[nm]]
  spec <- model_specs_emo[[nm]]

  rows_used  <- as.integer(rownames(model.frame(mod)))
  m0_sub     <- glm(core_emo, data = df[rows_used, ],
                    family = binomial, na.action = na.omit)
  lrt_result <- lrtest(m0_sub, mod)
  fit        <- glance(mod)

  focal_coefs <- tidy(mod, exponentiate = TRUE, conf.int = TRUE) %>%
    filter(!term %in% core_terms_emo)

  if (nrow(focal_coefs) == 0) return(NULL)

  focal_coefs %>%
    mutate(
      model_id = nm,
      label    = spec$label,
      group    = spec$group,
      n        = nobs(mod),
      mcfadden = mcfadden_r2(mod),
      delta_r2 = round(mcfadden_r2(mod) - mcfadden_r2(m0_sub), 3),
      AIC      = round(fit$AIC, 1),
      LRT_chi2 = round(lrt_result$Chisq[2], 2),
      LRT_df   = as.integer(lrt_result$Df[2]),
      LRT_p    = round(lrt_result$`Pr(>Chisq)`[2], 4)
    )
})

model_lrt_ps <- summary_raw_emo %>%
  distinct(model_id, LRT_p) %>%
  mutate(LRT_p_adj = round(p.adjust(LRT_p, method = "BH"), 4))

summary_emo <- summary_raw_emo %>%
  left_join(model_lrt_ps, by = c("model_id", "LRT_p")) %>%
  mutate(
    or_ci     = sprintf("%.2f [%.2f, %.2f]", estimate, conf.low, conf.high),
    p_coef    = case_when(
      is.na(p.value)  ~ "\u2014",
      p.value < 0.001 ~ "<0.001",
      TRUE            ~ sprintf("%.3f", p.value)
    ),
    p_lrt_fmt = case_when(
      is.na(LRT_p)  ~ "\u2014",
      LRT_p < 0.001 ~ "<0.001",
      TRUE          ~ sprintf("%.3f", LRT_p)
    ),
    p_adj_fmt = case_when(
      is.na(LRT_p_adj)  ~ "\u2014",
      LRT_p_adj < 0.001 ~ "<0.001",
      TRUE              ~ sprintf("%.3f", LRT_p_adj)
    )
  )

summary_emo %>%
  arrange(AIC) %>%
  select(label, term, n, mcfadden, delta_r2, AIC, or_ci,
         p_coef, LRT_chi2, p_lrt_fmt, p_adj_fmt) %>%
  flextable() %>%
  set_header_labels(
    label     = "Model",
    term      = "Predictor term",
    n         = "N",
    mcfadden  = "McFadden R\u00b2",
    delta_r2  = "\u0394 McFadden R\u00b2",
    AIC       = "AIC",
    or_ci     = "OR [95% CI]",
    p_coef    = "p (coef.)",
    LRT_chi2  = "LRT \u03c7\u00b2",
    p_lrt_fmt = "p (LRT)",
    p_adj_fmt = "p_adj (BH)"
  ) %>%
  set_table_properties(layout = "autofit", width = 1)
```

## All EMO Models: Publication Table {#emo-all-models}

```{r emo-all-models}
readable_terms_emo <- c(
  HRS_PER_WEEK               = "Hours of supply / week",
  CONT_INTIntermittent       = "Intermittent vs. continuous supply",
  SEASONDry                  = "Dry vs. rainy season",
  NB_COMPSame                = "Neighborhood: same as CDMX",
  NB_COMPWorse               = "Neighborhood: worse than CDMX",
  TIME_WATERYes              = "Excessive time managing water",
  SUFF_RATING                = "Sufficiency rating (1\u20134)",
  `W_PressureNo/Low`         = "Water pressure: No/Low",
  W_FrequencyDaily           = "Supply frequency: Daily",
  `W_Frequency4–6 days/week` = "Supply frequency: 4\u20136 days/wk",
  `W_Frequency1–3 days/week` = "Supply frequency: 1\u20133 days/wk",
  W_FrequencyIntermittent    = "Supply frequency: Intermittent",
  `W_WS_LOC_FWater Insecure` = "Neighborhood: Water Insecure",
  D_LOC_TIME                 = "Years in neighbourhood"
)

all_ids_emo <- summary_emo %>%
  distinct(model_id, AIC) %>%
  arrange(AIC) %>%
  mutate(rank = row_number())

top_emo <- summary_emo %>%
  left_join(select(all_ids_emo, model_id, rank), by = "model_id") %>%
  arrange(rank, estimate) %>%
  transmute(
    rank      = rank,
    Model     = label,
    Predictor = coalesce(readable_terms_emo[term], term),
    N         = n,
    or_ci     = or_ci,
    mcfadden  = mcfadden,
    delta_r2  = delta_r2,
    AIC       = AIC,
    p_coef    = p_coef,
    p_adj     = p_adj_fmt
  )

ft_emo <- top_emo %>%
  flextable() %>%
  set_header_labels(
    rank      = "#",
    Model     = "Model",
    Predictor = "Focal predictor",
    N         = "N",
    or_ci     = "OR [95% CI]",
    mcfadden  = "McFadden R\u00b2",
    delta_r2  = "\u0394 McFadden R\u00b2",
    AIC       = "AIC",
    p_coef    = "p (coef.)",
    p_adj     = "p_adj (BH)"
  ) %>%
  merge_v(j = c("rank", "Model", "N", "mcfadden", "delta_r2", "AIC", "p_adj")) %>%
  valign(j = c("rank", "Model", "N", "mcfadden", "delta_r2", "AIC", "p_adj"),
         valign = "top", part = "body") %>%
  set_caption(paste0(
    "All HWISE-adjusted Aim 2 emotion models ranked by AIC. ",
    "OR = odds ratio; 95% CI = confidence interval. ",
    "McFadden R\u00b2 = pseudo-R\u00b2 of the full model; ",
    "\u0394 McFadden R\u00b2 = change over core-covariate-only model. ",
    "All models adjusted for age, SES, household size, ",
    "perceived stress (PSS), chronic disease, and HWISE total score. ",
    "p_adj (BH) = LRT p-value adjusted using the Benjamini-Hochberg procedure ",
    "across all Aim 2 emotion models."
  )) %>%
  pretty_ft_aim2()

ft_emo

# Append to same output document (PSS table already written; re-open and add)
doc_emo <- read_docx(out_pss) %>%
  body_add_break() %>%
  body_add_par("Aim 2 (HWISE-Adjusted) \u2014 All Emotion Models (Logistic Regression)",
               style = "heading 1") %>%
  body_add_flextable(ft_emo)
print(doc_emo, target = out_pss)
cat("Updated:", out_pss, "\n")
```

## Forest Plot: Water Supply Predictors (HWISE-Adjusted) {#emo-forest}

```{r emo-forest, fig.width = 10, fig.height = 8}
supply_order_emo <- c(
  "Hours of supply per week",
  "Supply continuity: Intermittent (ref: Continuous)",
  "Excessive time managing water: Yes (ref: No)",
  "Sufficiency rating (1\u20134)",
  "Adequate water pressure: No/Low (ref: Yes)",
  "Season: Dry (ref: Rainy)",
  "Supply frequency: Daily (ref: Continuous)",
  "Supply frequency: 4\u20136 days/wk (ref: Continuous)",
  "Supply frequency: 1\u20133 days/wk (ref: Continuous)",
  "Supply frequency: Intermittent (ref: Continuous)",
  "Neighborhood comparison: Same as CDMX",
  "Neighborhood comparison: Worse than CDMX",
  "Neighborhood: Water Insecure (ref: Water Secure)",
  "Years in neighbourhood"
)

coefs_emo <- summary_emo %>%
  mutate(
    term_label = coalesce(term_labels[term], term),
    sig        = sig_sym(p.value)
  )

coefs_emo$term_label <- factor(coefs_emo$term_label,
                               levels = rev(supply_order_emo))

x_all_emo <- c(coefs_emo$conf.low, coefs_emo$conf.high)
x_lim_emo <- c(min(x_all_emo, na.rm = TRUE), max(x_all_emo, na.rm = TRUE))

p_emo <- ggplot(coefs_emo,
       aes(x = estimate, y = term_label,
           xmin = conf.low, xmax = conf.high,
           colour = p.value < 0.05)) +
  geom_vline(xintercept = 1, linetype = "dashed",
             colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.5) +
  geom_text(aes(label = sig), nudge_y = 0.35, size = 3.5,
            fontface = "bold", show.legend = FALSE) +
  scale_colour_manual(values = c("FALSE" = "grey60", "TRUE" = "#2471a3")) +
  scale_x_log10(expand = expansion(mult = c(0.05, 0.12))) +
  coord_cartesian(xlim = x_lim_emo) +
  labs(
    title    = "HWISE-Adjusted logistic regression \u2014 Negative emotional response (OR)",
    subtitle = "Adjusted for age, SES, household size, chronic disease, and HWISE total",
    caption  = "Blue = p < 0.05  |  Symbols: \u2020p<0.10  *p<0.05  **p<0.01  ***p<0.001\nLog scale  |  Each predictor from a separate model; 95% CI shown.",
    x        = "Odds Ratio (95% CI, log scale)",
    y        = NULL
  ) +
  pub_theme

p_emo
ggsave("./output/figures/09.aim2_adjusted_for_hwise_emo_forest.png", p_emo,
       width = 10, height = 8, dpi = 300, bg = "white")
cat("Saved: ./output/figures/09.aim2_adjusted_for_hwise_emo_forest.png\n")
```

## Per-Model Coefficient Tables (EMO) {#emo-coef-tables}

```{r emo-coef-tables, results="asis"}
core_label_list_emo <- list(
  D_AGE        ~ "Age (years)",
  SES_SC_Total ~ "SES composite score",
  D_HH_SIZE    ~ "Household size",
  PSS_TOTAL    ~ "PSS Total (perceived stress)",
  Chronic_dis  ~ "Chronic disease",
  HW_TOTAL     ~ "HWISE total score"
)

for (nm in names(fitted_emo)) {
  spec <- model_specs_emo[[nm]]
  mod  <- fitted_emo[[nm]]
  fit  <- glance(mod)

  cat("\n\n## ", spec$label, " {.unnumbered}\n\n", sep = "")
  cat("**N =**", nobs(mod),
      "| **McFadden R\u00b2 =**", mcfadden_r2(mod),
      "| **AIC =**", round(fit$AIC, 1), "\n\n")

  print(
    tbl_regression(
      mod,
      exponentiate = TRUE,
      label        = core_label_list_emo,
      conf.int     = TRUE
    ) %>%
      bold_p(t = 0.05) %>%
      italicize_levels() %>%
      modify_caption(paste("**", spec$label, "**  (OR, 95% CI)"))
  )
}
```
