---
title: "10. Full Models: All Significant Predictors"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(gtsummary)
library(flextable)
library(officer)
library(ggplot2)
library(ggstance)

dir.create("./output/figures", recursive = TRUE, showWarnings = FALSE)
dir.create("./output/tables",  recursive = TRUE, showWarnings = FALSE)
```

## Overview

This script builds **full models** for each of the two outcomes by simultaneously
including all predictors that showed a statistically significant focal coefficient
(p < 0.05, from the adjusted Aim 1 or Aim 2 models in files 07–09) in at least
one prior model.

| Outcome | Model type | Core covariates |
|---------|-----------|----------------|
| `PSS_TOTAL` | Linear regression | Age, SES, Household size, Chronic disease |
| `EMO_NEG` | Logistic regression | Age, SES, Household size, Chronic disease, PSS_TOTAL |

> **Collinearity note**: The HWISE items (HW_WORRY, HW_INTERR, etc.) are
> highly correlated with each other. Including multiple items simultaneously
> in the full model may inflate standard errors and produce unstable estimates.
> It is recommended to compare results with and without the individual HWISE
> items (e.g., substituting the HWISE reduced composite score instead). See
> Section 5 for a sensitivity analysis using `HWISE_REDUCED` in place of
> individual items.

> **Note on D_LOC_TIME**: Years in neighborhood (`D_LOC_TIME`) is treated as
> a focal predictor (not a core covariate) in these full models, consistent
> with its role in Aim 2.

---

# Section 1: Data Setup

```{r load-data}
MISSING_CODES <- c(999, 888, 777, 666, 555, 444, 333)

clean_num <- function(x) {
  x_num <- suppressWarnings(as.numeric(as.character(x)))
  ifelse(x_num %in% MISSING_CODES, NA_real_, x_num)
}

df_raw <- read_csv("./data/03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv",
                   show_col_types = FALSE)

cat("Dataset:", nrow(df_raw), "rows ×", ncol(df_raw), "cols\n")

new_cols <- c("HLTH_CDIS_CAT", "HW_TOTAL", "HW_WORRY_SUB", "HW_MEX_5SUB",
              "MX6_WPRESS", "MX4_WFREQ", "MX4_W_FREQ", "W_WS_LOC",
              "D_LOC_TIME_CLEAN")
cat("\nNew column availability:\n")
for (col in new_cols) {
  cat(sprintf("  %-20s %s\n", col,
              ifelse(col %in% names(df_raw), "FOUND", "*** MISSING ***")))
}
```

```{r prepare-variables}
hwise_items <- c("HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
                 "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
                 "HW_ANGRY", "HW_SLEEP",  "HW_NONE",    "HW_SHAME")

hwise_items_reduced <- c("HW_WORRY", "HW_INTERR", "HW_CLOTHES", "HW_PLANS",
                         "HW_FOOD",  "HW_HANDS",  "HW_BODY",    "HW_DRINK",
                         "HW_SLEEP", "HW_NONE")

df <- df_raw %>%
  mutate(
    # ── Outcomes ───────────────────────────────────────────────────────────────
    PSS_TOTAL = clean_num(PSS_TOTAL),

    EMO_NEG = {
      raw <- clean_num(MX26_EM_HHW_CAT)
      case_when(
        raw == 0 ~ 0L,
        raw == 1 ~ 1L,
        TRUE     ~ NA_integer_
      )
    },

    # ── Core covariates ────────────────────────────────────────────────────────
    D_AGE        = clean_num(D_AGE),
    SES_SC_Total = clean_num(SES_SC_Total),
    D_HH_SIZE    = clean_num(D_HH_SIZE),
    D_LOC_TIME   = clean_num(D_LOC_TIME_CLEAN),

    # ── HWISE individual items ─────────────────────────────────────────────────
    across(all_of(hwise_items), clean_num),

    # ── HWISE reduced score (10 items, excluding angry and shame) ─────────────
    HWISE_REDUCED = rowSums(across(all_of(hwise_items_reduced)), na.rm = FALSE),

    # ── Hours of water supply per week ─────────────────────────────────────────
    HRS_PER_WEEK = clean_num(MX5_WFREQ_HRS_PER_WEEK),

    # ── Continuous vs. intermittent supply ─────────────────────────────────────
    # 0 = Continuous (reference), 1 = Intermittent
    CONT_INT = {
      raw     <- as.character(MX3_CONT_INT)
      raw_num <- suppressWarnings(as.numeric(raw))
      factor(case_when(
        raw_num %in% MISSING_CODES                         ~ NA_character_,
        raw_num == 0                                       ~ "Continuous",
        raw_num == 1                                       ~ "Intermittent",
        str_detect(str_to_lower(raw), "contin")            ~ "Continuous",
        str_detect(str_to_lower(raw), "inter")             ~ "Intermittent",
        is.na(raw) | str_trim(raw) == "" | raw == "NA"    ~ NA_character_,
        TRUE                                               ~ raw
      ), levels = c("Continuous", "Intermittent"))
    },

    # ── Season ─────────────────────────────────────────────────────────────────
    # 0 = Season 1 (rainy, reference), 1 = Season 2 (dry)
    SEASON = factor(SEASON, levels = c(0, 1), labels = c("Rainy", "Dry")),

    # ── Neighborhood water comparison ──────────────────────────────────────────
    # 0 = Better (reference), 1 = Same, 2 = Worse
    NB_COMP = factor(
      case_when(
        clean_num(MX28_WQ_COMP_CAT) == 0 ~ "Better",
        clean_num(MX28_WQ_COMP_CAT) == 1 ~ "Same",
        clean_num(MX28_WQ_COMP_CAT) == 2 ~ "Worse",
        TRUE                             ~ NA_character_
      ),
      levels = c("Better", "Same", "Worse")
    ),

    # ── Excessive time handling water ──────────────────────────────────────────
    # 0 = No (reference), 1 = Yes
    TIME_WATER = factor(
      case_when(
        clean_num(MX19_TIME_CAT) == 0 ~ "No",
        clean_num(MX19_TIME_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── Sufficiency rating ─────────────────────────────────────────────────────
    # 1 = not sufficient, 4 = completely sufficient
    SUFF_RATING = clean_num(MX16_SCEN_99),

    # ── Chronic disease ────────────────────────────────────────────────────────
    # 0 = No (reference), 1 = Yes
    Chronic_dis = factor(
      case_when(
        clean_num(HLTH_CDIS_CAT) == 0 ~ "No",
        clean_num(HLTH_CDIS_CAT) == 1 ~ "Yes",
        TRUE                          ~ NA_character_
      ),
      levels = c("No", "Yes")
    ),

    # ── Water pressure ─────────────────────────────────────────────────────────
    W_Pressure = factor(
      case_when(
        str_detect(str_to_lower(as.character(MX6_WPRESS)),
                   "s[ií]|yes|buena|bien")  ~ "Yes",
        str_detect(str_to_lower(as.character(MX6_WPRESS)),
                   "^no|baja|poca")         ~ "No/Low",
        TRUE                                ~ NA_character_
      ),
      levels = c("Yes", "No/Low")
    ),

    # ── Supply frequency ───────────────────────────────────────────────────────
    W_Frequency = {
      freq_raw <- if ("MX4_W_FREQ" %in% names(df_raw)) {
        dplyr::coalesce(as.character(MX4_WFREQ), as.character(MX4_W_FREQ))
      } else {
        as.character(MX4_WFREQ)
      }
      raw <- str_to_lower(as.character(freq_raw))
      factor(
        case_when(
          str_detect(raw, "^contin")           ~ "Continuous",
          str_detect(raw, "^diari|every day")  ~ "Daily",
          str_detect(raw, "4 a 6|4-6")         ~ "4\u20136 days/week",
          str_detect(raw, "1 a 3|1-3")         ~ "1\u20133 days/week",
          str_detect(raw, "^intermi")          ~ "Intermittent",
          TRUE                                 ~ NA_character_
        ),
        levels = c("Continuous", "Daily", "4\u20136 days/week",
                   "1\u20133 days/week", "Intermittent")
      )
    },

    # ── Neighborhood water security ────────────────────────────────────────────
    W_WS_LOC_F = if ("W_WS_LOC" %in% names(df_raw)) {
      factor(W_WS_LOC, levels = c("WS", "WI"),
             labels = c("Water Secure", "Water Insecure"))
    } else {
      factor(NA_character_, levels = c("Water Secure", "Water Insecure"))
    }
  )
```

```{r check-outcomes}
cat("PSS_TOTAL summary:\n")
print(summary(df$PSS_TOTAL))
cat("\nEMO_NEG distribution:\n")
print(table(df$EMO_NEG, useNA = "always"))
cat("\nProportion negative (valid responses):",
    round(mean(df$EMO_NEG == 1, na.rm = TRUE), 3), "\n")
```

---

# Section 2: Identify Significant Predictors

This section re-runs all adjusted models from Aim 1 (07) and Aim 2 (08) inline,
extracts the focal coefficient p-values (unadjusted), and identifies predictors
with p < 0.05.

> **Variable naming note**: Aim 1 uses `SES_TOTAL` and `CHRONIC_DIS` (computed
> separately in that script), while Aim 2 uses `SES_SC_Total` and `Chronic_dis`.
> Here we use Aim 2 naming throughout. Aim 1 predictors
> (`HWISE_REDUCED`, `WI_NEIGHBORHOOD`, `SEASON`) are recreated using Aim 2
> variable prep (SEASON is already prepared above; HWISE_REDUCED and W_WS_LOC_F
> serve as the Aim 2 equivalents).

```{r aim1-models-pss}
# ── Aim 1: PSS models (Panel A) ───────────────────────────────────────────────
# Using Aim 2 variable names: SES_SC_Total, Chronic_dis, W_WS_LOC_F

aim1_pss_models <- list(
  a1 = lm(PSS_TOTAL ~ HWISE_REDUCED + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, na.action = na.omit),
  a2 = lm(PSS_TOTAL ~ HWISE_REDUCED * SEASON + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, na.action = na.omit),
  a3 = lm(PSS_TOTAL ~ W_WS_LOC_F + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, na.action = na.omit),
  a4 = lm(PSS_TOTAL ~ W_WS_LOC_F * SEASON + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, na.action = na.omit),
  a5 = lm(PSS_TOTAL ~ HWISE_REDUCED + SEASON + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, na.action = na.omit),
  a6 = lm(PSS_TOTAL ~ W_WS_LOC_F + SEASON + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, na.action = na.omit),
  a7 = lm(PSS_TOTAL ~ SEASON + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, na.action = na.omit),
  a8 = lm(PSS_TOTAL ~ HWISE_REDUCED * W_WS_LOC_F + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, na.action = na.omit),
  a9 = lm(PSS_TOTAL ~ HWISE_REDUCED + W_WS_LOC_F + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, na.action = na.omit)
)
```

```{r aim2-models-pss}
# ── Aim 2: PSS models (one predictor per model) ───────────────────────────────

hwise_specs_pss <- setNames(
  lapply(hwise_items, function(item) {
    lm(as.formula(paste("PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis +", item)),
       data = df, na.action = na.omit)
  }),
  paste0("m_a5_", tolower(hwise_items))
)

aim2_pss_models <- c(
  hwise_specs_pss,
  list(
    m_a6  = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + HRS_PER_WEEK,
               data = df, na.action = na.omit),
    m_a7  = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + CONT_INT,
               data = df, na.action = na.omit),
    m_a8  = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + SEASON,
               data = df, na.action = na.omit),
    m_a9  = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + NB_COMP,
               data = df, na.action = na.omit),
    m_a10 = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + TIME_WATER,
               data = df, na.action = na.omit),
    m_a11 = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + SUFF_RATING,
               data = df, na.action = na.omit),
    m_a16 = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + W_Pressure,
               data = df, na.action = na.omit),
    m_a17 = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + W_Frequency,
               data = df, na.action = na.omit),
    m_a18 = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + W_WS_LOC_F,
               data = df, na.action = na.omit),
    m_a_loc = lm(PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis + D_LOC_TIME,
                 data = df, na.action = na.omit)
  )
)
```

```{r aim1-models-emo}
# ── Aim 1: EMO models (Panel B) ───────────────────────────────────────────────

aim1_emo_models <- list(
  b1 = glm(EMO_NEG ~ HWISE_REDUCED + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, family = binomial, na.action = na.omit),
  b2 = glm(EMO_NEG ~ HWISE_REDUCED * SEASON + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, family = binomial, na.action = na.omit),
  b3 = glm(EMO_NEG ~ W_WS_LOC_F + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, family = binomial, na.action = na.omit),
  b4 = glm(EMO_NEG ~ W_WS_LOC_F * SEASON + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, family = binomial, na.action = na.omit),
  b5 = glm(EMO_NEG ~ HWISE_REDUCED + SEASON + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, family = binomial, na.action = na.omit),
  b6 = glm(EMO_NEG ~ W_WS_LOC_F + SEASON + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, family = binomial, na.action = na.omit),
  b7 = glm(EMO_NEG ~ SEASON + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, family = binomial, na.action = na.omit),
  b8 = glm(EMO_NEG ~ HWISE_REDUCED * W_WS_LOC_F + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, family = binomial, na.action = na.omit),
  b9 = glm(EMO_NEG ~ HWISE_REDUCED + W_WS_LOC_F + PSS_TOTAL + D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis,
           data = df, family = binomial, na.action = na.omit)
)
```

```{r aim2-models-emo}
# ── Aim 2: EMO models (one predictor per model) ───────────────────────────────

hwise_specs_emo <- setNames(
  lapply(hwise_items, function(item) {
    glm(as.formula(paste("EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis +", item)),
        data = df, family = binomial, na.action = na.omit)
  }),
  paste0("m_b5_", tolower(hwise_items))
)

aim2_emo_models <- c(
  hwise_specs_emo,
  list(
    m_b6  = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis + HRS_PER_WEEK,
                data = df, family = binomial, na.action = na.omit),
    m_b7  = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis + CONT_INT,
                data = df, family = binomial, na.action = na.omit),
    m_b8  = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis + SEASON,
                data = df, family = binomial, na.action = na.omit),
    m_b9  = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis + NB_COMP,
                data = df, family = binomial, na.action = na.omit),
    m_b10 = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis + TIME_WATER,
                data = df, family = binomial, na.action = na.omit),
    m_b11 = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis + SUFF_RATING,
                data = df, family = binomial, na.action = na.omit),
    m_b16 = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis + W_Pressure,
                data = df, family = binomial, na.action = na.omit),
    m_b17 = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis + W_Frequency,
                data = df, family = binomial, na.action = na.omit),
    m_b18 = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis + W_WS_LOC_F,
                data = df, family = binomial, na.action = na.omit),
    m_b_loc = glm(EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + D_LOC_TIME + PSS_TOTAL + Chronic_dis,
                  data = df, family = binomial, na.action = na.omit)
  )
)
```

```{r identify-sig-pss}
# Core terms to exclude from significance screening
core_terms_pss <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE",
                    "Chronic_disYes", "D_LOC_TIME", "PSS_TOTAL")

# Interaction terms to exclude (only main effects considered for full model)
is_interaction <- function(term) str_detect(term, ":")

# Helper: extract focal terms with p < 0.05 from a list of models
get_sig_terms <- function(model_list, core_terms) {
  map_dfr(names(model_list), function(nm) {
    tidy(model_list[[nm]], conf.int = FALSE) %>%
      filter(!term %in% core_terms, !is_interaction(term), p.value < 0.05) %>%
      mutate(model_id = nm)
  })
}

# Identify significant PSS predictors across Aim 1 + Aim 2
sig_pss_aim1 <- get_sig_terms(aim1_pss_models, core_terms_pss)
sig_pss_aim2 <- get_sig_terms(aim2_pss_models, core_terms_pss)

sig_pss_all <- bind_rows(
  sig_pss_aim1 %>% mutate(aim = "Aim 1"),
  sig_pss_aim2 %>% mutate(aim = "Aim 2")
)

cat("Significant PSS focal terms (p < 0.05, unadjusted):\n")
sig_pss_all %>%
  arrange(term, p.value) %>%
  select(aim, model_id, term, estimate, p.value) %>%
  print(n = Inf)
```

```{r identify-sig-emo}
core_terms_emo <- c("(Intercept)", "D_AGE", "SES_SC_Total", "D_HH_SIZE",
                    "Chronic_disYes", "D_LOC_TIME", "PSS_TOTAL")

sig_emo_aim1 <- get_sig_terms(aim1_emo_models, core_terms_emo)
sig_emo_aim2 <- get_sig_terms(aim2_emo_models, core_terms_emo)

sig_emo_all <- bind_rows(
  sig_emo_aim1 %>% mutate(aim = "Aim 1"),
  sig_emo_aim2 %>% mutate(aim = "Aim 2")
)

cat("Significant EMO focal terms (p < 0.05, unadjusted):\n")
sig_emo_all %>%
  arrange(term, p.value) %>%
  select(aim, model_id, term, estimate, p.value) %>%
  print(n = Inf)
```

```{r sig-summary-table}
# ── Summary table of all significant predictors ───────────────────────────────

sig_table <- bind_rows(
  sig_pss_all %>% mutate(outcome = "PSS_TOTAL"),
  sig_emo_all %>% mutate(outcome = "EMO_NEG")
) %>%
  group_by(outcome, term) %>%
  summarise(
    n_models_significant = n(),
    min_p  = min(p.value),
    aims   = paste(sort(unique(aim)), collapse = " & "),
    .groups = "drop"
  ) %>%
  arrange(outcome, min_p)

cat("\nSummary: predictors significant in ≥1 model\n")
print(sig_table, n = Inf)

# ── Extract the unique variable names (stripping level suffixes) ──────────────
# For factors, the term looks like "W_FrequencyDaily" — we need "W_Frequency"
# For HWISE items (numeric), term == variable name

extract_variable_name <- function(terms) {
  # Try to match factor variable names from df
  factor_vars <- c("CONT_INT", "SEASON", "NB_COMP", "TIME_WATER",
                   "W_Pressure", "W_Frequency", "W_WS_LOC_F",
                   "Chronic_dis", "HWISE_REDUCED")
  result <- character(length(terms))
  for (i in seq_along(terms)) {
    matched <- factor_vars[str_starts(terms[i], factor_vars)]
    if (length(matched) > 0) {
      result[i] <- matched[which.max(nchar(matched))]  # longest match
    } else {
      result[i] <- terms[i]  # numeric variables: term == variable name
    }
  }
  result
}

sig_vars_pss <- sig_pss_all %>%
  pull(term) %>%
  unique() %>%
  extract_variable_name() %>%
  unique()

sig_vars_emo <- sig_emo_all %>%
  pull(term) %>%
  unique() %>%
  extract_variable_name() %>%
  unique()

cat("\nUnique significant variable names for PSS full model:\n")
cat(paste(" -", sig_vars_pss), sep = "\n")

cat("\nUnique significant variable names for EMO full model:\n")
cat(paste(" -", sig_vars_emo), sep = "\n")
```

---

# Section 3: PSS Full Model

## Build and Fit

```{r pss-full-model}
# Remove core covariates from the significant predictors list
core_vars_pss <- c("D_AGE", "SES_SC_Total", "D_HH_SIZE", "Chronic_dis",
                   "PSS_TOTAL", "D_LOC_TIME")

sig_vars_pss_clean <- setdiff(sig_vars_pss, core_vars_pss)

if (length(sig_vars_pss_clean) == 0) {
  warning("No significant PSS predictors identified. Using HWISE_REDUCED as placeholder.")
  sig_vars_pss_clean <- "HWISE_REDUCED"
}

cat("PSS full model predictors (beyond core covariates):\n")
cat(paste(" -", sig_vars_pss_clean), sep = "\n")

pss_full_formula <- as.formula(
  paste("PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis +",
        paste(sig_vars_pss_clean, collapse = " + "))
)

cat("\nFull PSS formula:\n")
print(pss_full_formula)

pss_full <- lm(pss_full_formula, data = df, na.action = na.omit)
cat("\nN =", nobs(pss_full), "\n")
cat("Adj. R² =", round(summary(pss_full)$adj.r.squared, 3), "\n")
cat("AIC =", round(AIC(pss_full), 1), "\n")
```

## Regression Table (PSS)

```{r pss-full-table, results="asis"}
label_list_pss <- list(
  D_AGE        ~ "Age (years)",
  SES_SC_Total ~ "SES composite score",
  D_HH_SIZE    ~ "Household size",
  Chronic_dis  ~ "Chronic disease",
  D_LOC_TIME   ~ "Years in neighborhood",
  HWISE_REDUCED ~ "HWISE reduced score (0–40)",
  HRS_PER_WEEK  ~ "Hours of water supply/week",
  CONT_INT      ~ "Supply continuity (ref: Continuous)",
  SEASON        ~ "Season (ref: Rainy)",
  NB_COMP       ~ "Neighborhood comparison (ref: Better)",
  TIME_WATER    ~ "Excessive time managing water (ref: No)",
  SUFF_RATING   ~ "Sufficiency rating (1–4)",
  W_Pressure    ~ "Water pressure (ref: Yes)",
  W_Frequency   ~ "Supply frequency (ref: Continuous)",
  W_WS_LOC_F    ~ "Neighborhood water security (ref: Water Secure)"
)

# Keep only labels relevant to this model
model_vars_pss <- all.vars(pss_full_formula)
label_list_pss_filtered <- label_list_pss[
  sapply(label_list_pss, function(x) as.character(x[[2]]) %in% model_vars_pss)
]

tbl_pss <- tbl_regression(
  pss_full,
  label    = label_list_pss_filtered,
  conf.int = TRUE
) %>%
  bold_p(t = 0.05) %>%
  italicize_levels() %>%
  modify_caption(
    "**Full PSS Model**: All significant predictors from Aim 1 & Aim 2"
  ) %>%
  modify_footnote(
    update = everything() ~ paste0(
      "Linear regression. Adjusted R² = ",
      round(summary(pss_full)$adj.r.squared, 3),
      ". N = ", nobs(pss_full),
      ". p-values are unadjusted."
    )
  )

tbl_pss
```

## Coefficient Plot (PSS)

```{r pss-full-plot, fig.width=9, fig.height=7}
readable_terms <- c(
  HWISE_REDUCED              = "HWISE reduced score",
  HRS_PER_WEEK               = "Hours of supply / week",
  CONT_INTIntermittent       = "Supply continuity: Intermittent",
  SEASONDry                  = "Season: Dry",
  NB_COMPSame                = "Neighborhood comparison: Same as CDMX",
  NB_COMPWorse               = "Neighborhood comparison: Worse than CDMX",
  TIME_WATERYes              = "Excessive time managing water: Yes",
  SUFF_RATING                = "Sufficiency rating (1–4)",
  `W_PressureNo/Low`         = "Water pressure: No/Low",
  W_FrequencyDaily           = "Supply frequency: Daily",
  `W_Frequency4–6 days/week` = "Supply frequency: 4–6 days/wk",
  `W_Frequency1–3 days/week` = "Supply frequency: 1–3 days/wk",
  W_FrequencyIntermittent    = "Supply frequency: Intermittent",
  `W_WS_LOC_FWater Insecure` = "Neighborhood: Water Insecure",
  D_LOC_TIME                 = "Years in neighborhood",
  D_AGE                      = "Age (years)",
  SES_SC_Total               = "SES composite score",
  D_HH_SIZE                  = "Household size",
  Chronic_disYes             = "Chronic disease: Yes",
  HW_WORRY                   = "HWISE: Worry about water",
  HW_INTERR                  = "HWISE: Interrupted supply",
  HW_CLOTHES                 = "HWISE: Couldn't wash clothes",
  HW_PLANS                   = "HWISE: Plans disrupted",
  HW_FOOD                    = "HWISE: Food preparation affected",
  HW_HANDS                   = "HWISE: Couldn't wash hands",
  HW_BODY                    = "HWISE: Couldn't bathe",
  HW_DRINK                   = "HWISE: Not enough to drink",
  HW_ANGRY                   = "HWISE: Felt angry",
  HW_SLEEP                   = "HWISE: Went to sleep thirsty",
  HW_NONE                    = "HWISE: No water at all",
  HW_SHAME                   = "HWISE: Felt shame"
)

coef_pss_full <- tidy(pss_full, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    term_label = coalesce(readable_terms[term], term),
    is_core    = term %in% c("D_AGE", "SES_SC_Total", "D_HH_SIZE", "Chronic_disYes"),
    sig_col    = case_when(
      p.value < 0.05 & !is_core ~ "Significant (p<0.05)",
      !is_core                   ~ "Non-significant",
      TRUE                       ~ "Core covariate"
    )
  ) %>%
  arrange(is_core, estimate)

coef_pss_full$term_label <- factor(coef_pss_full$term_label,
                                   levels = unique(coef_pss_full$term_label))

p_pss_full <- ggplot(coef_pss_full,
                     aes(x = estimate, y = term_label,
                         xmin = conf.low, xmax = conf.high,
                         colour = sig_col, shape = is_core)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(height = 0.3, linewidth = 0.6) +
  geom_point(size = 3) +
  scale_colour_manual(
    name   = NULL,
    values = c(
      "Significant (p<0.05)" = "#c0392b",
      "Non-significant"      = "grey50",
      "Core covariate"       = "#2980b9"
    )
  ) +
  scale_shape_manual(
    name   = NULL,
    values = c("TRUE" = 17, "FALSE" = 16),
    labels = c("TRUE" = "Core covariate", "FALSE" = "Focal predictor")
  ) +
  labs(
    title    = "Full PSS Model — All significant predictors (Aim 1 & Aim 2)",
    subtitle = "Linear regression; outcome: Perceived Stress Score (PSS_TOTAL)",
    caption  = "Red = p < 0.05  |  Blue triangles = core covariates  |  95% CI shown",
    x        = "β coefficient (95% CI)",
    y        = NULL
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title         = element_text(face = "bold", size = 12),
    axis.text.y        = element_text(size = 9),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    legend.position    = "bottom"
  )

p_pss_full

ggsave("./output/figures/10.pss_full_model.png", p_pss_full,
       width = 9, height = 7, dpi = 300, bg = "white")
cat("Saved: ./output/figures/10.pss_full_model.png\n")
```

---

# Section 4: EMO Full Model

## Build and Fit

```{r emo-full-model}
core_vars_emo <- c("D_AGE", "SES_SC_Total", "D_HH_SIZE", "Chronic_dis",
                   "PSS_TOTAL", "D_LOC_TIME")

sig_vars_emo_clean <- setdiff(sig_vars_emo, core_vars_emo)

if (length(sig_vars_emo_clean) == 0) {
  warning("No significant EMO predictors identified. Using SEASON as placeholder.")
  sig_vars_emo_clean <- "SEASON"
}

cat("EMO full model predictors (beyond core covariates):\n")
cat(paste(" -", sig_vars_emo_clean), sep = "\n")

emo_full_formula <- as.formula(
  paste("EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis +",
        paste(sig_vars_emo_clean, collapse = " + "))
)

cat("\nFull EMO formula:\n")
print(emo_full_formula)

emo_full <- glm(emo_full_formula, data = df, family = binomial, na.action = na.omit)
cat("\nN =", nobs(emo_full), "\n")
cat("AIC =", round(AIC(emo_full), 1), "\n")
```

## Regression Table (EMO)

```{r emo-full-table, results="asis"}
label_list_emo <- list(
  D_AGE         ~ "Age (years)",
  SES_SC_Total  ~ "SES composite score",
  D_HH_SIZE     ~ "Household size",
  Chronic_dis   ~ "Chronic disease",
  PSS_TOTAL     ~ "PSS total score",
  D_LOC_TIME    ~ "Years in neighborhood",
  HWISE_REDUCED ~ "HWISE reduced score (0–40)",
  HRS_PER_WEEK  ~ "Hours of water supply/week",
  CONT_INT      ~ "Supply continuity (ref: Continuous)",
  SEASON        ~ "Season (ref: Rainy)",
  NB_COMP       ~ "Neighborhood comparison (ref: Better)",
  TIME_WATER    ~ "Excessive time managing water (ref: No)",
  SUFF_RATING   ~ "Sufficiency rating (1–4)",
  W_Pressure    ~ "Water pressure (ref: Yes)",
  W_Frequency   ~ "Supply frequency (ref: Continuous)",
  W_WS_LOC_F    ~ "Neighborhood water security (ref: Water Secure)"
)

model_vars_emo <- all.vars(emo_full_formula)
label_list_emo_filtered <- label_list_emo[
  sapply(label_list_emo, function(x) as.character(x[[2]]) %in% model_vars_emo)
]

tbl_emo <- tbl_regression(
  emo_full,
  label      = label_list_emo_filtered,
  conf.int   = TRUE,
  exponentiate = TRUE
) %>%
  bold_p(t = 0.05) %>%
  italicize_levels() %>%
  modify_header(estimate ~ "**OR**") %>%
  modify_caption(
    "**Full EMO Model**: All significant predictors from Aim 1 & Aim 2"
  ) %>%
  modify_footnote(
    update = everything() ~ paste0(
      "Logistic regression. Estimates are odds ratios. N = ", nobs(emo_full),
      ". p-values are unadjusted."
    )
  )

tbl_emo
```

## Coefficient Plot (EMO)

```{r emo-full-plot, fig.width=9, fig.height=7}
coef_emo_full <- tidy(emo_full, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    term_label = coalesce(readable_terms[term], term),
    is_core    = term %in% c("D_AGE", "SES_SC_Total", "D_HH_SIZE",
                              "Chronic_disYes", "PSS_TOTAL"),
    sig_col    = case_when(
      p.value < 0.05 & !is_core ~ "Significant (p<0.05)",
      !is_core                   ~ "Non-significant",
      TRUE                       ~ "Core covariate"
    )
  ) %>%
  arrange(is_core, estimate)

coef_emo_full$term_label <- factor(coef_emo_full$term_label,
                                   levels = unique(coef_emo_full$term_label))

p_emo_full <- ggplot(coef_emo_full,
                     aes(x = estimate, y = term_label,
                         xmin = conf.low, xmax = conf.high,
                         colour = sig_col, shape = is_core)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50", linewidth = 0.5) +
  geom_errorbarh(height = 0.3, linewidth = 0.6) +
  geom_point(size = 3) +
  scale_colour_manual(
    name   = NULL,
    values = c(
      "Significant (p<0.05)" = "#c0392b",
      "Non-significant"      = "grey50",
      "Core covariate"       = "#2980b9"
    )
  ) +
  scale_shape_manual(
    name   = NULL,
    values = c("TRUE" = 17, "FALSE" = 16),
    labels = c("TRUE" = "Core covariate", "FALSE" = "Focal predictor")
  ) +
  scale_x_log10() +
  labs(
    title    = "Full EMO Model — All significant predictors (Aim 1 & Aim 2)",
    subtitle = "Logistic regression; outcome: Negative water emotion (EMO_NEG)",
    caption  = "Red = p < 0.05  |  Blue triangles = core covariates  |  95% CI shown\nX-axis on log scale; OR = 1 (dashed line) indicates no association.",
    x        = "Odds Ratio (95% CI, log scale)",
    y        = NULL
  ) +
  theme_classic(base_size = 11) +
  theme(
    plot.title         = element_text(face = "bold", size = 12),
    axis.text.y        = element_text(size = 9),
    panel.grid.major.x = element_line(colour = "grey88", linewidth = 0.3),
    legend.position    = "bottom"
  )

p_emo_full

ggsave("./output/figures/10.emo_full_model.png", p_emo_full,
       width = 9, height = 7, dpi = 300, bg = "white")
cat("Saved: ./output/figures/10.emo_full_model.png\n")
```

---

# Section 4b: Sensitivity Analysis — HWISE Reduced Score

> **Collinearity note**: Individual HWISE items are highly correlated.
> Including multiple items in the full model simultaneously may produce
> inflated standard errors and unstable estimates. This sensitivity analysis
> substitutes the HWISE reduced composite score (`HWISE_REDUCED`) for any
> individual HWISE items that appeared in the main full model.

```{r sensitivity-pss}
# Replace individual HWISE items with HWISE_REDUCED composite
hwise_item_names <- hwise_items  # the 12 individual items

sig_vars_pss_sens <- sig_vars_pss_clean
has_hwise_items_pss <- any(sig_vars_pss_sens %in% hwise_item_names)

if (has_hwise_items_pss) {
  sig_vars_pss_sens <- c(
    setdiff(sig_vars_pss_sens, hwise_item_names),
    "HWISE_REDUCED"
  )
  cat("PSS sensitivity model: replaced individual HWISE items with HWISE_REDUCED\n")
} else {
  cat("PSS sensitivity: no individual HWISE items in main model — unchanged\n")
}

pss_sens_formula <- as.formula(
  paste("PSS_TOTAL ~ D_AGE + SES_SC_Total + D_HH_SIZE + Chronic_dis +",
        paste(sig_vars_pss_sens, collapse = " + "))
)
pss_sens <- lm(pss_sens_formula, data = df, na.action = na.omit)

cat("PSS sensitivity model Adj. R² =", round(summary(pss_sens)$adj.r.squared, 3),
    "| N =", nobs(pss_sens), "\n")

tbl_pss_sens <- tbl_regression(pss_sens, conf.int = TRUE) %>%
  bold_p(t = 0.05) %>%
  italicize_levels() %>%
  modify_caption("**PSS Sensitivity Model**: HWISE composite instead of individual items")
tbl_pss_sens
```

```{r sensitivity-emo}
sig_vars_emo_sens <- sig_vars_emo_clean
has_hwise_items_emo <- any(sig_vars_emo_sens %in% hwise_item_names)

if (has_hwise_items_emo) {
  sig_vars_emo_sens <- c(
    setdiff(sig_vars_emo_sens, hwise_item_names),
    "HWISE_REDUCED"
  )
  cat("EMO sensitivity model: replaced individual HWISE items with HWISE_REDUCED\n")
} else {
  cat("EMO sensitivity: no individual HWISE items in main model — unchanged\n")
}

emo_sens_formula <- as.formula(
  paste("EMO_NEG ~ D_AGE + SES_SC_Total + D_HH_SIZE + PSS_TOTAL + Chronic_dis +",
        paste(sig_vars_emo_sens, collapse = " + "))
)
emo_sens <- glm(emo_sens_formula, data = df, family = binomial, na.action = na.omit)

cat("EMO sensitivity model AIC =", round(AIC(emo_sens), 1),
    "| N =", nobs(emo_sens), "\n")

tbl_emo_sens <- tbl_regression(emo_sens, conf.int = TRUE, exponentiate = TRUE) %>%
  bold_p(t = 0.05) %>%
  italicize_levels() %>%
  modify_header(estimate ~ "**OR**") %>%
  modify_caption("**EMO Sensitivity Model**: HWISE composite instead of individual items")
tbl_emo_sens
```

---

# Section 4c: Export to Word

```{r export-word}
# ── Convert gtsummary tables to flextable ─────────────────────────────────────
ft_pss_full  <- as_flex_table(tbl_pss)
ft_emo_full  <- as_flex_table(tbl_emo)
ft_pss_sens  <- as_flex_table(tbl_pss_sens)
ft_emo_sens  <- as_flex_table(tbl_emo_sens)

# ── Significant predictors summary table ──────────────────────────────────────
ft_sig_summary <- sig_table %>%
  flextable() %>%
  set_header_labels(
    outcome             = "Outcome",
    term                = "Predictor term",
    n_models_significant = "# Models significant",
    min_p               = "Min p-value",
    aims                = "Found in"
  ) %>%
  colformat_double(j = "min_p", digits = 4) %>%
  bold(part = "header") %>%
  theme_booktabs() %>%
  set_table_properties(layout = "autofit", width = 1)

# ── Build Word document ───────────────────────────────────────────────────────
out_path <- "./output/tables/10.Tables_Full_Models.docx"

doc <- read_docx() %>%
  # Title
  body_add_par("10. Full Models: All Significant Predictors", style = "heading 1") %>%
  body_add_par(paste("Generated:", Sys.Date()), style = "Normal") %>%
  body_add_par("", style = "Normal") %>%

  # Section 2: Significant predictors
  body_add_par("Significant Predictors Summary", style = "heading 2") %>%
  body_add_par(
    paste0("The following predictors had p < 0.05 (unadjusted focal coefficient) ",
           "in at least one Aim 1 or Aim 2 adjusted model."),
    style = "Normal"
  ) %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_sig_summary) %>%
  body_add_par("", style = "Normal") %>%

  # Section 3: PSS full model
  body_add_par("PSS Full Model (Linear Regression)", style = "heading 2") %>%
  body_add_par(
    paste0("Outcome: PSS_TOTAL. Core covariates: Age, SES, Household size, ",
           "Chronic disease. Additional predictors: all terms significant in ",
           "Aim 1 or Aim 2 PSS models (p < 0.05)."),
    style = "Normal"
  ) %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_pss_full) %>%
  body_add_par("", style = "Normal") %>%

  # Section 4: EMO full model
  body_add_par("EMO Full Model (Logistic Regression)", style = "heading 2") %>%
  body_add_par(
    paste0("Outcome: EMO_NEG (1 = negative emotion). Core covariates: Age, SES, ",
           "Household size, Chronic disease, PSS_TOTAL. Additional predictors: ",
           "all terms significant in Aim 1 or Aim 2 EMO models (p < 0.05). ",
           "Estimates are odds ratios."),
    style = "Normal"
  ) %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_emo_full) %>%
  body_add_par("", style = "Normal") %>%

  # Sensitivity
  body_add_par("Sensitivity Analysis: HWISE Composite Score", style = "heading 2") %>%
  body_add_par(
    paste0("Individual HWISE items are highly correlated. These models replace ",
           "all individual HWISE items with the HWISE reduced composite score."),
    style = "Normal"
  ) %>%
  body_add_par("PSS Sensitivity Model", style = "heading 3") %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_pss_sens) %>%
  body_add_par("", style = "Normal") %>%
  body_add_par("EMO Sensitivity Model", style = "heading 3") %>%
  body_add_par("", style = "Normal") %>%
  body_add_flextable(ft_emo_sens) %>%
  body_add_par("", style = "Normal")

print(doc, target = out_path)
cat("Saved:", out_path, "\n")
```

---

# Section 5: Session Info

```{r session-info}
sessionInfo()
```
