---
title: "Sample Size & Power Analysis"
author: "Paloma"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Setup

```{r packages}
# Install packages if needed
if (!requireNamespace("pwr",       quietly = TRUE)) install.packages("pwr")
if (!requireNamespace("WebPower",  quietly = TRUE)) install.packages("WebPower")

library(pwr)
library(WebPower)
```

```{r load-data}
df <- read.csv("03.Merged_Q1-6_Q16_Q19_Q26_Q28.csv", stringsAsFactors = FALSE)
```

---

## 1. Observed sample characteristics

```{r sample-size}
# Total N
n_total <- nrow(df)
cat("Total N:", n_total, "\n")

# N with non-missing PSS_TOTAL
n_pss <- sum(!is.na(df$PSS_TOTAL) & df$PSS_TOTAL != "" & df$PSS_TOTAL != "NA")
cat("N with PSS_TOTAL:", n_pss, "\n")

# emo_type (MX26_EM_HHW_CAT): binary 0/1, exclude 444 (ambiguous)
emo_clean <- df$MX26_EM_HHW_CAT[df$MX26_EM_HHW_CAT %in% c("0", "1")]
n_emo    <- length(emo_clean)
p_neg    <- mean(emo_clean == "1")   # proportion reporting negative emotion
cat("N with emo_type (0/1 only):", n_emo, "\n")
cat("Proportion negative emotion (p_neg):", round(p_neg, 3), "\n")
```

---

## 2. Linear regression power (PSS_TOTAL outcome)

Using `pwr::pwr.f2.test()`. The model includes 1 predictor of interest + 3 core
covariates (SES, age, household size), so **u = 4** (total regression df).

```{r power-linear}
# --- Minimum detectable effect size given N ---
# v = residual df = N - u - 1
n_lin <- n_pss          # use N available for PSS
u     <- 4              # predictors (1 focal + 3 covariates)
v     <- n_lin - u - 1

pwr_lin <- pwr.f2.test(u = u, v = v, sig.level = 0.05, power = 0.80)

cat("---------------------------------------------\n")
cat("Linear regression (PSS_TOTAL)\n")
cat("N =", n_lin, "| alpha = 0.05 | power = 0.80\n")
cat("Minimum detectable f² =", round(pwr_lin$f2, 4), "\n")
cat("  (small = 0.02, medium = 0.15, large = 0.35)\n")
cat("---------------------------------------------\n")

# --- What power do we have for a small effect (f2 = 0.02)? ---
pwr_lin_small <- pwr.f2.test(u = u, v = v, f2 = 0.02, sig.level = 0.05)
cat("Power to detect f² = 0.02:", round(pwr_lin_small$power, 3), "\n")

# --- What N would we need for 80% power at f2 = 0.02? ---
pwr_lin_req <- pwr.f2.test(u = u, f2 = 0.02, sig.level = 0.05, power = 0.80)
cat("N required for 80% power at f² = 0.02:",
    ceiling(pwr_lin_req$v + u + 1), "\n")
```

---

## 3. Logistic regression power (emo_type outcome)

### 3a. Binary predictor (e.g., W_WS_LOC: water-secure vs. water-insecure neighborhood)

Using `WebPower::wp.logistic()`. We solve for the minimum detectable log-odds
ratio (converted to OR) assuming equal group sizes.

```{r power-logistic-binary}
n_log <- n_emo           # N available for emo_type models

# Baseline event rate (negative emotion) in reference group
# We use the overall p_neg as an approximation of the reference group rate
p0 <- p_neg

# Sequence of ORs to evaluate
or_seq <- seq(1.3, 3.0, by = 0.1)

power_by_or <- sapply(or_seq, function(or) {
  # Convert OR to p1
  p1 <- (or * p0) / (1 - p0 + or * p0)
  tryCatch(
    wp.logistic(n = n_log, p0 = p0, p1 = p1,
                alpha = 0.05, power = NULL,
                alternative = "two.sided")$power,
    error = function(e) NA
  )
})

results_bin <- data.frame(OR = or_seq, Power = round(power_by_or, 3))
print(results_bin)

# Minimum OR detectable at 80% power
min_or_80 <- or_seq[which(power_by_or >= 0.80)[1]]
cat("\nMinimum detectable OR at 80% power (binary predictor):",
    min_or_80, "\n")
```

### 3b. Continuous predictor (e.g., modified HWISE total score)

For continuous predictors in logistic regression, power depends on the predictor's
distribution. We use `WebPower::wp.logistic()` with the predictor standardized
to 1 SD, so the OR reflects a 1-SD increase in the predictor.

```{r power-logistic-continuous}
# Estimate SD of HWISE total from data (proxy for continuous predictor scale)
hw_vals <- as.numeric(df$HW_TOTAL[!is.na(df$HW_TOTAL) &
                                    df$HW_TOTAL != "" &
                                    df$HW_TOTAL != "NA"])
sd_hwise <- sd(hw_vals, na.rm = TRUE)
cat("SD of HW_TOTAL:", round(sd_hwise, 2), "\n\n")

# For a 1-SD increase in a continuous predictor, we treat it similarly to
# a binary predictor using the OR-based approximation
or_seq_cont <- seq(1.3, 3.0, by = 0.1)

power_cont <- sapply(or_seq_cont, function(or) {
  p1 <- (or * p0) / (1 - p0 + or * p0)
  tryCatch(
    wp.logistic(n = n_log, p0 = p0, p1 = p1,
                alpha = 0.05, power = NULL,
                alternative = "two.sided")$power,
    error = function(e) NA
  )
})

min_or_cont <- or_seq_cont[which(power_cont >= 0.80)[1]]
cat("Minimum detectable OR per 1-SD increase (continuous predictor) at 80% power:",
    min_or_cont, "\n")
```

---

## 4. Power note for Aim 2 (multiple models + FDR correction)

With FDR correction, the effective alpha per test is lower than 0.05. Here we
examine how power changes under a more conservative per-test alpha, to give an
honest characterisation of Aim 2 power.

```{r power-aim2}
# Approximate number of Aim 2 models per outcome
# (10 HWISE items + ~7 other predictors = ~17 models per outcome)
n_tests   <- 17
fdr_alpha <- 0.05   # BH controls FDR at 0.05, not per-comparison alpha;
                    # for illustration we check power at alpha = 0.01 as
                    # a conservative approximation

# Linear regression at alpha = 0.01
pwr_aim2_lin <- pwr.f2.test(u = u, v = v, f2 = 0.02, sig.level = 0.01)
cat("Aim 2 linear: power at f² = 0.02, alpha = 0.01:",
    round(pwr_aim2_lin$power, 3), "\n")

# Logistic at alpha = 0.01, OR = min_or_80 from above
p1_aim2 <- (min_or_80 * p0) / (1 - p0 + min_or_80 * p0)
pwr_aim2_log <- wp.logistic(n = n_log, p0 = p0, p1 = p1_aim2,
                             alpha = 0.01, power = NULL,
                             alternative = "two.sided")
cat("Aim 2 logistic: power at OR =", min_or_80, ", alpha = 0.01:",
    round(pwr_aim2_log$power, 3), "\n")
```

---

## 5. Summary for methods section

```{r summary}
cat("============================================================\n")
cat("NUMBERS TO INSERT INTO METHODS SECTION\n")
cat("============================================================\n\n")

cat("Total analytic N (PSS models):   ", n_lin, "\n")
cat("Total analytic N (emo_type models):", n_log, "\n")
cat("Event rate (negative emotion):   ", round(p_neg * 100, 1), "%\n\n")

cat("-- Linear regression (PSS_TOTAL) --\n")
cat("Minimum detectable f²:           ", round(pwr_lin$f2, 4), "\n")
cat("Power at f² = 0.02:              ", round(pwr_lin_small$power * 100, 1), "%\n\n")

cat("-- Logistic regression (emo_type) --\n")
cat("Min detectable OR (binary pred): ", min_or_80, "\n")
cat("Min detectable OR (continuous):  ", min_or_cont, "\n\n")

cat("============================================================\n")
```
