---
title: "HBA2025_results"
author: "Paloma"
date: "2025-01-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Here you will find the code used to obtain results shown in the annual meeting of the HBA, 2025. 

Abstract:

**Coping with water insecurity: women's strategies and emotional responses in Iztapalapa,
Mexico City**

Water insecurity in urban areas presents distinctive challenges, particularly in marginalized communities. While past studies have documented how households adapt to poor water services, many of these coping strategies come at a significant personal cost. Here we examine the **coping strategies and emotional impacts of unreliable water services** among 400 women in Iztapalapa, Mexico City. Data were collected through surveys over the Fall of 2022 and Spring of 2023. We assessed **household water access, water management practices, and emotional responses to local water services.** 

Results indicate that during acute water shortages, women can spend extended periods (several hours, or sometimes days) waiting for water trucks. Additionally, 57% of respondents reported feeling frustrated or angry about their water situation, while around 20% experienced family conflicts over water use or community-level conflicts around water management, often involving water vendors or government services. 

This study offers one of the first in-depth examinations of how water insecurity specifically affects women in Iztapalapa, a densely populated region of Mexico City with severe water access challenges. The findings highlight the urgent need for policy interventions that address water insecurity with a gender-sensitive approach, recognizing the disproportionate burden placed on women as primary water managers in their households.

```{r libraries, echo = FALSE, warning = FALSE, message= FALSE}

# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2) 
library(knitr)
# Load the dataset

data_path = "data"

data <- read.csv(file.path(data_path, "Cleaned_Dataset_Screening_HWISE_PSS_V3.csv"), 
         stringsAsFactors = FALSE,  
         na.strings = c("", "N/A", "NA", "pending"))


```

```{r HWISE categ}

# Ensure HW_TOTAL is numeric
data$HW_TOTAL <- as.numeric(data$HW_TOTAL)

# Categorize HW_TOTAL into four groups
data <- data %>%
  mutate(HW_TOTAL_category = case_when(
    HW_TOTAL >= 0 & HW_TOTAL <= 2  ~ "No-to-Marginal",
    HW_TOTAL >= 3 & HW_TOTAL <= 11 ~ "Low",
    HW_TOTAL >= 12 & HW_TOTAL <= 23 ~ "Moderate",
    HW_TOTAL >= 24 & HW_TOTAL <= 36 ~ "High",
    TRUE ~ NA_character_  # Assign NA if value is missing or out of range
  ))

# Convert to factor to maintain categorical order
data$HW_TOTAL_category <- factor(data$HW_TOTAL_category, 
                                 levels = c("No-to-Marginal", "Low", "Moderate", "High"))

# HWISE scores 
summary(data$HW_TOTAL)
# Check the new variable distribution
temp <- table(data$HW_TOTAL_category)
table(data$HW_TOTAL_category)

```

```{r echo=FALSE}

# Create a data frame with category counts
hw_total_counts <- data.frame(
  HW_TOTAL_category = factor(c("No-to-Marginal", "Low", "Moderate", "High"), 
                             levels = c("No-to-Marginal", "Low", "Moderate", "High")),
  Count = c(temp[1], temp[2], temp[3], temp[4])  # Given category frequencies
)

# Calculate the total count and percentage for each category
hw_total_counts <- hw_total_counts %>%
  mutate(Percentage = (Count / sum(Count)) * 100)

# Create the bar plot with percentages
ggplot(hw_total_counts, aes(x = HW_TOTAL_category, y = Percentage, fill = HW_TOTAL_category)) +
  geom_bar(stat = "identity") +  # Use stat="identity" for predefined percentages
  theme_minimal() +
  labs(title = "Distribution of HWISE Categories",
       x = "Water insecurity level (HWISE)", 
       y = "Percentage") +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#fdbf6f", "#ff7f00")) +  # Custom colors
  theme(legend.position = "none") +  # Remove legend
  geom_text(aes(label = paste0(round(Percentage, 1), "%")), vjust = -0.5) # Add percentage labels

```


```{r echo=FALSE}

# Convert SEASON to a factor with labels
data$SEASON <- factor(data$SEASON, levels = c(0, 1), labels = c("Dry", "Rainy"))

# Create HW_TOTAL categories
data <- data %>%
  mutate(HW_TOTAL_category = case_when(
    HW_TOTAL >= 0 & HW_TOTAL <= 2  ~ "No-to-Marginal",
    HW_TOTAL >= 3 & HW_TOTAL <= 11 ~ "Low",
    HW_TOTAL >= 12 & HW_TOTAL <= 23 ~ "Moderate",
    HW_TOTAL >= 24 & HW_TOTAL <= 36 ~ "High",
    TRUE ~ NA_character_  # Assign NA for missing or out-of-range values
  ))

# Convert to factor with proper ordering
data$HW_TOTAL_category <- factor(data$HW_TOTAL_category, 
                                 levels = c("No-to-Marginal", "Low", "Moderate", "High"))

# Calculate percentages within each season
hw_total_counts <- data %>%
  filter(!is.na(HW_TOTAL_category)) %>% 
  group_by(SEASON, HW_TOTAL_category) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(SEASON) %>%
  mutate(Percentage = (Count / sum(Count)) * 100)

# Create the bar plot with percentages, faceted by SEASON
ggplot(hw_total_counts, aes(x = HW_TOTAL_category, y = Percentage, fill = HW_TOTAL_category)) +
  geom_bar(stat = "identity") +  # Use predefined percentages
  theme_minimal() +
  labs(title = "Distribution of HWISE Categories by Season",
       x = "HWISE Category", 
       y = "Percentage") +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#fdbf6f", "#ff7f00")) +
  theme(legend.position = "none") +  # Remove legend
  geom_text(aes(label = paste0(round(Percentage, 1), "%")), vjust = -0.5) +  # Add percentage labels
  facet_wrap(~SEASON)  # Separate plots by season

# HWISE scores 
dry <- data[c(data$SEASON == "Dry"), ]
rainy <- data[c(data$SEASON == "Rainy"), ]
summary(dry$HW_TOTAL)
summary(rainy$HW_TOTAL)
# Check the new variable distribution
table(dry$HW_TOTAL_category)
table(rainy$HW_TOTAL_category)
```

# Rainy season to dry season

```{r}

data <- read.csv(file.path(data_path, "Cleaned_Dataset_Screening_HWISE_PSS_V3.csv"), 
         stringsAsFactors = FALSE,  
         na.strings = c("", "N/A", "NA", "pending"))

# Define Rainy (Fall = 1) and Dry (Spring = 0) seasons
data <- data %>%
  mutate(Season_Type = case_when(
    SEASON == 1 ~ "Rainy",
    SEASON == 0 ~ "Dry"
  ))

# Convert Season_Type to a factor
data$Season_Type <- factor(data$Season_Type, levels = c("Rainy", "Dry"))

# Categorize HW_TOTAL into four groups
data <- data %>%
  filter(!is.na(HW_TOTAL)) %>%  # Remove missing values
  mutate(HW_TOTAL_category = case_when(
    HW_TOTAL >= 0 & HW_TOTAL <= 2  ~ "No-to-Marginal",
    HW_TOTAL >= 3 & HW_TOTAL <= 11 ~ "Low",
    HW_TOTAL >= 12 & HW_TOTAL <= 23 ~ "Moderate",
    HW_TOTAL >= 24 & HW_TOTAL <= 36 ~ "High"
  ))

# Convert to factor with correct ordering
data$HW_TOTAL_category <- factor(data$HW_TOTAL_category, 
                                 levels = c("No-to-Marginal", "Low", "Moderate", "High"))

# Calculate the percentage for each category within each season
hw_season_counts <- data %>%
  filter(!is.na(HW_TOTAL_category)) %>%  # Ensure no NA categories
  group_by(Season_Type, HW_TOTAL_category) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(Season_Type) %>%
  mutate(Percentage = (Count / sum(Count)) * 100)

# Reshape data to calculate the percentage difference (Dry - Rainy)
hw_diff <- hw_season_counts %>%
  select(Season_Type, HW_TOTAL_category, Percentage) %>%
  spread(Season_Type, Percentage) %>%  # Convert to wide format
  mutate(Difference = Dry - Rainy)  # Compute difference from Rainy to Dry

# Create the bar plot showing the percentage difference
ggplot(hw_diff, aes(x = HW_TOTAL_category, y = Difference, fill = Difference > 0)) +
  geom_bar(stat = "identity") +  # Use precomputed differences
  theme_minimal() +
  labs(title = "Seasonal change: Rainy compared to Dry Season",
       x = "Water insecurity level (HWISE)", 
       y = "Percentage Difference (Rainy to Dry season)") +
  scale_fill_manual(values = c("#1f78b4", "#ff7f00"), labels = c("Decrease", "Increase")) +  # Assign colors
  theme(legend.position = "right") +  # Keep legend
  geom_text(aes(label = paste0(round(Difference, 1), "%")), vjust = -0.5)  # Add percentage labels

```

# BOXPLOTS

```{r}

data <- read.csv(file.path(data_path, "Cleaned_Dataset_Screening_HWISE_PSS_V3.csv"), 
         stringsAsFactors = FALSE,  
         na.strings = c("", "N/A", "NA", "pending"))

# Categorize HW_TOTAL into four groups
data <- data %>%
  filter(!is.na(HW_TOTAL)) %>%  # Remove missing values
  mutate(HW_TOTAL_category = case_when(
    HW_TOTAL >= 0 & HW_TOTAL <= 2  ~ "No-to-Marginal",
    HW_TOTAL >= 3 & HW_TOTAL <= 11 ~ "Low",
    HW_TOTAL >= 12 & HW_TOTAL <= 23 ~ "Moderate",
    HW_TOTAL >= 24 & HW_TOTAL <= 36 ~ "High"
  ))

# Convert to factor with proper order
data$HW_TOTAL_category <- factor(data$HW_TOTAL_category, 
                                 levels = c("No-to-Marginal", "Low", "Moderate", "High"))

# Count the number of data points per Water insecurity level (HWISE)
 summary_stats <- data %>%
  group_by(HW_TOTAL_category) %>%
  summarise(Count = n(), SD = sd(HRS_WEEK, na.rm = TRUE), .groups = 'drop')

 hrs.w <- data$HRS_WEEK
 means <- aggregate(hrs.w ~ HW_TOTAL_category, data, mean)
means$hrs.w <- round(means$hrs.w, 2)
 
# Define color palette
color_palette <- c("#1f78b4", "#a6cee3", "#fdbf6f", "#ff7f00")

# Create box-and-whisker plot with individual data points
ggplot(data, aes(x = HW_TOTAL_category, 
                 y = HRS_WEEK, 
                 fill = HW_TOTAL_category)) +
    geom_jitter(aes(color = HW_TOTAL_category), 
                size = 1, width = 0.25) +  
  # Jitter adds individual data points
    geom_violin(alpha = 0.6, width = 1.4) + # violin
  geom_boxplot(outlier.shape = 1, alpha = 0.5, 
               width = 0.08, color = "grey30") + # Boxplot
    geom_text(data = summary_stats, 
              aes(x = HW_TOTAL_category, 
                  y = max(data$HRS_WEEK, 
                          na.rm = TRUE) + 10, 
                  label = paste0("SD = ", round(SD, 1))), 
              size = 4, fontface = "bold") +    # Add SD text annotations above the boxes
   geom_text(data = means, 
             aes(label = hrs.w, y = hrs.w + 1, hjust=-1), 
             size = 3, color = "darkred") + #adds average labels
  theme_minimal() +
  labs(title = "Hours water of supply by Water Insecurity group (HWISE)",
       x = "HWISE Group", 
       y = "Hours of Water Supply per Week") +
  scale_fill_manual(values = color_palette) +  # Custom colors for boxes
  scale_color_manual(values = color_palette) +  # Custom colors for points +
   stat_summary(fun.y=mean, geom="point", shape=23, 
                size=5, color="darkred", fill="darkred") +
  theme(legend.position = "none", 
        axis.text = element_text(size = 10)) + # Remove legend for clarity 
  scale_x_discrete(labels = paste0(summary_stats$HW_TOTAL_category, 
                                 " (n=", summary_stats$Count, ")"))  # Add count to x-axis labels



```


```{r}

# Categorize HW_TOTAL into groups
data <- data %>%
  filter(!is.na(MX28_WQ_COMP)) %>%  # Remove missing values
  mutate(MX28_WQ_COMP_category = case_when(
    MX28_WQ_COMP == 0 ~ "Worse",
    MX28_WQ_COMP == 1 ~ "Same",
    MX28_WQ_COMP == 2 ~ "Better",
  ))

# Convert to factor with proper order
data$MX28_WQ_COMP_category <- factor(data$MX28_WQ_COMP_category, 
                                 levels = c("Worse", "Same", "Better"))

# Count the number of data points per Water insecurity level (HWISE)
 summary_stats <- data %>%
  group_by(MX28_WQ_COMP_category) %>%
  summarise(Count = n(), SD = sd(HRS_WEEK, na.rm = TRUE), .groups = 'drop')

 hrs.w <- data$HRS_WEEK
 means <- aggregate(hrs.w ~ MX28_WQ_COMP_category, data, mean)
means$hrs.w <- round(means$hrs.w, 2)

color_palette <- c("#ff7f00", "#a6cee3",  "#1f78b4")

# Create box-and-whisker plot with individual data points
ggplot(data, aes(x = MX28_WQ_COMP_category, 
                 y = HRS_WEEK, 
                 fill = MX28_WQ_COMP_category)) +
    geom_jitter(aes(color = MX28_WQ_COMP_category), 
                size = 1, width = 0.25) +  
  # Jitter adds individual data points
    geom_violin(alpha = 0.6, width = 1) + # violin
  geom_boxplot(outlier.shape = 1, alpha = 0.5, 
               width = 0.08, color = "grey30") + # Boxplot
      geom_text(data = means, 
             aes(label = hrs.w, y = hrs.w + 1, hjust=-0.8), 
             size = 3, color = "darkred") + #adds average labels
  theme_minimal() +
  labs(title = "Hours of water supply and comparison to other parts of the city",
       x = "Own water service is considered", 
       y = "Hours of Water Supply per Week") +
  scale_fill_manual(values = color_palette) +  # Custom colors for boxes
  scale_color_manual(values = color_palette) +  # Custom colors for points +
   stat_summary(fun.y=mean, geom="point", shape=23, 
                size=5, color="darkred", fill="darkred") +
  theme(legend.position = "none", 
        axis.text = element_text(size = 10)) +  
  scale_x_discrete(labels = paste0(summary_stats$MX28_WQ_COMP_category, 
                                 " (n=", summary_stats$Count, ")")) 
```

```{r}

# Count the number of data points per Water insecurity level (HWISE)
 summary_stats <- data %>%
  group_by(MX28_WQ_COMP_category) %>%
  summarise(Count = n(), SD = sd(HW_TOTAL, na.rm = TRUE), .groups = 'drop')

 hrs.w <- data$HW_TOTAL
 means <- aggregate(hrs.w ~ MX28_WQ_COMP_category, data, mean)
means$hrs.w <- round(means$hrs.w, 2)


# Create box-and-whisker plot with individual data points
ggplot(data, aes(x = MX28_WQ_COMP_category, 
                 y = HW_TOTAL, 
                 fill = MX28_WQ_COMP_category)) +
    geom_jitter(aes(color = MX28_WQ_COMP_category), 
               size = 1, width = 0.25) +  
  # Jitter adds individual data points
    geom_violin(alpha = 0.4, width = 1) + # violin
  geom_boxplot(outlier.shape = 1, alpha = 0.3, 
               width = 0.15, color = "grey30") + # Boxplot
    geom_text(data = means, 
             aes(label = hrs.w, y = hrs.w + 0.8, hjust=-0.85), 
             size = 3, color = "darkred") + #adds average labels
      geom_text(data = summary_stats, 
              aes(x = MX28_WQ_COMP_category, 
                  y = max(data$HW_TOTAL, 
                          na.rm = TRUE) + 2, 
                  label = paste0("SD = ", round(SD, 1))), 
              size = 4, fontface = "bold") +    # Add SD text annotations above the boxes
  theme_minimal() +
  labs(title =  "Hours of water supply and comparison to other parts of the city",
       x = "Own water service is considered", 
       y = "HWISE score") +
  scale_fill_manual(values = color_palette) +  # Custom colors for boxes
  scale_color_manual(values = color_palette) +  # Custom colors for points +
   stat_summary(fun.y=mean, geom="point", shape=23, 
               size=5, color="darkred", fill="darkred") +
  theme(legend.position = "none", 
        axis.text = element_text(size = 10)) +  
  scale_x_discrete(labels = paste0(summary_stats$MX28_WQ_COMP_category, 
                                 " (n=", summary_stats$Count, ")")) 
```





```{r}

# Load the dataset 
data <- read.csv(file.path(data_path, "Cleaned_Dataset_Screening_HWISE_PSS_V3.csv"), 
         stringsAsFactors = FALSE,  
         na.strings = c("", "N/A", "NA", "pending"))

# Keep only relevant columns (HW_ questions and HRS_WEEK)
data_long <- data %>%
  select(HW_TOTAL, HRS_WEEK) %>%
  pivot_longer(cols = HW_TOTAL, names_to = "HW_Question", values_to = "Response") %>%
  filter(!is.na(HRS_WEEK), !is.na(Response))  # Remove NAs

# Compute mean HRS_WEEK for each HW_Question & Response category
summary_stats <- data_long %>%
  group_by(HW_Question, Response) %>%
  summarise(Mean_HRS_WEEK = mean(HRS_WEEK, na.rm = TRUE), Count = n(), SD = sd(HRS_WEEK, na.rm = TRUE), .groups = 'drop')

# Generate boxplots with mean values
ggplot(data_long, aes(x = as.factor(Response), y = HRS_WEEK, fill = as.factor(Response))) +
  geom_jitter(aes(color = as.factor(Response)), size = 1.5, width = 0.2, alpha = 0.2) +  
   geom_boxplot(outlier.shape = NA, width = 0.5, alpha = 0.4, color = "gray70") +  # Thinner boxes
  geom_point(data = summary_stats, aes(x = as.factor(Response), y = Mean_HRS_WEEK), 
             color = "black", size = 1.5) +  # Add red dot for mean value
  geom_text(data = summary_stats, aes(x = as.factor(Response), y =  Mean_HRS_WEEK, 
             label = round(Mean_HRS_WEEK, 0)), color = "black", hjust =  -0.35, size = 2.8) +  # Add numeric mean label
  theme_minimal() +
  labs(title = "Distribution of Hours water of supply by total HWISE score",
       subtitle = "Black dots show mean num. hours per week",
       x = "HWISE score", 
       y = "Hours of Water Supply per Week") +
 # scale_fill_manual(values = color_palette) +  # Custom colors for boxes
 # scale_color_manual(values = color_palette) +  # Custom colors for points
  theme(legend.position = "none", axis.text.x = element_text(size = 5, angle = 45)) + 
   scale_x_discrete(labels = paste0(summary_stats$Response, 
                                 " (n=", summary_stats$Count, ")"))  





# Define color palette
#color_palette <- c("#1f78b4", "#a6cee3", "#fdbf6f", "#ff7f00")

# Generate boxplots
ggplot(data_long, aes(x = as.factor(Response), y = HRS_WEEK, fill = as.factor(Response))) +
  geom_boxplot(outlier.shape = NA, width = 0.5, alpha = 0.7) +  # Thinner boxes
  geom_jitter(aes(color = as.factor(Response)), size = 1.5, width = 0.2, alpha = 0.5) +  
  theme_minimal() +
  labs(title = "Distribution of Hours water of supply by total HWISE score",
       x = "Score", 
       y = "Hours of Water Supply per Week") +
#  scale_fill_manual(values = color_palette) +  # Custom colors for boxes
#  scale_color_manual(values = color_palette) +  # Custom colors for points
  theme(legend.position = "none")# +  # Remove legend for clarity
 

```


```{r}

# Load the dataset 
data <- read.csv(file.path(data_path, "Cleaned_Dataset_Screening_HWISE_PSS_V3.csv"), 
         stringsAsFactors = FALSE,  
         na.strings = c("", "N/A", "NA", "pending"))
data <- data %>% 
  select(-HW_TOTAL)

# Select all columns that start with "HW_"
hw_vars <- grep("^HW_", names(data), value = TRUE)

# Keep only relevant columns (HW_ questions and HRS_WEEK)
data_long <- data %>%
  select(all_of(hw_vars), HRS_WEEK) %>%
  pivot_longer(cols = all_of(hw_vars), names_to = "HW_Question", values_to = "Response") %>%
  filter(!is.na(HRS_WEEK), !is.na(Response))  # Remove NAs

# Compute mean HRS_WEEK for each HW_Question & Response category
mean_values <- data_long %>%
  group_by(HW_Question, Response) %>%
  summarise(Mean_HRS_WEEK = mean(HRS_WEEK, na.rm = TRUE), .groups = 'drop')


# Generate boxplots with mean values
ggplot(data_long, aes(x = as.factor(Response), y = HRS_WEEK, fill = as.factor(Response))) +
  geom_boxplot(outlier.shape = NA,  alpha = 0.4, varwidth = TRUE) +  # Thinner boxes
  geom_jitter(aes(color = as.factor(Response)), size = 1.5, width = 0.2, alpha = 0.4) +  
  geom_point(data = mean_values, aes(x = as.factor(Response), y = Mean_HRS_WEEK), 
             color = "red", size = 2) +  # Add red dot for mean value
 # geom_text(data = mean_values, aes(x = as.factor(Response), y = Mean_HRS_WEEK, 
            # label = round(Mean_HRS_WEEK, 1)), color = "red", vjust = -0.5, size = 3.5) +  # Add numeric mean label
  theme_minimal() +
  labs(title = "Distribution of Hours water of supply by HWISE Question",
       subtitle = "Red dots show mean num. hours per week",
       x = "Response", 
       y = "Hours of Water Supply per Week") +
  theme(legend.position = "none") +  # Remove legend for clarity
  facet_wrap(~HW_Question, scales = "free_y")  # Create separate plots for each HW_ question

```


