---
title: "HBA plots"
output: html_document
date: "2025-03-08"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, echo = FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(rstatix)
library(ggpubr)
library(coin)
library(tidyr)
library(MASS)  # For stepAIC
library(broom) # For model output formatting
library(dplyr) # For data manipulation
library(car)   # For VIF check


# Load the dataset
data_path = "data"

data <- read.csv(file.path(data_path, "Cleaned_Dataset_Screening_HWISE_PSS_V4.csv"), 
         stringsAsFactors = FALSE,  
         na.strings = c("", "N/A", "NA", "pending"))

```

#Categorical groups using HWISE
```{r}
# Categorize HW_TOTAL into four groups
data <- data %>%
  filter(!is.na(HRS_WEEK), !is.na(HW_TOTAL)) %>%
  mutate(HW_TOTAL_category = case_when(
    HW_TOTAL >= 0 & HW_TOTAL <= 2  ~ "No-to-Marginal",
    HW_TOTAL >= 3 & HW_TOTAL <= 11 ~ "Low",
    HW_TOTAL >= 12 & HW_TOTAL <= 23 ~ "Moderate",
    HW_TOTAL >= 24 & HW_TOTAL <= 36 ~ "High"
  ))

```
## Significant differences in hours of water supply per week
```{r echo = FALSE}
# Convert to factor
data$HW_TOTAL_category <- factor(data$HW_TOTAL_category, levels = c("No-to-Marginal", "Low", "Moderate", "High"))
cat("Number of samples falling in each category:")
print(table(data$HW_TOTAL_category))

# Check normality using Shapiro-Wilk test for each group
shapiro_results <- data %>%
  group_by(HW_TOTAL_category) %>%
  summarise(p_value = shapiro.test(HRS_WEEK)$p.value)
cat("Check normality using Shapiro-Wilk test for each group:")
print(shapiro_results)

```

```{r echo = FALSE}

# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(HRS_WEEK ~ HW_TOTAL_category, data = data)
cat("Results Kruskal-Wallis test:")
print(kruskal_result)

# Perform Wilcoxon pairwise comparisons
wilcox_results <- pairwise_wilcox_test(data, HRS_WEEK ~ HW_TOTAL_category, 
                                       p.adjust.method = "bonferroni")

# Rename columns for ggpubr compatibility
wilcox_results <- wilcox_results %>%
  rename(group1 = group1, group2 = group2, p.adj = p.adj) %>%
  mutate(p.adj.signif = case_when(
    p.adj < 0.001 ~ "***",
    p.adj < 0.01  ~ "**",
    p.adj < 0.05  ~ "*",
    TRUE          ~ "ns"
  ))

# Print results

cat("Results Wilcoxon pairwise comparisons")
print(wilcox_results)
```
Interpretation: 

    p.adj < 0.001 ~ "***",
    p.adj < 0.01  ~ "**",
    p.adj < 0.05  ~ "*",
    TRUE          ~ "ns"
  
```{r echo =FALSE}
  
# Ensure positions for p-value labels
wilcox_results <- wilcox_results %>%
  add_xy_position(x = "HW_TOTAL_category")

# Define y-axis positions for p-value labels (adjust as needed)
y_positions <- c(170, 180, 190, 200, 210, 220)  # Adjust these manually

# Add y position column manually
wilcox_results <- wilcox_results %>%
  mutate(y.position = y_positions[1:n()])  # Assign y positions

# Modify Wilcoxon results to include sample size (n1) in the annotation label
significant_wilcox_results <- wilcox_results %>%
  filter(p.adj < 0.05) %>%  # Remove non-significant results
  mutate(label = paste0("n=", n1, "\n", p.adj.signif))  # Combine sample size and significance

# Print the significant comparisons
cat("Printing significant results only:")
print(significant_wilcox_results)


# Count the number of data points per Water insecurity level (HWISE)
 summary_stats <- data %>%
  group_by(HW_TOTAL_category) %>%
  summarise(Count = n(), SD = sd(HRS_WEEK, na.rm = TRUE), .groups = 'drop')

 hrs.w <- data$HRS_WEEK
 means <- aggregate(hrs.w ~ HW_TOTAL_category, data, mean)
means$hrs.w <- round(means$hrs.w, 2)

# Create a boxplot with only significant Wilcoxon p-value annotations
ggplot(data, aes(x = HW_TOTAL_category, y = HRS_WEEK, fill = HW_TOTAL_category)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.5) +  # Boxplot
  geom_jitter(aes(color = HW_TOTAL_category), size = 1.5, width = 0.2, alpha = 0.5) +  # Individual points
  stat_pvalue_manual(significant_wilcox_results, label = "p.adj.signif", 
                     xmin = "xmin", xmax = "xmax", y.position = "y.position", 
                     tip.length = 0.02, inherit.aes = FALSE) +  # Annotate only significant p-values
  labs(title = "Comparing hours of water supply by HWISE category (Wilcoxon Post-Hoc)",
       x = "HWISE Category",
       y = "Hours of Water Supply per Week",
       subtitle = "Level of significance (*** for p < 0.001 ** for p < 0.01).") +
  theme_minimal() +
  theme(legend.position = "none") + 
    scale_x_discrete(labels = paste0(summary_stats$HW_TOTAL_category, 
                                 " (n=", summary_stats$Count, ")")) +
  stat_summary(fun.y=mean, geom="point", shape=23, 
                size=5, color="darkred", fill="darkred") 

```

```{r echo = TRUE}

# Ensure it's properly coded as 0 or 1
data$MX26_EM_HHW_TYPE <- as.numeric(as.factor(data$MX26_EM_HHW_TYPE)) - 1  
# Remove rows where MX26_EM_HHW_TYPE is 2
data <- data %>% filter(MX26_EM_HHW_TYPE != 2)

# Check unique values to confirm only 0 and 1 remain
table(data$MX26_EM_HHW_TYPE)
dim(data)

# Define dependent (outcome) variable
outcome_var <- "MX26_EM_HHW_TYPE"  # Binary outcome (0 = negative, 1 = positive)

# Define independent (predictor) variables
predictors <- c("HRS_WEEK", "W_WC_WI", "MX28_WQ_COMP", "W_WS_LOC",
                "SES_SC_Total", "HW_TOTAL", "D_HH_SIZE", "D_LOC_TIME",
                "PSS_TOTAL", "HLTH_CPAIN_CAT", "HLTH_CDIS_CAT", 
                "MX9_DRINK_W", "MX10_WSTORAGE", "SEASON")

# Run univariate logistic regression for each predictor
univariate_results <- list()
aic_values <- data.frame(Predictor = character(), AIC = numeric(), stringsAsFactors = FALSE)

data <- data %>% drop_na(predictors)
dim(data)

for (var in predictors) {
    # Build formula dynamically
    formula <- as.formula(paste(outcome_var, "~", var))
    
    # Fit univariate logistic regression model
    model <- glm(formula, data = data, family = binomial)
    
    # Store model summary
    univariate_results[[var]] <- summary(model)
    
    # Store AIC values
    aic_values <- rbind(aic_values, data.frame(Predictor = var, AIC = AIC(model)))
}

univariate_results

# Display AIC values sorted in ascending order (lower AIC = better fit)
aic_values <- aic_values %>% arrange(AIC)
print(aic_values)

# Select best predictors based on AIC (top 6 with lowest AIC)
top_predictors <- head(aic_values$Predictor, 6)

# Create multivariable model with top predictors
multivariable_formula <- as.formula(paste(outcome_var, "~", paste(top_predictors, collapse = " + ")))
multivariable_model <- glm(multivariable_formula, data = data, family = binomial)

# Display summary of multivariable model
summary(multivariable_model)

# Perform stepwise selection to find the best model
optimized_model <- stepAIC(multivariable_model, direction = "both")

# Display summary of optimized model
summary(optimized_model)

```

The Akaike Information Criterion (AIC) values indicate how well each predictor explains variations in MX26_EM_HHW_TYPE (perceived water service quality). Lower AIC values indicate better model fit.

1️⃣ HW_TOTAL (HWISE score) had the lowest AIC (261.93), suggesting that water insecurity is the strongest predictor of negative perceptions of water service.
2️⃣ W_WC_WI (intermittent vs. continuous water supply) was the second-best predictor (AIC = 299.52), reinforcing the idea that inconsistent water access influences perceptions.
3️⃣ Psychological stress (PSS_TOTAL) also emerged as a key factor (AIC = 306.81), highlighting a possible connection between water insecurity and mental well-being.
4️⃣ HRS_WEEK (hours of water supply per week) (AIC = 307.12) and MX28_WQ_COMP (perceived comparison to other areas) (AIC = 307.36) suggest that actual water availability and perceived relative quality both shape opinions.
5️⃣ W_WS_LOC (water insecurity classification by Mexico City authorities) had an AIC of 314.65, suggesting that official designations align with, but do not fully predict, personal experiences.


```{r}
# Load required packages
library(broom)

# Run multivariable logistic regression
final_model <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + W_WC_WI + PSS_TOTAL + 
                   HRS_WEEK + MX28_WQ_COMP, 
                   data = data, family = binomial)

# Extract coefficients and confidence intervals
odds_ratios <- tidy(final_model, exponentiate = TRUE, conf.int = TRUE)

# Plot odds ratios
ggplot(odds_ratios, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # OR = 1 reference line
  coord_flip() +
  labs(title = "Odds Ratios for Predictors of Perceived Water Service Quality",
       x = "Predictor", y = "Odds Ratio") +
  theme_minimal()

```




```{r}

# Load required libraries
library(ggplot2)
library(broom)
library(dplyr)
library(MASS)  # For stepAIC

# Remove missing values for selected variables
data_clean <- data %>%
  drop_na(MX26_EM_HHW_TYPE, HW_TOTAL, W_WC_WI, PSS_TOTAL, HRS_WEEK, MX28_WQ_COMP)

# Ensure outcome variable is a binary factor (0/1)
data_clean$MX26_EM_HHW_TYPE <- factor(data_clean$MX26_EM_HHW_TYPE, levels = c(0, 1))

# ---------------------------
# 1️⃣ Multivariable Logistic Regression Model
# ---------------------------
full_model <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + W_WC_WI + PSS_TOTAL + 
                  HRS_WEEK + MX28_WQ_COMP, 
                  data = data_clean, family = binomial)

# Display summary
summary(full_model)

# ---------------------------
# 2️⃣ Interaction Analysis: HW_TOTAL * PSS_TOTAL
# ---------------------------
interaction_model <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL * PSS_TOTAL + 
                         W_WC_WI + HRS_WEEK + MX28_WQ_COMP, 
                         data = data_clean, family = binomial)

summary(interaction_model)

# ---------------------------
# 3️⃣ Model Comparison Using AIC
# ---------------------------
# Compare AIC values
aic_values <- data.frame(
  Model = c("Full Model", "Interaction Model"),
  AIC = c(AIC(full_model), AIC(interaction_model))
)

print(aic_values)

# Stepwise model selection using stepAIC
optimized_model <- stepAIC(full_model, direction = "both")

# Display optimized model summary
summary(optimized_model)

# ---------------------------
# 4️⃣ Visualizing Odds Ratios
# ---------------------------
# Extract odds ratios from the optimized model
odds_ratios <- tidy(optimized_model, exponentiate = TRUE, conf.int = TRUE)

# Plot odds ratios with confidence intervals
ggplot(odds_ratios, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange() +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # Reference line at OR = 1
  coord_flip() +
  labs(title = "Odds Ratios for Predictors of Perceived Water Service Quality",
       x = "Predictor", y = "Odds Ratio") +
  theme_minimal()

```



```{r}

# Extract odds ratios and confidence intervals
odds_ratios <- tidy(optimized_model, exponentiate = TRUE, conf.int = TRUE)

# Plot odds ratios with confidence intervals
ggplot(odds_ratios, aes(x = term, y = estimate, ymin = conf.low, ymax = conf.high)) +
  geom_pointrange(color = "blue") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +  # OR = 1 reference line
  coord_flip() +
  labs(title = "Odds Ratios for Predictors of Perceived Water Hardship",
       x = "Predictor", y = "Odds Ratio") +
  theme_minimal()



```

```{r}

# Run a "full model" with previously excluded variables forced in
full_model <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + W_WC_WI + PSS_TOTAL + 
                  HRS_WEEK + MX28_WQ_COMP + SEASON + MX9_DRINK_W + 
                  MX10_WSTORAGE + SES_SC_Total, 
                  family = binomial, data = data)

# Compare with optimized model
optimized_model <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + W_WC_WI + PSS_TOTAL + 
                       HRS_WEEK + MX28_WQ_COMP, 
                       family = binomial, data = data)

# Display summaries of both models
#summary(full_model)
summary(optimized_model)

# Compare AIC values
AIC(full_model, optimized_model)

# Check significance of forced-in variables
tidy(full_model)

# Check if adding them improves model fit
anova(optimized_model, full_model, test = "Chisq") 

# Check for multicollinearity
vif(full_model)






# Run a full model forcing in previously excluded variables + interactions
full_model_interactions <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + W_WC_WI + PSS_TOTAL + 
                               HRS_WEEK + MX28_WQ_COMP + SEASON + MX9_DRINK_W + 
                               MX10_WSTORAGE + SES_SC_Total + 
                               HW_TOTAL * SEASON + HW_TOTAL * PSS_TOTAL + 
                               HRS_WEEK * W_WC_WI,  # Adding interactions
                               family = binomial, data = data)

# Compare with optimized model
optimized_model <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + W_WC_WI + PSS_TOTAL + 
                       HRS_WEEK + MX28_WQ_COMP, 
                       family = binomial, data = data)

# Display summaries of models
#summary(full_model_interactions)
summary(optimized_model)

# Compare AIC values
AIC(full_model_interactions, optimized_model)

# Check significance of forced-in variables and interactions
tidy(full_model_interactions)

# Compare models using ANOVA test (Chi-square)
anova(optimized_model, full_model_interactions, test = "Chisq") 

# Check for multicollinearity
vif(full_model_interactions)

# Visualize interaction effects using marginal effects plots
library(ggeffects)

# Plot interaction effects for HW_TOTAL * SEASON
plot(ggpredict(full_model_interactions, terms = c("HW_TOTAL", "SEASON")))

# Plot interaction effects for HW_TOTAL * PSS_TOTAL
plot(ggpredict(full_model_interactions, terms = c("HW_TOTAL", "PSS_TOTAL")))

# Plot interaction effects for HRS_WEEK * W_WC_WI
plot(ggpredict(full_model_interactions, terms = c("HRS_WEEK", "W_WC_WI")))

interaction_model <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + PSS_TOTAL + HRS_WEEK * W_WC_WI + MX28_WQ_COMP, 
                         family = binomial, data = data)

```

# Interactions
```{r}
library(MASS)  # For stepAIC
library(car)   # For VIF

# Baseline Model (Main Effects Only)
model1 <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + PSS_TOTAL + HRS_WEEK + W_WC_WI + MX28_WQ_COMP, 
              family = binomial, data = data)

# Model with Interaction between HRS_WEEK and W_WC_WI
model2 <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + PSS_TOTAL + HRS_WEEK * W_WC_WI + MX28_WQ_COMP, 
              family = binomial, data = data)

# Model with Interaction between HW_TOTAL and W_WC_WI
model3 <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL * W_WC_WI + PSS_TOTAL + HRS_WEEK + MX28_WQ_COMP, 
              family = binomial, data = data)

# Model with Interaction between HW_TOTAL and HRS_WEEK
model4 <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL * HRS_WEEK + PSS_TOTAL + W_WC_WI + MX28_WQ_COMP, 
              family = binomial, data = data)

# Model with Multiple Interactions
model5 <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL * W_WC_WI + HRS_WEEK * W_WC_WI + PSS_TOTAL + MX28_WQ_COMP, 
              family = binomial, data = data)

```
# Comparing models 
```{r}
# Compare AIC values
AIC_values <- data.frame(
  Model = c("Main Effects", "HRS_WEEK * W_WC_WI", "HW_TOTAL * W_WC_WI", 
            "HW_TOTAL * HRS_WEEK", "Multiple Interactions"),
  AIC = c(AIC(model1), AIC(model2), AIC(model3), AIC(model4), AIC(model5))
)

# Order models by AIC (lower is better)
AIC_values <- AIC_values[order(AIC_values$AIC), ]
print(AIC_values)


optimized_model <- stepAIC(model1, 
                           scope = list(lower = model1, upper = model5),
                           direction = "both", trace = TRUE)

summary(optimized_model)  # Final optimized model


vif(optimized_model)

```


# The chosen one
```{r odds ratio}
# Best model
model1 <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL + PSS_TOTAL + HRS_WEEK + MX28_WQ_COMP, 
              family = binomial, data = data)

# Compute Odds Ratios
odds_ratios <- exp(coef(model1))
odds_ratios

# Compute Confidence Intervals for ORs
conf_intervals <- exp(confint(model1))

# Create a DataFrame for Plotting
odds_df <- data.frame(
  Predictor = names(odds_ratios),
  OR = odds_ratios,
  Lower_CI = conf_intervals[, 1],
  Upper_CI = conf_intervals[, 2]
)

# Remove Intercept for Better Visualization
#odds_df <- odds_df[-1, ]

# Print Table of Odds Ratios
print(odds_df)

odds_df$Predictor <- factor(odds_df$Predictor, 
                                  levels = c("(Intercept)", "HW_TOTAL", "PSS_TOTAL", "HRS_WEEK", "MX28_WQ_COMP"),
                                  labels = c("Intercept",
                                             "HWISE\nscore", 
                                             "Perceived\nStress", 
                                             "Hours of\nWater Supply", 
                                             "Perception of water service\n(Same or Better than others)"))

# Save the plot as a high-resolution PNG for a poster
ggsave("odds_ratio_plot.png", width = 10, height = 5, dpi = 300)

ggplot(odds_df, aes(x = reorder(Predictor, OR), y = OR)) +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "#1f78b4") +
  geom_point(size = 4, color = "#ff7f00") +  # Points for ORs
  geom_hline(yintercept = 1, linetype = "dashed", color = "#fdbf6f") +  # Reference Line at OR = 1
  coord_flip() +  # Flip for better readability
  labs(title = "Odds ratios of having \na negative emotional response",
       x = "Predictors",
       y = "Odds Ratio (95% CI)") +
  theme_minimal() + 
  theme(
    axis.text.y = element_text(size = 14),  # Y-axis labels (predictors)
    axis.text.x = element_text(size = 14),  # X-axis labels (odds ratio values)
    axis.title.x = element_text(size = 16, face = "bold"),  # X-axis title
    axis.title.y = element_text(size = 16, face = "bold"),  # Y-axis title
    plot.title = element_text(size = 18, face = "bold", hjust = 0.5)  # Title (if any)
  )




```


Variable	Odds Ratio (OR)	Interpretation
HW_TOTAL	1.230	Higher water insecurity → More negative emotions (23% increase per unit).
PSS_TOTAL	1.078	More stress → More negative emotions (7.8% increase per unit).
MX28_WQ_COMP	0.449	Better perceived water service → Lower odds of negative emotion (55.1% lower odds per unit).
HRS_WEEK	0.9996	Water availability (hours) has no significant effect on emotion.