wflow---
title: "HBA2025_results"
author: "Paloma"
date: "2025-01-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Here you will find the code used to obtain results shown in the annual meeting of the HBA, 2025. 

Abstract:

**Coping with water insecurity: women's strategies and emotional responses in Iztapalapa,
Mexico City**

Water insecurity in urban areas presents distinctive challenges, particularly in marginalized communities. While past studies have documented how households adapt to poor water services, many of these coping strategies come at a significant personal cost. Here we examine the **coping strategies and emotional impacts of unreliable water services** among 400 women in Iztapalapa, Mexico City. Data were collected through surveys over the Fall of 2022 and Spring of 2023. We assessed **household water access, water management practices, and emotional responses to local water services.** 

Results indicate that during acute water shortages, women can spend extended periods (several hours, or sometimes days) waiting for water trucks. Additionally, 57% of respondents reported feeling frustrated or angry about their water situation, while around 20% experienced family conflicts over water use or community-level conflicts around water management, often involving water vendors or government services. 

This study offers one of the first in-depth examinations of how water insecurity specifically affects women in Iztapalapa, a densely populated region of Mexico City with severe water access challenges. The findings highlight the urgent need for policy interventions that address water insecurity with a gender-sensitive approach, recognizing the disproportionate burden placed on women as primary water managers in their households.

# Cleaning screening file
```{r}
# Load necessary libraries
library(dplyr)
library(tidyr)

# Load the dataset
file_path <- "../data/01.SCREENING.csv"
df <- read.csv(file_path, stringsAsFactors = FALSE, na.strings = c("", "N/A", "NA", "pending"))

# Convert all character columns to NA where values match missing patterns
df <- df %>%
  mutate(across(where(is.character), ~ ifelse(. %in% c("", "N/A", "NA", "pending"), NA, .)))

# Remove duplicate IDs and keep rows with at least 10 non-NA values
df_filtered <- df %>%
  mutate(Non_NA_Count = rowSums(!is.na(.))) %>%
  filter(!duplicated(ID) & Non_NA_Count >= 10) %>%
  select(-Non_NA_Count) # Remove helper column


# Display the filtered dataset
print(dim(df_filtered)) # Show dimensions for confirmation
write.csv(df_filtered, "./Filtered_Screening.csv", row.names = FALSE)

d <- df_filtered

# Confirm total number of unique participants
cat("Total unique participants:", length(unique(d$ID)), "\n") 

# Identify missing ID numbers in the first and second trip ranges
missing_first_trip <- setdiff(1:204, d$ID)
cat("Missing IDs in the first trip:", missing_first_trip, "\n")

missing_second_trip <- setdiff(301:497, d$ID)
cat("Missing IDs in the second trip:", missing_second_trip, "\n")

# Count participants within ID ranges
cat("Num of Participants first trip ≤ 250:", sum(d$ID <= 250, na.rm = TRUE), "\n")
cat("Num of Participants second trip ≥ 250:", sum(d$ID >= 250, na.rm = TRUE), "\n")

# Confirm total number of rows
cat("Total number of rows:", nrow(d), "\n")

# Set ID as row names (ensuring IDs are unique and non-NA)
if (!any(duplicated(d$ID)) && !any(is.na(d$ID))) {
    rownames(d) <- as.character(d$ID)
} else {
    warning("Some IDs are duplicated or NA, row names were not set.")
}

# Count and report rows containing NAs
num_na_rows <- sum(!complete.cases(d))
cat("Rows with missing values:", num_na_rows, "\n")

# Identify columns that start with "SES" and end with "SC"
ses_sc_columns <- grep("^SES.*SC$", names(d), value = TRUE)

# Count total number of NA values in each SES_SC column
na_counts <- colSums(is.na(d[ses_sc_columns]))

# Count total NA values across all SES_SC columns
total_na_count <- sum(na_counts)

# Print results
cat("Number of NA values per SES_SC column:\n")
print(na_counts)
cat("\nTotal NA values across all SES_SC columns:", total_na_count, "\n")

# Check if SES_SC columns exist before proceeding
if (length(ses_sc_columns) > 0) {
  
  # Convert SES_SC columns to numeric (handles factors and character types)
  d[ses_sc_columns] <- lapply(d[ses_sc_columns], function(x) as.numeric(as.character(x)))
  
  # Sum SES_SC columns row-wise, keeping NA if any SES_SC value is missing
  d <- d %>%
    rowwise() %>%
    mutate(SES_SC_Total = if_else(any(is.na(c_across(all_of(ses_sc_columns)))),
                                  NA_real_, 
                                  sum(c_across(all_of(ses_sc_columns)), na.rm = TRUE))) %>%
    ungroup()
  
  # Display confirmation message
  cat("Added SES_SC_Total column, maintaining NA where any SES_SC column has NA.\n")
  
} else {
  cat("No SES_SC columns found in the dataset.\n")
}

# Count the number of NAs in the SES_SC_Total column
num_na_ses_sc_total <- sum(is.na(d$SES_SC_Total))

# Print the count
cat("Number of NA values in SES_SC_Total:", num_na_ses_sc_total, "\n")
print(d[is.na(d$SES_SC_Total),c("ID",ses_sc_columns)], n=50)

# Calculate the percentage of missing values per column
missing_percentage <- colMeans(is.na(d)) * 100

# Get columns with more than 10% missing data
columns_to_remove <- names(missing_percentage[missing_percentage > 10])
length(columns_to_remove)
dim(d)

# Print the selected columns
cat("Columns with more than 10% missing values:\n")
print(columns_to_remove)

d <- d %>% select(-all_of(columns_to_remove))
dim(d)

head(d)

```

## Data type transformations
```{r}
# Function to convert non-numeric values to "other"
convert_to_other <- function(x) {
  ifelse(grepl("^[0-9]+$", x) | x == "other" | is.na(x), x, "other")
}

# Apply transformation from yes and no to 1 and 0, to some columns
selected_columns <- c("D_CHLD", "D_HH_SIZE")
d <- d %>%
  mutate(across(all_of(selected_columns), ~ convert_to_other(as.character(.))))

# Specify the columns where transformation should be applied
columns_to_transform <- c("HLTH_SMK", "HLTH_TTM", "HLTH_CPAIN")  # Replace with actual column names

# Convert "yes" to 1, "no" to 0, and everything else to "other"
d <- d %>%
  mutate(across(all_of(columns_to_transform), ~ case_when(
    . == "Yes" ~ "1",
    . == "No"  ~ "0",
    !is.na(.)  ~ "other",  # Assign "other" to all non-missing values that are not "yes" or "no"
    TRUE ~ NA_character_
  )))

# Convert numeric-coded columns to factors where applicable
d <- d %>%
  mutate(across(where(~ all(. %in% c("0", "1", "other", NA_character_))), as.factor))  # Ensure consistency

# Convert variables to numeric when appropriate
## Pending: D_YRBR, D_LOC_TIME, D_CHLD, D_HH_SIZE

##

print(d)
summary(d)
```

## Select cleaned columns

```{r}
#Keep columns relevant/cleaned
columns_to_keep2 <- c("ID", "D_YRBR", "D_LOC_TIME", "D_AGE", "D_HH_SIZE", "D_CHLD", "HLTH_SMK", "SES_SC_Total")
d <- d %>% select(all_of(columns_to_keep2))
write.csv(d,"../data/00.SCREENING_V2.csv")

```

# Merge with HWISE data

```{r}

file2 <- "../data/02.HWISE_PSS.csv"
file3 <- "../data/Chronic_pain_illness.csv"
df1 <- d
df2 <- read.csv(file2, stringsAsFactors = FALSE, na.strings = c("", "N/A", "NA", "pending"))
df3 <- read.csv(file3, stringsAsFactors = FALSE, na.strings = c("", "N/A", "NA", "pending"))

# Identify the common column for merging
common_column <- "ID"  # Change this if needed

# Ensure the common column is of the same type in all datasets
df1[[common_column]] <- as.character(df1[[common_column]])
df2[[common_column]] <- as.character(df2[[common_column]])
df3[[common_column]] <- as.character(df3[[common_column]])

# Merge datasets sequentially by the common column
merged_df <- df1 %>%
  full_join(df2, by = common_column) %>%
  full_join(df3, by = common_column)

cols_to_remove<- c("HW_INFO", "PSS_INFO", "PSS_TOTAL", "HW_TOTAL")

merged_df <- merged_df %>% select(-all_of(cols_to_remove))


##### Calculate final HW score
# Identify columns that start with "HW_"
hw_columns <- grep("^HW_", names(merged_df), value = TRUE)

# Sum up all HW_ columns row-wise, keeping NA if any HW_ column has NA
merged_df <- merged_df %>%
  rowwise() %>%
  mutate(HW_TOTAL = if_else(any(is.na(c_across(all_of(hw_columns)))),
                            NA_real_, 
                            sum(c_across(all_of(hw_columns)), 
                                na.rm = TRUE))) %>%
  ungroup()


# Print merged dataset dimensions
cat("Merged dataset dimensions:", dim(merged_df), "\n")
print(merged_df, n=20)
# Save the merged dataset
write.csv(merged_df, "./data/Cleaned_Dataset_Screening_HWISE_PSS.csv", row.names = FALSE)
cat("Merged dataset saved as 'Cleaned_Dataset_Screening_HWISE_PSS.csv'.\n")

```

# Merge dataset with info on hours of water supply

```{r}

data_path = "../data"
# merged dataset
df1 = read.csv(file.path(data_path, "Cleaned_Dataset_Screening_HWISE_PSS.csv"), stringsAsFactors = FALSE, na.strings = c("", "N/A", "NA", "pending"))
df2 = read.csv(file.path(data_path, "hours_water_supply.csv"), stringsAsFactors = FALSE, na.strings = c("", "N/A", "NA", "pending"))


# Merge datasets sequentially by the common column
merged_df <- df1 %>%
  full_join(df2, by = "ID") 

seas <- c("SEASON")
merged_df <- merged_df %>%
  mutate(across(all_of(seas), ~ case_when(
    . == "Oct-Dec" ~ "1",
    . == "Apr-Jun"  ~ "0",
    !is.na(.)  ~ "other",  # Assign "other" to all non-missing values that are not "yes" or "no"
    TRUE ~ NA_character_
  )))

#### THIS COL SHOULD BE REMOVED. TOO MANY NAs
#Specify the columns where transformation should be applied
columns_to_transform <- c("W_WC_WI") #  # Replace with actual column names

# Convert "intermitent" to 1, "continuous" to 0, and everything else to "other"
merged_df <- merged_df %>%
  mutate(across(all_of(columns_to_transform), ~ case_when(
    . == "Intermitente" ~ "1",
    . == "Continua"  ~ "0",
    !is.na(.)  ~ "other",  # Assign "other" to all non-missing values that are not "yes" or "no"
    TRUE ~ NA_character_
  )))

columns_to_transform <- c("W_WS_LOC")
# Convert "WI" to 1, "WS" to 0, and everything else to "other"
merged_df <- merged_df %>%
  mutate(across(all_of(columns_to_transform), ~ case_when(
    . == "WI" ~ "1",
    . == "WS"  ~ "0",
    !is.na(.)  ~ "other",  # Assign "other" to all non-missing values that are not "yes" or "no"
    TRUE ~ NA_character_
  )))

# Define the list of variables to convert to numeric
col_to_num <- c("D_YRBR", "D_LOC_TIME", "D_HH_SIZE", "D_CHLD", "HRS_WEEK", "HLTH_SMK", "W_WS_LOC", "SEASON", "HRS_WEEK", "W_WC_WI")

# Convert the specified columns to numeric
merged_df <- merged_df %>%
  mutate(across(all_of(col_to_num), ~as.numeric(as.character(.)), .names = "{.col}"))

```

```{r}

# Print merged dataset dimensions
cat("Merged dataset dimensions:", dim(merged_df), "\n")
head(merged_df, 20)
summary(merged_df)
dim(merged_df)
# Save the merged dataset
write.csv(merged_df, "../data/Cleaned_Dataset_Screening_HWISE_PSS_V2.csv", row.names = FALSE)
cat("Merged dataset saved as 'Cleaned_Dataset_Screening_HWISE_PSS_V2.csv'.\n")

```


