---
title: "HBA2025_MX28"
author: "Paloma C."
date: "2025-03-06"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Here we explore the survey responses for questions 26 and 28.

Question 28 of the QUAIL-Mex questionnaire reveals if participants think of their water service as worse, similar, or better than in other parts of Mexico City. Exploring this data and the factors that may influence this perception is an important dimension of how water insecurity is viewed by participants.

*warning*: some explanations and interpretations were written using Artificial Intelligence. They are in the process of being evaluated, filtered, and re-written by the author.

```{r load files, warning=FALSE, message=FALSE, echo = FALSE}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(corrplot)
library(rstatix)
library(rcompanion)

# Read the dataset
data.file <- "data/Cleaned_Dataset_Screening_HWISE_PSS_V3.csv"
data <- read.csv(data.file)

data$MX28_WQ_COMP <- as.numeric(data$MX28_WQ_COMP)

# select data
keep_columns <- c("MX28_WQ_COMP",
                "D_LOC_TIME",
                "D_AGE",
                "D_HH_SIZE",
                "D_CHLD",
                "HLTH_SMK",
                "SES_SC_Total",
                "SEASON",
                "HW_TOTAL",
                "W_WC_WI",
                "W_WS_LOC",
                "HRS_WEEK")

data <- data %>% 
  select(keep_columns)
```
# Q28
## Compute correlation of numeric variables in data set
```{r correlations,warning = FALSE, message=FALSE, echo = FALSE}

if("MX28_WQ_COMP" %in% colnames(data)) {
  
  # Select numeric columns for correlation analysis
  numeric_data <- data %>% select(where(is.numeric))
  
  # Compute correlation matrix
  correlation_matrix <- cor(numeric_data, use = "pairwise.complete.obs")
  
  # Extract correlations with MX28_WQ_COMP
  correlations <- as.data.frame(correlation_matrix["MX28_WQ_COMP", ])
  correlations$Variable <- rownames(correlations)
  colnames(correlations)[1] <- "Correlation"
  
  # Print correlation values sorted in descending order
  print(correlations %>% arrange(desc(Correlation)))
  
  # Visualize correlation matrix as a heatmap
  corrplot(correlation_matrix, method = "color", type = "upper", tl.col = "black", tl.srt = 45)
    
  }

```

## Barplots using two categorical variables (percentages)
```{r barplots q28 catvars, echo = FALSE, warning=FALSE}
# visualize data from Q28
# Convert selected variables to factors
categorical_vars <- c("MX28_WQ_COMP", "SEASON", "W_WC_WI", "W_WS_LOC")
data <- data %>%
  mutate(across(all_of(categorical_vars), as.factor))

cat("Verifying conversion of variables to factors")
str(data[categorical_vars])

# Define the main categorical variable
main_var <- "MX28_WQ_COMP"

# Rename levels of MX28_WQ_COMP
levels(data$MX28_WQ_COMP) <- c("Worse", "Same", "Better")
levels(data$SEASON) <- c("Fall", "Spring")
levels(data$W_WC_WI) <- c("Continous water supply", "Intermittent")
levels(data$W_WS_LOC) <- c("water secure neighborhood", "water insecure neighborhood")

# Verify that levels have been updated
print(levels(data$MX28_WQ_COMP))

# Exclude MX28_WQ_COMP from the list of categorical variables
cat_vars <- setdiff(categorical_vars, main_var)

# Generate bar plots with percentages
for (var in cat_vars) {
  p <- ggplot(data, aes_string(x = main_var, fill = var)) +
    geom_bar(aes(y = (..count..) / sum(..count..) * 100), position = "fill") +
    scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Convert to percentage format
    theme_minimal() +
    labs(title = paste("Perceived service in relation to others", "vs", var),
         x = "Perception of water service", y = "Percentage", fill = var)
  
  print(p)  # Display each bar plot
}

```

# Q26
Question 26 asks: how do participants feel when they think about their water service? This is a question asked towards the end of the interview, after participants have evaluated and reflected on multiple aspects of their water service. Here we ask: what aspects influence a positive or negative emotional response? What elements can buffer other aspects of water insecurity and reduce their negative feelings about the service? 

## Barplots Q26 and HWISE categories
In the following plots we can visualize how people feel when they have different levels of water insecurity
```{r barplots q26 hwise cat, echo = FALSE}
data <- read.csv(data.file)
# Ensure HW_TOTAL_category is correctly categorized
data <- data %>%
  mutate(HW_TOTAL_category = case_when(
    HW_TOTAL >= 0 & HW_TOTAL <= 2  ~ "No-to-Marginal",
    HW_TOTAL >= 3 & HW_TOTAL <= 11 ~ "Low",
    HW_TOTAL >= 12 & HW_TOTAL <= 23 ~ "Moderate",
    HW_TOTAL >= 24 & HW_TOTAL <= 36 ~ "High"
  ))

data <- data %>%
  mutate(MX26_EM_HHW_TYPE = case_when(
    MX26_EM_HHW_TYPE == "0" ~ "positive",
    MX26_EM_HHW_TYPE == "1" ~ "negative",
    MX26_EM_HHW_TYPE == !is.na(.) ~ "other",  # Assign "other" to all non-missing values that are not "yes" or "no"
    TRUE ~ NA_character_
  ))

data <- data %>% 
  filter(!is.na(HW_TOTAL_category), !is.na(MX26_EM_HHW_TYPE))

# Convert to factor for correct ordering
data$HW_TOTAL_category <- factor(data$HW_TOTAL_category, levels = c("No-to-Marginal", "Low", "Moderate", "High"))
data$MX26_EM_HHW_TYPE <- as.factor(data$MX26_EM_HHW_TYPE)  # Convert to factor

# Barplot of MX26_EM_HHW_TYPE by HW_TOTAL_category
ggplot(data, aes(x = HW_TOTAL_category, fill = MX26_EM_HHW_TYPE)) +
  geom_bar(position = "dodge") +
  labs(title = "Proportion of Positive and negative emotional responses by HW Category",
       x = "HWISE Category",
       y = "Count",
       fill = "Emotions") +
  theme_minimal()

# Calculate proportions
data_prop <- data %>%
  group_by(HW_TOTAL_category, MX26_EM_HHW_TYPE) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(percentage = count / sum(count) * 100)

# Create percentage barplot
ggplot(data_prop, aes(x = HW_TOTAL_category, y = percentage, fill = MX26_EM_HHW_TYPE)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Proportion of Positive and negative emotional responses by HW Category",
       x = "HWISE Category",
       y = "Percentage",
       fill = "Emotions") +
  theme_minimal() 

```

```{r stacked plot , echo = FALSE}

# Stacked percentage barplot
ggplot(data_prop, aes(x = HW_TOTAL_category, y = percentage, fill = MX26_EM_HHW_TYPE)) +
  geom_bar(stat = "identity", position = "fill") +  # "fill" makes it a 100% stacked barplot
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +  # Convert y-axis to percentage
  labs(title = "Proportion of Household Water Type by HW Category",
       x = "HWISE Category",
       y = "Percentage",
       fill = "Household Water Type") +
  theme_minimal()

```

## Testing associations: Cramer's V 
Contingency table comparing cases in each category 
```{r echo = FALSE}
# Create a contingency table
contingency_table <- table(data$HW_TOTAL_category, data$MX26_EM_HHW_TYPE)

# Print the table
print(contingency_table)

  fisher_test <- fisher.test(contingency_table)
  print(fisher_test)  # Print Fisher's test results

# Compute Cramér's V for effect size
  cat("Compute Cramér's V for effect size:\n 0.1 → Small association\n 0.3 → Moderate association\n 0.5 → Strong association")
cramer_v <- cramerV(contingency_table)
print(cramer_v)

```

The Cramér’s V value of 0.3961 suggests a moderate association between HW_TOTAL_category and MX26_EM_HHW_TYPE. Cramér’s V ranges from 0 to 1, where 0 indicates no association, and 1 indicates a perfect association. A value of 0.3961 falls in the moderate range (0.3 - 0.5), implying that while there is a statistically significant relationship between these two categorical variables, it is not extremely strong. This means that the emotional response to water service (MX26_EM_HHW_TYPE) varies meaningfully across different levels of water insecurity (HW_TOTAL_category), but other factors might also be influencing this relationship. Further analysis, such as post-hoc tests or visualizing the data with stacked barplots, could provide more insights into how the categories differ and whether certain water supply types are more prevalent in specific HWISE categories.

## Testing associations: Pairwise Fisher's Exact Test
Using Bonferroni correction for multiple comparisons
```{r fisher, echo=FALSE}
# Run Fisher’s Exact Test
fisher_test <- fisher.test(contingency_table)

# Print results
print(fisher_test)
# Get unique categories of HW_TOTAL_category
categories <- levels(data$HW_TOTAL_category)

# Create an empty list to store results
pairwise_results <- list()

# Loop through all pairs of HW_TOTAL_category
for (i in 1:(length(categories) - 1)) {
  for (j in (i + 1):length(categories)) {
    cat1 <- categories[i]
    cat2 <- categories[j]
    
    # Subset the data for the two categories
    subset_data <- data %>%
      filter(HW_TOTAL_category %in% c(cat1, cat2))
    
    # Create a 2xN contingency table
    sub_table <- table(subset_data$HW_TOTAL_category, subset_data$MX26_EM_HHW_TYPE)
    
    # Run Fisher’s Exact Test
    fisher_res <- fisher.test(sub_table)
    
    # Store results in a data frame
    pairwise_results[[paste(cat1, "vs", cat2)]] <- data.frame(
      Group1 = cat1,
      Group2 = cat2,
      p_value = fisher_res$p.value
    )
  }
}

# Combine results into a single data frame
pairwise_fisher_results <- bind_rows(pairwise_results)

# Apply Bonferroni correction for multiple comparisons
pairwise_fisher_results <- pairwise_fisher_results %>%
  mutate(p_adj = p.adjust(p_value, method = "bonferroni"))

# Print results
print(pairwise_fisher_results)
```
If p_adj < 0.05, it indicates a significant difference between the two HW_TOTAL_category groups.

If p_adj > 0.05, the difference is not statistically significant after Bonferroni correction



✅ Significant Differences (p_adj < 0.05):
Comparison	Interpretation
No-to-Marginal vs. Low	Significant difference (p_adj = 4.66e-05), meaning that households in the No-to-Marginal category have significantly different water types compared to the Low category.
No-to-Marginal vs. Moderate	Highly significant difference (p_adj = 2.19e-13), suggesting that households with No-to-Marginal water insecurity have a very different water type distribution than Moderate water-insecure households.
No-to-Marginal vs. High	Significant difference (p_adj = 0.0181), indicating that No-to-Marginal households also differ from High water-insecure households.
Low vs. Moderate	Significant difference (p_adj = 7.44e-05), suggesting that Low water-insecure households also have different water types compared to Moderate water-insecure households.
❌ Non-Significant Differences (p_adj > 0.05):
Comparison	Interpretation
Low vs. High	p_adj = 1.000 (No significant difference) – meaning water types in Low and High HWISE categories are similar.
Moderate vs. High	p_adj = 1.000 (No significant difference) – meaning water types in Moderate and High HWISE categories are similar.

📝 Key Takeaways
There are significant differences between No-to-Marginal households and all other categories (Low, Moderate, High).
Low and Moderate categories are also significantly different.
High water-insecure households do not significantly differ from Low or Moderate categories, meaning their water sources are similar.

## Plots Emotion/HWISE category
```{r plots emotion/hwise, echo = FALSE}

# Create the stacked barplot
ggplot(data, aes(x = HW_TOTAL_category, fill = MX26_EM_HHW_TYPE)) +
  geom_bar(position = "fill") +  # Fill ensures bars are proportional
  labs(title = "Proportion of Household Water Types by HW_TOTAL_category",
       x = "HW_TOTAL Category",
       y = "Proportion",
       fill = "Water Type") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()

```

# Regressions

Why Logistic Regression?
Dependent Variable (MX26_EM_HHW_TYPE) is categorical (Nominal variable with multiple categories).
Independent Variable (HW_TOTAL_category) is categorical.
Logistic regression is appropriate for categorical outcomes, unlike linear regression which assumes continuous data.

Need to run binary logistic regression.

```{r}
# Load necessary libraries
library(dplyr)

# Ensure variables are factors
data <- data %>%
  mutate(HW_TOTAL_category = as.factor(HW_TOTAL_category),
         MX26_EM_HHW_TYPE = as.factor(MX26_EM_HHW_TYPE))

# Run Binary Logistic Regression
logit_model <- glm(MX26_EM_HHW_TYPE ~ HW_TOTAL_category, 
                   data = data, 
                   family = binomial)

# Show summary of model
summary(logit_model)

```


```{r}
# Compute odds ratios
exp(coef(logit_model))
```
✔ Odds ratios < 1 → Decreased likelihood
✔ Odds ratios > 1 → Increased likelihood

Suggests that any level of water insecurity decreases the probability of having positive feelings about water supply. 

Predictor	Odds Ratio (OR)
(Intercept)	1.846
HW_TOTAL_categoryLow	0.277
HW_TOTAL_categoryModerate	0.0655
HW_TOTAL_categoryHigh	9.41e-08

Reference category is MX26_EM_HHW_TYPE == positive
HW_TOTAL_categoryLow OR = 0.277 → Individuals in the "Low" category are 72.3% less likely (1 - 0.277) to belong to MX26_EM_HHW_TYPE compared to the reference.
HW_TOTAL_categoryModerate OR = 0.0655 → Individuals in the "Moderate" category are 93.45% less likely (1 - 0.0655).
HW_TOTAL_categoryHigh OR = 9.41e-08 → Extremely small, suggesting that this group is almost never associated with the reference category.

The "High" HWISE category (OR = 9.41e-08) has the largest effect, meaning it is almost completely absent from the reference water type.
Moderate category (OR = 0.0655) has a strong negative effect as well.
Low category (OR = 0.277) has a smaller negative effect than Moderate and High.
Intercept is just the baseline, not a comparison.


Check model fit

```{r}
# Compute pseudo R-squared (Nagelkerke)
library(pscl)
pR2(logit_model)

```
If the High category has very few cases, consider merging it with another category.
✔ Higher values (~0.2–0.5) indicate a better model.

```{r}
# Create a data frame with odds ratios, confidence intervals, and p-values
effect_sizes <- data.frame(
  Category = c("Intercept", "Low", "Moderate", "High"),
  Odds_Ratio = c(1.846, 0.277, 0.0655, 9.41e-08),
  Lower_CI = c(1.2, 0.18, 0.03, 1e-10),  # Replace with actual confidence interval values
  Upper_CI = c(2.8, 0.42, 0.15, 1e-05),  # Replace with actual confidence interval values
  p_value = c(0.012, 0.0001, 0.00001, 0.99)  # Replace with actual p-values
)

# Convert Category into a factor for ordered plotting
effect_sizes$Category <- factor(effect_sizes$Category, levels = rev(effect_sizes$Category))  # Reverse for horizontal plotting

ggplot(effect_sizes, aes(x = Odds_Ratio, y = Category)) +
  geom_point(size = 4, color = "blue") +  # Plot odds ratios as points
  geom_errorbarh(aes(xmin = Lower_CI, xmax = Upper_CI), height = 0.2, color = "black") +  # Add confidence intervals (horizontal)
  scale_x_log10() +  # Log-scale for better visualization
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", linewidth = 1) +  # Add reference line at OR = 1
  geom_text(aes(label = paste0("p = ", formatC(p_value, format = "e", digits = 2))), 
            hjust = 2, vjust = 1, size = 5, color = "black") +  # Annotate p-values
  labs(title = "Effect Sizes (Odds Ratios) of HW_TOTAL_category",
       x = "Odds Ratio (Log Scale)",
       y = "HWISE Category") +
  theme_minimal()


```



```{r}
data <- data %>% filter(!is.na(HRS_WEEK) & !is.na(MX26_EM_HHW_TYPE))

wilcox.test(HRS_WEEK ~ MX26_EM_HHW_TYPE, data = data)
library(effsize)
cliff.delta(HRS_WEEK ~ MX26_EM_HHW_TYPE, data = data)

```

Interpretation of Cliff's Delta:

|d| < 0.147 → Negligible effect
0.147 ≤ |d| < 0.33 → Small effect
0.33 ≤ |d| < 0.474 → Medium effect
|d| ≥ 0.474 → Large effect

```{r}
logit_model <- glm(MX26_EM_HHW_TYPE ~ HRS_WEEK, 
                   data = data, 
                   family = binomial)
summary(logit_model)


data$predicted <- predict(logit_model, type = "response")

# Ensure correct row alignment
if(nrow(data) != length(data$predicted)) {
    stop("Mismatch in row counts: Check for missing values or inconsistent dataset.")
}

```

✅ Interpreting Logistic Regression:
Significant HRS_WEEK (p < 0.05) → Water supply duration is linked to positive/negative perceptions.


```{r}
exp(coef(logit_model))

```
When HRS_WEEK = 0 (i.e., baseline scenario), the odds of having a positive perception of water service (MX26_EM_HHW_TYPE) are 0.2699.
This means the probability of a positive perception is quite low when water supply is absent. In terms of probability, 21% prob. of positive perception


HRS_WEEK (OR = 1.0067)
For every additional hour of water supply per week (HRS_WEEK), the odds of having a positive perception (MX26_EM_HHW_TYPE) increase by ~0.67%.
Since OR > 1, more hours of water supply are associated with a higher likelihood of positive perception

Practical Interpretation
A small effect: Since the OR is very close to 1, an increase in HRS_WEEK has a very small impact on perception.
Example Calculation:
If HRS_WEEK increases by 10 hours, the odds of positive perception change by: 1.0067^{10} = 1.0693 \ (\text{~6.93% increase in odds})
If HRS_WEEK increases by 50 hours, the odds increase by: 1.0067^{50} = 1.398 \ (\text{~39.8% increase in odds})


Water supply duration (HRS_WEEK) positively influences water service perception (MX26_EM_HHW_TYPE).
✔ The effect is small, meaning other factors likely play a role in perception (e.g., water quality, socioeconomic status).
✔ Policy Implication: While increasing water supply may improve perception, other factors (such as reliability, quality, and accessibility) should be considered.


```{r}

data <- data %>% 
  filter(!is.na(W_WC_WI), !is.na(SES_SC_Total), !is.na(SEASON))
# add covariates
logit_model <- glm(MX26_EM_HHW_TYPE ~ HRS_WEEK + SES_SC_Total + W_WC_WI + SEASON, 
                   data = data, family = binomial)
exp(coef(logit_model))  # Get updated odds ratios


data$predicted <- predict(logit_model, type = "response")

ggplot(data, aes(x = HRS_WEEK, y = predicted)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), color = "blue") +
  labs(title = "Effect of HRS_WEEK on Perceived Water Service",
       x = "Hours of Water Supply per Week",
       y = "Predicted Probability of Positive Perception") +
  theme_minimal()

```




```{r}
library(ggplot2)
ggplot(data, aes(x = MX26_EM_HHW_TYPE, y = HRS_WEEK, fill = MX26_EM_HHW_TYPE)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(title = "Distribution of HRS_WEEK by Perceived Water Service",
       x = "Perception of Water Service",
       y = "Hours of Water Supply per Week") +
  theme_minimal()


data$predicted <- predict(logit_model, type = "response")

# Ensure correct row alignment
if(nrow(data) != length(data$predicted)) {
    stop("Mismatch in row counts: Check for missing values or inconsistent dataset.")
}




ggplot(data, aes(x = HRS_WEEK, y = predicted)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), color = "blue") +
  labs(title = "Effect of HRS_WEEK on Water Perception",
       x = "Hours of Water Supply per Week",
       y = "Predicted Probability of Positive Perception") +
  theme_minimal()


ggplot(data, aes(x = MX26_EM_HHW_TYPE, y = HRS_WEEK, fill = MX26_EM_HHW_TYPE)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  labs(title = "Distribution of HRS_WEEK by Perceived Water Service",
       x = "Perception of Water Service",
       y = "Hours of Water Supply per Week") +
  theme_minimal()
```


```{r}
logit_model <- glm(MX26_EM_HHW_TYPE ~ HRS_WEEK + SES_SC_Total + SEASON, 
                   data = data, 
                   family = binomial)
summary(logit_model)

```
Null Deviance = 387.61 → The model with no predictors.
Residual Deviance = 371.68 → The model with predictors.
AIC = 379.68 → Lower AIC values indicate a better model fit.
42 observations deleted due to missing values → Consider imputing or removing predictors with excessive missingness.
📌 The model explains some variation in perception, but additional covariates may improve the fit.


Interpretation of Coefficients
Intercept (-1.509241)
When HRS_WEEK, SES_SC_Total, and SEASON = 0, the odds of a positive perception of water service (MX26_EM_HHW_TYPE = 1) are: 0.221
This means that in the baseline condition, the probability of positive perception is very low.

Effect of HRS_WEEK (0.006903, p = 0.000119)
✔ Statistically significant at p < 0.001 (*).**
✔ For every additional hour of water supply per week (HRS_WEEK), the odds of a positive perception increase by: 
This means a 0.69% increase in odds per additional hour of water supply.
While significant, the effect size is small.
Example: A 50-hour increase in water supply:
𝑒
0.006903
×
50
=
1.37

37% increase in odds of a positive perception.
More hours of water supply are associated with a higher likelihood of positive perception. However, the effect size is small, suggesting that other factors may play a more significant role

Effect of SES_SC_Total (0.002020, p = 0.469056)
❌ Not statistically significant (p = 0.469).
❌ Socioeconomic status (SES) does not significantly predict positive water service perception in this model.
💡 This suggests that water perception is not strongly related to SES after adjusting for water supply hours.


Effect of SEASON (-0.096373, p = 0.705301)
❌ Not statistically significant (p = 0.705).
❌ Season (Fall vs. Spring) does not significantly predict positive water perception.
💡 This suggests that water perception remains stable across seasons in your dataset.

Conclusion
More water supply hours (HRS_WEEK) significantly increase the odds of positive water perception.
Socioeconomic status (SES_SC_Total) and season (SEASON) do not have a statistically significant effect.
The effect of HRS_WEEK is small, meaning other factors may be more influential.
To improve the model, consider adding more variables (e.g., water quality concerns, reliability of service, household stress).


```{r}

library(ggplot2)
library(broom)

# Convert coefficients to odds ratios
odds_ratios <- tidy(logit_model, exponentiate = TRUE, conf.int = TRUE)

ggplot(odds_ratios, aes(x = term, y = estimate)) +
  geom_point(size = 4, color = "blue") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "red") +
  coord_flip() +
  labs(title = "Odds Ratios for Predicting Positive Water Service Perception",
       y = "Odds Ratio", x = "Predictor Variable") +
  theme_minimal()

```


```{r}


library(ggeffects)
predicted_data <- ggpredict(logit_model, terms = "HRS_WEEK [all]")

ggplot(predicted_data, aes(x = x, y = predicted)) +
  geom_line(color = "blue") +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  labs(title = "Effect of Water Supply Hours on Water Perception",
       x = "Hours of Water Supply per Week",
       y = "Predicted Probability of Positive Perception") +
  theme_minimal()

```

✔ Try adding interaction effects (e.g., HRS_WEEK * SES_SC_Total) to see if SES influences water perception at different supply levels.
✔ Include more predictors related to water quality, infrastructure, or trust in tap water.
✔ Run diagnostics to check model assumptions.