---
title: "Plots_Latinx_research_week"
author: "Paloma"
date: "2025-01-27"
output: workflowr::wflow_html
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r warnings = FALSE, echo = FALSE, message=FALSE}
# Load libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(broom)
library(ggpubr)
library(patchwork)
library(corrplot)

# Load your data
df <- read.csv("./data/Cleaned_Dataset_Screening_HWISE_PSS_V3.csv")

# View summary
glimpse(df)
```

### Data cleaning

```{r screening data}

#Loading file and setting empty cells as "NAs"
d <- read.csv("./data/01.SCREENING.csv", stringsAsFactors = TRUE, na.strings = c("", " ","NA", "N/A"))


 # extract duplicates and compare values (removed manually for now)
dup <- d$ID[duplicated(d$ID)] 
length(dup) # 38
print(dup)

# remove duplicates
d <- d[!duplicated(d$ID),]
# confirm total number of participants
length(unique(d$ID)) # 433 rows, but 394 participants


# What numbers are missing? 
ID <- as.ordered(d$ID)
# first trip
setdiff(1:204, ID)
nrow(d[d$ID <=250,])

# second trip
setdiff(301:497, ID)
nrow(d[d$ID >=250,])

nrow(d)


 #make ID number the row names
rownames(d) <- d$ID

# Code NAs
d[d=='NA'] <- NA
d <- d %>% 
    replace_na()

# Count rows with NAs
nrow(d[rowSums(is.na(d)) > 0,]) # 12 rows
nrow(d[!complete.cases(d),]) # 12 rows

#NAs <- d[rowSums(is.na(d))> 0,]
#write.table(NAs, "240301_SES_Age_NAs.csv", row.names=FALSE)

# Select useful information
d <- d %>% 
  select(ID, SES_EDU_SC, SES_BTHR_SC, SES_CAR_SC, SES_INT_SC, SES_WRK_SC, SES_BEDR_SC)

# transform factors to numbers
for (i in c(2:length(d))) {
    d[,i] <- as.numeric(as.character(d[,i]))
}

# Keep rows with no missing data
d <- d[complete.cases(d),] # - 9 (dup)  # total 382 complete cases, total 394
nrow(d)
head(d)
#write.csv(d, paste("./cleaned/", date, "_SES_clean.csv", sep=""))


# Calcular total SES score
d$SES_score <- rowSums(d[2:7], na.rm = TRUE)

# SES categories
d$SES <- ifelse(d$SES_score <= 47,"E", 
                ifelse(d$SES_score <= 89, "D-/E", 
                       ifelse(d$SES_score <= 111, "D", 
                              ifelse(d$SES_score <= 135, "C-",
                                     ifelse(d$SES_score <= 165, "C",
                                            ifelse(d$SES_score <= 204, "C+",
                                                   "A/B"))))))
head(d)
```

###  combining data
```{r }

#Loading file and setting empty cells as "NAs"
c <- read.csv("./data/Chronic_pain_illness.csv", stringsAsFactors = TRUE, na.strings = c("", " ","NA", "N/A"))

c$chronic <- as.factor(ifelse(c$HLTH_CPAIN_CAT == 1 | c$HLTH_CDIS_CAT == 1, "Yes", "No"))

m <- merge(d,c, by="ID")
dim(m)
m[m =='NA'] <- NA
m <- m %>% 
    replace_na()
m <- m %>% 
  drop_na()

m %>%
  group_by(SES) %>%
  summarise(n_total = n())


agg<- count(m, SES, chronic)
agg2 <- pivot_wider(agg, 
            names_from = chronic,
            values_from = n)
agg2$Total <- agg2$Yes/(agg2$Yes + agg2$No)*100

#png(file= "HLTH_counts_by_SES.png", width = 700, height = 800 )
ggplot(agg2, aes(x = SES, y = Total)) +
    geom_bar(stat= "identity", fill = "orchid4", alpha = 0.9) +
     ylim(0, 55) + 
   theme_minimal()  +
   geom_text(aes(label = paste(round(Total, 1), "%", sep="")), 
            vjust = -0.5, 
            colour = "gray40", 
            size = 9) +
   theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 32, face = "bold"),
    axis.title = element_text(size = 22),  
    axis.text = element_text(size = 26)  
  ) + 
  labs(title="Proportion of participants suffering from \n chronic pain or chronic disease") +
  xlab("Socioeconomic status") +
  ylab("Percent (%)")  

#dev.off()

data <- m
# Step 1: Create a frequency table by SES and chronic condition
plot_data <- data %>%
  count(SES, chronic) %>%
  group_by(SES) %>%
  mutate(percent = 100 * n / sum(n))

# Step 2: Plot as grouped bar chart (percentage)
ggplot(plot_data, aes(x = SES, y = percent, fill = chronic)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("No" = "steelblue", "Yes" = "firebrick")) +
  labs(
    title = "Chronic Condition by Socioeconomic Status",
    x = "Socioeconomic Status",
    y = "Percentage",
    fill = "Chronic Condition"
  ) +
  theme_minimal(base_size = 13)


```

# models
```{r}
# Sample dataframe (assumes it's already loaded as `data`)
data <- m
data$chronic <- factor(data$chronic, levels = c("No", "Yes"))
# Model 1: Overall SES score
m1 <- glm(chronic ~ SES_score, data = data, family = "binomial")

# Model 2: Overall SES category
m2 <- glm(chronic ~ SES, data = data, family = "binomial")

# Model 3: Individual SES components
m3 <- glm(chronic ~ SES_EDU_SC + SES_BTHR_SC + SES_CAR_SC + SES_INT_SC +
            SES_WRK_SC + SES_BEDR_SC, data = data, family = "binomial")

# Model 4: Components + categorical SES
m4 <- glm(chronic ~ SES + SES_EDU_SC + SES_BTHR_SC + SES_CAR_SC + SES_INT_SC +
            SES_WRK_SC + SES_BEDR_SC, data = data, family = "binomial")

# Model 5: SES components + interactions
m5 <- glm(chronic ~ SES_EDU_SC * SES_WRK_SC +
                    SES_BTHR_SC * SES_INT_SC +
                    SES_CAR_SC +
                    SES_BEDR_SC, 
          data = data, family = "binomial")

# Each model includes one SES component
m6 <- glm(chronic ~ SES_EDU_SC, data = data, family = "binomial")
m7 <- glm(chronic ~ SES_BTHR_SC, data = data, family = "binomial")
m8 <- glm(chronic ~ SES_CAR_SC, data = data, family = "binomial")
m9 <- glm(chronic ~ SES_INT_SC, data = data, family = "binomial")
m10 <- glm(chronic ~ SES_WRK_SC, data = data, family = "binomial")
m11 <- glm(chronic ~ SES_BEDR_SC, data = data, family = "binomial")
m12 <- glm(chronic ~ SES_EDU_SC + SES_BTHR_SC + SES_CAR_SC +
                     SES_INT_SC + SES_WRK_SC + SES_BEDR_SC,
           data = data, family = "binomial")


aic_table <- AIC(m1, m2, m3, m4, m5, m6, m7, m8, m9, m10 , m11, m12) %>%
  as.data.frame() %>%
  arrange(AIC)

print(aic_table)

m7
m11
m4
m1

best_model <- m7  # or whichever had lowest AIC

tidy_model <- tidy(best_model, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)")

ggplot(tidy_model, aes(x = reorder(term, estimate), y = estimate)) +
  geom_point(size = 4, color = "darkred") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2, color = "gray40") +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray60") +
  coord_flip() +
  labs(title = "Odds Ratios from Logistic Regression",
       subtitle = "Predicting Chronic Condition from SES Factors",
       x = "Predictor",
       y = "Odds Ratio (95% CI)") +
  theme_minimal(base_size = 13)

```

#### SES and PSS - work in progress

```{r}
f<-read.csv("./data/02.HWISE_PSS.csv")
# Add more info to file
dim(f)
f$ID <- as.factor(f$ID)

# Calculate total score PSS per participant
f$t_pss <- 0
# change positive statements to negative values
pss <- c(20, 21, 22, 23, 25, 26, 29)
for (i in pss) {f[i] <- f[i]*-1 }
    for (i in 1:nrow(f)) { 
    sum(f[i, c(18:31)]) -> f$Total_PSS[i]
}
summary(f$Total_PSS)

pss <- merge(m,f, by="ID")
dim(pss)
pss[pss =='NA'] <- NA

agg<- count(pss, SES, chronic, Total_PSS)
agg2 <- pivot_wider(agg, 
            names_from = chronic,
            values_from = n)
agg2$Total <- agg2$Yes/(agg2$Yes + agg2$No)*100

#png(file= "HLTH_counts_by_SES.png", width = 700, height = 800 )
ggplot(agg2, aes(x = SES, y = Total)) +
    geom_bar(stat= "identity", fill = "orchid4", alpha = 0.9) +
     ylim(0, 55) + 
   theme_minimal()  +
   geom_text(aes(label = paste(round(Total, 1), "%", sep="")), 
            vjust = -0.5, 
            colour = "gray40", 
            size = 9) +
   theme(
    plot.title = element_text(hjust = 0.5, 
                              size = 32, face = "bold"),
    axis.title = element_text(size = 22),  
    axis.text = element_text(size = 26)  
  ) + 
  labs(title="Proportion of participants suffering from \n chronic pain or chronic disease") +
  xlab("Socioeconomic status") +
  ylab("Percent (%)")  
#dev.off()


```


# 2. Create Combined Chronic Condition Variable
```{r}
# Create binary variable: 1 = has chronic pain or disease
data <- df %>%
  mutate(chronic_any = ifelse(HLTH_CPAIN_CAT == 1 | HLTH_CDIS_CAT == 1, 1, 0)) %>%
  mutate(chronic_any = factor(chronic_any, labels = c("No chronic condition", "Chronic condition")))
data <- data %>% 
  filter(!is.na(chronic_any)) 
dim(data)
```

# 3. Descriptive stats
```{r}
# Summary stats by chronic condition
data %>%
  group_by(chronic_any) %>%
  summarise(
    mean_pss = mean(PSS_TOTAL, na.rm = TRUE),
    sd_pss = sd(PSS_TOTAL, na.rm = TRUE),
    mean_ses = mean(SES_SC_Total, na.rm = TRUE),
    sd_ses = sd(SES_SC_Total, na.rm = TRUE),
    n = n()
  )

```

4. T-tests / Non-parametric Tests


```{r}
# PSS comparison
wilcox.test(PSS_TOTAL ~ chronic_any, data = data)

# SES comparison
wilcox.test(SES_SC_Total ~ chronic_any, data = data)

```

There is a statistically significant difference in **stress levels** between people with and without chronic conditions.

There is no statistically significant difference in SES between people with and without chronic conditions.

5. Visualizations: Boxplots

```{r}

# Stress and SES by chronic condition
p1 <- ggplot(data, aes(x = chronic_any, y = PSS_TOTAL, fill = chronic_any)) +
  geom_point(aes(alpha = 0.8,  color = chronic_any), position = position_dodge2(width = 0.6)) +
  geom_violin(alpha = 0.3) +
  geom_boxplot(alpha = 0.8, width = 0.15) +
  labs(title = "Perceived Stress (PSS)", y = "PSS Score", x = NULL) +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(data, aes(x = chronic_any, y = SES_SC_Total, fill = chronic_any)) +
  geom_point(aes(alpha = 0.8,  color = chronic_any), position = position_dodge2(width = 0.6)) +
  geom_violin(alpha = 0.3) +
  geom_boxplot(alpha = 0.8, width = 0.15) +
  labs(title = "Socioeconomic Status", y = "SES Score", x = NULL) +
  theme_minimal() +
  theme(legend.position = "none")

# Combine plots
p1 + p2


```

6. Correlations
```{r}
# Correlation matrix
cor_data <- data %>%
  select(PSS_TOTAL, SES_SC_Total, HW_TOTAL, D_AGE) %>%
  drop_na()

cor_matrix <- cor(cor_data)
corrplot(cor_matrix, method = "circle", type = "upper", tl.col = "black")

```


7. Logistic Regression: Chronic condition ~ PSS + SES + Age
```{r}
data <- data %>%
  mutate(chronic_any_binary = ifelse(chronic_any == "Chronic condition", 1, 0))

model <- glm(chronic_any_binary ~ PSS_TOTAL + SES_SC_Total + D_AGE, data = data, family = "binomial")

summary(model)
exp(cbind(OR = coef(model), confint(model)))

```

8. Plot Odds Ratios
```{r}

library(broom)
model_tidy <- tidy(model, conf.int = TRUE, exponentiate = TRUE)

ggplot(model_tidy, aes(x = reorder(term, estimate), y = estimate)) +
  geom_point(size = 4, color = "darkred") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_flip() +
  labs(title = "Odds Ratios for Chronic Condition",
       x = "Predictor", y = "Odds Ratio (95% CI)") +
  theme_minimal()

```

9. Optional: Interaction Term (Stress × SES)

```{r}
interaction_model <- glm(chronic_any_binary ~ PSS_TOTAL * SES_SC_Total + D_AGE,
                         data = data, family = "binomial")

summary(interaction_model)

```




```{r}
library(dplyr)
library(ggplot2)
library(broom)

# Combine chronic pain + chronic disease
data <- data %>%
  mutate(chronic_any = ifelse(HLTH_CPAIN_CAT == 1 | HLTH_CDIS_CAT == 1, 1, 0),
         chronic_any = factor(chronic_any, levels = c(0,1)),
         chronic_any_binary = as.numeric(as.character(chronic_any)))  # for glm

# Model 1: Stress only
m1 <- glm(chronic_any_binary ~ PSS_TOTAL, data = data, family = "binomial")

# Model 2: SES only
m2 <- glm(chronic_any_binary ~ SES_SC_Total, data = data, family = "binomial")

# Model 3: Stress + SES
m3 <- glm(chronic_any_binary ~ PSS_TOTAL + SES_SC_Total, data = data, family = "binomial")

# Model 4: Add age
m4 <- glm(chronic_any_binary ~ PSS_TOTAL + SES_SC_Total + D_AGE, data = data, family = "binomial")

# Model 5: Add water insecurity
m5 <- glm(chronic_any_binary ~ PSS_TOTAL + SES_SC_Total + D_AGE + HW_TOTAL, data = data, family = "binomial")


AIC_table <- AIC(m1, m2, m3, m4, m5) %>%
  as.data.frame() %>%
  arrange(AIC)

print(AIC_table)


best_model <- m5  # or choose based on AIC_table

model_summary <- tidy(best_model, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)")

ggplot(model_summary, aes(x = reorder(term, estimate), y = estimate)) +
  geom_point(size = 4, color = "darkblue") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray60") +
  coord_flip() +
  labs(
    title = "Odds Ratios for Predicting Chronic Condition",
    subtitle = "Best model based on AIC",
    x = "Predictor",
    y = "Odds Ratio (95% CI)"
  ) +
  theme_minimal()

# Interaction: Stress × SES
m_interact <- glm(chronic_any_binary ~ PSS_TOTAL * SES_SC_Total + D_AGE + HW_TOTAL,
                  data = data, family = "binomial")
summary(m_interact)


tidy_interact <- tidy(m_interact, conf.int = TRUE, exponentiate = TRUE) %>%
  filter(term != "(Intercept)")

ggplot(tidy_interact, aes(x = reorder(term, estimate), y = estimate)) +
  geom_point(size = 4, color = "darkred") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray60") +
  coord_flip() +
  labs(
    title = "Odds Ratios with Stress × SES Interaction",
    x = "Predictor",
    y = "Odds Ratio (95% CI)"
  ) +
  theme_minimal()

# Create Predicted Probabilities Across SES Levels
#We’ll generate predicted probabilities of chronic condition at low, medium, and #high SES across a range of PSS (stress) values.

# Create a new data pr predictions
newdata <- expand.grid(
  PSS_TOTAL = seq(min(data$PSS_TOTAL, na.rm = TRUE),
                  max(data$PSS_TOTAL, na.rm = TRUE), length.out = 100),
  SES_SC_Total = quantile(data$SES_SC_Total, probs = c(0.25, 0.5, 0.75), na.rm = TRUE),
  D_AGE = mean(data$D_AGE, na.rm = TRUE),
  HW_TOTAL = mean(data$HW_TOTAL, na.rm = TRUE)
)

# Add a label for SES group
newdata$SES_group <- factor(newdata$SES_SC_Total,
                            labels = c("Low SES", "Medium SES", "High SES"))

# Predict probabilities
newdata$predicted <- predict(m_interact, newdata, type = "response")
ggplot(newdata, aes(x = PSS_TOTAL, y = predicted, color = SES_group)) +
  geom_line(size = 1.2) +
  labs(
    title = "Interaction of Stress and SES on Chronic Condition Risk",
    subtitle = "Predicted probability of chronic condition",
    x = "Perceived Stress Score",
    y = "Predicted Probability",
    color = "SES Level"
  ) +
  theme_minimal(base_size = 14) +
  scale_color_manual(values = c("Low SES" = "firebrick", "Medium SES" = "steelblue", "High SES" = "forestgreen")) +
  theme(legend.position = "top")


```

