---
title: "UROP_Brittany"
output: html_document
date: "2025-03-17"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r echo = FALSE, message = FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(knitr)
library(rstatix)
library(ggpubr)
library(coin)
library(tidyr)
library(MASS)  # For stepAIC
library(broom) # For model output formatting
library(dplyr) # For data manipulation
library(car)   # For VIF check
library(corrplot)
library(rcompanion)
```

```{r}
# loading data set
d <- read.csv("Cleaned_Dataset_Screening_HWISE_PSS_V2.csv")

# taking a quick look at first rows
head(d)

d$HLTH_CDIS_CAT <- as.factor(d$HLTH_CDIS_CAT) # HLTH_CDIS_CAT = chronic disease
# HW_TOTAL = Total HWISE score
# HRS_WEEK = Hours of water supply per week
```

```{r}
d <- d %>%
    filter(!is.na(HLTH_CDIS_CAT)) %>%
  mutate(HLTH_CDIS_CAT = case_when(
    HLTH_CDIS_CAT == "0" ~ "Absent",
    HLTH_CDIS_CAT == "1" ~ "Present",
    HLTH_CDIS_CAT == !is.na(.) ~ "other",  # Assign "other" to all non-missing values that are not "yes" or "no"
    TRUE ~ NA_character_
  ))


# Count the number of data points per Water insecurity level (HWISE)
 summary_stats <- d %>%
  group_by(HLTH_CDIS_CAT) %>%
  summarise(Count = n(), na.rm = TRUE)

 hw.total <- d$HW_TOTAL
 means <- aggregate(hw.total ~ HLTH_CDIS_CAT, d, mean)
means$hw.total <- round(means$hw.total, 2)

color_palette <- c("#ff7f00", "#a6cee3",  "#1f78b4", "orange")


ggplot(d, aes(x = HLTH_CDIS_CAT, 
                 y = HW_TOTAL, 
                 fill = HLTH_CDIS_CAT)) +
  geom_jitter(aes(color = HLTH_CDIS_CAT), 
                size = 1, width = 0.25) +  
  # Jitter adds individual data points
  geom_violin(alpha = 0.6, width = 1) + 
  geom_boxplot(outlier.shape = 1, alpha = 0.5, 
               width = 0.15, color = "grey30") + # Boxplot
  geom_text(data = means, 
             aes(label = hw.total, y = hw.total + 0.5, hjust=-0.7), 
             size = 5, color = "darkred") + #adds average labels
  theme_minimal() +
  labs(#title = "Hours water of supply by water Insecurity group\n(HWISE)",
       x = "Chronic Disease", 
       y = "HWISE score") +
  scale_fill_manual(values = color_palette) +  # Custom colors for boxes
  scale_color_manual(values = color_palette) +  # Custom colors for points +
   stat_summary(fun.y=mean, geom="point", shape=23, 
                size=4, color="darkred", fill="darkred") +
    theme(legend.position = "none", 
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 14),  
        plot.title = element_text(size = 18)) +  
  scale_x_discrete(labels = paste0(summary_stats$HLTH_CDIS_CAT, 
                                 "\n(n=", summary_stats$Count, ")"))

```

```{r}


# Count the number of data points per Water insecurity level (HWISE)
 summary_stats <- d %>%
  group_by(HLTH_CDIS_CAT) %>%
  summarise(Count = n(), na.rm = TRUE)

 hw.total <- d$HRS_WEEK
 means <- aggregate(hw.total ~ HLTH_CDIS_CAT, d, mean)
means$hw.total <- round(means$hw.total, 2)

ggplot(d, aes(x = HLTH_CDIS_CAT, 
                 y = HRS_WEEK, 
                 fill = HLTH_CDIS_CAT)) +
  geom_jitter(aes(color = HLTH_CDIS_CAT), 
                size = 1, width = 0.25) +  
  # Jitter adds individual data points
  geom_violin(alpha = 0.6, width = 1) + 
  geom_boxplot(outlier.shape = 1, alpha = 0.5, 
               width = 0.1, color = "grey30") + # Boxplot
  geom_text(data = means, 
             aes(label = hw.total, y = hw.total + 0.5, hjust=-0.7), 
             size = 5, color = "darkred") + #adds average labels
  theme_minimal() +
  labs(#title = "Hours water of supply by water Insecurity group\n(HWISE)",
       x = "Chronic Disease", 
       y = "Hours of water supply per week") +
  scale_fill_manual(values = color_palette) +  # Custom colors for boxes
  scale_color_manual(values = color_palette) +  # Custom colors for points +
   stat_summary(fun.y=mean, geom="point", shape=23, 
                size=4, color="darkred", fill="darkred") +
    theme(legend.position = "none", 
        axis.text = element_text(size = 13), 
        axis.title = element_text(size = 14),  
        plot.title = element_text(size = 18)) +  
  scale_x_discrete(labels = paste0(summary_stats$HLTH_CDIS_CAT, 
                                 "\n(n=", summary_stats$Count, ")"))  

```


```{r threshold 16}
# Categorize HW_TOTAL into two groups
d <- d %>%
  mutate(HW_TOTAL_category16 = case_when(
    HW_TOTAL >= 16 ~ "High",
    HW_TOTAL < 16 ~ "Low",
  ))

```


```{r}
summary_stats <- d %>%
  group_by(HLTH_CDIS_CAT) %>%
  summarise(Count = n(), na.rm = TRUE, .groups = 'drop')

color_palette <- c("#1f78b4",  "#a6cee3", "#fdbf6f", "#ff7f00")
ggsave("stacked_barplot.png", width = 6, height = 3, dpi = 300) 

ggplot(d, aes(x = HLTH_CDIS_CAT, fill = HW_TOTAL_category16)) +
  geom_bar(position = "fill", alpha = 0.8, width=0.5) +  # "fill" makes the bars proportional (percentages)
  labs(title = "Chronic disease vs. HWISE",
       subtitle = "using HWISE = 16 as threshold",
       x = "Presence/absence disease",
       y = "Percentage",
       fill = "HWISE\nWater insecurity\ncategory") +
   scale_fill_manual(values = color_palette) +
    scale_y_continuous(labels = scales::percent_format(scale = 100)) + 
  theme_minimal() +
   theme( 
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 18),
        plot.margin = margin(1, 1, 1, 1)
       )  +
    scale_x_discrete(labels = paste0(summary_stats$HLTH_CDIS_CAT, 
                                 "\n(n=", summary_stats$Count, ")")) +
  coord_flip() 



summary_stats <- d %>%
  group_by(HW_TOTAL_category16) %>%
  summarise(Count = n(), na.rm = TRUE, .groups = 'drop')

ggplot(d, aes(x = HW_TOTAL_category16, fill = HLTH_CDIS_CAT)) +
  geom_bar(position = "fill", alpha = 0.8, width=0.5) +  # "fill" makes the bars proportional (percentages)
  labs(title = "Chronic disease by water sec\ncategory",
       subtitle = "Threshold is HWISE = 16)",
       x = "Water security levels",
       y = "Percentage",
       fill = "Presence of\nchronic disease") +
   scale_fill_manual(values = color_palette) +
    scale_y_continuous(labels = scales::percent_format(scale = 100)) + 
  theme_minimal() +
   theme( 
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 18),
        plot.margin = margin(1, 1, 1, 1)
       )  +
  
    scale_x_discrete(labels = paste0(summary_stats$HW_TOTAL_category16, 
                                 "\n(n=", summary_stats$Count, ")")) +
  coord_flip() 


```


```{r threshold paper}
d <- d %>%
  filter(!is.na(HW_TOTAL)) %>%  # Remove missing values
  mutate(HW_TOTAL_category4 = case_when(
    HW_TOTAL >= 0 & HW_TOTAL <= 2  ~ "No-to-Marginal",
    HW_TOTAL >= 3 & HW_TOTAL <= 11 ~ "Low",
    HW_TOTAL >= 12 & HW_TOTAL <= 23 ~ "Moderate",
    HW_TOTAL >= 24 & HW_TOTAL <= 36 ~ "High"
  ))

d$HW_TOTAL_category4 <- factor(d$HW_TOTAL_category4, levels = c("No-to-Marginal", "Low", "Moderate", "High"))
```


```{r}
summary_stats <- d %>%
  group_by(HLTH_CDIS_CAT) %>%
  summarise(Count = n(), na.rm = TRUE, .groups = 'drop')

color_palette <- c("#1f78b4",  "#a6cee3", "#fdbf6f", "#ff7f00")
ggsave("stacked_barplot.png", width = 6, height = 3, dpi = 300) 

ggplot(d, aes(x = HLTH_CDIS_CAT, fill = HW_TOTAL_category4)) +
  geom_bar(position = "fill", alpha = 0.8, width=0.5) +  # "fill" makes the bars proportional (percentages)
  labs(title = "Chronic disease vs. HWISE categories",
       subtitle = "Using HWISE 4 categories",
       x = "Presence/absence disease",
       y = "Percentage",
       fill = "HWISE\nWater insecurity\ncategory") +
   scale_fill_manual(values = color_palette) +
    scale_y_continuous(labels = scales::percent_format(scale = 100)) + 
  theme_minimal() +
   theme( 
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 18),
        plot.margin = margin(1, 1, 1, 1)
       )  +
  
    scale_x_discrete(labels = paste0(summary_stats$HLTH_CDIS_CAT, 
                                 "\n(n=", summary_stats$Count, ")")) +
  coord_flip() 



summary_stats <- d %>%
  group_by(HW_TOTAL_category4) %>%
  summarise(Count = n(), na.rm = TRUE, .groups = 'drop')

ggplot(d, aes(x = HW_TOTAL_category4, fill = HLTH_CDIS_CAT)) +
  geom_bar(position = "fill", alpha = 0.8, width=0.5) +  
  labs(title = "HWISE category vs chronic disease",
       subtitle = "",
       x = "Water security levels",
       y = "Percentage",
       fill = "Presence of\nchronic disease") +
   scale_fill_manual(values = color_palette) +
    scale_y_continuous(labels = scales::percent_format(scale = 100)) + 
  theme_minimal() +
   theme( 
        axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        plot.title = element_text(size = 18),
        plot.margin = margin(1, 1, 1, 1)
       )  +
      scale_x_discrete(labels = paste0(summary_stats$HW_TOTAL_category4, 
                                 "\n(n=", summary_stats$Count, ")")) +
  coord_flip() 
```

```{r Picking HWISE categories, echo = FALSE}
#change types of grouping hwise

# Categorize HW_TOTAL into two groups

d <- d %>%
  mutate(HW_TOTAL_category8 = case_when(
    HW_TOTAL >= 8 ~ "High",
    HW_TOTAL < 8 ~ "Low",
  ))

d$HW_TOTAL_category16 <- factor(d$HW_TOTAL_category16, levels = c("Low", "High"))
d$HW_TOTAL_category8 <- factor(d$HW_TOTAL_category8, levels = c("Low", "High"))
d$HW_TOTAL_category4 <- factor(d$HW_TOTAL_category4, levels = c("No-to-Marginal", "Low", "Moderate", "High"))
```

# Testing associations
Contingency table comparing cases in each category 

```{r}
# Create a contingency table
contingency_table16 <- table(d$HW_TOTAL_category16, d$HLTH_CDIS_CAT)
contingency_table8 <- table(d$HW_TOTAL_category8, d$HLTH_CDIS_CAT)
contingency_table4 <- table(d$HW_TOTAL_category4, d$HLTH_CDIS_CAT)
# Print the table
print(contingency_table16)
print(contingency_table8)
print(contingency_table4)
```

## Cramer's V
```{r}
# Compute Cramér's V for effect size
  cat("Compute Cramér's V for effect size:\n 0.1 → Small association\n 0.3 → Moderate association\n 0.5 → Strong association")
cramerV(contingency_table16)
cramerV(contingency_table8)
cramerV(contingency_table4)
```

## Pairwise Fisher's Exact Test
Using Bonferroni correction for multiple comparisons
```{r fisher, echo=TRUE}
# Run Fisher’s Exact Test
fisher_test16 <- fisher.test(contingency_table16)
fisher_test8 <- fisher.test(contingency_table8)
fisher_test4 <- fisher.test(contingency_table4)

# Print results
print(fisher_test16)
print(fisher_test8)
print(fisher_test4)
# Get unique categories of HW_TOTAL_category
categories16 <- levels(d$HW_TOTAL_category16)
categories8 <- levels(d$HW_TOTAL_category8)
categories4 <- levels(d$HW_TOTAL_category4)
# Create an empty list to store results
pairwise_results16 <- list()
pairwise_results8 <- list()
pairwise_results4 <- list()

# Loop through all pairs of HW_TOTAL_category
for (i in 1:(length(categories16) - 1)) {
  for (j in (i + 1):length(categories16)) {
    cat1 <- categories16[i]
    cat2 <- categories16[j]
    
    # Subset the data for the two categories
    subset_data <- d %>%
      filter(HW_TOTAL_category16 %in% c(cat1, cat2))
    
    # Create a 2xN contingency table
    sub_table <- table(subset_data$HW_TOTAL_category16, subset_data$HLTH_CDIS_CAT)
    
    # Run Fisher’s Exact Test
    fisher_res <- fisher.test(sub_table)
    
    # Store results in a data frame
    pairwise_results16[[paste(cat1, "vs", cat2)]] <- data.frame(
      Group1 = cat1,
      Group2 = cat2,
      p_value = fisher_res$p.value
    )
  }
}

# Combine results into a single data frame
pairwise_fisher_results16 <- bind_rows(pairwise_results16)

# Apply Bonferroni correction for multiple comparisons
pairwise_fisher_results16 <- pairwise_fisher_results16 %>%
  mutate(p_adj = p.adjust(p_value, method = "bonferroni"))

# Print results
print(pairwise_fisher_results16)
```
If p_adj < 0.05, it indicates a significant difference 

If p_adj > 0.05, the difference is not statistically significant after Bonferroni correction


```{r}
# Loop through all pairs of HW_TOTAL_category
for (i in 1:(length(categories8) - 1)) {
  for (j in (i + 1):length(categories8)) {
    cat1 <- categories8[i]
    cat2 <- categories8[j]
    
    # Subset the data for the two categories
    subset_data <- d %>%
      filter(HW_TOTAL_category8 %in% c(cat1, cat2))
    
    # Create a 2xN contingency table
    sub_table <- table(subset_data$HW_TOTAL_category8, subset_data$HLTH_CDIS_CAT)
    
    # Run Fisher’s Exact Test
    fisher_res <- fisher.test(sub_table)
    
    # Store results in a data frame
    pairwise_results8[[paste(cat1, "vs", cat2)]] <- data.frame(
      Group1 = cat1,
      Group2 = cat2,
      p_value = fisher_res$p.value
    )
  }
}

# Combine results into a single data frame
pairwise_fisher_results8 <- bind_rows(pairwise_results8)

# Apply Bonferroni correction for multiple comparisons
pairwise_fisher_results8 <- pairwise_fisher_results8 %>%
  mutate(p_adj = p.adjust(p_value, method = "bonferroni"))

# Print results
print(pairwise_fisher_results8)
```


```{r}
# Loop through all pairs of HW_TOTAL_category
for (i in 1:(length(categories4) - 1)) {
  for (j in (i + 1):length(categories4)) {
    cat1 <- categories4[i]
    cat2 <- categories4[j]
    
    # Subset the data for the two categories
    subset_data <- d %>%
      filter(HW_TOTAL_category4 %in% c(cat1, cat2))
    
    # Create a 2xN contingency table
    sub_table <- table(subset_data$HW_TOTAL_category4, subset_data$HLTH_CDIS_CAT)
    
    # Run Fisher’s Exact Test
    fisher_res <- fisher.test(sub_table)
    
    # Store results in a data frame
    pairwise_results4[[paste(cat1, "vs", cat2)]] <- data.frame(
      Group1 = cat1,
      Group2 = cat2,
      p_value = fisher_res$p.value
    )
  }
}

# Combine results into a single data frame
pairwise_fisher_results4 <- bind_rows(pairwise_results4)

# Apply Bonferroni correction for multiple comparisons
pairwise_fisher_results4 <- pairwise_fisher_results4 %>%
  mutate(p_adj = p.adjust(p_value, method = "bonferroni"))

# Print results
print(pairwise_fisher_results4)
```

## Plots Emotion/HWISE category
```{r plots emotion/hwise, echo = FALSE}

# Create the stacked barplot
ggplot(d, aes(x = HW_TOTAL_category16, fill = as.factor(HLTH_CDIS_CAT))) +
  geom_bar(position = "fill") +  # Fill ensures bars are proportional
  labs(title = "Proportion of participants reporting chronic disease, by HWISE category",
       subtitle = "using 16 as threshold",
       x = "HWISE Category",
       y = "Proportion",
       fill = "Disease presence") +
#  scale_y_continuous(labels = scales::percent) +
  theme_minimal()

# Create the stacked barplot
ggplot(d, aes(x = HW_TOTAL_category8, fill = as.factor(HLTH_CDIS_CAT))) +
  geom_bar(position = "fill") +  # Fill ensures bars are proportional
  labs(title = "Proportion of participants reporting chronic disease, by HWISE category",
              subtitle = "using 8 as threshold",
       x = "HWISE Category",
       y = "Proportion",
       fill = "Disease presence") +
#  scale_y_continuous(labels = scales::percent) +
  theme_minimal()

# Create the stacked barplot
ggplot(d, aes(x = HW_TOTAL_category4, fill = as.factor(HLTH_CDIS_CAT))) +
  geom_bar(position = "fill") +  # Fill ensures bars are proportional
  labs(title = "Proportion of participants reporting chronic disease, by HWISE category",
              subtitle = "using 4 HWISE groups",
       x = "HWISE Category",
       y = "Proportion",
       fill = "Disease presence") +
#  scale_y_continuous(labels = scales::percent) +
  theme_minimal()

```

# Regressions

Why Logistic Regression?
Dependent Variable is categorical (Nominal variable with multiple categories).
Independent Variable (HW_TOTAL_category) is categorical.
Logistic regression is appropriate for categorical outcomes, unlike linear regression which assumes continuous data.

Need to run binary logistic regression.

```{r}
# Ensure variables are factors

#d <- d %>%
 # mutate(HW_TOTAL_category = as.factor(HW_TOTAL_category),
  #       HLTH_CDIS_CAT = as.factor(HLTH_CDIS_CAT))

d <- d %>%
  mutate(HLTH_CDIS_CAT = case_when(
    HLTH_CDIS_CAT == "Absent" ~ "0",
    HLTH_CDIS_CAT == "Present" ~ "1",
    HLTH_CDIS_CAT == !is.na(.) ~ "other",  # Assign "other" to all non-missing values that are not "yes" or "no"
    TRUE ~ NA_character_
  ))

d$HLTH_CDIS_CAT <- as.numeric(d$HLTH_CDIS_CAT)

# Run Binary Logistic Regression
logit_model1.16 <- glm(HLTH_CDIS_CAT ~ HW_TOTAL_category16, 
                   data = d, 
                   family = binomial)

# Show summary of model
summary(logit_model1.16)


# Run Binary Logistic Regression
logit_model1.8 <- glm(HLTH_CDIS_CAT ~ HW_TOTAL_category8, 
                   data = d, 
                   family = binomial)

# Show summary of model
summary(logit_model1.8)


# Run Binary Logistic Regression
logit_model1.4 <- glm(HLTH_CDIS_CAT ~ HW_TOTAL_category4, 
                   data = d, 
                   family = binomial)

# Show summary of model
summary(logit_model1.4)



logit_model2 <- glm(HLTH_CDIS_CAT ~ HRS_WEEK + SES_SC_Total + SEASON, 
                   data = d, 
                   family = binomial)
summary(logit_model2)


logit_model3 <- glm(HLTH_CDIS_CAT ~ HRS_WEEK, 
                   data = d, 
                   family = binomial)
summary(logit_model3)

```


```{r}
# Compute odds ratios
exp(coef(logit_model1.16))
exp(coef(logit_model1.8))
exp(coef(logit_model1.4))
exp(coef(logit_model2)) # these are not valid (non-categorical predictors)
exp(coef(logit_model3)) # these are not valid
``` 
✔ Odds ratios < 1 → Decreased likelihood

✔ Odds ratios > 1 → Increased likelihood




# Regressions

## 1.b Variable descriptions for quick reference

Ordered alphabetically

```{r table info variables, echo=FALSE, results='asis'}
# Load necessary package

# Create a data frame with variable descriptions, classes, and additional details
variables_info <- data.frame(
  Variable = c("D_AGE", "D_CHLD", "D_HH_SIZE", "D_LOC_TIME", "HLTH_CDIS_CAT", "HLTH_CPAIN_CAT", 
               "HLTH_SMK", "HRS_WEEK", "HW_TOTAL", "MX28_WQ_COMP", "PSS_TOTAL", "SEASON", 
               "SES_SC_Total", "W_WS_LOC", "W_WC_WI"),
  Description = c("Participants' age", 
                  "Number of children participant has birthed",
                  "Household size",
                  "For how long have you lived in this neighborhood?",
                  "Presence of chronic disease",
                  "Presence of chronic pain",
                  "Tobacco smoker",
                  "Hours of water supply in the household per week",
                  "Sum of all 12-items in HWISE questionnaire",
                  "Perception of water service as worse, same, or better than rest of Mexico City",
                  "Total Perceived Stress Score",
                  "Fall or Spring (when data collection happened)",
                  "Socioeconomic status score",
                  "Classification of neighborhoods as water secure or insecure",
                  "Classification of water supply as continuous or intermittent"),
  Class = c("Numeric", "Numeric", "Numeric", "Numeric", "Categorical (Binary)", "Categorical (Binary)", 
            "Categorical (Binary)", "Numeric", "Numeric", "Categorical (Ordinal)", "Numeric", "Categorical (Binary)",
            "Numeric", "Categorical (Binary)", "Categorical (Binary)"),
  Values = c("18:49", 
              "0:8", 
              "2:40", 
              "1:46 (years)", 
              "1 = yes, 0 = no", 
              "1 = yes, 0 = no", 
              "1 = yes, 0 = no", 
              "0:168", 
              "0:27", 
              "0 = worse, 1 = same, 2 = better", 
              "-19:19", 
              "Fall = 1, Spring = 0", 
              "25:263", 
              "1 = water insecure, 0 = water secure", 
              "1 = intermittent, 0 = continuous")
)

# Print the table using kable
kable(variables_info, caption = "Variable Descriptions, Classes, and Additional Details")

```

## 2 Data preparation

1. We remove rows with missing data.

2. HW_TOTAL is calculated by adding up all the HWISE scores; PSS_TOTAL is calculated by adding up PSS 1,2,3, 8, 11, 12, 14, and substracting 4,5,6,7,9,10, and 13.

```{r function to count missing info}
# Function to count missing information per varaible
create_na_table <- function(data) {
  # Create a data frame with variable names and count of missing values
  na_info <- data.frame(
    Variable = names(data),
    Missing_Values = colSums(is.na(data))
  ) %>%
    arrange(desc(Missing_Values))
    # Generate a kable table
  kable(na_info, caption = "Number of Missing Values per Variable")
}
```

```{r read-data}

# Load the dataset
data_path = "data"

data <- read.csv(file.path(data_path, "Cleaned_Dataset_Screening_HWISE_PSS_V4.csv"), 
         stringsAsFactors = FALSE,  
         na.strings = c("", "N/A", "NA", "pending"))

```

## 3 Results

### 3.1 HWISE scores, variable set 1

The regression results for HW is summarized as follows.

```{r echo = TRUE}

# Check unique values to confirm only 0 and 1 remain
table(data$HLTH_CDIS_CAT)
dim(data)

# Define dependent (outcome) variable
outcome_var <- "HLTH_CDIS_CAT"  # Binary outcome (0 = negative, 1 = positive)

# Define independent (predictor) variables
predictors <- c("MX8_TRUST", "MX28_WQ_COMP", "MX26_EM_HHW_TYPE", "D_LOC_TIME", "D_AGE", "D_HH_SIZE", "HLTH_SMK", "D_CHLD", "SES_SC_Total", "SEASON", "W_WS_LOC", "HW_TOTAL", "W_WC_WI", "HRS_WEEK", "PSS_TOTAL", "HW_WORRY")

# Run univariate logistic regression for each predictor
univariate_results <- list()
aic_values <- data.frame(Predictor = character(), AIC = numeric(), stringsAsFactors = FALSE)

data <- data %>% drop_na(all_of(predictors))
dim(data)

for (var in predictors) {
    # Build formula dynamically
    formula <- as.formula(paste(outcome_var, "~", var))
    
    # Fit univariate logistic regression model
    model <- glm(formula, data = data, family = binomial)
    
    # Store model summary
    univariate_results[[var]] <- summary(model)
    
    # Store AIC values
    aic_values <- rbind(aic_values, data.frame(Predictor = var, AIC = AIC(model)))
}

univariate_results

# Display AIC values sorted in ascending order (lower AIC = better fit)
aic_values <- aic_values %>% arrange(AIC)
print(aic_values)

# Select best predictors based on AIC (top 6 with lowest AIC)
top_predictors <- head(aic_values$Predictor, 9)

# Create multivariable model with top predictors
multivariable_formula <- as.formula(paste(outcome_var, "~", paste(top_predictors, collapse = " + ")))
multivariable_model <- glm(multivariable_formula, data = data, family = binomial)

# Display summary of multivariable model
summary(multivariable_model)

# Perform stepwise selection to find the best model
optimized_model <- stepAIC(multivariable_model, direction = "both")

# Display summary of optimized model
summary(optimized_model)

```


